### 2010-02-25 Paper/Paper.tex
preparing ISMIR paper; writing things down helps me

% -----------------------------------------------
% Template for ISMIR 2010
% (based on earlier ISMIR templates)
% -----------------------------------------------

\documentclass{article}
\usepackage{ismir2010,amsmath,cite}
\usepackage{graphicx}

% Title.
% ------
\title{PAPER TEMPLATE FOR ISMIR 2010}

% Single address
% To use with only one author or several with the same address
% ---------------
%\oneauthor
% {Names should be omitted for double-blind reviewing}
% {Affiliations should be omitted for double-blind reviewing}

% Two addresses
% --------------
%\twoauthors
%  {First author} {School \\ Department}
%  {Second author} {Company \\ Address}

% Three addresses
% --------------
\threeauthors
  {First author} {Affiliation1 \\ {\tt author1@ismir.edu}}
  {Second author} {\bf Retain these fake authors in\\\bf submission to preserve the formatting}
  {Third author} {Affiliation3 \\ {\tt author3@ismir.edu}}

\begin{document}
%
\maketitle
%
\begin{abstract}
The abstract should be placed at the top left column and should contain about 150-200 words.
\end{abstract}
%
\section{Introduction}\label{sec:introduction}

This template includes all the information about formatting manuscripts for the ISMIR 2010. 
Please follow these guidelines to give the final proceedings a uniform look. 
If you have any questions, please contact the Conference Management.
This template can be downloaded from the ISMIR 2010 web site (http://ismir2010.ismir.net).

\section{Page Size}\label{sec:page_size}

The proceedings will be printed on 
 \underline{portrait A4-size paper} \underline{(21.0cm x 29.7cm)}. 
All material on each page should fit within a rectangle of 17.2cm x 25.2cm, 
centered on the page, beginning 2.0cm 
from the top of the page and ending with 2.5cm from the bottom. 
The left and right margins should be 1.9cm. 
The text should be in two 8.2cm columns with a 0.8cm gutter. 
All text must be in a two-column format. 
Text must be fully justified.

\section{Typeset Text}\label{sec:typeset_text}

\subsection{Normal or Body Text}\label{subsec:body}

Please use a 10pt (point) Times font. Sans-serif or non-proportional fonts 
can be used only for special purposes, such as distinguishing source code text.

The first paragraph in each section should not be indented, but all other paragraphs should be.

\subsection{Title and Authors}

The title is 14pt Times, bold, caps, upper case, centered.
Authors' names are omitted when submitting for double-blind reviewing. 
The following is for making a camera-ready version. 
Authors' names are centered. 
The lead author's name is to be listed first (left-most), and the co-authors' names after. 
If the addresses for all authors are the same, include the address only once, centered. 
If the authors have different addresses, put the addresses, evenly spaced, under each authors' name.

\subsection{First Page Copyright Notice}

Please include the copyright notice exactly as it appears here in the lower left-hand corner of the page. 
It is set in 8pt Times.

\subsection{Page Numbering, Headers and Footers}

Do not include headers, footers or page numbers in your submission. 
These will be added when the publications are assembled.

\section{First Level Headings}

First level headings are in Times 10pt bold, 
centered with 1 line of space above the section head, and 1/2 space below it. 
For a section header immediately followed by a subsection header, the space should be merged.

\subsection{Second Level Headings}

Second level headings are in Times 10pt bold, flush left, 
with 1 line of space above the section head, and 1/2 space below it. 
The first letter of each significant word is capitalized.

\subsubsection{Third and Further Level Headings}

Third level headings are in Times 10pt italic, flush left, 
with 1/2 line of space above the section head, and 1/2 space below it. 
The first letter of each significant word is capitalized.

Using more than three levels of headings is highly discouraged.

\section{Footnotes and Figures}

\subsection{Footnotes}

Indicate footnotes with a number in the text.\footnote{This is a footnote.} 
Use 8pt type for footnotes. Place the footnotes at the bottom of the page on which they appear. 
Precede the footnote with a 0.5pt horizontal rule.

\subsection{Figures, Tables and Captions}

All artwork must be centered, neat, clean, and legible. 
All lines should be very dark for purposes of reproduction and art work should not be hand-drawn. 
The proceedings are not in color, and therefore all figures must make sense in black-and-white form. 
Figure and table numbers and captions always appear below the figure. 
Leave 1 line space between the figure or table and the caption. 
Each figure or table is numbered consecutively. Captions should be Times 10pt. 
Place tables/figures in text as close to the reference as possible. 
References to tables and figures should be capitalized, for example: 
see \figref{fig:example} and \tabref{tab:example}. 
Figures and tables may extend across both columns to a maximum width of 17.2cm.

\begin{table}
 \begin{center}
 \begin{tabular}{|l|l|}
  \hline
  String value & Numeric value \\
  \hline
  Hello ISMIR  & 2010 \\
  \hline
 \end{tabular}
\end{center}
 \caption{Table captions should be placed below the table.}
 \label{tab:example}
\end{table}

\begin{figure}
 \centerline{\framebox{
 \includegraphics[width=\columnwidth]{figure.png}}}
 \caption{Figure captions should be placed below the figure.}
 \label{fig:example}
\end{figure}

\section{Equations}

Equations should be placed on separated lines and numbered.
The number should be on the right side, in parentheses.

\begin{equation}
E=mc^{2}
\end{equation}

\section{Citations}

All bibliographical references should be listed at the end, 
inside a section named ``REFERENCES,'' numbered and in alphabetical order. 
Also, all references listed should be cited in the text. 
When referring to a document, type the numbering square brackets
\cite{Author:00} or \cite{Author:00,Someone:10,Someone:04}.

\begin{thebibliography}{citations}

\bibitem {Author:00} 
E. Author:
``The Title of the Conference Paper,''
{\it Proceedings of the International Symposium 
on Music Information Retrieval}, pp.~000--111, 2000.

\bibitem{Someone:10} 
A. Someone, B. Someone, and C. Someone:
``The Title of the Journal Paper,''
{\it Journal of New Music Research}, 
Vol.~A, No.~B, pp.~111--222, 2010.

\bibitem{Someone:04} 
X. Someone and Y. Someone:
{\it Title of the Book},
Editorial Acme, Utrecht, 2004.

\end{thebibliography}

%\bibliography{ismir2010template}

\end{document}

### 2010-02-25 Paper/Paper.tex
putting pieces together

@@
 \usepackage{graphicx}
+\usepackage{url}
 
@@
 % ------
-\title{PAPER TEMPLATE FOR ISMIR 2010}
+\title{Large-scale harmonic patterns clustering}
 
@@
 \begin{abstract}
-The abstract should be placed at the top left column and should contain about 150-200 words.
+We present a method for discovering typical patterns in music.
+Similar to the \textit{shingle} idea, but done on a larger scale through
+Echo Nest API, we cluster harmonic patterns from 43K songs. Patterns are
+made of ordered chroma features, and not of summarization like gaussian
+mixtures coefficients or NMF weights. We explain
+how to operate on such large datasets, we analyze the obained clusters,
+and we discuss there use for retrieval and encoding. All code is made
+available.
 \end{abstract}
@@
 
-\section{Page Size}\label{sec:page_size}
+\section{Previous Work}\label{sec:prevwork}
+The idea is closely related to the \textit{Shingles} described by
+Slaney and Casey \cite{Casey2006,Casey2007,Casey2008}. LSH: \cite{E2LSH}.
+Barrington et al. recently studied music texture using a set of ordered
+chroma vectors \cite{Barrington2009a}. 
+Nearest neighbor for music also includes \cite{Cano2004} and ...
 
-The proceedings will be printed on 
- \underline{portrait A4-size paper} \underline{(21.0cm x 29.7cm)}. 
-All material on each page should fit within a rectangle of 17.2cm x 25.2cm, 
-centered on the page, beginning 2.0cm 
-from the top of the page and ending with 2.5cm from the bottom. 
-The left and right margins should be 1.9cm. 
-The text should be in two 8.2cm columns with a 0.8cm gutter. 
-All text must be in a two-column format. 
-Text must be fully justified.
+\section{Data}\label{sec:data}
 
-\section{Typeset Text}\label{sec:typeset_text}
+\subsection{Echo Nest features}
+Echo Nest analyze API: \cite{EchoNest}.
 
-\subsection{Normal or Body Text}\label{subsec:body}
+\subsection{Cowbell Dataset}
 
-Please use a 10pt (point) Times font. Sans-serif or non-proportional fonts 
-can be used only for special purposes, such as distinguishing source code text.
+\subsection{uspop2002}
+uspop: \cite{uspop2002}.
 
-The first paragraph in each section should not be indented, but all other paragraphs should be.
+\section{Algorithm}\label{sec:algo}
 
-\subsection{Title and Authors}
+Vector quantization. Online learning.
 
-The title is 14pt Times, bold, caps, upper case, centered.
-Authors' names are omitted when submitting for double-blind reviewing. 
-The following is for making a camera-ready version. 
-Authors' names are centered. 
-The lead author's name is to be listed first (left-most), and the co-authors' names after. 
-If the addresses for all authors are the same, include the address only once, centered. 
-If the authors have different addresses, put the addresses, evenly spaced, under each authors' name.
+\section{Experiments}\label{sec:experiments}
 
-\subsection{First Page Copyright Notice}
+\section{Conclusion}
+We are awesome. Here's how:
+\begin{itemize}
+\item large-scale
+\item smarter shingles, based on beats and bars
+\item clusters good for encoding
+\item clusters good for something else?
+\item free online API, code available
+\end{itemize}
 
-Please include the copyright notice exactly as it appears here in the lower left-hand corner of the page. 
-It is set in 8pt Times.
+\small
+\section{Acknowledgements}
+Thierry is NSERC graduate fellow, or some title like that.
 
-\subsection{Page Numbering, Headers and Footers}
+%\begin{thebibliography}{citations}
+%\bibitem{Someone:04} 
+%X. Someone and Y. Someone:
+%{\it Title of the Book},
+%Editorial Acme, Utrecht, 2004.
+%\end{thebibliography}
 
-Do not include headers, footers or page numbers in your submission. 
-These will be added when the publications are assembled.
-
-\section{First Level Headings}
-
-First level headings are in Times 10pt bold, 
-centered with 1 line of space above the section head, and 1/2 space below it. 
-For a section header immediately followed by a subsection header, the space should be merged.
-
-\subsection{Second Level Headings}
-
-Second level headings are in Times 10pt bold, flush left, 
-with 1 line of space above the section head, and 1/2 space below it. 
-The first letter of each significant word is capitalized.
-
-\subsubsection{Third and Further Level Headings}
-
-Third level headings are in Times 10pt italic, flush left, 
-with 1/2 line of space above the section head, and 1/2 space below it. 
-The first letter of each significant word is capitalized.
-
-Using more than three levels of headings is highly discouraged.
-
-\section{Footnotes and Figures}
-
-\subsection{Footnotes}
-
-Indicate footnotes with a number in the text.\footnote{This is a footnote.} 
-Use 8pt type for footnotes. Place the footnotes at the bottom of the page on which they appear. 
-Precede the footnote with a 0.5pt horizontal rule.
-
-\subsection{Figures, Tables and Captions}
-
-All artwork must be centered, neat, clean, and legible. 
-All lines should be very dark for purposes of reproduction and art work should not be hand-drawn. 
-The proceedings are not in color, and therefore all figures must make sense in black-and-white form. 
-Figure and table numbers and captions always appear below the figure. 
-Leave 1 line space between the figure or table and the caption. 
-Each figure or table is numbered consecutively. Captions should be Times 10pt. 
-Place tables/figures in text as close to the reference as possible. 
-References to tables and figures should be capitalized, for example: 
-see \figref{fig:example} and \tabref{tab:example}. 
-Figures and tables may extend across both columns to a maximum width of 17.2cm.
-
-\begin{table}
- \begin{center}
- \begin{tabular}{|l|l|}
-  \hline
-  String value & Numeric value \\
-  \hline
-  Hello ISMIR  & 2010 \\
-  \hline
- \end{tabular}
-\end{center}
- \caption{Table captions should be placed below the table.}
- \label{tab:example}
-\end{table}
-
-\begin{figure}
- \centerline{\framebox{
- \includegraphics[width=\columnwidth]{figure.png}}}
- \caption{Figure captions should be placed below the figure.}
- \label{fig:example}
-\end{figure}
-
-\section{Equations}
-
-Equations should be placed on separated lines and numbered.
-The number should be on the right side, in parentheses.
-
-\begin{equation}
-E=mc^{2}
-\end{equation}
-
-\section{Citations}
-
-All bibliographical references should be listed at the end, 
-inside a section named ``REFERENCES,'' numbered and in alphabetical order. 
-Also, all references listed should be cited in the text. 
-When referring to a document, type the numbering square brackets
-\cite{Author:00} or \cite{Author:00,Someone:10,Someone:04}.
-
-\begin{thebibliography}{citations}
-
-\bibitem {Author:00} 
-E. Author:
-``The Title of the Conference Paper,''
-{\it Proceedings of the International Symposium 
-on Music Information Retrieval}, pp.~000--111, 2000.
-
-\bibitem{Someone:10} 
-A. Someone, B. Someone, and C. Someone:
-``The Title of the Journal Paper,''
-{\it Journal of New Music Research}, 
-Vol.~A, No.~B, pp.~111--222, 2010.
-
-\bibitem{Someone:04} 
-X. Someone and Y. Someone:
-{\it Title of the Book},
-Editorial Acme, Utrecht, 2004.
-
-\end{thebibliography}
-
-%\bibliography{ismir2010template}
+\bibliography{tbm_bib}
 

### 2010-02-25 Paper/Paper.tex
algo added

@@
 \usepackage{url}
+\usepackage{algorithm,algorithmic}
 
@@
   {Third author} {Affiliation3 \\ {\tt author3@ismir.edu}}
+% what order do we use? Thierry, Ron, Dan? is Juan on?
 
@@
 \subsection{Cowbell Dataset}
+We have $3,720,091$ non zero bars, a bar usually contains
+$4$ beats.
 
@@
 Vector quantization. Online learning.
+
+\subsection{Codebook Learning}
+We initialize the codebook by choosing $K$ random points from our dataset.
+
+\begin{algorithm}
+\caption{Pseudocode of Vector Quantization}
+\label{alg:vq}
+\begin{algorithmic}
+\STATE$l$ learning rate
+\STATE$\{P_n\}$ set of patches
+\STATE$\{C_k\}$ codebook of $K$ codes
+%\STATE $m \leftarrow b$
+\REQUIRE $0 < l \leq 1$
+\FOR{$nIters$}
+\FOR{$p \in \{P_n\}$}
+\STATE$c \leftarrow min_{c \in C_k} dist(p,c)$
+\STATE$c \leftarrow c + (p - c) * l$
+\ENDFOR
+\ENDFOR
+\RETURN $u$
+\end{algorithmic}
+\end{algorithm}
 
@@
 \item clusters good for something else?
-\item free online API, code available
+\item free online API, code available\footnote{code not released yet to preserve submission's anonymity}
+\item we have lab t-shirts
 \end{itemize}

### 2010-02-25 PythonSrc/VQutils.py
removed assert

@@ -24,7 +24,6 @@ def euclidean_dist(a,b):
     #return np.sqrt(np.square(a-b).sum())
     # following faster, but does not work on boar?
     #return np.sqrt(np.dot((a-b),(a-b).T)[0][0])
-    assert((a-b).size == a.size)
     return np.sqrt(np.square(a-b).sum())
 
 def euclidean_norm(a):
@@ -71,7 +70,6 @@ def encode_oneiter(signal,codebook):
     dists = [euclidean_dist(signal,r) for r in codebook]
     # return the index, scaling, and distance for the MIN DISTANCE
     idxs = np.argsort(dists)
-    assert( dists[idxs[0]] == min(dists) )
     return idxs,dists
 
 

### 2010-02-26 PythonSrc/VQutils.py
scaling now false by default

@@ -287,7 +287,7 @@ def get_codes_ordering(best_code_per_p, nCodes):
     return np.flipud(np.argsort(freqs.flatten())), freqs.flatten()
 
 
-def find_best_code_per_pattern(dataset,codebook,scale=True):
+def find_best_code_per_pattern(dataset,codebook,scale=False):
     """
     Find the best code for each pattenr in the dataset.
     Dataset: one code per row (samething for codebook)
@@ -308,7 +308,7 @@ def find_best_code_per_pattern(dataset,codebook,scale=True):
     return best_code_per_p
 
 
-def online_vq(feats,K,lrate,nIter=10,thresh=0.0000001,maxRise=.05,scale=True,repulse=False):
+def online_vq(feats,K,lrate,nIter=10,thresh=0.0000001,maxRise=.05,scale=False,repulse=False):
     """
     Online vector quantization
     INPUT:

### 2010-02-26 PythonSrc/VQutils.py
LLE visualization works, some fancy stuff to be added, like specifying which code should be displayed; but proof of concept works

@@ -4,7 +4,7 @@ Set of functions to perform vector quantization,
 meaning find a set of 'typical patterns', get them in
 a codebook, then reencode new vectors
 
-Relies in part on the scipy.cluster.vq package
+Relies less and less on the scipy.cluster.vq package
 and scipy.spatial.distance
 
 T. Bertin-Mahieux (2010) Columbia University
@@ -306,6 +306,87 @@ def find_best_code_per_pattern(dataset,codebook,scale=False):
         best_code_per_p[k] = idxs[0]
     # done
     return best_code_per_p
+
+
+def add_image(P,im,x,y,width=.15):
+    """
+    Used by LLE_my_codebook to add a specific image to a given plot
+    I should start my plotting library... or still Ron's...
+
+    Image is a matrix, (x,y) is in data coord, image is centered
+
+    INPUT:
+      - P      pylab object (P in LLE_my_codebook)
+      - im     matrix representing the image
+      - x      x position in data coord
+      - y      y position in data coord
+      - width  width in fig size, height automatically found
+    """
+    # current axes
+    curr_axes = P.gca()
+    # current axis in data coord
+    (minX,maxX,minY,maxY) = P.axis()
+    # get placement of mainx axes in figure
+    bbox = curr_axes.bbox._bbox
+    fX = bbox.x0
+    fY = bbox.y0
+    fWidth = bbox.x1 - bbox.x0
+    fHeight = bbox.y1 - bbox.y0
+    # find pos relative to main axes
+    relX = (x - minX) * 1. / (maxX-minX)
+    relY = (y - minY) * 1. / (maxY-minY)
+    # find pos relative to main figure
+    relX = (relX * fWidth) + fX
+    relY = (relY * fHeight) + fY
+    # height relative to width
+    height = im.shape[0] * 1. / im.shape[1] * width
+    # create new axis
+    a = P.axes([relX-width/2.,relY-height/2.,width,height])
+    # set to x and y ticks
+    P.setp(a, xticks=[], yticks=[])
+    # plot image, new axes is the current default
+    P.imshow(im,interpolation='nearest',origin='lower')
+    # set back axes
+    P.axes(curr_axes)
+
+
+
+def LLE_my_codebook(codebook,nNeighbors=5):
+    """
+    Performs LLE on the codebook
[…]

### 2010-02-26 Paper/Paper.tex
improving...

@@
 
-This template includes all the information about formatting manuscripts for the ISMIR 2010. 
-Please follow these guidelines to give the final proceedings a uniform look. 
-If you have any questions, please contact the Conference Management.
-This template can be downloaded from the ISMIR 2010 web site (http://ismir2010.ismir.net).
+The original goal of the \textit{Shingles} was retrieval, i.e. find very
+similar song segments based on their harmonic pattern. Part of the
+challenge was to handle tons of overlapping song segments represented
+by a rather long feature vector (a serie of chroma features).
+The LSH algorithm proved a great solution for performing a $k$-nearest
+neighbors ($k$nn) search in such a space. However, the paradigm of $k$nn has
+an intrinsec scaling limit.
+
+Finding a closest match is a tangible goal within a personal collection
+of a few thousands songs. It can be used to discover cover songs or
+find the seamless transitions in a playlist. However, the closest match
+in millions and millions of music excerpts do not help us organize such
+an amount of media. Some tasks are still important, for instance cover
+recognition, and improvements in $k$nn methods are always welcome.
+But it does not the describe the music space, nor does it necessarily
+make good recommendations. A similar argument in the recommendation
+domain is that precision becomes more important than recall passed a certain
+scale of data for many practical uses.
+
+We propose to move the \textit{Shingle} idea into the larger clustering paradigm.
+The goal is to find typical patterns that can be seen as cluster centroids.
+Those typical patterns help us understand a music collection from an
+harmonic point of view. Segments of songs can be described in regard to their
+proximity to the different centroids, thus creating some notion of distance
+among the song segments. These cluster of similar patterns can have many usages:
+\begin{enumerate}
+\item describing the music space
+\item finding potential similar songs and/or covers
+\item eliminating quickly very wrong matches in a query system
+\end{enumerate}
+
+We present a practical way to do such large-scale clustering using vector
+quantization. We present clustering results of an online algorithm on a data 
+set of $43$K songs, but the method applies to any data size. We show how to
+select different parameters such as the size of the \textit{Shingles} 
+using an encoding scheme on a test set of $4$K songs. Finally, we illustrate
+the method with some specific examples.
 
@@
 The idea is closely related to the \textit{Shingles} described by
-Slaney and Casey \cite{Casey2006,Casey2007,Casey2008}. LSH: \cite{E2LSH}.
+Slaney and Casey \cite{Casey2006,Casey2007,Casey2008}. 
+LSH implementation: \cite{E2LSH}.
 Barrington et al. recently studied music texture using a set of ordered
-chroma vectors \cite{Barrington2009a}. 
+MFCC vectors \cite{Barrington2009a}. 
 Nearest neighbor for music also includes \cite{Cano2004} and ...
+
+Other large-scale work use summarization of the features. We mention 
+gaussian mixtures \cite{Mandel2005}, aggregate features \cite{Bergstra2006a}
+and linear dynamical systems \cite{Barrington2009a}. The chroma representation
+is close from an harmonic point of view to the actual music, in particular
+we can ``listen'' to the patches (see Section \ref{sec:experiments}).
+This is an advantage for analyzing a large data set, and in accordance with our goal
+we restrain ourselves from using generative models to compress our features.
 
@@
 \item clusters good for something else?
-\item free online API, code available\footnote{code not released yet to preserve submission's anonymity}
+\item free online API, code available\footnote{code not released yet to preserve 
+submission's anonymity}
 \item we have lab t-shirts
 \end{itemize}
+
+As for enhancements, we specifically did not compress our features using
+gaussian mixtures or other generative model. That being said, these methods
+could be use to develop better distance measures between \textit{Shingles}.
+The use of the euclidean distance is arbitrary. Summarizing patches
+with gaussians, and then comparing the distance between those gaussians,
+could reduce the influence of the noise in the distance measure.
 

### 2010-02-26 Paper/Paper.tex
improved

@@
 Slaney and Casey \cite{Casey2006,Casey2007,Casey2008}. 
-LSH implementation: \cite{E2LSH}.
+LSH: \cite{Datar2004}, LSH implementation: \cite{E2LSH}.
 Barrington et al. recently studied music texture using a set of ordered
@@
 \subsection{Echo Nest features}
-Echo Nest analyze API: \cite{EchoNest}.
+We use the Echo Nest analyze API \cite{EchoNest} throughout this work.
+The API gives us a chroma vector (length $12$) for every music event, 
+or ``segment'', for any song we upload to their platform. 
+We also get an estimate of where the beats and the bars are. 
+We transform the chroma vectors per segment into per beat vector using a 
+simple average. We can then stack beat vectors for a number of bars (usually 1 or 2). 
+We get a fixed-size patch by resampling the number of chroma vectors. Typically,
+there are $4$ beats per bar. Thus an appropriate patch size for one bar
+is $12$x$4$ or $12$x$8$.
+
+Note that we do not claim that any of these informations (segments, beats, bars)
+are perfectly accurate. Practise showed us that they are reasonable, and the
+size of the data set should make up for the imperfections or noise.
+We also believe that patches sizes based on a number of beats or bars are more
+meaningful than an arbitrary time length. More on this in the experiments
+(Section \ref{sec:experiments}).
+
+\begin{figure}[htb]
+\begin{center}
+\includegraphics[width=.8\columnwidth]{code}
+\end{center}
+\caption{{A typical code from a code book of size $200$. Patterns represented
+$2$ bars and the pattern length was set to $16$.}}
+\label{fig:code}
+\end{figure}
 
@@
 \subsection{uspop2002}
-uspop: \cite{uspop2002}.
+We had acces to a low quality (32kbps) version of the songs in the uspop 2002 data
+set \cite{uspop2002}.
+This well-known data show a great diversity of pop songs, but is not particular
+otherwise. We also tried to use Tzanetakis dataset \cite{Tzanetakis2002a}, but 
+the $30$ seconds segments seemed hard to analyze by the Echo Nest API, most did
+not contain any bar.
 
@@
 
-Vector quantization. Online learning.
+Vector quantization \cite{Gersho1991}. Online learning.
 
@@
 \caption{Pseudocode of Vector Quantization}
-\label{alg:vq}
+\label{algo:vq}
 \begin{algorithmic}
@@
 \ENDFOR
-\RETURN $u$
+\RETURN $\{C_k\}$
 \end{algorithmic}
@@
 \section{Experiments}\label{sec:experiments}
+
+\subsection{Setting}
+We take one or two bars, normalize the patches to size 4, 8, or 16.
+We roll the patch to be invariant to the key (on the patch level, not on
+the song level). We learn a codebook of size $K$ over the cowbell dataset 
+using the VQ algorithm (Algorithm \ref{algo:vq}). A typical learning rate 
+is $1e-2$ for $20$ iterations over the whole dataset.
+
+We use the codebook to encode our testing set (see Section \ref{sec:data}).
+Each pattern is encoded with only one code. We can measure the average
+distance between a pattern and its encoding. We can also measure the use
+of the different codes.
+
+\subsection{Pattern length}
+
+\subsection{Visualization}
+
+\begin{figure}[htb]
+\begin{center}
+\includegraphics[width=.9\columnwidth]{codes_lle}
+\end{center}
+\caption{{LLE visualization of the code book.}}
+\label{fig:lle}
+\end{figure}
+
+\subsection{Specific Examples}
 

### 2010-02-26 PythonSrc/demos.py
better get data function

@@ -4,9 +4,21 @@ More of a demo than a useful code
 
 
 def get_data_maxener_16_true_false_bars2():
+    return get_data_maxener(pSize=16,keyInv=True,downBeatInv=False,bars=2)
+
+def get_data_maxener_8_true_false_bars2():
+    return get_data_maxener(pSize=8,keyInv=True,downBeatInv=False,bars=2)
+
+
+def get_data_maxener(pSize=16,keyInv=True,downBeatInv=False,bars=2):
     """
     Util function for something we do all the time
     Remove the empty patterns
+    INPUT:
+        - pSize        (default: 16)
+        - keyInv       (default: True)
+        - downBeatInv  (default: False)
+        - bars         (default: 2)
     """
     import data_iterator
     import feats_utils as FU
@@ -16,13 +28,15 @@ def get_data_maxener_16_true_false_bars2():
     tstart = time.time()
     # get maltab files
     allfiles = FU.get_all_matfiles('.')
+    print len(allfiles),' .mat files found'
     # create and set iterator
     data_iter = data_iterator.DataIterator()
     data_iter.setMatfiles(allfiles) # set matfiles
-    data_iter.useBars(2)            # a pattern spans two bars
+    data_iter.useBars( bars )            # a pattern spans two bars
     data_iter.stopAfterOnePass(True)# stop after one full iteration
     # get features
-    featsNorm = [FU.normalize_pattern_maxenergy(p,16,True,False).flatten() for p in data_iter]
+    featsNorm = [FU.normalize_pattern_maxenergy(p,pSize,keyInv,downBeatInv).flatten() for p in data_iter]
+    print 'found ', len(featsNorm),' patterns before removing empty ones'
     # make it an array
     featsNorm = np.array(featsNorm)
     # remove empyt patterns

### 2010-02-26 Paper/Paper.tex
iproved

@@
 LSH: \cite{Datar2004}, LSH implementation: \cite{E2LSH}.
+
 Barrington et al. recently studied music texture using a set of ordered
-MFCC vectors \cite{Barrington2009a}. 
+MFCC vectors \cite{Barrington2009a}. This could be seen as complementary
+to our work. Chromas capture the harmonic information and some of the
+rythm while the MFCC are great timbral features.
+
 Nearest neighbor for music also includes \cite{Cano2004} and ...
@@
 Note that we do not claim that any of these informations (segments, beats, bars)
-are perfectly accurate. Practise showed us that they are reasonable, and the
-size of the data set should make up for the imperfections or noise.
+are perfectly accurate (e.g. in \cite{Barrington2009a} authors improve on the
+song segmentation features). Practise showed us that they are reasonable, 
+and the size of the data set should make up for the imperfections or noise.
 We also believe that patches sizes based on a number of beats or bars are more
@@
 
+uspop2002 serves as testing set to measure how well a code book learned on
+the Cowbell data set can represent new songs.
+
 \section{Algorithm}\label{sec:algo}
+In this section we present the creation of the harmonic patterns from
+the Echo Nest data. Then, we explain the online vector quantization algorithm
+used to learn a code book of those patterns.
 
-Vector quantization \cite{Gersho1991}. Online learning.
+\subsection{Pattern Creation}
+Here we should explain key invariance, resizing, etc, and not in the
+experiment section.
 
 \subsection{Codebook Learning}
+We use an online version of the vector quantization algorithm 
+\cite{Gersho1991}. This can also be seen as an online Kmeans.
+The idea is as follow: take a sample from your data, find the closest
+code in the codebook, bring that code closer to the sample by some amount.
+The details are explained as Algorithm \ref{algo:vq}. note that this
+algorithm, though not optimal, is $O(n)$ with $n$ number of songs seen.
+Thus it scales linearly to any size of data set. The setting of the learning
+rate might become trickier, but this is not fundamentally a problem.
+
 We initialize the codebook by choosing $K$ random points from our dataset.
@@
 using the VQ algorithm (Algorithm \ref{algo:vq}). A typical learning rate 
-is $1e-2$ for $20$ iterations over the whole dataset.
+is $0.01$ for $20$ iterations over the whole dataset.
 
@@
 \subsection{Visualization}
+We can use an algorithm like locally linear embedding \cite{Roweis2000} (LLE) 
+or multidimensional scaling \cite{Kruskal1964} to visualize the data in
+2 dimensions. We use the former\footnote{implementation: 
+\url{http://www.astro.washington.edu/users/vanderplas/coding/LLE/}}, the
+result can be seen in Figure \ref{fig:lle}.
 

### 2010-02-26 PythonSrc/VQutils.py
VQutils.get_best_code_per_p now also returns the distance from each pattern to the closest code. Can be use to encode the test set...

@@ -141,6 +141,7 @@ def encode_scale(signal,codebook,thresh,cbIsNormalized=False,maxIter=1e8):
     return weights, signal
 
 
+
 def encode_dataset_scale(data,codebook,thresh,cbIsNormalized=False):
     """
     Iteratively encode a whole dataset, where each row is a signal
@@ -294,8 +295,12 @@ def find_best_code_per_pattern(dataset,codebook,scale=False):
     If scale=True, we assume codebook normalized
     return a vector, length=dataset size, contains the index
     of the closest code for each pattern (or -1 if there was a problem)
+    RETURN:
+       - index of the best code per pattern
+       - distance for pattern to the best code
     """
     best_code_per_p = np.ones([dataset.shape[0],1]) * -1
+    p_dists = np.ones([dataset.shape[0],1]) * -1
     for k in range(dataset.shape[0]):
         pattern = dataset[k,:].reshape(1,dataset.shape[1])
         if scale:
@@ -304,8 +309,9 @@ def find_best_code_per_pattern(dataset,codebook,scale=False):
         else:
             idxs,dists = encode_oneiter(pattern,codebook)
         best_code_per_p[k] = idxs[0]
+        p_dists[k] = dists[0]
     # done
-    return best_code_per_p
+    return best_code_per_p, p_dists
 
 
 def add_image(P,im,x,y,width=.15):

### 2010-02-26 Paper/Paper.tex
improved

@@
 
-Nearest neighbor for music also includes \cite{Cano2004} and ...
+Nearest neighbor for music also includes \cite{Cano2004} and
+\cite{Holzapfel2009}.
 
@@
 
-\subsection{Setting}
+\subsection{Setting}\label{ssec:setting}
 We take one or two bars, normalize the patches to size 4, 8, or 16.
@@
 \subsection{Pattern length}
+
+\begin{table}
+\begin{center}
+\begin{tabular}{|l|c|c|c|c|}
+\hline
+size & {\bf 1 bar (4)} & {\bf 2 bars (16)} & {\bf 4 beats} & {\bf 7 beats} \\ 
+\hline \hline
+avg. dist & 1 & 2 & 3 & 4 \\ \hline
+something & 5 & 6 & 7 & 8 \\ \hline
+\end{tabular}
+\end{center}
+\caption{{Encoding score for different pattern sizes. We can select the
+size of the patterns based on bars. In this case, we put the actual size
+of the pattern in parentheses. If the pattern size is based on beats, the
+size is the number of beats.}}
+\label{tab:psize}
+\end{table}
 
@@
 \label{fig:lle}
+\end{figure}
+
+Another way is simply to present the codes in order of cluster size.
+We encode the training or testing set as explained in Subsection 
+\ref{ssec:setting} and we count how many patterns were closer to a particular
+code than any other code. In Figure \ref{fig:firstcodes}, we show the $25$
+most used codes out of a code book of size $200$.
+Code from Figure \ref{fig:code} is the 37th most used code in that
+code book. In Figure \ref{fig:closep}, we show a sample of patterns that
+fall into its cluster.
+
+\begin{figure}[htb]
+\begin{center}
+\includegraphics[width=.9\columnwidth]{codes1}
+\end{center}
+\caption{{$25$ most used codes on the training set.}}
+\label{fig:firstcodes}
+\end{figure}
+
+The most common pattern is a sustained note, as we can see in Figure
+\ref{fig:firstcodes}. Another common pattern is a sustained note
+with an harmonic, as seen in the LLE visualization (Figure \ref{fig:lle}).
+
+\begin{figure}[htb]
+\begin{center}
+\includegraphics[width=.9\columnwidth]{close_patterns1}
+\end{center}
+\caption{{Sample of the close patterns to code in Figure \ref{fig:code}.}}
+\label{fig:closep}
 \end{figure}

### 2010-02-26 PythonSrc/demos.py
added size based on beats

@@ -10,7 +10,7 @@ def get_data_maxener_8_true_false_bars2():
     return get_data_maxener(pSize=8,keyInv=True,downBeatInv=False,bars=2)
 
 
-def get_data_maxener(pSize=16,keyInv=True,downBeatInv=False,bars=2):
+def get_data_maxener(pSize=16,keyInv=True,downBeatInv=False,bars=2,beats=0):
     """
     Util function for something we do all the time
     Remove the empty patterns
@@ -32,7 +32,12 @@ def get_data_maxener(pSize=16,keyInv=True,downBeatInv=False,bars=2):
     # create and set iterator
     data_iter = data_iterator.DataIterator()
     data_iter.setMatfiles(allfiles) # set matfiles
-    data_iter.useBars( bars )            # a pattern spans two bars
+    if bars > 0:
+        data_iter.useBars( bars )            # a pattern spans two bars
+    elif beats > 0:
+        data_iter.setFeatsize( beats )       # a pattern spans two bars
+    else:
+        print 'bars or beats should be > 0'
     data_iter.stopAfterOnePass(True)# stop after one full iteration
     # get features
     featsNorm = [FU.normalize_pattern_maxenergy(p,pSize,keyInv,downBeatInv).flatten() for p in data_iter]

### 2010-02-26 PythonSrc/demos.py
added size based on beats

@@ -10,7 +10,7 @@ def get_data_maxener_8_true_false_bars2():
     return get_data_maxener(pSize=8,keyInv=True,downBeatInv=False,bars=2)
 
 
-def get_data_maxener(pSize=16,keyInv=True,downBeatInv=False,bars=2,beats=0):
+def get_data_maxener(pSize=16,keyInv=True,downBeatInv=False,bars=2):
     """
     Util function for something we do all the time
     Remove the empty patterns
@@ -34,10 +34,8 @@ def get_data_maxener(pSize=16,keyInv=True,downBeatInv=False,bars=2,beats=0):
     data_iter.setMatfiles(allfiles) # set matfiles
     if bars > 0:
         data_iter.useBars( bars )            # a pattern spans two bars
-    elif beats > 0:
-        data_iter.setFeatsize( beats )       # a pattern spans two bars
     else:
-        print 'bars or beats should be > 0'
+        data_iter.setFeatsize( pSize )       # a pattern spans two bars
     data_iter.stopAfterOnePass(True)# stop after one full iteration
     # get features
     featsNorm = [FU.normalize_pattern_maxenergy(p,pSize,keyInv,downBeatInv).flatten() for p in data_iter]

### 2010-02-27 PythonSrc/data_iterator.py
some debugging when the patch size is in beats

@@ -158,6 +158,10 @@ class DataIterator:
             if type(self.barbts) == type(0.0): # weird problem sometimes
                 self.fidx = self.fidx + 1
                 continue
+            if self.currfeats == []:
+                print 'currfeats is []'
+                self.fidx = self.fidx + 1
+                continue
             if self.usebars == 0 and self.currfeats.shape[1] < self.featsize :
                 self.fidx = self.fidx + 1
                 continue

### 2010-02-27 PythonSrc/demos.py
better verbose

@@ -33,9 +33,10 @@ def get_data_maxener(pSize=16,keyInv=True,downBeatInv=False,bars=2):
     data_iter = data_iterator.DataIterator()
     data_iter.setMatfiles(allfiles) # set matfiles
     if bars > 0:
-        data_iter.useBars( bars )            # a pattern spans two bars
+        data_iter.useBars( bars )            # a pattern spans 'bars' bars
     else:
-        data_iter.setFeatsize( pSize )       # a pattern spans two bars
+        data_iter.useBars(0)                 # important to set it to zero!
+        data_iter.setFeatsize( pSize )       # a pattern is a num. of beats
     data_iter.stopAfterOnePass(True)# stop after one full iteration
     # get features
     featsNorm = [FU.normalize_pattern_maxenergy(p,pSize,keyInv,downBeatInv).flatten() for p in data_iter]

### 2010-02-27 PythonSrc/data_iterator.py
fixed the fix

@@ -158,7 +158,7 @@ class DataIterator:
             if type(self.barbts) == type(0.0): # weird problem sometimes
                 self.fidx = self.fidx + 1
                 continue
-            if self.currfeats == []:
+            if self.currfeats.size == 0:
                 print 'currfeats is []'
                 self.fidx = self.fidx + 1
                 continue

### 2010-02-27 PythonSrc/data_iterator.py
verbose reduced

@@ -159,7 +159,6 @@ class DataIterator:
                 self.fidx = self.fidx + 1
                 continue
             if self.currfeats.size == 0:
-                print 'currfeats is []'
                 self.fidx = self.fidx + 1
                 continue
             if self.usebars == 0 and self.currfeats.shape[1] < self.featsize :

### 2010-03-01 PythonSrc/VQutils.py
one more output to find_best_code_per_pattern, which affect demos also

@@ -298,9 +298,11 @@ def find_best_code_per_pattern(dataset,codebook,scale=False):
     RETURN:
        - index of the best code per pattern
        - distance for pattern to the best code
+       - average distance for pattern to the best code (per pixel)
     """
     best_code_per_p = np.ones([dataset.shape[0],1]) * -1
     p_dists = np.ones([dataset.shape[0],1]) * -1
+    avg_p_dists = np.ones([dataset.shape[0],1]) * -1
     for k in range(dataset.shape[0]):
         pattern = dataset[k,:].reshape(1,dataset.shape[1])
         if scale:
@@ -310,8 +312,9 @@ def find_best_code_per_pattern(dataset,codebook,scale=False):
             idxs,dists = encode_oneiter(pattern,codebook)
         best_code_per_p[k] = idxs[0]
         p_dists[k] = dists[0]
+        avg_p_dists[k] = np.sum(np.abs(pattern - codebook[idxs[0]]))*1./pattern.size
     # done
-    return best_code_per_p, p_dists
+    return best_code_per_p, p_dists, avg_p_dist
 
 
 def add_image(P,im,x,y,width=.15):

### 2010-03-01 PythonSrc/VQutils.py
typo fixed

@@ -314,7 +314,7 @@ def find_best_code_per_pattern(dataset,codebook,scale=False):
         p_dists[k] = dists[0]
         avg_p_dists[k] = np.sum(np.abs(pattern - codebook[idxs[0]]))*1./pattern.size
     # done
-    return best_code_per_p, p_dists, avg_p_dist
+    return best_code_per_p, p_dists, avg_p_dists
 
 
 def add_image(P,im,x,y,width=.15):

### 2010-03-01 Paper/Paper.tex
improved

@@
 and we discuss there use for retrieval and encoding. All code is made
-available.
+available.\\
+TAKE-HOME MESSAGE: \textit{Shingles} idea can go large-scale?
 \end{abstract}
@@
 But it does not the describe the music space, nor does it necessarily
-make good recommendations. A similar argument in the recommendation
-domain is that precision becomes more important than recall passed a certain
-scale of data for many practical uses.
+make good recommendations. We do not want to look for one neighbor, but
+rather a cloud of potential close patterns. A similar argument in the 
+recommendation domain is that precision becomes more important than recall 
+passed a certain scale of data for many practical uses.
 
@@
 
+Our contribution is... large-scale clustering method,
+
 We present a practical way to do such large-scale clustering using vector
@@
 Slaney and Casey \cite{Casey2006,Casey2007,Casey2008}.
+
 LSH: \cite{Datar2004}, LSH implementation: \cite{E2LSH}.
+The E$^2$LSH packages can efficiently handle about $50$K or $100$K
+vectors of size $200$. Even with a better implementation, the LSH
+model still has some scaling problem. A large portion of the data
+must be known to decide how many projections to use, and what guarantee
+of finding the nearest neighbor it can provides. Also, we could fix the
+number of projections and consider as clusters all the elements that
+fall into the same bin. However, this does not give us a centroid, a
+typical pattern representing that cluster.
 
@@
 
-Nearest neighbor for music also includes \cite{Cano2004} and
-\cite{Holzapfel2009}.
+Nearest neighbor and clustering methods are common in music similarity 
+research, for instance \cite{Cano2004,Holzapfel2009}. The difference here 
+are the actual patterns that we cluster.
 
@@
 \section{Data}\label{sec:data}
+In this section we present our use of the Echo Nest features and the
+datasets we use for training and testing our model.
 
@@
 \subsection{Cowbell Dataset}
-We have $3,720,091$ non zero bars, a bar usually contains
-$4$ beats.
+We have $43,300$ songs, giving  $3,720,091$ non zero bars, a bar usually 
+contains $4$ beats.
 
 \subsection{uspop2002}
-We had acces to a low quality (32kbps) version of the songs in the uspop 2002 data
-set \cite{uspop2002}.
+We had acces to a low quality (32kbps) version of the songs in the uspop 2002 
+data set \cite{uspop2002}.
 This well-known data show a great diversity of pop songs, but is not particular
@@
 uspop2002 serves as testing set to measure how well a code book learned on
-the Cowbell data set can represent new songs.
+the Cowbell data set can represent new songs. We get Echo Nest features
+for $8651$ songs from that dataset.
 
@@
 \subsection{Pattern length}
+See encoding experiments in Table \ref{tab:psize}.
 
+% RESULTS
+% 200codes, 2 bars, pSize=16, maxener, True, False, not scaled = 3.788
+%                                             averaged on size = 
+% 200codes, 2 bars, pSize=8, maxener, True, False, not scaled  = 3.135
+%                                             averaged on size = 
+% 200codes, 1 bars, pSize=4, maxener, True, False, not scaled  = 
+%                                             averaged on size = 
+% 200codes, 7 beats, pSize=7, maxener, True, False, not scaled = 2.771
+%                                             averaged on size = 
+% 200codes, 5 beats, pSize=5, maxener, True, False, not scaled = 2.397
+%                                             averaged on size = 
+% 200codes, 3 beats, pSize=3, maxener, True, False, not scaled = 
+%                                             averaged on size = 
 \begin{table}
 \begin{center}
-\begin{tabular}{|l|c|c|c|c|}
+\begin{tabular}{|l|c|}
 \hline
-size & {\bf 1 bar (4)} & {\bf 2 bars (16)} & {\bf 4 beats} & {\bf 7 beats} \\ 
-\hline \hline
-avg. dist & 1 & 2 & 3 & 4 \\ \hline
-something & 5 & 6 & 7 & 8 \\ \hline
+size & avg. dist \\ \hline \hline
+1 bar (4) & \\
+2 bars (8) & $$\\
+2 bars (12) & $$\\
+2 bars (16) & $$ \\ \hline
+3 beats & \\
+5 beats & $$\\
+7 beats & $$\\ \hline
 \end{tabular}
@@
 of the pattern in parentheses. If the pattern size is based on beats, the
-size is the number of beats.}}
+size is the number of beats. Results are on the uspop dataset, not seen
+during training.}}
 \label{tab:psize}
 \end{table}
+
+\subsection{Data Size}
+We are interesting to see how the amount of training data changes
+the quality of our code.
+We experiment with different data sizes for a fixed size of codebook and a 
+fixed code length. The data points are chose randomly from the whole
+data set.
+Code length: 1 bar resized to size 4. Code book size: 100
+We measure the distortion on our test set. We hope for some convergence.
+See results in Table \ref{tab:datasize}.
+
+Our learning can overfit, which explain the bad scores of small datasets.
+The larger the dataset, the smallest the chance to overfit.
+We ar elooking for a size of dataset that guarantees us a good learning
+process, while being unnecessary large.
+
+\begin{table}
+\begin{center}
+\begin{tabular}{|c|c|c|c|}
+\hline
+$10$K & $100$K & $500$K & $2$M  \\ \hline
+$0$ & $1$ & $2$ & $3$\\ \hline
+\end{tabular}
+\end{center}
+\caption{{Distortion for a code book of size $100$ encoding one bar
+at a time, a bar being represented by $4$ chroma feature vector.
+Therefore, a code or a pattern is of size $12$ x $4$ = $48$.
+Distortion measure on the test data set. Data size tested range
+from $10$ thousands to $2$ millions. Patterns were selected at
+random from the dataset of approximately $3,7$ million patterns.}}
+\label{tab:datasize}
+\end{table}
+% RESULTS:    avg.       mean dist
+% 10K         
+% 100K
+% 500K
+% 2M
+
+\subsection{Codebook Size in Function of Pattern Length}
+If the signal was totally random, and we were using 2 code to encode
+every beat, we would need.... 4 codes for 2 beats? and 16 codes for
+4 beats? and 256 codes for 8 beats?
+
+Think about it in binary: two code to encode (0) and (1), but for
+(0,0),(0,1),(1,0),(1,1) we need four codes for the same accuracy.
+
+Reasonable size: 2 - 4 - 16 - 256. If we improve distance, music
+is not random!
+
+Then... 3 beats, 5 beats, 7 beats don't encode as well!
+
+Special: 4 beats versus 1 bar - 4 beats, show the problem of being
+not in sync with bars.
 

### 2010-03-01 Paper/Paper.tex
improved

@@
 
+Working with large data amount of audio is a goal and a challenge at the
+same time. One of the most common task is to find the nearest neighbor,
+based on some distance, to a given segment of music. This can be viewed
+as \textit{retrieval} as you are looking for some specific target
+based on some query. $k$-nearest neighbors ($k$nn) algorithms have been
+designed for that purpose, and recent improvements let them scale to
+considerable data set sizes \cite{Datar2004}. The \textit{Shingles}
+were introduced in that framework \cite{Casey2006}. A \textit{Shingle}
+is a pattern of stacked chroma features. Similarity between
+\textit{Shingles} coresspond to comparing harmonic patterns. For this
+reason, they were used for cover or remix recognition \cite{Casey2007}.
+
+That being said, nearest neighbor search is only one aspect of the
+analysis of large-scale music data sets. Finding $k$ neighbors for a
+particular query does not tell you much about the music space.
+Clustering is a larger and more flexible framework where the goal is
+to divide the space in meaningful blocks. The starting point is not
+a query, but all the data set itself. In this work, we bring the
+\textit{Shingle} idea into that framework in order to discover
+meaningful and typical music harmonic patterns. The underlying assumption
+is that the music space is articulated around a certain number of
+frequent patterns.
+
+Note that clustering and neighbor search is not totally unrelated.
+In practise, we would hope that close neighbors live in one or two
+close clusters, thus clustering can be a preprocess step. Also,
+knowing all $k$ neighbors of each sample let you do graph clustering,
+a subset of the clustering algorithms.
+
+\iffalse % ignored section
 The original goal of the \textit{Shingles} was retrieval, i.e. find very
@@
 
-We propose to move the \textit{Shingle} idea into the larger clustering paradigm.
+We propose to move the \textit{Shingle} idea into the larger clustering 
+paradigm.
+\fi % end of ignored section
+
 The goal is to find typical patterns that can be seen as cluster centroids.
@@
 \begin{center}
-\begin{tabular}{|c|c|c|c|}
+\begin{tabular}{|c|c|c|c|c|}
 \hline
-$10$K & $100$K & $500$K & $2$M  \\ \hline
-$0$ & $1$ & $2$ & $3$\\ \hline
+$10$K & $50$K & $100$K & $500$K & $2$M  \\ \hline
+$0.143$ & $$ & $$ & $$ & $$\\ \hline
 \end{tabular}
@@
 % RESULTS:    avg.       mean dist
-% 10K         
+% 10K         2.012      0.143
+% 50K
 % 100K

### 2010-03-01 PythonSrc/VQutils.py
tiny code improvement

@@ -312,7 +312,8 @@ def find_best_code_per_pattern(dataset,codebook,scale=False):
             idxs,dists = encode_oneiter(pattern,codebook)
         best_code_per_p[k] = idxs[0]
         p_dists[k] = dists[0]
-        avg_p_dists[k] = np.sum(np.abs(pattern - codebook[idxs[0]]))*1./pattern.size
+        #avg_p_dists[k] = np.sum(np.abs(pattern - codebook[idxs[0]]))*1./pattern.size
+        avg_p_dists[k] = np.average(np.abs(pattern - codebook[idxs[0]]))
     # done
     return best_code_per_p, p_dists, avg_p_dists
 

### 2010-03-02 PythonSrc/VQutils.py
remove best_code_per_p from learning, that was ridiculous

@@ -412,7 +412,6 @@ def online_vq(feats,K,lrate,nIter=10,thresh=0.0000001,maxRise=.05,scale=False,re
     OUTPUT:
       - codebook (one code per row)
       - average distance between a feature and it's encoding
-      - list of which code to use for which data sample
     Inspired by the algorithm here:
     http://en.wikipedia.org/w/index.php?title=Vector_quantization&oldid=343764861
     Codes are normalized, and can be scaled as to better feat a segment
@@ -445,8 +444,8 @@ def online_vq(feats,K,lrate,nIter=10,thresh=0.0000001,maxRise=.05,scale=False,re
     # keep the best result
     best_sum_dist = -1
     # know which code goes with each pattern
-    best_code_per_pattern = np.ones([nFeats,1])
-    best_code_per_pattern *= -1
+    #best_code_per_pattern = np.ones([nFeats,1])
+    #best_code_per_pattern *= -1
     # not scaled? artificial weights
     if not scale:
         weights = np.ones(K)
@@ -481,7 +480,8 @@ def online_vq(feats,K,lrate,nIter=10,thresh=0.0000001,maxRise=.05,scale=False,re
             if scale:
                 codebook[idx,:] = normalize(codebook[idx,:])
             # remember that code for that pattern
-            best_code_per_pattern[whichPattern] = idx
+            #best_code_per_pattern[whichPattern] = idx
+            # bad idea, codes will change
 
             ######################
             # TEST on repulsiveness
@@ -510,6 +510,6 @@ def online_vq(feats,K,lrate,nIter=10,thresh=0.0000001,maxRise=.05,scale=False,re
         prev_sum_dist = sum_distance
         
     # return codebook, average distance
-    return codebook,(sum_distance * 1. / nFeats), best_code_per_pattern
+    return codebook,(sum_distance * 1. / nFeats)#, best_code_per_pattern
 
 

### 2010-03-04 Paper/generate_datasize_curves.py
create fig with curves about data size

@@ -1,0 +1,95 @@
+"""
+Script to generate one figure from our Shingles paper
+
+T. Bertin-Mahieux (2010) Columbia University
+tb2332@columbia.edu
+"""
+
+import os
+import sys
+import time
+import numpy as np
+import pylab as P
+
+
+
+def create_fig():
+
+    # DATA
+
+    sizes = [1000,5000,10000,50000,100000,250000,500000,1000000,2000000]
+    labels = ['1K','5K','10K','50K','100K','250K','500K','1M','2M']
+    
+    # data 1 bar 4 beats
+    bar1beat4 = [0.150370,0.145327,0.143567,0.142257,0.142195,0.141796,0.142081,0.142118,0.141596]
+    # data 2 bars 8 beats
+    bar2beat8 = [0.171300,0.165960,0.165063,0.162439,0.161588,0.161910,0.161233,0.161057,0.161068]
+    # data 1 bar 4 beats not random! first K in data
+    bar1beat4FIRST = [0.157142,0.150045,0.148357,0.146069,0.145765,0.145819,0.145772,0.145924,0]
+    # data 0 bar 4 beats
+    bar0beat4 = [0.155474,0.149619,0.148056,0.146328,0.146004,0,0,0,0]
+
+
+    # START DISPLAYING STUFF
+    
+    P.figure()
+    P.hold(True)
+    # axis
+    xmin = 0
+    xmax = len(labels)
+    ymin = 0.13
+    ymax = 0.17
+    P.axis([xmin,xmax,ymin,ymax])
+
+    # plot1
+    P.plot(bar1beat4,'o-',label='1 bar 4 beats')
+    # plot2
+    P.plot(bar2beat8,'x-',label='2 bars 8 beats')
+    # plot3
+    P.plot(bar1beat4FIRST,'--',label='1 bar 4 beats non random')
+    # plot4
+    P.plot(bar0beat4,'-.',label='0 bars 4 beats')
+    
+    # labels
+    P.xticks(P.arange(len(labels)),list(labels))
+    
+    # x titles and y titles
+    P.xlabel('data size')
+    P.ylabel('average error')
+
+    # legend
+    P.legend()
+
+    # main title
+    P.title('Encoding error per training data size for certain conditions')
+
+    P.show()
+
+
+
+
+
+
+
+
+
+
+
+
+
+
+
+
+def die_with_usage():
+    """ help menu """
+    print 'create figure'
[…]

### 2010-03-05 Paper/Paper.tex
improved, new table and fig

@@
 
-Working with large data amount of audio is a goal and a challenge at the
+Working with large amount of audio data is a goal and a challenge at the
 same time. One of the most common task is to find the nearest neighbor,
@@
 is a pattern of stacked chroma features. Similarity between
-\textit{Shingles} coresspond to comparing harmonic patterns. For this
+\textit{Shingles} coresponds to comparing harmonic patterns. For this
 reason, they were used for cover or remix recognition \cite{Casey2007}.
@@
 to divide the space in meaningful blocks. The starting point is not
-a query, but all the data set itself. In this work, we bring the
+a query, but all the data itself. In this work, we bring the
 \textit{Shingle} idea into that framework in order to discover
@@
 
-Note that clustering and neighbor search is not totally unrelated.
+Note that clustering and neighbor search are not unrelated.
 In practise, we would hope that close neighbors live in one or two
@@
 a subset of the clustering algorithms.
+
+A major component of this work is scalability. We experiment with a training
+set of $43$ thousands song, but everything scales linear in time
+relative to the number of samples. We could imagine, as an extension
+of this work, an algorithm ``listening'' to all music it can find online.
+Time would be an issue, but not memory.
 
@@
 the method with some specific examples.
+
+\subsection{Definitions}
+A pattern, or a patch, is a \textit{Shingle} taken from an actual song.
+A code is similar to a pattern, but is the result of the clustering
+algorithm and most likely does not come from an actual song.
+Distortion is our measure of distance between a pattern and a code.
 
@@
 \begin{algorithm}
-\caption{Pseudocode of Vector Quantization}
+%\caption{Pseudocode of Vector Quantization}
 \label{algo:vq}
@@
 \RETURN $\{C_k\}$
+\caption{{Pseudocode of Online Vector Quantization. Note that we can replace
+the number of iteration by a threshold on the distortion over some test set.}}
 \end{algorithmic}
@@
 We measure the distortion on our test set. We hope for some convergence.
-See results in Table \ref{tab:datasize}.
+See results in Figure \ref{fig:sizes}.
 
@@
 The larger the dataset, the smallest the chance to overfit.
-We ar elooking for a size of dataset that guarantees us a good learning
+We are looking for a size of dataset that guarantees us a good learning
 process, while being unnecessary large.
 
+$500$ samples per code in the code book seems a good data size, see
+Figure \ref{fig:sizes} for size $50$K.
+
+\begin{figure}[htb]
+\begin{center}
+\includegraphics[width=.9\columnwidth]{data_sizes}
+\end{center}
+\caption{{Distortion for a code book of size $100$ encoding one bar
+at a time, a bar being represented by $4$ chroma feature vector.
+Therefore, a code or a pattern is of size $12$ x $4$ = $48$.
+Distortion measure on the test data set. Data size tested range
+from $10$ thousands to $2$ millions. Patterns were selected at
+random from the dataset of approximately $3,7$ million patterns.}}
+\label{fig:sizes}
+\end{figure}
+
+\iffalse
 \begin{table}
 \begin{center}
-\begin{tabular}{|c|c|c|c|c|}
+\begin{tabular}{|l|c|c|c|c|c|c|}
 \hline
-$10$K & $50$K & $100$K & $500$K & $2$M  \\ \hline
-$0.143$ & $$ & $$ & $$ & $$\\ \hline
+size & $5$K & $10$K & $50$K & $100$K & $500$K & $2$M  \\ \hline
+distortion & $$ & $0.143$ & $0.142$ & $0.142$ & $0.142$ & $$\\ \hline
 \end{tabular}
@@
 \end{table}
-% RESULTS:    avg.       mean dist
-% 10K         2.012      0.143
-% 50K
-% 100K
-% 500K
-% 2M
+\fi
+% RESULTS:    
+% 1bar s4     avg.       mean dist
+% 1K          2.30777    0.150370
+% 5K          1.88822    0.145327
+% 10K         2.00349    0.143567
+% 50K         2.59714    0.142257
+% 100K        1.89185    0.142195
+% 250K        2.07822    0.141796
+% 500K        1.88141    0.142081
+% 1M          1.91990    0.142118
+% 2M          2.01141    0.141596
+
+% 1bar s4 first X samples
+% 1K          2.05621    0.157142
+% 5K          2.01956    0.150045
+% 10K         1.95797    0.148357
+% 50K         1.91257    0.146069
+% 100K        1.98744    0.145765
+% 250K        1.95362    0.145819
+% 500K        2.00899    0.145772
+% 1M          2.09856    0.145924
+% 2M          1.91375    0.146065
+
+% 0bar s4 (and test set same, we could do 1bar s4 in test set...)
+% 1K          2.10588    0.155474
+% 5K          1.92358    0.149619
+% 10K         2.00653    0.148056
+% 50K         1.89443    0.146328
+% 100K        1.92187    0.146004
+% 250K        1.93067    0.146024
+% 500K        1.91913    0.145962
+% 1M          2.24550    0.145873
+% 2M          2.01653    0.145955
+
+% 2bar s8     
+% 1K          3.27198    0.171300
+% 5K          3.56894    0.165960
+% 10K         2.61057    0.165063
+% 50K         3.03945    0.162439
+% 100K        2.89310    0.161588
+% 250K        2.90660    0.161910
+% 500K        2.69134    0.161233
+% 1M          3.05344    0.161057
+% 2M          2.74493    0.161068
 
@@
 not in sync with bars.
+
+See code book size experiments in Table \ref{tab:cbsize}.
+
+\begin{table}
+\begin{center}
+\begin{tabular}{|c|c|c|}
+\hline
+\# codes & \# samples & avg. dist \\ \hline \hline
+1 & $500$ & \\
+10 & $5$K &\\
+50 & $25$K &\\
+100 & $50$K &\\
+200 & $100$K & \\
+500 & $250$K & \\
+1000 & $500$K & \\
+5000 & $2.5$M & \\ \hline
+\end{tabular}
+\end{center}
+\caption{{Code book size.}}
+\label{tab:cbsize}
+\end{table}
 
@@
 \small
-\section{Acknowledgements}
-Thierry is NSERC graduate fellow, or some title like that.
+%\section{Acknowledgements}
+%Thierry is NSERC graduate fellow, or some title like that.
 

### 2010-03-07 PythonSrc/demos.py
removed best_code_per_p for load and encode, it is useless since the data is randomized and lost

@@ -12,7 +12,7 @@ def load_and_encode_data(codebook,pSize=16,keyInv=True,
                          downBeatInv=False,bars=2):
     """
     Load a dataset, and encode it with codebook
-    Return best_code_per_p, dists
+    Return dists, avg_dists
     """
     assert(codebook.shape[1] == pSize * 12)
     import VQutils
@@ -20,7 +20,8 @@ def load_and_encode_data(codebook,pSize=16,keyInv=True,
     featsNorm = get_data_maxener(pSize=pSize,keyInv=keyInv,
                                  downBeatInv=downBeatInv,bars=bars)
     # encode
-    return VQutils.find_best_code_per_pattern(featsNorm,codebook,scale=False)
+    best_code_per_p, dists, avg_dists = VQutils.find_best_code_per_pattern(featsNorm,codebook,scale=False)
+    return dists, avg_dists
 
 
 def get_data_maxener_16_true_false_bars2():

### 2010-03-08 PythonSrc/VQutils.py
major changes to VQ, especially the way we measure stuff... and better threshold during learning

@@ -250,7 +250,7 @@ def online_encoding_learn(feats,K,lrate=1.,nIter=10,nEncode=-1,
             # add residual
             sum_residual += euclidean_norm(residual)
 
-        # update best sum reesidual
+        # update best sum residual
         if best_sum_residual < 0 or best_sum_residual > sum_residual:
             best_sum_residual = sum_residual
         # verbose
@@ -312,8 +312,8 @@ def find_best_code_per_pattern(dataset,codebook,scale=False):
             idxs,dists = encode_oneiter(pattern,codebook)
         best_code_per_p[k] = idxs[0]
         p_dists[k] = dists[0]
-        #avg_p_dists[k] = np.sum(np.abs(pattern - codebook[idxs[0]]))*1./pattern.size
-        avg_p_dists[k] = np.average(np.abs(pattern - codebook[idxs[0]]))
+        avg_p_dists[k] = np.average(np.square(pattern - codebook[idxs[0]]))
+        assert abs(avg_p_dists[k]*codebook[0].size - p_dists[k] * p_dists[k]) < 1e-10,'distance and average distance do not behave as expected'
     # done
     return best_code_per_p, p_dists, avg_p_dists
 
@@ -399,7 +399,7 @@ def LLE_my_codebook(codebook,nNeighbors=5):
 
 
 
-def online_vq(feats,K,lrate,nIter=10,thresh=0.0000001,maxRise=.05,scale=False,repulse=False):
+def online_vq(feats,K,lrate,nIter=10,thresh=1e-7,maxRise=.05,scale=False):
     """
     Online vector quantization
     INPUT:
@@ -443,9 +443,6 @@ def online_vq(feats,K,lrate,nIter=10,thresh=0.0000001,maxRise=.05,scale=False,re
     nFeats = feats.shape[0]
     # keep the best result
     best_sum_dist = -1
-    # know which code goes with each pattern
-    #best_code_per_pattern = np.ones([nFeats,1])
-    #best_code_per_pattern *= -1
     # not scaled? artificial weights
     if not scale:
         weights = np.ones(K)
@@ -479,19 +476,7 @@ def online_vq(feats,K,lrate,nIter=10,thresh=0.0000001,maxRise=.05,scale=False,re
             codebook[idx,:] += (pattern / weight - codebook[idx,:]) * lrate
             if scale:
                 codebook[idx,:] = normalize(codebook[idx,:])
-            # remember that code for that pattern
-            #best_code_per_pattern[whichPattern] = idx
[…]

### 2010-03-08 PythonSrc/VQutils.py
major bug solved

@@ -311,9 +311,9 @@ def find_best_code_per_pattern(dataset,codebook,scale=False):
         else:
             idxs,dists = encode_oneiter(pattern,codebook)
         best_code_per_p[k] = idxs[0]
-        p_dists[k] = dists[0]
-        avg_p_dists[k] = np.average(np.square(pattern - codebook[idxs[0]]))
-        assert abs(avg_p_dists[k]*codebook[0].size - p_dists[k] * p_dists[k]) < 1e-10,'distance and average distance do not behave as expected'
+        p_dists[k] = dists[idxs[0]]
+        #avg_p_dists[k] = np.average(np.square(pattern - codebook[idxs[0]]))
+        avg_p_dists[k] = np.square(p_dists[k]) * 1. / codebook[0].size
     # done
     return best_code_per_p, p_dists, avg_p_dists
 

### 2010-03-08 PythonSrc/VQutils.py
oups bad copypaste

@@ -298,7 +298,8 @@ def find_best_code_per_pattern(dataset,codebook,scale=False):
     RETURN:
        - index of the best code per pattern
        - distance for pattern to the best code
-       - average distance for pattern to the best code (per pixel)
+       - average distance for pattern to the best code
+         (squared distance per pixel)
     """
     best_code_per_p = np.ones([dataset.shape[0],1]) * -1
     p_dists = np.ones([dataset.shape[0],1]) * -1
@@ -491,14 +492,6 @@ def online_vq(feats,K,lrate,nIter=10,thresh=1e-7,maxRise=.05,scale=False):
                 print 'online_vq stops because of thresholding after iter: ' + str(iteration+1)
                 break
         if best_sum_dist * (1. + maxRise) < sum_distance:
-            break
-
-        # check threshold
-        if prev_sum_residual >= 0:
-            if abs((sum_residual - prev_sum_residual) * 1./nFeats) < thresh:
-                print 'online_encode_learn stops because of thresholding after iter: ' + str(iteration+1)
-                break
-        if best_sum_residual * (1. + maxRise) < sum_residual:
             break
 
         prev_sum_dist = sum_distance

### 2010-03-09 Paper/Paper.tex
recent improvs
result organization
NMF stuff started

@@
 \begin{center}
-\includegraphics[width=.9\columnwidth]{data_sizes}
+\includegraphics[width=.99\columnwidth]{data_sizes}
 \end{center}
@@
 % RESULTS:    
-% 1bar s4     avg.       mean dist
-% 1K          2.30777    0.150370
-% 5K          1.88822    0.145327
-% 10K         2.00349    0.143567
-% 50K         2.59714    0.142257
-% 100K        1.89185    0.142195
-% 250K        2.07822    0.141796
-% 500K        1.88141    0.142081
-% 1M          1.91990    0.142118
-% 2M          2.01141    0.141596
+% 1bar s4     avg_distOLD    avg_dist
+% 1K          0.150370
+% 5K          0.145327
+% 10K         0.143567
+% 50K         0.142257
+% 100K        0.142195
+% 250K        0.141796
+% 500K        0.142081
+% 1M          0.142118
+% 2M          0.141596
+% 1bar s4 first X samples
+% 1K          0.157142
+% 5K          0.150045
+% 10K         0.148357
+% 50K         0.146069
+% 100K        0.145765
+% 250K        0.145819
+% 500K        0.145772
+% 1M          0.145924
+% 2M          0.146065
+% 0bar s4 (and test set same, we could do 1bar s4 in test set...)
+% 1K          0.155474
+% 5K          0.149619
+% 10K         0.148056
+% 50K         0.146328
+% 100K        0.146004
+% 250K        0.146024
+% 500K        0.145962
+% 1M          0.145873
+% 2M          0.145955
+% 2bar s8     
+% 1K          0.171300
+% 5K          0.165960
+% 10K         0.165063
+% 50K         0.162439
+% 100K        0.161588
+% 250K        0.161910
+% 500K        0.161233
+% 1M          0.161057
+% 2M          0.161068
 
-% 1bar s4 first X samples
-% 1K          2.05621    0.157142
-% 5K          2.01956    0.150045
-% 10K         1.95797    0.148357
-% 50K         1.91257    0.146069
-% 100K        1.98744    0.145765
-% 250K        1.95362    0.145819
-% 500K        2.00899    0.145772
-% 1M          2.09856    0.145924
-% 2M          1.91375    0.146065
+\subsection{Codebook Size}
+Results in Table \ref{tab:cbsize}.
 
-% 0bar s4 (and test set same, we could do 1bar s4 in test set...)
-% 1K          2.10588    0.155474
-% 5K          1.92358    0.149619
-% 10K         2.00653    0.148056
-% 50K         1.89443    0.146328
-% 100K        1.92187    0.146004
-% 250K        1.93067    0.146024
-% 500K        1.91913    0.145962
-% 1M          2.24550    0.145873
-% 2M          2.01653    0.145955
-
-% 2bar s8     
-% 1K          3.27198    0.171300
-% 5K          3.56894    0.165960
-% 10K         2.61057    0.165063
-% 50K         3.03945    0.162439
-% 100K        2.89310    0.161588
-% 250K        2.90660    0.161910
-% 500K        2.69134    0.161233
-% 1M          3.05344    0.161057
-% 2M          2.74493    0.161068
+\begin{table}
+\begin{center}
+\begin{tabular}{|l|l|c|}
+\hline
+\# codes & \# samples & avg. dist \\ \hline \hline
+1 & $500$ & $0.195939$ \\
+10 & $5$K & $0.161812$ \\
+50 & $25$K & $0.147472$ \\
+100 & $50$K & $0.142194$ \\
+200 & $100$K & $0.137548$ \\
+500 & $250$K & $0.131211$ \\
+1000 & $500$K & \\ \hline
+\end{tabular}
+\end{center}
+\caption{{Code book size.}}
+\label{tab:cbsize}
+\end{table}
+% RESULTS:      avg_distsOLD    avg_dist
+% code 1        0.195939
+% code 10       0.161812
+% code 50       0.147472
+% code 100      0.142194
+% code 200      0.137548
+% code 500      0.131211
 
@@
 \begin{center}
-\begin{tabular}{|c|c|c|}
+\begin{tabular}{|l|c|c|}
 \hline
-\# codes & \# samples & avg. dist \\ \hline \hline
-1 & $500$ & \\
-10 & $5$K &\\
-50 & $25$K &\\
-100 & $50$K &\\
-200 & $100$K & \\
-500 & $250$K & \\
-1000 & $500$K & \\
-5000 & $2.5$M & \\ \hline
+\# codes & \# bars & avg. dist \\ \hline \hline
+2 & $1$ & $0.058435$ \\
+4 & $2$ & $0.054792$ \\
+16 & $4$ & $0.055489$ \\
+256 & $8$ & $0.054860$ \\ \hline
+2 & $2$ beats & $0.046955$ \\
+4 & $1$ & $0.$ \\
+16 & $2$ & $0.$ \\
+256 & $4$ & $0.$ \\ \hline
 \end{tabular}
 \end{center}
-\caption{{Code book size.}}
-\label{tab:cbsize}
+\caption{{Code book size in function of the number of bars to encode.
+All experiments done wtih $256 \times 500 = 128$K samples randomly chosen.}}
+\label{tab:size_pattern}
 \end{table}
+% RESULTS:           avg_distsOLD   avg_dist
+%2 codes 1 bar       0.180750       0.058435
+%4 codes 2 bars      0.183631       0.054792
+%16 codes 4 bars     0.184078       0.055489
+%256 codes 8 bars    0.183464       0.054860
+%---------------
+%2 codes 2 beats     0.166840       0.046955
+%4 codes 1 bar       0.171495       0
+%16 codes 2 bars     0.173767
+%256 codes 4 bars    0.174169       0
 
@@
 
+\section{Links with NMF}\label{sec:nmf}
+In this section we discuss links with non-negative matrix factorization
+\cite{Lee2000} (NMF), another method to build a codebook representing some data.
+NMF finds two matrices $H$ and $W$ that approximate some data $D$ (one
+sample per line) by
+\[
+D = W \times H + \epsilon
+\]
+where $\epsilon$ is some noise and $W$ and $H$ only have non-negative 
+elements. If $W$ is $\#samples \times n$, $n$ is the codebook size.
+
+Note that NMF does not solve the same problem. It tries to explain each pattern
+as a combination of codes. Thus it does not perform clustering, though you
+could cluster patterns using their encoding weights.
+
+In all our experiments we use Matlab's implementation of NMF:
+nmflib\footnote{implementation: \url{http://www.ee.columbia.edu/~grindlay/code.html}}.
+
+\subsection{Training on our Codebook}
+NMF provides us with a way of measuring how much information is retained
+in our code book. More specifically, we can train a code book of size $500$
+using our online VQ algorithm, then decompose that code into $10$ basis
+using NMF. We can compare that to $10$ basis from training NMF on the
+whole data.
+
+The idea is simple, if we encode our test set using basis from the
+codebook as well as if we were using basis from the whole data, our codes
+retain much of the information. That being said, the typical NMF implementations
+are not expected to scale to data that does not fit in memory. Online VQ
+could be an interesting preprocessing step. Future work include looking
+at online NMF \cite{Thurau2009}.
+
 \section{Conclusion}

### 2010-03-10 Paper/Paper.tex
improved

@@
 %\caption{Pseudocode of Vector Quantization}
-\label{algo:vq}
 \begin{algorithmic}
@@
 \caption{{Pseudocode of Online Vector Quantization. Note that we can replace
-the number of iteration by a threshold on the distortion over some test set.}}
+the number of iteration by a threshold on the distortion over some test set.}
+\label{algo:vq}}
 \end{algorithmic}
@@
 % RESULTS:      avg_distsOLD    avg_dist
-% code 1        0.195939
-% code 10       0.161812
-% code 50       0.147472
-% code 100      0.142194
-% code 200      0.137548
-% code 500      0.131211
+% code 1        0.195939        0.066079
+% code 10       0.161812        0.045690
+% code 50       0.147472        0.038456
+% code 100      0.142194        0.036725
+% code 200      0.137548        0.
+% code 500      0.131211        0.
 
@@
 256 & $8$ & $0.054860$ \\ \hline
-2 & $2$ beats & $0.046955$ \\
-4 & $1$ & $0.$ \\
-16 & $2$ & $0.$ \\
-256 & $4$ & $0.$ \\ \hline
+2 & $2$ beats* & $0.066416$ \\
+4 & $1$ & $0.050712$ \\
+16 & $2$ & $0.050172$ \\
+256 & $4$ & $0.04989$ \\ \hline
 \end{tabular}
@@
 \caption{{Code book size in function of the number of bars to encode.
-All experiments done wtih $256 \times 500 = 128$K samples randomly chosen.}}
+All experiments done wtih $256 \times 500 = 128$K samples randomly chosen.
+* For the $2$ beats experiment, we take the first half of every beats.
+It is different from ignoring bars and taking every $2$ beats. It is more
+coherent with the other settings ($1-2-4-8$ bars).}}
 \label{tab:size_pattern}
@@
 %2 codes 2 beats     0.166840       0.046955
-%4 codes 1 bar       0.171495       0
-%16 codes 2 bars     0.173767
-%256 codes 4 bars    0.174169       0
+%2 codes 2 beats                    0.066416
+% based on bars
+%4 codes 1 bar       0.171495       0.050712
+%16 codes 2 bars     0.173767       0.050172
+%256 codes 4 bars    0.174169       0.049894
 
@@
 as a combination of codes. Thus it does not perform clustering, though you
-could cluster patterns using their encoding weights.
+could cluster patterns using their encoding weights. Also, we do not fully
+compare our methods with NMF, it would be a paper on its own.
 
@@
 at online NMF \cite{Thurau2009}.
+
+\subsection{Experiment}
+Put figure, compare basis, yeah!
 

### 2010-03-10 Paper/generate_expo_codes_curves.py
to generate graph on codebook size when pattern size double

@@ -1,0 +1,128 @@
+"""
+Code to generate the curve where the number of codes is squared
+when the pattern size is doubled
+"""
+
+
+
+import os
+import sys
+import time
+import numpy as np
+import pylab as P
+
+
+
+def create_fig():
+
+    # DATA
+    labels = ['2','4','16','256']
+
+    # 2 beats - 1 bar - 2 bars - 4 bars
+    data1 = [0.066416,0.050712,0.050172,0.049894]
+    # 1 bar - 2 bars - 4 bars - 8 bars
+    data2 = [0.058435,0.054792,0.055489,0.054860]
+
+    
+    # START DISPLAYING STUFF
+    
+    P.figure()
+    P.hold(True)
+    # axis
+    xmin = 0
+    xmax = len(labels)
+    ymin = 0.49
+    ymax = 0.67
+    P.axis([xmin,xmax,ymin,ymax])
+
+    # plot1
+    P.plot(data1,'o-',label='2 beats, 1 bar, 2 bars, 4 bars')
+    # plot2
+    P.plot(data2,'x-',label='1 bar, 2 bars, 4 bars, 8 bars')
+
+    # ANNOTATION
+    gcf = P.gca()
+    # data1 pt1
+    gcf.annotate('2 beats', xy=(0, data1[0]),  xycoords='data',
+                xytext=(50, 30), textcoords='offset points',
+                arrowprops=dict(arrowstyle="->",
+                                connectionstyle="arc3,rad=.2")
+                )
+    # data1 pt2
+    gcf.annotate('1 bar', xy=(1, data1[1]),  xycoords='data',
+                xytext=(50, 30), textcoords='offset points',
+                arrowprops=dict(arrowstyle="->",
+                                connectionstyle="arc3,rad=.2")
+                )
+    # data1 pt3
+    gcf.annotate('2 bars', xy=(2, data1[2]),  xycoords='data',
+                xytext=(50, 30), textcoords='offset points',
+                arrowprops=dict(arrowstyle="->",
+                                connectionstyle="arc3,rad=.2")
+                )
+    # data1 pt4
+    gcf.annotate('4 bars', xy=(3, data1[3]),  xycoords='data',
+                xytext=(-50, -30), textcoords='offset points',
+                arrowprops=dict(arrowstyle="->",
+                                connectionstyle="arc3,rad=.2")
+                )
+    # data2 pt1
[…]

### 2010-03-10 PythonSrc/for_ron.py
easy way to do mass experiments

@@ -1,0 +1,113 @@
+"""
+Simple code to run a typical experiment
+MUST change the two lines starting with cd
+"""
+
+
+import os
+import sys
+import VQutils
+import demos
+import numpy as np
+
+
+featsDir = 
+testFeatsDir = 
+
+
+def do_experiment(beats,bars,nCodes,nSamples=0,useFirsts=False):
+    """
+    Performs an independant experiment!!!!
+    """
+
+    # TRAINING
+    # go to the folder of features (per beat)
+    cd featsDir
+
+    # load everything, unit: 1 bar, resized to 4 beats
+    # key invariant, not downbeatinvariant
+    featsNorm = demos.get_data_maxener(pSize=beats,keyInv=True,downBeatInv=False,bars=bars)
+
+    # select 50K random samples out of it
+    if nSamples == 0:
+        nSamples = featsNorm.shape[0]
+    if useFirsts:
+        featsNorm = featsNorm[:nSamples]
+        r = range(featsNorm.shape[0]) # still randomize
+        np.random.shuffle(r)
+        featsNorm = featsNorm[r[:]]
+    else:
+        r = range(featsNorm.shape[0])
+        np.random.shuffle(r)
+        featsNorm = featsNorm[r[:nSamples]]
+
+    # train a codebook of size 100
+    codebook,dists = VQutils.online_vq(featsNorm,100,lrate=1e-2,nIter=200):
+        
+    # TESTING
+    # go to the folder of test features (per beat)
+    cd testFeatsDir
+    
+    # load and test
+    dists,avg_dists = demos.load_and_encode_data(codebook,pSize=beats,
+                                                 keyInv=True,
+                                                 downBeatInv=False,bars=bars)
+    
+    # report result (average sqaure distance per ... pixel?
+    # with print outs to know what we are doing
+    print 'beats:',beats,', bars:',bars,', nCodes:',nCodes,', nSamples:',nSamples
+    if useFirsts:
+        print 'we use firsts ', nCodes, ' samples'
+    print 'np.average(avg_dists):',np.average(avg_dists)
+
+
+
+
+def die_with_usage():
+    print 'launch all experiments set in main'
+    print 'DONT FORGET TO HARDCODE PATHS'
+    print 'usage:'
[…]

### 2010-03-10 PythonSrc/for_ron.py
all experiments should be there

@@ -11,8 +11,8 @@ import demos
 import numpy as np
 
 
-featsDir = 
-testFeatsDir = 
+featsDir = /proj/hog7/cowbell43k/beatFeats
+testFeatsDir = ~/uspop_mat
 
 
 def do_experiment(beats,bars,nCodes,nSamples=0,useFirsts=False):
@@ -55,10 +55,12 @@ def do_experiment(beats,bars,nCodes,nSamples=0,useFirsts=False):
     
     # report result (average sqaure distance per ... pixel?
     # with print outs to know what we are doing
+    print 'EXPERIMENT REPORT ******************************'
     print 'beats:',beats,', bars:',bars,', nCodes:',nCodes,', nSamples:',nSamples
     if useFirsts:
         print 'we use firsts ', nCodes, ' samples'
     print 'np.average(avg_dists):',np.average(avg_dists)
+    print '************************************************'
 
 
 
@@ -71,9 +73,16 @@ def die_with_usage():
     sys.exit()
 
 
+
+if __name__ == '__main__':
+
+    if len(sys.argv) < 2:
+        die_with_usage()
+
     # EXPERIMENT SET 1: 1 bar 4 beats change data size
     # exp1
     do_experiment(4,1,100,nSamples=1000)
+    sys.exit(0)
     # exp2
     do_experiment(4,1,100,nSamples=5000)
     # exp3
@@ -110,4 +119,46 @@ def die_with_usage():
     do_experiment(8,2,100,nSamples=1000000)
     # exp9
     do_experiment(8,2,100,nSamples=2000000)
+
+
+    # EXPERIMENT SET 3: 0 bar 4 beats change data size 
+    # exp1
+    do_experiment(4,0,100,nSamples=1000)
+    # exp2
+    do_experiment(4,0,100,nSamples=5000)
+    # exp3
+    do_experiment(4,0,100,nSamples=10000)
+    # exp4
+    do_experiment(4,0,100,nSamples=50000)
+    # exp5
+    do_experiment(4,0,100,nSamples=100000)
+    # exp6
+    do_experiment(4,0,100,nSamples=250000)
+    # exp7
+    do_experiment(4,0,100,nSamples=500000)
+    # exp8
+    do_experiment(4,0,100,nSamples=1000000)
+    # exp9
+    do_experiment(4,0,100,nSamples=2000000)
+
+
[…]

### 2010-03-10 PythonSrc/for_ron.py
typo fixed

@@ -11,8 +11,8 @@ import demos
 import numpy as np
 
 
-featsDir = /proj/hog7/cowbell43k/beatFeats
-testFeatsDir = ~/uspop_mat
+featsDir = '/proj/hog7/cowbell43k/beatFeats'
+testFeatsDir = '~/uspop_mat'
 
 
 def do_experiment(beats,bars,nCodes,nSamples=0,useFirsts=False):

### 2010-03-10 PythonSrc/for_ron.py
typo fixed

@@ -22,7 +22,7 @@ def do_experiment(beats,bars,nCodes,nSamples=0,useFirsts=False):
 
     # TRAINING
     # go to the folder of features (per beat)
-    cd featsDir
+    os.chdir(featsDir)
 
     # load everything, unit: 1 bar, resized to 4 beats
     # key invariant, not downbeatinvariant
@@ -46,7 +46,7 @@ def do_experiment(beats,bars,nCodes,nSamples=0,useFirsts=False):
         
     # TESTING
     # go to the folder of test features (per beat)
-    cd testFeatsDir
+    os.chdir(testFeatsDir)
     
     # load and test
     dists,avg_dists = demos.load_and_encode_data(codebook,pSize=beats,

### 2010-03-10 PythonSrc/for_ron.py
typo fixed

@@ -42,7 +42,7 @@ def do_experiment(beats,bars,nCodes,nSamples=0,useFirsts=False):
         featsNorm = featsNorm[r[:nSamples]]
 
     # train a codebook of size 100
-    codebook,dists = VQutils.online_vq(featsNorm,100,lrate=1e-2,nIter=200):
+    codebook,dists = VQutils.online_vq(featsNorm,100,lrate=1e-2,nIter=200)
         
     # TESTING
     # go to the folder of test features (per beat)

### 2010-03-10 PythonSrc/for_ron.py
typo fixed

@@ -42,7 +42,7 @@ def do_experiment(beats,bars,nCodes,nSamples=0,useFirsts=False):
         featsNorm = featsNorm[r[:nSamples]]
 
     # train a codebook of size 100
-    codebook,dists = VQutils.online_vq(featsNorm,100,lrate=1e-2,nIter=200)
+    codebook,dists = VQutils.online_vq(featsNorm,nCodes,lrate=1e-2,nIter=200)
         
     # TESTING
     # go to the folder of test features (per beat)

### 2010-03-10 PythonSrc/for_ron.py
typo fixed

@@ -12,7 +12,7 @@ import numpy as np
 
 
 featsDir = '/proj/hog7/cowbell43k/beatFeats'
-testFeatsDir = '~/uspop_mat'
+testFeatsDir = '/home/empire6/thierry/uspop_mat'
 
 
 def do_experiment(beats,bars,nCodes,nSamples=0,useFirsts=False):

### 2010-03-10 PythonSrc/for_ron.py
tiny code improvement

@@ -33,9 +33,10 @@ def do_experiment(beats,bars,nCodes,nSamples=0,useFirsts=False):
         nSamples = featsNorm.shape[0]
     if useFirsts:
         featsNorm = featsNorm[:nSamples]
-        r = range(featsNorm.shape[0]) # still randomize
-        np.random.shuffle(r)
-        featsNorm = featsNorm[r[:]]
+        #r = range(featsNorm.shape[0]) # still randomize
+        #np.random.shuffle(r)
+        #featsNorm = featsNorm[r[:]]
+        np.random.shuffle(featsNorm)
     else:
         r = range(featsNorm.shape[0])
         np.random.shuffle(r)

### 2010-03-10 PythonSrc/VQutils.py
better error measure during training, but should affect much the current algorithm

@@ -473,6 +473,7 @@ def online_vq(feats,K,lrate,nIter=10,thresh=1e-7,maxRise=.05,scale=False):
             idx = idxs[0]
             weight = weights[idx]
             dist = dists[idx]
+            dist = dist * dist # squared distance, closer to what we measure
             # get that code closer by some learning rate
             codebook[idx,:] += (pattern / weight - codebook[idx,:]) * lrate
             if scale:
@@ -497,6 +498,6 @@ def online_vq(feats,K,lrate,nIter=10,thresh=1e-7,maxRise=.05,scale=False):
         prev_sum_dist = sum_distance
         
     # return codebook, average distance
-    return codebook,(sum_distance * 1. / nFeats)#, best_code_per_pattern
+    return codebook,(sum_distance * 1. / nFeats)
 
 

### 2010-03-10 PythonSrc/feats_utils.py
made pytables import optional (this will cause bugs in the code that requires pytables)

@@ -21,8 +21,11 @@ import scipy.signal
 #import matplotlib.pyplot as P 
 import numpy as N
 import glob
-import tables
 
+try:
+    import tables
+except ImportError:
+    print 'Cannot import pytables'
 
 try:
     from pyechonest import config

### 2010-03-10 PythonSrc/for_ron.py
updated experiment script to use multiprocessing and save output

@@ -3,22 +3,34 @@ Simple code to run a typical experiment
 MUST change the two lines starting with cd
 """
 
-
 import os
+import multiprocessing
 import sys
+
+import numpy as np
+import scipy as sp
+import scipy.io
+
 import VQutils
 import demos
-import numpy as np
 
+featsDir = os.path.expanduser('~/projects/ismir10-patterns/beatFeats')
+testFeatsDir = os.path.expanduser('~/projects/ismir10-patterns/uspop_mat')
+outputDir = os.path.expanduser('~/projects/ismir10-patterns/experiments')
 
-featsDir = '/proj/hog7/cowbell43k/beatFeats'
-testFeatsDir = '/home/empire6/thierry/uspop_mat'
-
-
-def do_experiment(beats,bars,nCodes,nSamples=0,useFirsts=False):
+def do_experiment(experiment_dir,beats,bars,nCodes,nSamples=0,useFirsts=False):
     """
     Performs an independant experiment!!!!
     """
+    try:
+        os.makedirs(experiment_dir)
+    except OSError:
+        pass
+
+    args = dict(experiment_dir=experiment_dir, beats=beats, bars=bars,
+                nCodes=nCodes, nSamples=nSamples, useFirsts=useFirsts)
+    sp.io.savemat(os.path.join(experiment_dir, 'args.mat'), args)
+
 
     # TRAINING
     # go to the folder of features (per beat)
@@ -44,7 +56,9 @@ def do_experiment(beats,bars,nCodes,nSamples=0,useFirsts=False):
 
     # train a codebook of size 100
     codebook,dists = VQutils.online_vq(featsNorm,nCodes,lrate=1e-2,nIter=200)
-        
+    sp.io.savemat(os.path.join(experiment_dir, 'codebook.mat'),
+                  dict(codebook=codebook, dists=dists))
+            
     # TESTING
     # go to the folder of test features (per beat)
     os.chdir(testFeatsDir)
@@ -53,16 +67,23 @@ def do_experiment(beats,bars,nCodes,nSamples=0,useFirsts=False):
     dists,avg_dists = demos.load_and_encode_data(codebook,pSize=beats,
                                                  keyInv=True,
                                                  downBeatInv=False,bars=bars)
+    sp.io.savemat(os.path.join(experiment_dir, 'test.mat'),
[…]

### 2010-03-10 Paper/Paper.tex
improved
details

@@
 is a pattern of stacked chroma features. Similarity between
-\textit{Shingles} coresponds to comparing harmonic patterns. For this
+\textit{Shingles} coresponds to similarity between harmonic patterns. For this
 reason, they were used for cover or remix recognition \cite{Casey2007}.
@@
 
+We investigate a vector quantization \cite{Datar2004} (VQ) method to
+cluster the \textit{Shingles}. We create a codebook of typical patterns
+that have many usages:
+\begin{enumerate}
+\item describing the music space
+\item finding potential similar songs and/or covers
+\item eliminating quickly very wrong matches in a query system
+\end{enumerate}
+
 A major component of this work is scalability. We experiment with a training
@@
 
-\iffalse % ignored section
-The original goal of the \textit{Shingles} was retrieval, i.e. find very
-similar song segments based on their harmonic pattern. Part of the
-challenge was to handle tons of overlapping song segments represented
-by a rather long feature vector (a serie of chroma features).
-The LSH algorithm proved a great solution for performing a $k$-nearest
-neighbors ($k$nn) search in such a space. However, the paradigm of $k$nn has
-an intrinsec scaling limit.
-
-Finding a closest match is a tangible goal within a personal collection
-of a few thousands songs. It can be used to discover cover songs or
-find the seamless transitions in a playlist. However, the closest match
-in millions and millions of music excerpts do not help us organize such
-an amount of media. Some tasks are still important, for instance cover
-recognition, and improvements in $k$nn methods are always welcome.
-But it does not the describe the music space, nor does it necessarily
-make good recommendations. We do not want to look for one neighbor, but
-rather a cloud of potential close patterns. A similar argument in the 
-recommendation domain is that precision becomes more important than recall 
-passed a certain scale of data for many practical uses.
-
-We propose to move the \textit{Shingle} idea into the larger clustering 
-paradigm.
-\fi % end of ignored section
-
-The goal is to find typical patterns that can be seen as cluster centroids.
-Those typical patterns help us understand a music collection from an
-harmonic point of view. Segments of songs can be described in regard to their
-proximity to the different centroids, thus creating some notion of distance
-among the song segments. These cluster of similar patterns can have many usages:
-\begin{enumerate}
-\item describing the music space
-\item finding potential similar songs and/or covers
-\item eliminating quickly very wrong matches in a query system
-\end{enumerate}
-
-Our contribution is... large-scale clustering method,
-
-We present a practical way to do such large-scale clustering using vector
-quantization. We present clustering results of an online algorithm on a data 
-set of $43$K songs, but the method applies to any data size. We show how to
-select different parameters such as the size of the \textit{Shingles} 
-using an encoding scheme on a test set of $4$K songs. Finally, we illustrate
-the method with some specific examples.
+In this paper, we present a practical way to do such large-scale 
+clustering using online vector quantization. We present clustering results 
+of an online algorithm on a data set of $43$K songs, but the method applies 
+to any data size. We show how to select different parameters such as the size 
+of the \textit{Shingles} using an encoding scheme on a test set of $8$K songs. 
+Finally, we illustrate the method with some specific examples.
 
@@
 algorithm and most likely does not come from an actual song.
-Distortion is our measure of distance between a pattern and a code.
+Distortion is our measure of distance between a pattern and a code
+(or between two patterns, see equation \ref{eq:dist} in Section 
+\ref{sec:experiments}).
 
@@
 or ``segment'', for any song we upload to their platform. 
-We also get an estimate of where the beats and the bars are. 
-We transform the chroma vectors per segment into per beat vector using a 
-simple average. We can then stack beat vectors for a number of bars (usually 1 or 2). 
+We also get an estimate of where the beats and the bars are. Beats span over
+multiple segments, and bars span over multiple beats. 
+We transform the chroma vectors per segment into per beat vectors using a 
+simple average. We can then stack beat vectors for a number of bars 
+(usually 1 or 2). 
 We get a fixed-size patch by resampling the number of chroma vectors. Typically,
-there are $4$ beats per bar. Thus an appropriate patch size for one bar
+there are $4$ beats per bar. Thus an appropriate pattern size for one bar
 is $12$x$4$ or $12$x$8$.
@@
 Note that we do not claim that any of these informations (segments, beats, bars)
-are perfectly accurate (e.g. in \cite{Barrington2009a} authors improve on the
-song segmentation features). Practise showed us that they are reasonable, 
+are perfectly accurate (e.g. in \cite{Barrington2009a} authors obtain a better
+song segmentation than Echonest). Practise showed us that they are reasonable, 
 and the size of the data set should make up for the imperfections or noise.
@@
 We have $43,300$ songs, giving  $3,720,091$ non zero bars, a bar usually 
-contains $4$ beats.
+contains $4$ beats. WHERE DOES THIS DATA COMES FROM? GIFT FROM EN?
 
@@
 experiment section.
+
+We try to normalize the patterns regarding the key. We roll the matrix
+so the row with the more energy is always the lowest one. This can easily be
+seen in Figure \ref{fig:code}. We considered the same technique for rolling
+along the other axis and be invariant regarding the downbeat. However, it
+seems less robust to noise. Note that we roll each pattern independentely.
+Also, we imply that the pitch envelope is important rather than the actual 
+pitches.
 
@@
 \section{Experiments}\label{sec:experiments}
+In this section we present general experiments that describe our method.
+We start by giving some details on how we selected our training set and
+how to measure error. Secondly, we experiment with different pattern sizes.
+Thirdly, we look at the effect of the size of the
+data set on the training. Then, we show that the patterns are not random,
+which is expected from music data. Finally, we show examples of patterns,
+code books and clusters.
 
@@
 
+As a distance measure (or distortion measure) between patterns, we use
+the average squared euclidean distance. In details, if a pattern $p1$
+is composed of elements $p1_{i,j}$, and similarly for a pattern $p2$,
+the distance between $p1$ and $p2$ is:
+\begin{eqnarray}
+  dist(p1,p2) = \sum_{i,j} (p1_{i,j} - p2_{i,j})^2 / size(p1)  \label{eq:dist}
+\end{eqnarray}
+We assume $p1$ and $p2$ have the same size.
+
 \subsection{Pattern length}
@@
 % RESULTS
-% 200codes, 2 bars, pSize=16, maxener, True, False, not scaled = 3.788
-%                                             averaged on size = 
-% 200codes, 2 bars, pSize=8, maxener, True, False, not scaled  = 3.135
-%                                             averaged on size = 
+% 200codes, 2 bars, pSize=16, maxener, True, False, not scaled = 
+% 200codes, 2 bars, pSize=8, maxener, True, False, not scaled  = 0.
 % 200codes, 1 bars, pSize=4, maxener, True, False, not scaled  = 
-%                                             averaged on size = 
-% 200codes, 7 beats, pSize=7, maxener, True, False, not scaled = 2.771
-%                                             averaged on size = 
-% 200codes, 5 beats, pSize=5, maxener, True, False, not scaled = 2.397
-%                                             averaged on size = 
+% 200codes, 7 beats, pSize=7, maxener, True, False, not scaled = 
+% 200codes, 5 beats, pSize=5, maxener, True, False, not scaled = 
 % 200codes, 3 beats, pSize=3, maxener, True, False, not scaled = 
-%                                             averaged on size = 
 \begin{table}
@@
 \# codes & \# samples & avg. dist \\ \hline \hline
-1 & $500$ & $0.195939$ \\
-10 & $5$K & $0.161812$ \\
-50 & $25$K & $0.147472$ \\
-100 & $50$K & $0.142194$ \\
-200 & $100$K & $0.137548$ \\
-500 & $250$K & $0.131211$ \\
+1 & $500$ & $0.066079$ \\
+10 & $5$K & $0.045690$ \\
+50 & $25$K & $0.038456$ \\
+100 & $50$K & $0.036725$ \\
+200 & $100$K & $0.033923$ \\
+500 & $250$K & $$ \\
 1000 & $500$K & \\ \hline
@@
 % code 100      0.142194        0.036725
-% code 200      0.137548        0.
+% code 200      0.137548        0.033923
 % code 500      0.131211        0.
@@
 16 & $2$ & $0.050172$ \\
-256 & $4$ & $0.04989$ \\ \hline
+256 & $4$ & $0.049894$ \\ \hline
 \end{tabular}
@@
 %256 codes 4 bars    0.174169       0.049894
+
+% see generate_expo_codes_curves.py to generate the following figure:
+\begin{figure}[htb]
+\begin{center}
+\includegraphics[width=.99\columnwidth]{codesize_patternsize}
+\end{center}
+\caption{{Possible figure to replace Table \ref{tab:size_pattern}.}}
+\label{fig:size_pattern}
+\end{figure}
 
@@
 %Thierry is NSERC graduate fellow, or some title like that.
+% Graham for the NMF stuff, unless is an author.
 

### 2010-03-10 PythonSrc/for_ron.py
fixed typo

@@ -120,6 +120,6 @@ if __name__ == '__main__':
         die_with_usage()
 
     nprocesses = int(sys.argv[2])
-    pool = multiprocessing.Pool(processes=nprocesses)s
+    pool = multiprocessing.Pool(processes=nprocesses)
     pool.map(do_experiment_wrapper, experiment_args)
 

### 2010-03-10 Paper/Paper.tex
improved

@@
 \end{center}
-\caption{{A typical code from a code book of size $200$. Patterns represented
+\caption{{A typical code from a codebook of size $200$. Patterns represented
 $2$ bars and the pattern length was set to $16$.}}
@@
 
-uspop2002 serves as testing set to measure how well a code book learned on
+uspop2002 serves as testing set to measure how well a codebook learned on
 the Cowbell data set can represent new songs. We get Echo Nest features
@@
 the Echo Nest data. Then, we explain the online vector quantization algorithm
-used to learn a code book of those patterns.
+used to learn a codebook of those patterns.
 
@@
 Thirdly, we look at the effect of the size of the
-data set on the training. Then, we show that the patterns are not random,
+data set on the training. The next experiment look at the size of the
+codebook. Then, we show that the patterns are not random,
 which is expected from music data. Finally, we show examples of patterns,
-code books and clusters.
+codebooks and clusters.
 
@@
 data set.
-Code length: 1 bar resized to size 4. Code book size: 100
+Code length: 1 bar resized to size 4. Codebook size: 100
 We measure the distortion on our test set. We hope for some convergence.
@@
 
-$500$ samples per code in the code book seems a good data size, see
+$500$ samples per code in the codebook seems a good data size, see
 Figure \ref{fig:sizes} for size $50$K.
@@
 \end{center}
-\caption{{Distortion for a code book of size $100$ encoding one bar
+\caption{{Distortion for a codebook of size $100$ encoding one bar
 at a time, a bar being represented by $4$ chroma feature vector.
@@
 \subsection{Codebook Size}
-Results in Table \ref{tab:cbsize}.
+We experiment with the size of the codebook and its affect on the
+performance of the encoding. Results are shown in Table \ref{tab:cbsize}.
+Note that we trained on a number of samples depending on the codebook
+size, i.e. we used $500$ samples per code to learn. This is consistent
+with our findings in Figure \ref{fig:sizes}.
 
@@
 \end{center}
-\caption{{Code book size.}}
+\caption{{Codebook size.}}
 \label{tab:cbsize}
@@
 
-See code book size experiments in Table \ref{tab:cbsize}.
+See codebook size experiments in Table \ref{tab:cbsize}.
 
@@
 \end{center}
-\caption{{Code book size in function of the number of bars to encode.
+\caption{{Codebook size in function of the number of bars to encode.
 All experiments done wtih $256 \times 500 = 128$K samples randomly chosen.
@@
 \end{center}
-\caption{{LLE visualization of the code book.}}
+\caption{{LLE visualization of the codebook.}}
 \label{fig:lle}
@@
 code than any other code. In Figure \ref{fig:firstcodes}, we show the $25$
-most used codes out of a code book of size $200$.
+most used codes out of a codebook of size $200$.
 Code from Figure \ref{fig:code} is the 37th most used code in that
-code book. In Figure \ref{fig:closep}, we show a sample of patterns that
+codebook. In Figure \ref{fig:closep}, we show a sample of patterns that
 fall into its cluster.
@@
 NMF provides us with a way of measuring how much information is retained
-in our code book. More specifically, we can train a code book of size $500$
+in our codebook. More specifically, we can train a codebook of size $500$
 using our online VQ algorithm, then decompose that code into $10$ basis

### 2010-03-10 PythonSrc/for_ron.py
added random seed

@@ -18,7 +18,7 @@ featsDir = os.path.expanduser('~/projects/ismir10-patterns/beatFeats')
 testFeatsDir = os.path.expanduser('~/projects/ismir10-patterns/uspop_mat')
 outputDir = os.path.expanduser('~/projects/ismir10-patterns/experiments')
 
-def do_experiment(experiment_dir,beats,bars,nCodes,nSamples=0,useFirsts=False):
+def do_experiment(experiment_dir,beats,bars,nCodes,nSamples=0,useFirsts=False,seed=0):
     """
     Performs an independant experiment!!!!
     """
@@ -27,8 +27,10 @@ def do_experiment(experiment_dir,beats,bars,nCodes,nSamples=0,useFirsts=False):
     except OSError:
         pass
 
+    np.random.seed(seed)
+
     args = dict(experiment_dir=experiment_dir, beats=beats, bars=bars,
-                nCodes=nCodes, nSamples=nSamples, useFirsts=useFirsts)
+                nCodes=nCodes, nSamples=nSamples, useFirsts=useFirsts,seed=seed)
     sp.io.savemat(os.path.join(experiment_dir, 'args.mat'), args)
 
 

### 2010-03-11 PythonSrc/for_ron.py
fixed another typo

@@ -81,7 +81,7 @@ def do_experiment(experiment_dir,beats,bars,nCodes,nSamples=0,useFirsts=False,se
         report.append(['we use firsts %s samples' % nCodes])
     report.extend(['np.average(avg_dists): %s' % np.average(avg_dists),
                    '************************************************'])
-    reportstr = '\n'.join(reportstr)
+    reportstr = '\n'.join(report)
     print reportstr
     f = open(os.path.join(experiment_dir, 'report.txt'))
     f.write(reportstr)

### 2010-03-11 PythonSrc/for_ron.py
fixed another bug

@@ -60,6 +60,8 @@ def do_experiment(experiment_dir,beats,bars,nCodes,nSamples=0,useFirsts=False,se
     codebook,dists = VQutils.online_vq(featsNorm,nCodes,lrate=1e-2,nIter=200)
     sp.io.savemat(os.path.join(experiment_dir, 'codebook.mat'),
                   dict(codebook=codebook, dists=dists))
+
+    del featsNorm
             
     # TESTING
     # go to the folder of test features (per beat)
@@ -83,7 +85,7 @@ def do_experiment(experiment_dir,beats,bars,nCodes,nSamples=0,useFirsts=False,se
                    '************************************************'])
     reportstr = '\n'.join(report)
     print reportstr
-    f = open(os.path.join(experiment_dir, 'report.txt'))
+    f = open(os.path.join(experiment_dir, 'report.txt'), 'w')
     f.write(reportstr)
     f.close()
 

### 2010-03-11 PythonSrc/for_ron.py
update experiment code to run one "experiment set" at a time

@@ -95,35 +95,38 @@ def die_with_usage():
     print 'launch all experiments set in main'
     print 'DONT FORGET TO HARDCODE PATHS'
     print 'usage:'
-    print '   python for_ron -go nprocesses'
+    print '   python for_ron -go nprocesses experiment_set_number[1-4]'
     sys.exit()
 
-
-data_sizes = [1000, 5000, 10000, 50000, 100000, 250000, 500000, 1000000,
-              2000000]
-experiment_args = []
-# EXPERIMENT SET 1: 1 bar 4 beats change data size
-experiment_args.extend([(os.path.join(outputDir, 'set1exp%d' % n), 4,1,100,x)
-                        for n,x in enumerate(data_sizes)])
-# EXPERIMENT SET 2: 2 bar 8 beats change data size
-experiment_args.extend([(os.path.join(outputDir, 'set2exp%d' % n), 8,2,100,x)
-                        for n,x in enumerate(data_sizes)])
-# EXPERIMENT SET 3: 0 bar 4 beats change data size 
-experiment_args.extend([(os.path.join(outputDir, 'set3exp%d' % n), 4,0,100,x)
-                        for n,x in enumerate(data_sizes)])
-# EXPERIMENT SET 4: 1 bar 4 beats change data size, use first samples
-experiment_args.extend([(os.path.join(outputDir, 'set4exp%d' % n), 4,1,100,x,True)
-                        for n,x in enumerate(data_sizes)])
 
 def do_experiment_wrapper(args):
     return do_experiment(*args)
 
-if __name__ == '__main__':
 
-    if len(sys.argv) < 3:
+data_sizes = [1000, 5000, 10000, 50000, 100000, 250000, 500000, 1000000,
+              2000000]
+experiment_args = [
+    # EXPERIMENT SET 1: 1 bar 4 beats change data size
+    [(os.path.join(outputDir, 'set1exp%d' % n), 4,1,100,x)
+     for n,x in enumerate(data_sizes)],
+    # EXPERIMENT SET 2: 2 bar 8 beats change data size
+    [(os.path.join(outputDir, 'set2exp%d' % n), 8,2,100,x)
+     for n,x in enumerate(data_sizes)],
+    # EXPERIMENT SET 3: 0 bar 4 beats change data size 
+    [(os.path.join(outputDir, 'set3exp%d' % n), 4,0,100,x)
+     for n,x in enumerate(data_sizes)],
[…]

### 2010-03-11 PythonSrc/for_ron.py
added newline at end of experiment report

@@ -82,7 +82,7 @@ def do_experiment(experiment_dir,beats,bars,nCodes,nSamples=0,useFirsts=False,se
     if useFirsts:
         report.append(['we use firsts %s samples' % nCodes])
     report.extend(['np.average(avg_dists): %s' % np.average(avg_dists),
-                   '************************************************'])
+                   '************************************************', ''])
     reportstr = '\n'.join(report)
     print reportstr
     f = open(os.path.join(experiment_dir, 'report.txt'), 'w')

### 2010-03-11 PythonSrc/for_ron.py
added newexperiments

@@ -95,7 +95,7 @@ def die_with_usage():
     print 'launch all experiments set in main'
     print 'DONT FORGET TO HARDCODE PATHS'
     print 'usage:'
-    print '   python for_ron -go nprocesses experiment_set_number[1-4]'
+    print '   python for_ron -go nprocesses experiment_set_number(1-5)'
     sys.exit()
 
 
@@ -117,7 +117,16 @@ experiment_args = [
      for n,x in enumerate(data_sizes)],
     # EXPERIMENT SET 4: 1 bar 4 beats change data size, use first samples
     [(os.path.join(outputDir, 'set4exp%d' % n), 4,1,100,x,True)
-     for n,x in enumerate(data_sizes)]]
+     for n,x in enumerate(data_sizes)],
+    # EXPERIMENT SET 5: table 1 in the paper 
+    [(os.path.join(outputDir, 'set5exp0'), 3, 0, 100, 50000),
+     (os.path.join(outputDir, 'set5exp1'), 5, 0, 100, 50000),
+     (os.path.join(outputDir, 'set5exp2'), 7, 0, 100, 50000),
+     (os.path.join(outputDir, 'set5exp3'), 4, 1, 100, 50000),
+     (os.path.join(outputDir, 'set5exp4'), 8, 2, 100, 50000),
+     (os.path.join(outputDir, 'set5exp5'), 12, 2, 100, 50000),
+     (os.path.join(outputDir, 'set5exp6'), 16, 2, 100, 50000)],
+    ]
     
 if __name__ == '__main__':
     if len(sys.argv) < 4:

### 2010-03-11 Paper/Paper.tex

@@
 % 200codes, 2 bars, pSize=16, maxener, True, False, not scaled = 
-% 200codes, 2 bars, pSize=8, maxener, True, False, not scaled  = 0.
+% 200codes, 2 bars, pSize=8, maxener, True, False, not scaled  = 0.042432
 % 200codes, 1 bars, pSize=4, maxener, True, False, not scaled  = 
@@
 We are interesting to see how the amount of training data changes
-the quality of our code.
+the quality of our codebook.
 We experiment with different data sizes for a fixed size of codebook and a 
@@
 data set.
-Code length: 1 bar resized to size 4. Codebook size: 100
+In the experiment a pattern size correspond to $1$ bar resized to $4$
+beats, thus a pattern size of $12$x$4$. The codebook size is $100$.
 We measure the distortion on our test set. We hope for some convergence.
@@
 % 1bar s4     avg_distOLD    avg_dist
-% 1K          0.150370
-% 5K          0.145327
-% 10K         0.143567
-% 50K         0.142257
+% 1K          0.150370       0.041941
+% 5K          0.145327       0.038449
+% 10K         0.143567       0.036892
+% 50K         0.142257       0.036722
 % 100K        0.142195       
@@
 % 0bar s4 (and test set same, we could do 1bar s4 in test set...)
-% 1K          0.155474
+% 1K          0.155474       0.042596
 % 5K          0.149619
@@
 % 2bar s8     
-% 1K          0.171300
-% 5K          0.165960
-% 10K         0.165063
+% 1K          0.171300       0.049282
+% 5K          0.165960       0.046441
+% 10K         0.165063       0.045507
 % 50K         0.162439
@@
 \end{center}
-\caption{{Codebook size.}}
+\caption{{Experiment with the codebook size. We train on patterns
+corresponding to $1$ bar resized to $4$ beats. The number of training
+samples is set to $500$ times the number of codes, which is sufficient
+according to Figure \ref{fig:sizes}.}}
 \label{tab:cbsize}
@@
 \subsection{Codebook Size in Function of Pattern Length}
-If the signal was totally random, and we were using 2 code to encode
-every beat, we would need.... 4 codes for 2 beats? and 16 codes for
-4 beats? and 256 codes for 8 beats?
+We experiment on the randomness of the \textit{Shingles}, assuming
+that there is some correlation between beats and bars.
+The intuition is as follow: if the signal was totally random, every
+time the signal size double, the codebook size must be squared.
+To illustrate this, imagine a binary array of size $1$. We need two
+codes to encode perfectly: $(0)$ and $(1)$. If the binary is now of
+size $2$, we need four codes: $(0,0)$, $(0,1)$, $(1,1)$ and $(1,0)$.
 
-Think about it in binary: two code to encode (0) and (1), but for
-(0,0),(0,1),(1,0),(1,1) we need four codes for the same accuracy.
+In Table \ref{tab:size_pattern}, we test that hypothesis. As expected,
+the distortion improves with larger pattern sizes. The details
+are difficult to analyze. The error jump from two beats to one bar
+can probably be explained by the obvious case of sustained notes.
+We also believe that smaller codebooks are more affected by their
+initialization. Fortunately, we can conclude that music is not random,
+which is always a good sanity check!
 
-Reasonable size: 2 - 4 - 16 - 256. If we improve distance, music
-is not random!
-
-Then... 3 beats, 5 beats, 7 beats don't encode as well!
+Then... 3 beats, 5 beats, 7 beats don't encode as well?
 

### 2010-03-11 Paper/Paper.tex
improvements

@@
 time the signal size double, the codebook size must be squared.
-To illustrate this, imagine a binary array of size $1$. We need two
-codes to encode perfectly: $(0)$ and $(1)$. If the binary is now of
-size $2$, we need four codes: $(0,0)$, $(0,1)$, $(1,1)$ and $(1,0)$.
+To illustrate this, imagine a binary array of length $1$. We need two
+codes to encode perfectly: $(0)$ and $(1)$. If the binary array is now of
+length $2$, we need four codes: $(0,0)$, $(0,1)$, $(1,1)$ and $(1,0)$.
 
@@
 \subsection{Visualization}
+Interpreting a codebook of size larger than $100$ is not obvious.
 We can use an algorithm like locally linear embedding \cite{Roweis2000} (LLE) 
@@
 
-In all our experiments we use Matlab's implementation of NMF:
-nmflib\footnote{implementation: \url{http://www.ee.columbia.edu/~grindlay/code.html}}.
+In all our experiments we use the following NMF
+implementation\footnote{implementation: \url{http://www.ee.columbia.edu/~grindlay/code.html}}.
 
-\subsection{Training on our Codebook}
+\subsection{Training NMF on our Codebook}
 NMF provides us with a way of measuring how much information is retained
@@
 \section{Conclusion}
-We are awesome. Here's how:
-\begin{itemize}
-\item large-scale
-\item smarter shingles, based on beats and bars
-\item clusters good for encoding
-\item clusters good for something else?
-\item free online API, code available\footnote{code not released yet to preserve 
-submission's anonymity}
-\item we have lab t-shirts
-\end{itemize}
+We presented a practical method to perform large-scale clustering of
+harmonic patterns. We assessed the basic properties of the method through
+experiments on a large collection of music. We suggested many ways
+to interpret the data through visualization (LLE) and further encoding
+(NMF). We discussed the possibility to move to even larger scales
+and we provided our source code\footnote{Code not yet released to preserve
+submission's anonymity}.
 

### 2010-03-11 Paper/Paper.tex
improvements

@@
   {Third author} {Affiliation3 \\ {\tt author3@ismir.edu}}
-% what order do we use? Thierry, Ron, Dan? is Juan on?
+% what order do we use? Thierry, Ron, Dan? is Juan on? is Graham on?
 
@@
 We present a method for discovering typical patterns in music.
-Similar to the \textit{shingle} idea, but done on a larger scale through
-Echo Nest API, we cluster harmonic patterns from 43K songs. Patterns are
-made of ordered chroma features, and not of summarization like gaussian
-mixtures coefficients or NMF weights. We explain
-how to operate on such large datasets, we analyze the obained clusters,
-and we discuss there use for retrieval and encoding. All code is made
-available.\\
+Following up on the \textit{Shingle} idea, we bring these harmonic patterns
+into the clustering framework. We use EchoNest API and a training set
+of $43$K songs to assert our method. We restrain ourselves to an online
+algorithm, thus it can scale to arbitrary data sizes.
+We assess the basic good behaviours of our algorithm on this harmonic
+data. We explain how to interpret the centroids of the clusters through
+visualization and further encoding. We discuss ways to move to even larger
+scales and we provide our code for reproducing the results.
 TAKE-HOME MESSAGE: \textit{Shingles} idea can go large-scale?
@@
 of this work, an algorithm ``listening'' to all music it can find online.
-Time would be an issue, but not memory.
+Training time would be an issue, but not memory.
 

### 2010-03-11 Paper/Paper.tex
improvements

@@
 we can ``listen'' to the patches (see Section \ref{sec:experiments}).
-This is an advantage for analyzing a large data set, and in accordance with our goal
+This is an advantage for analyzing a large data set, thus
 we restrain ourselves from using generative models to compress our features.
@@
 
-We try to normalize the patterns regarding the key. We roll the matrix
+We try to normalize the patterns in respect to the key. We roll the matrix
 so the row with the more energy is always the lowest one. This can easily be
@@
 seems less robust to noise. Note that we roll each pattern independentely.
-Also, we imply that the pitch envelope is important rather than the actual 
+Also, this implies that the pitch envelope is important rather than the actual 
 pitches.
@@
 We use an online version of the vector quantization algorithm 
-\cite{Gersho1991}. This can also be seen as an online Kmeans.
+\cite{Gersho1991}. This can also be seen as an online K-means.
 The idea is as follow: take a sample from your data, find the closest
 code in the codebook, bring that code closer to the sample by some amount.
-The details are explained as Algorithm \ref{algo:vq}. note that this
+The details are explained in Algorithm \ref{algo:vq}. Note that this
 algorithm, though not optimal, is $O(n)$ with $n$ number of songs seen.
@@
 distance between a pattern and its encoding. We can also measure the use
-of the different codes.
+of the different codes, i.e. the size of the cluster around a code.
 
@@
 beats, thus a pattern size of $12$x$4$. The codebook size is $100$.
-We measure the distortion on our test set. We hope for some convergence.
+We measure the distortion on our test set. 
 See results in Figure \ref{fig:sizes}.
@@
 size, i.e. we used $500$ samples per code to learn. This is consistent
-with our findings in Figure \ref{fig:sizes}.
+with our findings from Figure \ref{fig:sizes}.
 
@@
 or multidimensional scaling \cite{Kruskal1964} to visualize the data in
-2 dimensions. We use the former\footnote{implementation: 
+2 dimensions. Here we use the former\footnote{implementation: 
 \url{http://www.astro.washington.edu/users/vanderplas/coding/LLE/}}, the
@@
 sample per line) by
-\[
-D = W \times H + \epsilon
-\]
+\begin{eqnarray}
+D \approx W^{m \times k} H^{k \times n}
+\end{eqnarray}
 where $\epsilon$ is some noise and $W$ and $H$ only have non-negative 
-elements. If $W$ is $\#samples \times n$, $n$ is the codebook size.
+elements. If $W$ is $\#samples \times k$, $k$ is the codebook size.
 
@@
 
+Moving on to larger scales, we could imagine merging codes together and
+splitting some on two if the clusters become too small or too large.We could
+also cluster the codebook itself, in a similar fashion as hierarchical
+gaussian mixtures \cite{Vasconcelos2001}.
+
 \small

### 2010-03-11 Paper/Paper.tex
improvements

@@
 \section{Previous Work}\label{sec:prevwork}
-The idea is closely related to the \textit{Shingles} described by
-Slaney and Casey \cite{Casey2006,Casey2007,Casey2008}.
+Our work can be seen as an extension of the \textit{Shingles} idea
+described by Casey and Slaney \cite{Casey2006,Casey2007,Casey2008}.
+The two main differences are the clustering framework and the use of
+beats and bars to select the position of the \textit{Shingles}.
 
-LSH: \cite{Datar2004}, LSH implementation: \cite{E2LSH}.
-The E$^2$LSH packages can efficiently handle about $50$K or $100$K
+Casey and Slaney \cite{Casey2006} use LSH \cite{Datar2004} to perform
+fast approximate nearest neighbor search. Their implementation is E$^2$LSH
+\cite{E2LSH}. This package can efficiently handle about $50$K or $100$K
 vectors of size $200$. Even with a better implementation, the LSH
@@
 
+Regarding similar representations of music,
 Barrington et al. recently studied music texture using a set of ordered

### 2010-03-12 Paper/Paper.tex
were up-to-date

@@
   {Third author} {Affiliation3 \\ {\tt author3@ismir.edu}}
-% what order do we use? Thierry, Ron, Dan? is Juan on? is Graham on?
+% what order do we use? Thierry, Ron, Dan? is Graham on?
 
@@
 
+Nearest neighbor and clustering methods are common in music similarity 
+research, for instance \cite{Cano2004,Holzapfel2009}. The difference here 
+are the actual patterns that we cluster.
+
 Regarding similar representations of music,
@@
 
-Nearest neighbor and clustering methods are common in music similarity 
-research, for instance \cite{Cano2004,Holzapfel2009}. The difference here 
-are the actual patterns that we cluster.
-
 Other large-scale work use summarization of the features. We mention 
-gaussian mixtures \cite{Mandel2005}, aggregate features \cite{Bergstra2006a}
-and linear dynamical systems \cite{Barrington2009a}. The chroma representation
-is close from an harmonic point of view to the actual music, in particular
+gaussian mixtures \cite{Mandel2005}, 
+aggregate features \cite{Bertin-Mahieux2008}
+and linear dynamical systems \cite{Barrington2009a}. 
+The advantage of \textit{Shingles} representation (stacked chroma vectors)
+is that it is close from an harmonic point of view to the actual music, 
+in particular
 we can ``listen'' to the patches (see Section \ref{sec:experiments}).
@@
 200 & $100$K & $0.033923$ \\
-500 & $250$K & $$ \\
+500 & $250$K & $0.030973$ \\
 1000 & $500$K & \\ \hline
@@
 % code 200      0.137548        0.033923
-% code 500      0.131211        0.
+% code 500      0.131211        0.030973
 
@@
 
+\begin{table}
+\begin{center}
+\begin{tabular}{|l|l|c|c||c|}
+\hline
+\# codes & \# bars & full data & codebook & test set \\ \hline \hline
+10 & 1 & $$ & $$ & \\ \hline
+20 & 1 & $$ & $$ & \\ \hline
+10 & 2 & $\bf{1774788}$ & $$ & $1782164$\\ \hline
+20 & 2 & $$ & $$ & \\ \hline
+\end{tabular}
+\end{center}
+\caption{{Encoding error of the test data set using NMF basis learned
+on the whole data and on the codebook. We also encode the test set
+directly to give us an upper bound.}}
+\label{tab:nmf}
+\end{table}
+% RESULTS:         whole data             codebook     test set
+%10 codes 2 bars   1774788.2566 (200 it)               1739639.7304 (1000 it)
+%10 codes 1 bar
+
 \section{Conclusion}

### 2010-03-12 PythonSrc/for_ron.py
added experiment for codebook size vs distortion with fixed number of training samples

@@ -126,6 +126,10 @@ experiment_args = [
      (os.path.join(outputDir, 'set5exp4'), 8, 2, 100, 50000),
      (os.path.join(outputDir, 'set5exp5'), 12, 2, 100, 50000),
      (os.path.join(outputDir, 'set5exp6'), 16, 2, 100, 50000)],
+    # EXPERIMENT SET 6: table 2 in paper (codebook size vs distortion with
+    #  constant numbre of training samples)
+    [(os.path.join(outputDir, 'set6exp%d' % n), 4, 1, x, 500000)
+     for n,x in enumerate([1, 10, 50, 100, 200, 500, 1000])],
     ]
     
 if __name__ == '__main__':

### 2010-03-12 PythonSrc/for_ron.py
extended experiment 5 (but called it experiment 7)

@@ -130,6 +130,15 @@ experiment_args = [
     #  constant numbre of training samples)
     [(os.path.join(outputDir, 'set6exp%d' % n), 4, 1, x, 500000)
      for n,x in enumerate([1, 10, 50, 100, 200, 500, 1000])],
+    # EXPERIMENT SET 7: exp 5 (table 1) extended: bar alignment vs taking every n-beats
+    [(os.path.join(outputDir, 'set7exp0'), 2, 0, 100, 50000),
+     (os.path.join(outputDir, 'set7exp1'), 4, 0, 100, 50000),
+     (os.path.join(outputDir, 'set7exp2'), 6, 0, 100, 50000),
+     (os.path.join(outputDir, 'set7exp3'), 8, 0, 100, 50000),
+     (os.path.join(outputDir, 'set7exp4'), 9, 0, 100, 50000),
+     (os.path.join(outputDir, 'set7exp5'), 10, 0, 100, 50000),
+     (os.path.join(outputDir, 'set7exp6'), 11, 0, 100, 50000),   
+     (os.path.join(outputDir, 'set7exp7'), 12, 0, 100, 50000)],   
     ]
     
 if __name__ == '__main__':

### 2010-03-12 PythonSrc/demos.py
added function to visualize one song + encoding

@@ -73,6 +73,43 @@ def get_data_maxener(pSize=16,keyInv=True,downBeatInv=False,bars=2):
     return featsNorm
 
 
+
+def encode_one_song(filename,codebook,pSize=8,keyInv=True,
+                    downBeatInv=False,bars=2):
+    """
+    returns: song, encoding, song as MAT, encoding as MAT
+    """
+    import feats_utils as FU
+    import numpy as np
+    import data_iterator
+    import VQutils
+
+    # create data iterator
+    data_iter = data_iterator.DataIterator()
+    data_iter.setMatfiles([filename]) # set matfiles
+    if bars > 0:
+        data_iter.useBars( bars )            # a pattern spans 'bars' bars
+    else:
+        data_iter.useBars(0)                 # important to set it to zero!
+        data_iter.setFeatsize( pSize )       # a pattern is a num. of beats
+    data_iter.stopAfterOnePass(True)
+    # load data
+    featsNorm = [FU.normalize_pattern_maxenergy(p,pSize,keyInv,downBeatInv).flatten() for p in data_iter]
+    featsNorm = np.array(featsNorm)
+    res = [np.sum(r) > 0 for r in featsNorm]
+    res2 = np.where(res)
+    featsNorm = featsNorm[res2]
+    # find code per pattern
+    best_code_per_p, dists, avg_dists = VQutils.find_best_code_per_pattern(featsNorm,codebook)
+    tmp = [int(p) for p in best_code_per_p]
+    encoding = codebook[tmp]
+    # transform into 2 matrices
+    featsNormMAT = np.concatenate([x.reshape(12,pSize) for x in featsNorm],axis=1)
+    encodingMAT = np.concatenate([x.reshape(12,pSize) for x in encoding],axis=1)
+    # return
+    featsNorm,encoding,featsNormMAT,encodingMAT
+
+
 def get_all_barfeats():
     """
     Returns all barfeats, we assume we're at the top of beatFeats dir

### 2010-03-12 PythonSrc/demos.py
cool plotting and hist in demos

@@ -101,13 +101,20 @@ def encode_one_song(filename,codebook,pSize=8,keyInv=True,
     featsNorm = featsNorm[res2]
     # find code per pattern
     best_code_per_p, dists, avg_dists = VQutils.find_best_code_per_pattern(featsNorm,codebook)
-    tmp = [int(p) for p in best_code_per_p]
-    encoding = codebook[tmp]
+    best_code_per_p = np.asarray([int(x) for x in best_code_per_p])
+    encoding = codebook[best_code_per_p]
     # transform into 2 matrices
     featsNormMAT = np.concatenate([x.reshape(12,pSize) for x in featsNorm],axis=1)
     encodingMAT = np.concatenate([x.reshape(12,pSize) for x in encoding],axis=1)
     # return
-    featsNorm,encoding,featsNormMAT,encodingMAT
+    return best_code_per_p,featsNorm,encoding,featsNormMAT,encodingMAT
+
+
+def get_codeword_histogram(codewords, ncodewords):
+    hist = np.zeros(ncodewords)
+    for cw in codewords:
+        hist[cw] += 1
+    return hist
 
 
 def get_all_barfeats():

### 2010-03-13 PythonSrc/data_iterator.py
offset implemented

@@ -20,21 +20,21 @@ import feats_utils as FU
 
 class DataIterator:
 
-    stopAfterOnePass = 1
-    matfiles = []
-    featsize = 4
-    usebars = 1
-    fidx = 0 # which file are we at
-    pidx = 0 # which pattern are we at
-    currfeats = []  # current features
-    nPatternSeen = 0 # for stats
-    passSize = -1 # number of element in one pass, unknown to start with
-    barbts = [] # index of first beats of bars
-
     def __init__(self):
         """ Constructor """
-        print 'creating an instance of DataIterator'
+        #print 'creating an instance of DataIterator'
         self.idx = 0
+        self._stopAfterOnePass = 1
+        self.matfiles = []
+        self.featsize = 4
+        self.usebars = 1
+        self.fidx = 0 # which file are we at
+        self.pidx = 0 # which pattern are we at
+        self.currfeats = []  # current features
+        self.nPatternSeen = 0 # for stats
+        self.passSize = -1 # number of element in one pass, unknown to start with
+        self.barbts = [] # index of first beats of bars
+        self.offset = 0 # percentage of a bar to offset
 
     def setMatfiles(self,mfiles):
         """
@@ -43,6 +43,14 @@ class DataIterator:
         """
         self.matfiles = mfiles
         np.random.shuffle(self.matfiles)
+
+    def setOffset(self,val):
+        """
+        Set the offset, value between 0 and 1
+        """
+        assert(val>=0)
+        assert(val<1)
+        self.offset = val;
 
     def useBars(self,n):
         """
@@ -79,9 +87,9 @@ class DataIterator:
     def stopAfterOnePass(self,boolean):
         """ Whether we stop after one pass, or never stop, default: True """
         if boolean:
-            self.stopAfterOnePass = 1
+            self._stopAfterOnePass = 1
         else:
-            self.stopAfterOnepass = 0
+            self._stopAfterOnepass = 0
 
     def passSize(self):
         """
@@ -124,13 +132,27 @@ class DataIterator:
                 else :
                     x1 = self.pidx
                     idx1 = np.where(self.barbts == self.pidx)[0][0]
-                    if idx1 + self.usebars < len(self.barbts):
+                    if self.offset > 0 and idx1 + self.usebars + 1 >= len(self.barbts):
+                        # we're at the end, we have an offset, cant make it
+                        self.fidx = self.fidx + 1
+                        self.pidx = 0
+                        self.barbts = []
[…]

### 2010-03-13 PythonSrc/data_iterator.py
partial bar now set

@@ -35,6 +35,7 @@ class DataIterator:
         self.passSize = -1 # number of element in one pass, unknown to start with
         self.barbts = [] # index of first beats of bars
         self.offset = 0 # percentage of a bar to offset
+        self.partialbar = 1
 
     def setMatfiles(self,mfiles):
         """
@@ -58,11 +59,23 @@ class DataIterator:
         The length of the pattern
         is therefore variable. Default: true.
         """
+        if n > 1:
+            assert(self.partialbar == 1)
         self.usebars = n
 
     def getBars(self):
         """ Returns the number of bars we use, or 0 """
         return self.usebars
+
+    def usePartialBar(self,val):
+        """
+        Returns a percentage of the bar, does nothing if 1. Value between
+        0 and 1. self.useBars() must be set to 1.
+        """
+        assert(val>0)
+        assert(val<=1)
+        assert(self.usebars == 1)
+        self.partialbar = val
 
     def setFeatsize(self,n):
         """
@@ -129,7 +142,7 @@ class DataIterator:
                     self.fidx = self.fidx + 1
                     self.pidx = 0
                     self.barbts = []
-                else :
+                else : # we use bars
                     x1 = self.pidx
                     idx1 = np.where(self.barbts == self.pidx)[0][0]
                     if self.offset > 0 and idx1 + self.usebars + 1 >= len(self.barbts):
@@ -145,6 +158,10 @@ class DataIterator:
                         if self.offset > 0:
                             curroffset = np.round((x2-x1)*self.offset/self.usebars)
                         self.nPatternSeen = self.nPatternSeen + 1
+                        if self.partialbar < 1:
+                            x2 = x1 + np.round((x2-x1)*self.partialbar)
+                            if x2 == x1:
+                                x2 = x1+1
                         return self.currfeats[:,x1+curroffset:x2+curroffset]
                     else :
                         # end of the song case, no offset
@@ -152,6 +169,10 @@ class DataIterator:
                         x2 = self.currfeats.shape[1]
                         self.pidx = x2
                         self.nPatternSeen = self.nPatternSeen + 1
+                        if self.partialbar < 1:
+                            x2 = x1 + np.round((x2-x1)*self.partialbar)
+                            if x2 == x1:
+                                x2 = x1+1
                         return self.currfeats[:,x1:x2]
         # NEW FILE
         # in case song does not contain features long enough
@@ -212,7 +233,12 @@ class DataIterator:
[…]

### 2010-03-13 PythonSrc/demos.py
partial bar and offset added

@@ -8,8 +8,8 @@ More of a demo than a useful code
 
 
 
-def load_and_encode_data(codebook,pSize=16,keyInv=True,
-                         downBeatInv=False,bars=2):
+def load_and_encode_data(codebook,pSize=8,keyInv=True,
+                         downBeatInv=False,bars=1,partialbar=1,offset=0):
     """
     Load a dataset, and encode it with codebook
     Return dists, avg_dists
@@ -18,7 +18,8 @@ def load_and_encode_data(codebook,pSize=16,keyInv=True,
     import VQutils
     # get data
     featsNorm = get_data_maxener(pSize=pSize,keyInv=keyInv,
-                                 downBeatInv=downBeatInv,bars=bars)
+                                 downBeatInv=downBeatInv,bars=bars,
+                                 partialbar=partialbar,offset=offset)
     # encode
     best_code_per_p, dists, avg_dists = VQutils.find_best_code_per_pattern(featsNorm,codebook,scale=False)
     return dists, avg_dists
@@ -30,7 +31,7 @@ def get_data_maxener_16_true_false_bars2():
 def get_data_maxener_8_true_false_bars2():
     return get_data_maxener(pSize=8,keyInv=True,downBeatInv=False,bars=2)
 
-def get_data_maxener(pSize=16,keyInv=True,downBeatInv=False,bars=2):
+def get_data_maxener(pSize=4,keyInv=True,downBeatInv=False,bars=1,partialbar=1,offset=0):
     """
     Util function for something we do all the time
     Remove the empty patterns
@@ -54,9 +55,14 @@ def get_data_maxener(pSize=16,keyInv=True,downBeatInv=False,bars=2):
     data_iter.setMatfiles(allfiles) # set matfiles
     if bars > 0:
         data_iter.useBars( bars )            # a pattern spans 'bars' bars
+        if partialbar < 1:
+            assert(bars==1)
+            data_iter.usePartialBar( partialbar )
     else:
         data_iter.useBars(0)                 # important to set it to zero!
         data_iter.setFeatsize( pSize )       # a pattern is a num. of beats
+    if offset > 0:
+        data_iter.setOffset(offset)
     data_iter.stopAfterOnePass(True)# stop after one full iteration
     # get features
     featsNorm = [FU.normalize_pattern_maxenergy(p,pSize,keyInv,downBeatInv).flatten() for p in data_iter]

### 2010-03-13 PythonSrc/demos.py
default changes

@@ -8,7 +8,7 @@ More of a demo than a useful code
 
 
 
-def load_and_encode_data(codebook,pSize=8,keyInv=True,
+def load_and_encode_data(codebook,pSize=4,keyInv=True,
                          downBeatInv=False,bars=1,partialbar=1,offset=0):
     """
     Load a dataset, and encode it with codebook

### 2010-03-13 PythonSrc/for_ron.py
added offset and partial bar

@@ -18,7 +18,7 @@ featsDir = os.path.expanduser('~/projects/ismir10-patterns/beatFeats')
 testFeatsDir = os.path.expanduser('~/projects/ismir10-patterns/uspop_mat')
 outputDir = os.path.expanduser('~/projects/ismir10-patterns/experiments')
 
-def do_experiment(experiment_dir,beats,bars,nCodes,nSamples=0,useFirsts=False,seed=0):
+def do_experiment(experiment_dir,beats,bars,nCodes,nSamples=0,useFirsts=False,seed=0,offset=0,partialbar=1):
     """
     Performs an independant experiment!!!!
     """
@@ -30,7 +30,8 @@ def do_experiment(experiment_dir,beats,bars,nCodes,nSamples=0,useFirsts=False,se
     np.random.seed(seed)
 
     args = dict(experiment_dir=experiment_dir, beats=beats, bars=bars,
-                nCodes=nCodes, nSamples=nSamples, useFirsts=useFirsts,seed=seed)
+                nCodes=nCodes, nSamples=nSamples, useFirsts=useFirsts,seed=seed,
+                offset=offset,partialbar=partialbar)
     sp.io.savemat(os.path.join(experiment_dir, 'args.mat'), args)
 
 
@@ -40,9 +41,9 @@ def do_experiment(experiment_dir,beats,bars,nCodes,nSamples=0,useFirsts=False,se
 
     # load everything, unit: 1 bar, resized to 4 beats
     # key invariant, not downbeatinvariant
-    featsNorm = demos.get_data_maxener(pSize=beats,keyInv=True,downBeatInv=False,bars=bars)
+    featsNorm = demos.get_data_maxener(pSize=beats,keyInv=True,downBeatInv=False,bars=bars,offset=offset,partialbar=partialbar)
 
-    # select 50K random samples out of it
+    # select nSamples random samples out of it
     if nSamples == 0:
         nSamples = featsNorm.shape[0]
     if useFirsts:

### 2010-03-13 PythonSrc/tzanetakis_utils.py
pylab removed from tzanetakis, useless

@@ -10,7 +10,7 @@ tb2332@columbia.edu
 import os
 import sys
 import numpy as np
-import pylab as P
+#import pylab as P
 import scipy as sp
 import scipy
 import scipy.io

### 2010-03-15 PythonSrc/for_ron.py
new parmas added to output

@@ -78,8 +78,8 @@ def do_experiment(experiment_dir,beats,bars,nCodes,nSamples=0,useFirsts=False,se
     # report result (average sqaure distance per ... pixel?
     # with print outs to know what we are doing
     report = ['EXPERIMENT REPORT ******************************',
-              'beats: %s , bars: %s , nCodes: %s , nSamples: %s'
-              % (beats, bars, nCodes, nSamples)]
+              'beats: %s , bars: %s , nCodes: %s , nSamples: %s , offset: %s , partialbar: %s'
+              % (beats, bars, nCodes, nSamples, offset, partialbar)]
     if useFirsts:
         report.append(['we use firsts %s samples' % nCodes])
     report.extend(['np.average(avg_dists): %s' % np.average(avg_dists),

### 2010-03-15 Paper/Paper.tex
more results added

@@
 size & avg. dist \\ \hline \hline
-1 bar (4) & \\
-2 bars (8) & $$\\
-2 bars (12) & $$\\
-2 bars (16) & $$ \\ \hline
-3 beats & \\
-5 beats & $$\\
-7 beats & $$\\ \hline
+1 bar (4) & $0.036722$\\
+2 bars (8) & $0.044301$\\
+2 bars (12) & $0.043583$\\
+2 bars (16) & $0.044086$ \\ \hline
+3 beats & $0.032688$\\
+5 beats & $0.040858$\\
+7 beats & $0.045515$\\ \hline
 \end{tabular}
@@
 % 50K         0.142257       0.036722
-% 100K        0.142195       
-% 250K        0.141796
-% 500K        0.142081
+% 100K        0.142195       0.035973
+% 250K        0.141796       0.035910
+% 500K        0.142081       0.035904
 % 1M          0.142118
@@
 % 1K          0.155474       0.042596
-% 5K          0.149619
-% 10K         0.148056
-% 50K         0.146328
-% 100K        0.146004
-% 250K        0.146024
-% 500K        0.145962
+% 5K          0.149619       0.038971
+% 10K         0.148056       0.038114
+% 50K         0.146328       0.037181
+% 100K        0.146004       0.037055
+% 250K        0.146024       0.037014
+% 500K        0.145962       0.037004
 % 1M          0.145873
@@
 % 10K         0.165063       0.045507
-% 50K         0.162439
-% 100K        0.161588
-% 250K        0.161910
-% 500K        0.161233
+% 50K         0.162439       0.044301
+% 100K        0.161588       0.044183
+% 250K        0.161910       0.043896
+% 500K        0.161233       0.043906
 % 1M          0.161057
@@
 % code 500      0.131211        0.030973
+% REDONE WITH FIX NUMBER OF FEATURES - 500K (new correct dist)
+% code 1        
+% code 10       0.045579
+% code 50       0.038302
+% code 100      
+% code 200      
+% code 500
 

### 2010-03-15 PythonSrc/for_ron.py
added support for picking a single experiment out of a set

@@ -167,5 +167,11 @@ if __name__ == '__main__':
 
     # Python indexes from 0, the argument indexes from 1.
     experiment_set_number = int(sys.argv[3]) - 1
-    pool.map(do_experiment_wrapper, experiment_args[experiment_set_number])
+    args = experiment_args[experiment_set_number]
+    try:
+        args = [args[int(sys.argv[4]) - 1]]
+    except IndexError:
+        pass
+    
+    pool.map(do_experiment_wrapper, args)
 

### 2010-03-15 PythonSrc/for_ron.py
two new sets of experiments

@@ -156,8 +156,33 @@ experiment_args = [
      (os.path.join(outputDir, 'set9exp1'),8,2,500,50000),
      (os.path.join(outputDir, 'set9exp2'),16,4,500,50000),
      (os.path.join(outputDir, 'set9exp3'),32,8,500,50000)],
+    # EXPERIMENT SET 10: more on offset with partial bars
+    [(os.path.join(outputDir, 'set10exp0'),4,1,100,50000,False,0,0.00,.50),
+     (os.path.join(outputDir, 'set10exp1'),4,1,100,50000,False,0,0.25,.50),
+     (os.path.join(outputDir, 'set10exp2'),4,1,100,50000,False,0,0.50,.50),
+     (os.path.join(outputDir, 'set10exp3'),4,1,100,50000,False,0,0.75,.50),
+     (os.path.join(outputDir, 'set10exp4'),4,1,100,50000,False,0,0.00,.75),
+     (os.path.join(outputDir, 'set10exp5'),4,1,100,50000,False,0,0.25,.75),
+     (os.path.join(outputDir, 'set10exp6'),4,1,100,50000,False,0,0.50,.75),
+     (os.path.join(outputDir, 'set10exp7'),4,1,100,50000,False,0,0.75,.75)],
+    # EXPERIMENT SET 11: redoing pattern doubles => nCodes squares
+    [(os.path.join(outputDir, 'set11exp0'),1,1,2,128000,False,0,0.00,.25),
+     (os.path.join(outputDir, 'set11exp1'),1,1,2,128000,False,0,0.25,.25),
+     (os.path.join(outputDir, 'set11exp2'),1,1,2,128000,False,0,0.50,.25),
+     (os.path.join(outputDir, 'set11exp3'),1,1,2,128000,False,0,0.75,.25),
+     (os.path.join(outputDir, 'set11exp4'),2,1,4,128000,False,0,0.00,.50),
+     (os.path.join(outputDir, 'set11exp5'),2,1,4,128000,False,0,0.25,.50),
+     (os.path.join(outputDir, 'set11exp6'),2,1,4,128000,False,0,0.50,.50),
+     (os.path.join(outputDir, 'set11exp7'),2,1,4,128000,False,0,0.75,.50),
+     (os.path.join(outputDir, 'set11exp7'),4,1,16,128000,False,0),
+     (os.path.join(outputDir, 'set11exp7'),8,2,256,128000,False,0),
[…]

### 2010-03-17 PythonSrc/for_ron.py
improved set 7 where we test beats without knowledge of bars

@@ -139,7 +139,11 @@ experiment_args = [
      (os.path.join(outputDir, 'set7exp4'), 9, 0, 100, 50000),
      (os.path.join(outputDir, 'set7exp5'), 10, 0, 100, 50000),
      (os.path.join(outputDir, 'set7exp6'), 11, 0, 100, 50000),   
-     (os.path.join(outputDir, 'set7exp7'), 12, 0, 100, 50000)],
+     (os.path.join(outputDir, 'set7exp7'), 12, 0, 100, 50000),
+     (os.path.join(outputDir, 'set7exp8'), 1, 0, 100, 50000),
+     (os.path.join(outputDir, 'set7exp9'), 3, 0, 100, 50000),
+     (os.path.join(outputDir, 'set7exp10'), 5, 0, 100, 50000),
+     (os.path.join(outputDir, 'set7exp11'), 7, 0, 100, 50000)],
     # EXPERIMENT SET 8: testing offsets
     [(os.path.join(outputDir, 'set8exp0'),4,1,100,50000,False,0,0,1),
      (os.path.join(outputDir, 'set8exp1'),4,1,100,50000,False,0,0.25,1),

### 2010-03-17 PythonSrc/demos.py
encoding of artists, i.e. finding the most used patters for them, added to demos

@@ -232,7 +232,7 @@ def get_all_barfeats():
 
 
 
-def add_image(P,im,x,y,width=.15):
+def add_image(P,im,x,y,width=.15,height=0,gray=True):
     """
     Used by LLE_my_codebook to add a specific image to a given plot
     I should start my plotting library... or still Ron's...
@@ -244,7 +244,8 @@ def add_image(P,im,x,y,width=.15):
       - im     matrix representing the image
       - x      x position in data coord
       - y      y position in data coord
-      - width  width in fig size, height automatically found
+      - width  width in fig size, height automatically if height=0
+      - height height in fig size, if 0 based on width
     """
     # current axes
     curr_axes = P.gca()
@@ -263,26 +264,33 @@ def add_image(P,im,x,y,width=.15):
     relX = (relX * fWidth) + fX
     relY = (relY * fHeight) + fY
     # height relative to width
-    height = im.shape[0] * 1. / im.shape[1] * width
+    if height == 0:
+        height = im.shape[0] * 1. / im.shape[1] * width
     # create new axis
     a = P.axes([relX-width/2.,relY-height/2.,width,height])
     # set to x and y ticks
     P.setp(a, xticks=[], yticks=[])
     # plot image, new axes is the current default
-    P.imshow(im,interpolation='nearest',origin='lower')
+    if gray:
+        P.imshow(im,interpolation='nearest',origin='lower',cmap=P.cm.gray_r)
+    else:
+        P.imshow(im,interpolation='nearest',origin='lower')
     # set back axes
     P.axes(curr_axes)
 
 
 
-def LLE_my_codebook(codebook,nNeighbors=5):
+def LLE_my_codebook(codebook,nNeighbors=5,nRand=5):
     """
     Performs LLE on the codebook
     Display the result
     LLE code not mine, see code for reference.
+    nRand=number of random images added
     """
     import pylab as P
     import LLE
+    import numpy as np
+    import VQutils as VQU
     # compute LLE, goal is 2D
     LLEres = LLE.LLE(codebook.T,nNeighbors,2)
     # plot that result
@@ -290,12 +298,14 @@ def LLE_my_codebook(codebook,nNeighbors=5):
     P.hold(True)
     # prepare to plot
     patch_size = codebook[0,:].size / 12
[…]

### 2010-03-17 PythonSrc/demos.py
bug solved

@@ -408,7 +408,7 @@ def freqs_for_my_artists(filenames,codebook,pSize=8,keyInv=True,
         if not res.has_key(artist):
             res[artist] = np.zeros([1,nCodes])
         for code in best_code_per_p:
-            res[artist][int(code)] += 1
+            res[artist][0,int(code)] += 1
     # done, return dictionary
     return res
         

### 2010-03-17 PythonSrc/demos.py
bug solved

@@ -116,6 +116,7 @@ def encode_one_song(filename,codebook,pSize=8,keyInv=True,
     # find code per pattern
     best_code_per_p, dists, avg_dists = VQutils.find_best_code_per_pattern(featsNorm,codebook)
     best_code_per_p = np.asarray([int(x) for x in best_code_per_p])
+    assert best_code_per_p.shape[0] > 0, "empty song, we should catch that"
     encoding = codebook[best_code_per_p]
     # transform into 2 matrices, with derolling!!!!!!!!!
     assert(featsNorm.shape[0] == encoding.shape[0])

### 2010-03-17 PythonSrc/demos.py
bug solved

@@ -105,6 +105,8 @@ def encode_one_song(filename,codebook,pSize=8,keyInv=True,
     keyroll = np.array([x[1] for x in featsNorm])
     dbroll = np.array([x[2] for x in featsNorm])
     featsNorm = [x[0].flatten() for x in featsNorm]
+    if len(featsNorm) == 0: # empty song
+        return [],[],[],[],[]
     featsNorm = np.array(featsNorm)
     res = [np.sum(r) > 0 for r in featsNorm]
     res2 = np.where(res)
@@ -116,7 +118,7 @@ def encode_one_song(filename,codebook,pSize=8,keyInv=True,
     # find code per pattern
     best_code_per_p, dists, avg_dists = VQutils.find_best_code_per_pattern(featsNorm,codebook)
     best_code_per_p = np.asarray([int(x) for x in best_code_per_p])
-    assert best_code_per_p.shape[0] > 0, "empty song, we should catch that"
+    assert best_code_per_p.shape[0] > 0,'empty song, we should have caught that'
     encoding = codebook[best_code_per_p]
     # transform into 2 matrices, with derolling!!!!!!!!!
     assert(featsNorm.shape[0] == encoding.shape[0])

### 2010-03-17 PythonSrc/demos.py
added experiment on artist recognition in demos

@@ -381,7 +381,31 @@ def LLE_my_codebook2(codebook,nLines=5,nCols=15,nNeighbors=5):
     P.hold(False)
     P.show()
 
-            
+
+
+def freqs_my_songs(filenames,codebook,pSize=8,keyInv=True,
+                   downBeatInv=False,bars=2):
+    """
+    Returns a list of numpy.array containing frequency for each
+    code in the codebook for each file in filenames
+    """
+    import numpy as np
+    res = []
+    for f in filenames:
+        # encode song
+        a,b,c,d,e = encode_one_song(f,codebook,pSize=pSize,keyInv=keyInv,
+                                    downBeatInv=downBeatInv,bars=bars)
+        best_code_per_p,featsNorm,encoding,featsNormMAT,encodingMAT = a,b,c,d,e
+        # get freqs
+        freqs = np.zeros([1,nCodes])
+        for code in best_code_per_p:
+            freqs[0,int(code)] += 1
+        res.append(freqs)
+    # done, return res
+    return res
+
+
+
 
 def freqs_for_my_artists(filenames,codebook,pSize=8,keyInv=True,
                          downBeatInv=False,bars=2):
@@ -415,3 +439,80 @@ def freqs_for_my_artists(filenames,codebook,pSize=8,keyInv=True,
     # done, return dictionary
     return res
         
+
+
+def knn_from_freqs_on_artists(filenames,codebook,pSize=8,keyInv=True,
+                              downBeatInv=False,bars=2):
+    """
+    Performs a leave-one-out experiments where we try to guess the artist
+    from it's nearest neighbors in frequencies
+    We use squared euclidean distance.
+
+    filenames are expected to be: */artist/album/*.mat
+    """
+    import numpy as np
+    import os
+    import VQutils as VQU
+    # get frequencies for all songs
+    freqs = freqs_my_songs(filenames,codebook,pSize=pSize,keyInv=keyInv,
+                           downBeatInv=downBeatInv,bars=bars)
+    print 'all frequencies computed'
+    # get artists for all songs
+    artists = []
+    for f in filenames:
+        tmp, song = os.path.split(f)
+        tmp,album = os.path.split(tmp)
+        tmp,artist = os.path.split(tmp)
+        artists.append(artist)
+    artists = np.array(artists)
+
+    # compute how many songs per artist
+    #nSongs = {}
+    #for artist in artists:
+    #    if nSongs.has_key(artist):
[…]

### 2010-03-17 PythonSrc/demos.py
bug solved

@@ -391,6 +391,7 @@ def freqs_my_songs(filenames,codebook,pSize=8,keyInv=True,
     """
     import numpy as np
     res = []
+    nCodes = codebook.shape[0]
     for f in filenames:
         # encode song
         a,b,c,d,e = encode_one_song(f,codebook,pSize=pSize,keyInv=keyInv,

### 2010-03-17 PythonSrc/demos.py
bug solved

@@ -486,7 +486,7 @@ def knn_from_freqs_on_artists(filenames,codebook,pSize=8,keyInv=True,
                 dists[l,c] = np.inf
                 dists[c,l] = np.inf
                 continue
-            dists[l,c] = VQU.eculidean_distance(freqs[l],freqs[c])
+            dists[l,c] = VQU.euclidean_distance(freqs[l],freqs[c])
             dists[c,l] = dists[l,c]
     for l in range(nFiles): # fill diag with inf
         dists[l,l] = np.inf

### 2010-03-17 Paper/Paper.tex
better lle pic
results added

@@
 
-% RESULTS
-% 200codes, 2 bars, pSize=16, maxener, True, False, not scaled = 
-% 200codes, 2 bars, pSize=8, maxener, True, False, not scaled  = 0.042432
-% 200codes, 1 bars, pSize=4, maxener, True, False, not scaled  = 
-% 200codes, 7 beats, pSize=7, maxener, True, False, not scaled = 
-% 200codes, 5 beats, pSize=5, maxener, True, False, not scaled = 
-% 200codes, 3 beats, pSize=3, maxener, True, False, not scaled = 
 \begin{table}
@@
 \end{table}
+% RESULTS
+% 200codes, 2 bars, pSize=16, maxener, True, False, not scaled = 
+% 200codes, 2 bars, pSize=8, maxener, True, False, not scaled  = 0.042432
+% 200codes, 1 bars, pSize=4, maxener, True, False, not scaled  = 
+% 200codes, 7 beats, pSize=7, maxener, True, False, not scaled = 
+% 200codes, 5 beats, pSize=5, maxener, True, False, not scaled = 
+% 200codes, 3 beats, pSize=3, maxener, True, False, not scaled = 
+% NEW STUFF - experiment set 7, 100 codes! 
+% 2 beats 0 bars       0.025346
+% 4 beats 0 bars       0.037181
+% 6 beats 0 bars       0.043237
+% 8 beats 0 bars       0.046832
+% 9 beats 0 bars       0.048579
+% 10 beats 0 bars      0.049530
+% 11 beats 0 bars      0.050717
+% 12 beats 0 bars      0.052020
 
@@
 % 1bar s4     avg_distOLD    avg_dist
+% 0                          0.052247
 % 1K          0.150370       0.041941
@@
 % 1bar s4 first X samples
-% 1K          0.157142       
-% 5K          0.150045
-% 10K         0.148357
-% 50K         0.146069
-% 100K        0.145765
-% 250K        0.145819
-% 500K        0.145772
+% 0
+% 1K          0.157142       0.049282
+% 5K          0.150045       0.046441
+% 10K         0.148357       0.045507
+% 50K         0.146069       0.044301
+% 100K        0.145765       0.044183
+% 250K        0.145819       0.043896
+% 500K        0.145772       0.043906
 % 1M          0.145924       
@@
 % 0bar s4 (and test set same, we could do 1bar s4 in test set...)
+% 0
 % 1K          0.155474       0.042596
@@
 % 2bar s8
+% 0                          0.066864179424
 % 1K          0.171300       0.049282
@@
 % code 50       0.038302
-% code 100      
+% code 100      0.035904
 % code 200      
@@
 
+\subsection{Bar offset}
+We study how great are patterns based on bars by playing with an offset.
+The offset can be $0$ (normal case) or a quarter of bar, half a bar, and
+three quarters of a bar.
+
+\begin{table}
+\begin{center}
+\begin{tabular}{|l|c|c|}
+\hline
+offset & avg. dist \\ \hline \hline
+$0$ & $$ \\
+$.25$ & $0.037130$ \\
+$.50$ & $0.037241$ \\
+$.75$ & $0.037337$ \\ \hline
+\end{tabular}
+\end{center}
+\caption{{Results with different offset. Always $100$ codes.
+\label{tab:offset}}}
+\end{table}
+% RESULTS             avg_results
+% 0 offset            0.036722 (to be confirmed by set8exp0)
+% .25 ofset           0.037130
+% .50 ofset           0.037241
+% .75 ofset           0.037337
+% RESULTS FOR PARTIALBAR TOO
+% 0 off .25 part      0.046304
+% .25 off .25 part    0.044950
+% .50 off .25 part    0.044395
+% .75 off .25 part    0.044473
+
 \subsection{Visualization}
@@
 
-\section{Links with NMF}\label{sec:nmf}
+\section{Links with NMF TO BE REMOVED}\label{sec:nmf}
 In this section we discuss links with non-negative matrix factorization

### 2010-03-17 PythonSrc/demos.py
bug solved

@@ -486,7 +486,7 @@ def knn_from_freqs_on_artists(filenames,codebook,pSize=8,keyInv=True,
                 dists[l,c] = np.inf
                 dists[c,l] = np.inf
                 continue
-            dists[l,c] = VQU.euclidean_distance(freqs[l],freqs[c])
+            dists[l,c] = VQU.euclidean_dist(freqs[l],freqs[c])
             dists[c,l] = dists[l,c]
     for l in range(nFiles): # fill diag with inf
         dists[l,l] = np.inf

### 2010-03-17 PythonSrc/demos.py
improvement, can normalize the freqs

@@ -384,12 +384,13 @@ def LLE_my_codebook2(codebook,nLines=5,nCols=15,nNeighbors=5):
 
 
 def freqs_my_songs(filenames,codebook,pSize=8,keyInv=True,
-                   downBeatInv=False,bars=2):
+                   downBeatInv=False,bars=2,normalize=False):
     """
     Returns a list of numpy.array containing frequency for each
     code in the codebook for each file in filenames
     """
     import numpy as np
+    import VQutils as VQU
     res = []
     nCodes = codebook.shape[0]
     for f in filenames:
@@ -401,6 +402,8 @@ def freqs_my_songs(filenames,codebook,pSize=8,keyInv=True,
         freqs = np.zeros([1,nCodes])
         for code in best_code_per_p:
             freqs[0,int(code)] += 1
+        if normalize and best_code_per_p.shape[0] > 0:
+            freqs *= 1./ VQU.euclidean_norm(freqs)
         res.append(freqs)
     # done, return res
     return res
@@ -443,7 +446,7 @@ def freqs_for_my_artists(filenames,codebook,pSize=8,keyInv=True,
 
 
 def knn_from_freqs_on_artists(filenames,codebook,pSize=8,keyInv=True,
-                              downBeatInv=False,bars=2):
+                              downBeatInv=False,bars=2,normalize=True):
     """
     Performs a leave-one-out experiments where we try to guess the artist
     from it's nearest neighbors in frequencies
@@ -456,7 +459,8 @@ def knn_from_freqs_on_artists(filenames,codebook,pSize=8,keyInv=True,
     import VQutils as VQU
     # get frequencies for all songs
     freqs = freqs_my_songs(filenames,codebook,pSize=pSize,keyInv=keyInv,
-                           downBeatInv=downBeatInv,bars=bars)
+                           downBeatInv=downBeatInv,bars=bars,
+                           normalize=normalize)
     print 'all frequencies computed'
     # get artists for all songs
     artists = []
@@ -466,15 +470,6 @@ def knn_from_freqs_on_artists(filenames,codebook,pSize=8,keyInv=True,
         tmp,artist = os.path.split(tmp)
         artists.append(artist)
     artists = np.array(artists)
-
-    # compute how many songs per artist
-    #nSongs = {}
-    #for artist in artists:
-    #    if nSongs.has_key(artist):
-    #        nSongs[artist] += 1
-    #    else:
-    #        nSongs[artist] = 1
-
     # sanity check
[…]

### 2010-03-17 PythonSrc/demos.py
bug solved

@@ -402,7 +402,7 @@ def freqs_my_songs(filenames,codebook,pSize=8,keyInv=True,
         freqs = np.zeros([1,nCodes])
         for code in best_code_per_p:
             freqs[0,int(code)] += 1
-        if normalize and best_code_per_p.shape[0] > 0:
+        if normalize and len(best_code_per_p) > 0:
             freqs *= 1./ VQU.euclidean_norm(freqs)
         res.append(freqs)
     # done, return res

### 2010-03-17 PythonSrc/demos.py
confusion matrix added

@@ -446,13 +446,16 @@ def freqs_for_my_artists(filenames,codebook,pSize=8,keyInv=True,
 
 
 def knn_from_freqs_on_artists(filenames,codebook,pSize=8,keyInv=True,
-                              downBeatInv=False,bars=2,normalize=True):
+                              downBeatInv=False,bars=2,normalize=True,
+                              confMatrix=True):
     """
     Performs a leave-one-out experiments where we try to guess the artist
     from it's nearest neighbors in frequencies
     We use squared euclidean distance.
 
     filenames are expected to be: */artist/album/*.mat
+    if confMatrix=True, plot it.
+    Always returns confusion matrix
     """
     import numpy as np
     import os
@@ -470,6 +473,9 @@ def knn_from_freqs_on_artists(filenames,codebook,pSize=8,keyInv=True,
         tmp,artist = os.path.split(tmp)
         artists.append(artist)
     artists = np.array(artists)
+    # names of artists
+    artist_names = np.unique(np.sort(artists))
+    nArtists = artist_names.shape[0]
     # sanity check
     assert(len(filenames)==len(artists))
     # compute distance between all songs
@@ -486,7 +492,8 @@ def knn_from_freqs_on_artists(filenames,codebook,pSize=8,keyInv=True,
     for l in range(nFiles): # fill diag with inf
         dists[l,l] = np.inf
     print 'distances computed between frequency vectors'
-
+    # confusion matrix
+    confMat = np.zeros([nArtists,nArtists])
     # performs leave-one-out KNN
     nExps = 0
     nGood = 0
@@ -506,9 +513,25 @@ def knn_from_freqs_on_artists(filenames,codebook,pSize=8,keyInv=True,
         # get stats
         nExps += 1
         randScore += nGoodMatches * 1. / nMatches
-        if artists[orderedMatches[0]] == artist:
+        pred_artist = artists[orderedMatches[0]]
+        if pred_artist == artist:
             nGood += 1
+        # fill confusion matrix
+        real_artist_id =np.where(artist_names==artist)[0][0]
+        pred_artist_id =np.where(artist_names==pred_artist)[0][0]
+        confMat[real_artist,pred_artist] += 1
     # done, print out
     print 'nExps:',nExps
     print 'rand accuracy:',(randScore*1./nExps)
     print 'accuracy:',(nGood*1./nExps)
+    # plot confusion matrix
+    if confMatrix:
[…]

### 2010-03-17 PythonSrc/demos.py
bug solved

@@ -460,11 +460,13 @@ def knn_from_freqs_on_artists(filenames,codebook,pSize=8,keyInv=True,
     import numpy as np
     import os
     import VQutils as VQU
+    import time
     # get frequencies for all songs
+    tstart = time.time()
     freqs = freqs_my_songs(filenames,codebook,pSize=pSize,keyInv=keyInv,
                            downBeatInv=downBeatInv,bars=bars,
                            normalize=normalize)
-    print 'all frequencies computed'
+    print 'all frequencies computed in',(time.time()-tstart),' seconds.'
     # get artists for all songs
     artists = []
     for f in filenames:
@@ -519,7 +521,7 @@ def knn_from_freqs_on_artists(filenames,codebook,pSize=8,keyInv=True,
         # fill confusion matrix
         real_artist_id =np.where(artist_names==artist)[0][0]
         pred_artist_id =np.where(artist_names==pred_artist)[0][0]
-        confMat[real_artist,pred_artist] += 1
+        confMat[real_artist_id,pred_artist_id] += 1
     # done, print out
     print 'nExps:',nExps
     print 'rand accuracy:',(randScore*1./nExps)

### 2010-03-17 PythonSrc/demos.py
bug solved

@@ -532,8 +532,8 @@ def knn_from_freqs_on_artists(filenames,codebook,pSize=8,keyInv=True,
         import pylab as P
         P.imshow(confMat,interpolation='nearest',cmap=P.cm.gray_r,
                  origin='lower')
-        P.xticks(P.arange(artist_names.shape[0]),list(artist_names))
-        P.yticks(P.arange(artist_names.shape[0]),list(short_names))
+        P.yticks(P.arange(artist_names.shape[0]),list(artist_names))
+        P.xticks(P.arange(artist_names.shape[0]),list(short_names))
         P.title('confusion matrix (real/predicted)')
     # return confusion matrix
     return confMat

### 2010-03-17 PythonSrc/demos.py
colorbar added

@@ -535,5 +535,6 @@ def knn_from_freqs_on_artists(filenames,codebook,pSize=8,keyInv=True,
         P.yticks(P.arange(artist_names.shape[0]),list(artist_names))
         P.xticks(P.arange(artist_names.shape[0]),list(short_names))
         P.title('confusion matrix (real/predicted)')
+        P.colorbar()
     # return confusion matrix
     return confMat

### 2010-03-17 PythonSrc/demos.py
labels added

@@ -535,6 +535,8 @@ def knn_from_freqs_on_artists(filenames,codebook,pSize=8,keyInv=True,
         P.yticks(P.arange(artist_names.shape[0]),list(artist_names))
         P.xticks(P.arange(artist_names.shape[0]),list(short_names))
         P.title('confusion matrix (real/predicted)')
+        P.ylabel('TRUE')
+        P.xlabel('RECOG')
         P.colorbar()
     # return confusion matrix
     return confMat

### 2010-03-17 PythonSrc/demos.py
return added

@@ -455,7 +455,10 @@ def knn_from_freqs_on_artists(filenames,codebook,pSize=8,keyInv=True,
 
     filenames are expected to be: */artist/album/*.mat
     if confMatrix=True, plot it.
-    Always returns confusion matrix
+
+    RETURNS:
+    - confusion matrix
+    - freqs per file
     """
     import numpy as np
     import os
@@ -539,4 +542,4 @@ def knn_from_freqs_on_artists(filenames,codebook,pSize=8,keyInv=True,
         P.xlabel('RECOG')
         P.colorbar()
     # return confusion matrix
-    return confMat
+    return confMat,freqs

### 2010-03-17 PythonSrc/demos.py
return added

@@ -459,6 +459,7 @@ def knn_from_freqs_on_artists(filenames,codebook,pSize=8,keyInv=True,
     RETURNS:
     - confusion matrix
     - freqs per file
+    - artist per file
     """
     import numpy as np
     import os
@@ -542,4 +543,4 @@ def knn_from_freqs_on_artists(filenames,codebook,pSize=8,keyInv=True,
         P.xlabel('RECOG')
         P.colorbar()
     # return confusion matrix
-    return confMat,freqs
+    return confMat,freqs,artists

### 2010-03-17 PythonSrc/demos.py
l0 norm added

@@ -445,9 +445,19 @@ def freqs_for_my_artists(filenames,codebook,pSize=8,keyInv=True,
         
 
 
+
+def l0_dist(a,b):
+    """
+    Compute the number of common non nul elems
+    """
+    import numpy as np
+    non_nul_a = np.where(a.flatten()>0)[0]
+    return np.where(b.flatten()[non_nul_a]>0)[0].shape[0]
+
+
 def knn_from_freqs_on_artists(filenames,codebook,pSize=8,keyInv=True,
                               downBeatInv=False,bars=2,normalize=True,
-                              confMatrix=True):
+                              confMatrix=True,use_l0_dist=False):
     """
     Performs a leave-one-out experiments where we try to guess the artist
     from it's nearest neighbors in frequencies
@@ -493,7 +503,10 @@ def knn_from_freqs_on_artists(filenames,codebook,pSize=8,keyInv=True,
                 dists[l,c] = np.inf
                 dists[c,l] = np.inf
                 continue
-            dists[l,c] = VQU.euclidean_dist(freqs[l],freqs[c])
+            if use_l0_dist:
+                dists[l,c] = l0_dist(freqs[l],freqs[c])
+            else:
+                dists[l,c] = VQU.euclidean_dist(freqs[l],freqs[c])
             dists[c,l] = dists[l,c]
     for l in range(nFiles): # fill diag with inf
         dists[l,l] = np.inf

### 2010-03-17 PythonSrc/demos.py
bug solved

@@ -452,8 +452,8 @@ def l0_dist(a,b):
     """
     import numpy as np
     non_nul_a = np.where(a.flatten()>0)[0]
-    return np.where(b.flatten()[non_nul_a]>0)[0].shape[0]
-
+    nCommon = np.where(b.flatten()[non_nul_a]>0)[0].shape[0]
+    return a.flatten().shape[0] - nCommon
 
 def knn_from_freqs_on_artists(filenames,codebook,pSize=8,keyInv=True,
                               downBeatInv=False,bars=2,normalize=True,

### 2010-03-17 PythonSrc/demos.py
added use_artists feature

@@ -449,6 +449,7 @@ def freqs_for_my_artists(filenames,codebook,pSize=8,keyInv=True,
 def l0_dist(a,b):
     """
     Compute the number of common non nul elems
+    returns (size - nCommon elems) so it is a distance (smaller=closer)
     """
     import numpy as np
     non_nul_a = np.where(a.flatten()>0)[0]
@@ -457,7 +458,7 @@ def l0_dist(a,b):
 
 def knn_from_freqs_on_artists(filenames,codebook,pSize=8,keyInv=True,
                               downBeatInv=False,bars=2,normalize=True,
-                              confMatrix=True,use_l0_dist=False):
+                              confMatrix=True,use_l0_dist=False,use_artists=False):
     """
     Performs a leave-one-out experiments where we try to guess the artist
     from it's nearest neighbors in frequencies
@@ -465,6 +466,7 @@ def knn_from_freqs_on_artists(filenames,codebook,pSize=8,keyInv=True,
 
     filenames are expected to be: */artist/album/*.mat
     if confMatrix=True, plot it.
+    if use_artists, song are matched to artist, not other songs
 
     RETURNS:
     - confusion matrix
@@ -475,6 +477,8 @@ def knn_from_freqs_on_artists(filenames,codebook,pSize=8,keyInv=True,
     import os
     import VQutils as VQU
     import time
+
+    nCodes = codebook.shape[0]
     # get frequencies for all songs
     tstart = time.time()
     freqs = freqs_my_songs(filenames,codebook,pSize=pSize,keyInv=keyInv,
@@ -496,21 +500,47 @@ def knn_from_freqs_on_artists(filenames,codebook,pSize=8,keyInv=True,
     assert(len(filenames)==len(artists))
     # compute distance between all songs
     nFiles = len(filenames)
-    dists = np.zeros([nFiles,nFiles])
-    for l in range(nFiles):
-        for c in range(l+1,nFiles):
-            if len(freqs[l])==0 or len(freqs[c])==0:
-                dists[l,c] = np.inf
-                dists[c,l] = np.inf
-                continue
-            if use_l0_dist:
-                dists[l,c] = l0_dist(freqs[l],freqs[c])
-            else:
-                dists[l,c] = VQU.euclidean_dist(freqs[l],freqs[c])
-            dists[c,l] = dists[l,c]
-    for l in range(nFiles): # fill diag with inf
-        dists[l,l] = np.inf
[…]

### 2010-03-17 PythonSrc/demos.py
bug solved

@@ -525,7 +525,7 @@ def knn_from_freqs_on_artists(filenames,codebook,pSize=8,keyInv=True,
             artistFreqs = {}
             for k in artist_names:
                 cntArtists[k] = 0
-                artistFreqs = np.zeros([1,nCodes])
+                artistFreqs[k] = np.zeros([1,nCodes])
             for k in range(artists.shape[0]):
                 if k == l:
                     continue

### 2010-03-17 PythonSrc/demos.py
better verbose

@@ -500,6 +500,7 @@ def knn_from_freqs_on_artists(filenames,codebook,pSize=8,keyInv=True,
     assert(len(filenames)==len(artists))
     # compute distance between all songs
     nFiles = len(filenames)
+    tstart = time.time()
     if not use_artists:
         dists = np.zeros([nFiles,nFiles])
         for l in range(nFiles):
@@ -515,7 +516,6 @@ def knn_from_freqs_on_artists(filenames,codebook,pSize=8,keyInv=True,
                     dists[c,l] = dists[l,c]
         for l in range(nFiles): # fill diag with inf
             dists[l,l] = np.inf
-        print 'distances computed between frequency vectors'
     else:
         # create a matrix songs * nArtists
         dists = np.zeros([nFiles,nArtists])
@@ -527,7 +527,7 @@ def knn_from_freqs_on_artists(filenames,codebook,pSize=8,keyInv=True,
                 cntArtists[k] = 0
                 artistFreqs[k] = np.zeros([1,nCodes])
             for k in range(artists.shape[0]):
-                if k == l:
+                if k == l: # skip current file/song
                     continue
                 art = artists[k]
                 cntArtists[art] += 1
@@ -541,6 +541,7 @@ def knn_from_freqs_on_artists(filenames,codebook,pSize=8,keyInv=True,
                     dists[l,c] = l0_dist(freqs[l],artistFreqs[art])
                 else:
                     dists[l,c] = VQU.euclidean_dist(freqs[l],artistFreqs[art])
+    print 'distances computed in',(time.time()-tstart),' seconds.'
     # confusion matrix
     confMat = np.zeros([nArtists,nArtists])
     # performs leave-one-out KNN

### 2010-03-17 PythonSrc/demos.py
faster, more verbose

@@ -477,6 +477,7 @@ def knn_from_freqs_on_artists(filenames,codebook,pSize=8,keyInv=True,
     import os
     import VQutils as VQU
     import time
+    import copy
 
     nCodes = codebook.shape[0]
     # get frequencies for all songs
@@ -484,7 +485,7 @@ def knn_from_freqs_on_artists(filenames,codebook,pSize=8,keyInv=True,
     freqs = freqs_my_songs(filenames,codebook,pSize=pSize,keyInv=keyInv,
                            downBeatInv=downBeatInv,bars=bars,
                            normalize=normalize)
-    print 'all frequencies computed in',(time.time()-tstart),' seconds.'
+    print 'all frequencies computed in',(time.time()-tstart),'seconds.'
     # get artists for all songs
     artists = []
     for f in filenames:
@@ -519,29 +520,35 @@ def knn_from_freqs_on_artists(filenames,codebook,pSize=8,keyInv=True,
     else:
         # create a matrix songs * nArtists
         dists = np.zeros([nFiles,nArtists])
+        # precompute cntArtists and artistFreqs, not normalized
+        cntArtists = {}
+        artistFreqs = {}
+        for k in artist_names:
+            cntArtists[k] = 0
+            artistFreqs[k] = np.zeros([1,nCodes])
+        for k in range(artists.shape[0]):
+            if k == l: # skip current file/song
+                continue
+            art = artists[k]
+            cntArtists[art] += 1
+            artistFreqs[art] += freqs[k]
+        # iterate over files
         for l in range(nFiles):
-            # compute artist freqs but skipping this current file
-            cntArtists = {}
-            artistFreqs = {}
-            for k in artist_names:
-                cntArtists[k] = 0
-                artistFreqs[k] = np.zeros([1,nCodes])
-            for k in range(artists.shape[0]):
-                if k == l: # skip current file/song
-                    continue
-                art = artists[k]
-                cntArtists[art] += 1
-                artistFreqs[art] += freqs[k]
-            for k in artistFreqs.keys(): # normalize
-                artistFreqs[k] *= 1. / cntArtists[k]
+            currArtist = artists[l]
+            currCntArtist = copy.deepcopy(cntArtist)
+            currCntArtist[currArtist] -= 1
+            currArtistFreqs = copy.deepcopy(artistFreqs)
+            currArtistFreqs[currArist] -= freqs[l]
[…]

### 2010-03-17 PythonSrc/demos.py
bug solved

@@ -564,6 +564,8 @@ def knn_from_freqs_on_artists(filenames,codebook,pSize=8,keyInv=True,
         # artist
         artist = artists[songid]
         nMatches = orderedMatches.shape[0]
+        if use_artists:
+            assert nMatches == nArtists
         # get stats
         nExps += 1
         if not use_artists:

### 2010-03-17 PythonSrc/demos.py
bug solved

@@ -527,8 +527,6 @@ def knn_from_freqs_on_artists(filenames,codebook,pSize=8,keyInv=True,
             cntArtists[k] = 0
             artistFreqs[k] = np.zeros([1,nCodes])
         for k in range(artists.shape[0]):
-            if k == l: # skip current file/song
-                continue
             art = artists[k]
             cntArtists[art] += 1
             artistFreqs[art] += freqs[k]

### 2010-03-17 PythonSrc/demos.py
bug solved

@@ -533,10 +533,10 @@ def knn_from_freqs_on_artists(filenames,codebook,pSize=8,keyInv=True,
         # iterate over files
         for l in range(nFiles):
             currArtist = artists[l]
-            currCntArtist = copy.deepcopy(cntArtist)
-            currCntArtist[currArtist] -= 1
+            currCntArtists = copy.deepcopy(cntArtists)
+            currCntArtists[currArtist] -= 1
             currArtistFreqs = copy.deepcopy(artistFreqs)
-            currArtistFreqs[currArist] -= freqs[l]
+            currArtistFreqs[currArtist] -= freqs[l]
             for k in currArtistFreqs.keys(): # normalize
                 currArtistFreqs[k] *= 1. / currCntArtists[k]
             # fill in the line in dists

### 2010-03-20 Paper/Paper.tex
restarting from scratch

--- 2010-03-12 Paper/Paper.tex
@@
   {Third author} {Affiliation3 \\ {\tt author3@ismir.edu}}
-% what order do we use? Thierry, Ron, Dan? is Graham on?
+% what order do we use? Thierry, Ron, Dan?
 
@@
 \begin{abstract}
-We present a method for discovering typical patterns in music.
-Following up on the \textit{Shingle} idea, we bring these harmonic patterns
-into the clustering framework. We use EchoNest API and a training set
-of $43$K songs to assert our method. We restrain ourselves to an online
-algorithm, thus it can scale to arbitrary data sizes.
-We assess the basic good behaviours of our algorithm on this harmonic
-data. We explain how to interpret the centroids of the clusters through
-visualization and further encoding. We discuss ways to move to even larger
-scales and we provide our code for reproducing the results.
-TAKE-HOME MESSAGE: \textit{Shingles} idea can go large-scale?
+
 \end{abstract}
@@
 \section{Introduction}\label{sec:introduction}
-
-Working with large amount of audio data is a goal and a challenge at the
-same time. One of the most common task is to find the nearest neighbor,
-based on some distance, to a given segment of music. This can be viewed
-as \textit{retrieval} as you are looking for some specific target
-based on some query. $k$-nearest neighbors ($k$nn) algorithms have been
-designed for that purpose, and recent improvements let them scale to
-considerable data set sizes \cite{Datar2004}. The \textit{Shingles}
-were introduced in that framework \cite{Casey2006}. A \textit{Shingle}
-is a pattern of stacked chroma features. Similarity between
-\textit{Shingles} coresponds to similarity between harmonic patterns. For this
-reason, they were used for cover or remix recognition \cite{Casey2007}.
-
-That being said, nearest neighbor search is only one aspect of the
-analysis of large-scale music data sets. Finding $k$ neighbors for a
-particular query does not tell you much about the music space.
-Clustering is a larger and more flexible framework where the goal is
-to divide the space in meaningful blocks. The starting point is not
-a query, but all the data itself. In this work, we bring the
-\textit{Shingle} idea into that framework in order to discover
-meaningful and typical music harmonic patterns. The underlying assumption
-is that the music space is articulated around a certain number of
-frequent patterns.
-
-Note that clustering and neighbor search are not unrelated.
-In practise, we would hope that close neighbors live in one or two
-close clusters, thus clustering can be a preprocess step. Also,
-knowing all $k$ neighbors of each sample let you do graph clustering,
-a subset of the clustering algorithms.
-
-We investigate a vector quantization \cite{Datar2004} (VQ) method to
-cluster the \textit{Shingles}. We create a codebook of typical patterns
-that have many usages:
-\begin{enumerate}
-\item describing the music space
-\item finding potential similar songs and/or covers
-\item eliminating quickly very wrong matches in a query system
-\end{enumerate}
-
-A major component of this work is scalability. We experiment with a training
-set of $43$ thousands song, but everything scales linear in time
-relative to the number of samples. We could imagine, as an extension
-of this work, an algorithm ``listening'' to all music it can find online.
-Training time would be an issue, but not memory.
-
-In this paper, we present a practical way to do such large-scale 
-clustering using online vector quantization. We present clustering results 
-of an online algorithm on a data set of $43$K songs, but the method applies 
-to any data size. We show how to select different parameters such as the size 
-of the \textit{Shingles} using an encoding scheme on a test set of $8$K songs. 
-Finally, we illustrate the method with some specific examples.
-
-\subsection{Definitions}
-A pattern, or a patch, is a \textit{Shingle} taken from an actual song.
-A code is similar to a pattern, but is the result of the clustering
-algorithm and most likely does not come from an actual song.
-Distortion is our measure of distance between a pattern and a code
-(or between two patterns, see equation \ref{eq:dist} in Section 
-\ref{sec:experiments}).
-
-\section{Previous Work}\label{sec:prevwork}
-Our work can be seen as an extension of the \textit{Shingles} idea
-described by Casey and Slaney \cite{Casey2006,Casey2007,Casey2008}.
-The two main differences are the clustering framework and the use of
-beats and bars to select the position of the \textit{Shingles}.
-
-Casey and Slaney \cite{Casey2006} use LSH \cite{Datar2004} to perform
-fast approximate nearest neighbor search. Their implementation is E$^2$LSH
-\cite{E2LSH}. This package can efficiently handle about $50$K or $100$K
-vectors of size $200$. Even with a better implementation, the LSH
-model still has some scaling problem. A large portion of the data
-must be known to decide how many projections to use, and what guarantee
-of finding the nearest neighbor it can provides. Also, we could fix the
-number of projections and consider as clusters all the elements that
-fall into the same bin. However, this does not give us a centroid, a
-typical pattern representing that cluster.
-
-Nearest neighbor and clustering methods are common in music similarity 
-research, for instance \cite{Cano2004,Holzapfel2009}. The difference here 
-are the actual patterns that we cluster.
-
-Regarding similar representations of music,
-Barrington et al. recently studied music texture using a set of ordered
-MFCC vectors \cite{Barrington2009a}. This could be seen as complementary
-to our work. Chromas capture the harmonic information and some of the
-rythm while the MFCC are great timbral features.
-
-Other large-scale work use summarization of the features. We mention 
-gaussian mixtures \cite{Mandel2005}, 
-aggregate features \cite{Bertin-Mahieux2008}
-and linear dynamical systems \cite{Barrington2009a}. 
-The advantage of \textit{Shingles} representation (stacked chroma vectors)
-is that it is close from an harmonic point of view to the actual music, 
-in particular
-we can ``listen'' to the patches (see Section \ref{sec:experiments}).
-This is an advantage for analyzing a large data set, thus
-we restrain ourselves from using generative models to compress our features.
-
-\section{Data}\label{sec:data}
-In this section we present our use of the Echo Nest features and the
-datasets we use for training and testing our model.
-
-\subsection{Echo Nest features}
-We use the Echo Nest analyze API \cite{EchoNest} throughout this work.
-The API gives us a chroma vector (length $12$) for every music event, 
-or ``segment'', for any song we upload to their platform. 
-We also get an estimate of where the beats and the bars are. Beats span over
-multiple segments, and bars span over multiple beats. 
-We transform the chroma vectors per segment into per beat vectors using a 
-simple average. We can then stack beat vectors for a number of bars 
-(usually 1 or 2). 
-We get a fixed-size patch by resampling the number of chroma vectors. Typically,
-there are $4$ beats per bar. Thus an appropriate pattern size for one bar
-is $12$x$4$ or $12$x$8$.
-
-Note that we do not claim that any of these informations (segments, beats, bars)
-are perfectly accurate (e.g. in \cite{Barrington2009a} authors obtain a better
-song segmentation than Echonest). Practise showed us that they are reasonable, 
-and the size of the data set should make up for the imperfections or noise.
-We also believe that patches sizes based on a number of beats or bars are more
-meaningful than an arbitrary time length. More on this in the experiments
-(Section \ref{sec:experiments}).
-
-\begin{figure}[htb]
-\begin{center}
-\includegraphics[width=.8\columnwidth]{code}
-\end{center}
-\caption{{A typical code from a codebook of size $200$. Patterns represented
-$2$ bars and the pattern length was set to $16$.}}
-\label{fig:code}
-\end{figure}
-
-\subsection{Cowbell Dataset}
-We have $43,300$ songs, giving  $3,720,091$ non zero bars, a bar usually 
-contains $4$ beats. WHERE DOES THIS DATA COMES FROM? GIFT FROM EN?
-
-\subsection{uspop2002}
-We had acces to a low quality (32kbps) version of the songs in the uspop 2002 
-data set \cite{uspop2002}.
-This well-known data show a great diversity of pop songs, but is not particular
-otherwise. We also tried to use Tzanetakis dataset \cite{Tzanetakis2002a}, but 
-the $30$ seconds segments seemed hard to analyze by the Echo Nest API, most did
-not contain any bar.
-
-uspop2002 serves as testing set to measure how well a codebook learned on
-the Cowbell data set can represent new songs. We get Echo Nest features
-for $8651$ songs from that dataset.
-
-\section{Algorithm}\label{sec:algo}
-In this section we present the creation of the harmonic patterns from
-the Echo Nest data. Then, we explain the online vector quantization algorithm
-used to learn a codebook of those patterns.
-
-\subsection{Pattern Creation}
-Here we should explain key invariance, resizing, etc, and not in the
-experiment section.
-
-We try to normalize the patterns in respect to the key. We roll the matrix
-so the row with the more energy is always the lowest one. This can easily be
-seen in Figure \ref{fig:code}. We considered the same technique for rolling
-along the other axis and be invariant regarding the downbeat. However, it
-seems less robust to noise. Note that we roll each pattern independentely.
-Also, this implies that the pitch envelope is important rather than the actual 
-pitches.
-
-\subsection{Codebook Learning}
-We use an online version of the vector quantization algorithm 
-\cite{Gersho1991}. This can also be seen as an online K-means.
-The idea is as follow: take a sample from your data, find the closest
-code in the codebook, bring that code closer to the sample by some amount.
-The details are explained in Algorithm \ref{algo:vq}. Note that this
-algorithm, though not optimal, is $O(n)$ with $n$ number of songs seen.
-Thus it scales linearly to any size of data set. The setting of the learning
-rate might become trickier, but this is not fundamentally a problem.
-
-We initialize the codebook by choosing $K$ random points from our dataset.
-
-\begin{algorithm}
-%\caption{Pseudocode of Vector Quantization}
-\begin{algorithmic}
-\STATE$l$ learning rate
-\STATE$\{P_n\}$ set of patches
-\STATE$\{C_k\}$ codebook of $K$ codes
-%\STATE $m \leftarrow b$
-\REQUIRE $0 < l \leq 1$
-\FOR{$nIters$}
-\FOR{$p \in \{P_n\}$}
-\STATE$c \leftarrow min_{c \in C_k} dist(p,c)$
-\STATE$c \leftarrow c + (p - c) * l$
-\ENDFOR
-\ENDFOR
-\RETURN $\{C_k\}$
-\caption{{Pseudocode of Online Vector Quantization. Note that we can replace
-the number of iteration by a threshold on the distortion over some test set.}
-\label{algo:vq}}
-\end{algorithmic}
-\end{algorithm}
-
-\section{Experiments}\label{sec:experiments}
-In this section we present general experiments that describe our method.
-We start by giving some details on how we selected our training set and
-how to measure error. Secondly, we experiment with different pattern sizes.
-Thirdly, we look at the effect of the size of the
-data set on the training. The next experiment look at the size of the
-codebook. Then, we show that the patterns are not random,
-which is expected from music data. Finally, we show examples of patterns,
-codebooks and clusters.
-
-\subsection{Setting}\label{ssec:setting}
-We take one or two bars, normalize the patches to size 4, 8, or 16.
-We roll the patch to be invariant to the key (on the patch level, not on
-the song level). We learn a codebook of size $K$ over the cowbell dataset 
-using the VQ algorithm (Algorithm \ref{algo:vq}). A typical learning rate 
-is $0.01$ for $20$ iterations over the whole dataset.
-
-We use the codebook to encode our testing set (see Section \ref{sec:data}).
-Each pattern is encoded with only one code. We can measure the average
-distance between a pattern and its encoding. We can also measure the use
-of the different codes, i.e. the size of the cluster around a code.
-
-As a distance measure (or distortion measure) between patterns, we use
-the average squared euclidean distance. In details, if a pattern $p1$
-is composed of elements $p1_{i,j}$, and similarly for a pattern $p2$,
-the distance between $p1$ and $p2$ is:
-\begin{eqnarray}
-  dist(p1,p2) = \sum_{i,j} (p1_{i,j} - p2_{i,j})^2 / size(p1)  \label{eq:dist}
-\end{eqnarray}
-We assume $p1$ and $p2$ have the same size.
-
-\subsection{Pattern length}
-See encoding experiments in Table \ref{tab:psize}.
-
-% RESULTS
-% 200codes, 2 bars, pSize=16, maxener, True, False, not scaled = 
-% 200codes, 2 bars, pSize=8, maxener, True, False, not scaled  = 0.042432
-% 200codes, 1 bars, pSize=4, maxener, True, False, not scaled  = 
-% 200codes, 7 beats, pSize=7, maxener, True, False, not scaled = 
-% 200codes, 5 beats, pSize=5, maxener, True, False, not scaled = 
-% 200codes, 3 beats, pSize=3, maxener, True, False, not scaled = 
-\begin{table}
-\begin{center}
-\begin{tabular}{|l|c|}
-\hline
-size & avg. dist \\ \hline \hline
-1 bar (4) & \\
-2 bars (8) & $$\\
-2 bars (12) & $$\\
-2 bars (16) & $$ \\ \hline
-3 beats & \\
-5 beats & $$\\
-7 beats & $$\\ \hline
-\end{tabular}
-\end{center}
-\caption{{Encoding score for different pattern sizes. We can select the
-size of the patterns based on bars. In this case, we put the actual size
-of the pattern in parentheses. If the pattern size is based on beats, the
-size is the number of beats. Results are on the uspop dataset, not seen
-during training.}}
-\label{tab:psize}
-\end{table}
-
-\subsection{Data Size}
-We are interesting to see how the amount of training data changes
-the quality of our codebook.
-We experiment with different data sizes for a fixed size of codebook and a 
-fixed code length. The data points are chose randomly from the whole
-data set.
-In the experiment a pattern size correspond to $1$ bar resized to $4$
-beats, thus a pattern size of $12$x$4$. The codebook size is $100$.
-We measure the distortion on our test set. 
-See results in Figure \ref{fig:sizes}.
-
-Our learning can overfit, which explain the bad scores of small datasets.
-The larger the dataset, the smallest the chance to overfit.
-We are looking for a size of dataset that guarantees us a good learning
-process, while being unnecessary large.
-
-$500$ samples per code in the codebook seems a good data size, see
-Figure \ref{fig:sizes} for size $50$K.
-
-\begin{figure}[htb]
-\begin{center}
-\includegraphics[width=.99\columnwidth]{data_sizes}
-\end{center}
-\caption{{Distortion for a codebook of size $100$ encoding one bar
-at a time, a bar being represented by $4$ chroma feature vector.
-Therefore, a code or a pattern is of size $12$ x $4$ = $48$.
-Distortion measure on the test data set. Data size tested range
-from $10$ thousands to $2$ millions. Patterns were selected at
-random from the dataset of approximately $3,7$ million patterns.}}
-\label{fig:sizes}
-\end{figure}
-
-\iffalse
-\begin{table}
-\begin{center}
-\begin{tabular}{|l|c|c|c|c|c|c|}
-\hline
-size & $5$K & $10$K & $50$K & $100$K & $500$K & $2$M  \\ \hline
-distortion & $$ & $0.143$ & $0.142$ & $0.142$ & $0.142$ & $$\\ \hline
-\end{tabular}
-\end{center}
-\caption{{Distortion for a codebook of size $100$ encoding one bar
-at a time, a bar being represented by $4$ chroma feature vector.
-Therefore, a code or a pattern is of size $12$ x $4$ = $48$.
-Distortion measure on the test data set. Data size tested range
-from $10$ thousands to $2$ millions. Patterns were selected at
-random from the dataset of approximately $3,7$ million patterns.}}
-\label{tab:datasize}
-\end{table}
-\fi
-% RESULTS:    
-% 1bar s4     avg_distOLD    avg_dist
-% 1K          0.150370       0.041941
-% 5K          0.145327       0.038449
-% 10K         0.143567       0.036892
-% 50K         0.142257       0.036722
-% 100K        0.142195       
-% 250K        0.141796
-% 500K        0.142081
-% 1M          0.142118
-% 2M          0.141596
-% 1bar s4 first X samples
-% 1K          0.157142       
-% 5K          0.150045
-% 10K         0.148357
-% 50K         0.146069
-% 100K        0.145765
-% 250K        0.145819
-% 500K        0.145772
-% 1M          0.145924
-% 2M          0.146065       
-% 0bar s4 (and test set same, we could do 1bar s4 in test set...)
-% 1K          0.155474       0.042596
-% 5K          0.149619
-% 10K         0.148056
-% 50K         0.146328
-% 100K        0.146004
-% 250K        0.146024
-% 500K        0.145962
-% 1M          0.145873
-% 2M          0.145955
-% 2bar s8     
-% 1K          0.171300       0.049282
-% 5K          0.165960       0.046441
-% 10K         0.165063       0.045507
-% 50K         0.162439
-% 100K        0.161588
-% 250K        0.161910
-% 500K        0.161233
-% 1M          0.161057
-% 2M          0.161068
-
-\subsection{Codebook Size}
-We experiment with the size of the codebook and its affect on the
-performance of the encoding. Results are shown in Table \ref{tab:cbsize}.
-Note that we trained on a number of samples depending on the codebook
-size, i.e. we used $500$ samples per code to learn. This is consistent
-with our findings from Figure \ref{fig:sizes}.
-
-\begin{table}
-\begin{center}
-\begin{tabular}{|l|l|c|}
-\hline
-\# codes & \# samples & avg. dist \\ \hline \hline
-1 & $500$ & $0.066079$ \\
-10 & $5$K & $0.045690$ \\
-50 & $25$K & $0.038456$ \\
-100 & $50$K & $0.036725$ \\
-200 & $100$K & $0.033923$ \\
-500 & $250$K & $0.030973$ \\
-1000 & $500$K & \\ \hline
-\end{tabular}
-\end{center}
-\caption{{Experiment with the codebook size. We train on patterns
-corresponding to $1$ bar resized to $4$ beats. The number of training
-samples is set to $500$ times the number of codes, which is sufficient
-according to Figure \ref{fig:sizes}.}}
-\label{tab:cbsize}
-\end{table}
-% RESULTS:      avg_distsOLD    avg_dist
-% code 1        0.195939        0.066079
-% code 10       0.161812        0.045690
-% code 50       0.147472        0.038456
-% code 100      0.142194        0.036725
-% code 200      0.137548        0.033923
-% code 500      0.131211        0.030973
-
-\subsection{Codebook Size in Function of Pattern Length}
-We experiment on the randomness of the \textit{Shingles}, assuming
-that there is some correlation between beats and bars.
-The intuition is as follow: if the signal was totally random, every
-time the signal size double, the codebook size must be squared.
-To illustrate this, imagine a binary array of length $1$. We need two
-codes to encode perfectly: $(0)$ and $(1)$. If the binary array is now of
-length $2$, we need four codes: $(0,0)$, $(0,1)$, $(1,1)$ and $(1,0)$.
-
-In Table \ref{tab:size_pattern}, we test that hypothesis. As expected,
-the distortion improves with larger pattern sizes. The details
-are difficult to analyze. The error jump from two beats to one bar
-can probably be explained by the obvious case of sustained notes.
-We also believe that smaller codebooks are more affected by their
-initialization. Fortunately, we can conclude that music is not random,
-which is always a good sanity check!
-
-Then... 3 beats, 5 beats, 7 beats don't encode as well?
-
-Special: 4 beats versus 1 bar - 4 beats, show the problem of being
-not in sync with bars.
-
-See codebook size experiments in Table \ref{tab:cbsize}.
-
-\begin{table}
-\begin{center}
-\begin{tabular}{|l|c|c|}
-\hline
-\# codes & \# bars & avg. dist \\ \hline \hline
-2 & $1$ & $0.058435$ \\
-4 & $2$ & $0.054792$ \\
-16 & $4$ & $0.055489$ \\
-256 & $8$ & $0.054860$ \\ \hline
-2 & $2$ beats* & $0.066416$ \\
-4 & $1$ & $0.050712$ \\
-16 & $2$ & $0.050172$ \\
-256 & $4$ & $0.049894$ \\ \hline
-\end{tabular}
-\end{center}
-\caption{{Codebook size in function of the number of bars to encode.
-All experiments done wtih $256 \times 500 = 128$K samples randomly chosen.
-* For the $2$ beats experiment, we take the first half of every beats.
-It is different from ignoring bars and taking every $2$ beats. It is more
-coherent with the other settings ($1-2-4-8$ bars).}}
-\label{tab:size_pattern}
-\end{table}
-% RESULTS:           avg_distsOLD   avg_dist
-%2 codes 1 bar       0.180750       0.058435
-%4 codes 2 bars      0.183631       0.054792
-%16 codes 4 bars     0.184078       0.055489
-%256 codes 8 bars    0.183464       0.054860
-%---------------
-%2 codes 2 beats     0.166840       0.046955
-%2 codes 2 beats                    0.066416
-% based on bars
-%4 codes 1 bar       0.171495       0.050712
-%16 codes 2 bars     0.173767       0.050172
-%256 codes 4 bars    0.174169       0.049894
-
-% see generate_expo_codes_curves.py to generate the following figure:
-\begin{figure}[htb]
-\begin{center}
-\includegraphics[width=.99\columnwidth]{codesize_patternsize}
-\end{center}
-\caption{{Possible figure to replace Table \ref{tab:size_pattern}.}}
-\label{fig:size_pattern}
-\end{figure}
-
-\subsection{Visualization}
-Interpreting a codebook of size larger than $100$ is not obvious.
-We can use an algorithm like locally linear embedding \cite{Roweis2000} (LLE) 
-or multidimensional scaling \cite{Kruskal1964} to visualize the data in
-2 dimensions. Here we use the former\footnote{implementation: 
-\url{http://www.astro.washington.edu/users/vanderplas/coding/LLE/}}, the
-result can be seen in Figure \ref{fig:lle}.
-
-\begin{figure}[htb]
-\begin{center}
-\includegraphics[width=.9\columnwidth]{codes_lle}
-\end{center}
-\caption{{LLE visualization of the codebook.}}
-\label{fig:lle}
-\end{figure}
-
-Another way is simply to present the codes in order of cluster size.
-We encode the training or testing set as explained in Subsection 
-\ref{ssec:setting} and we count how many patterns were closer to a particular
-code than any other code. In Figure \ref{fig:firstcodes}, we show the $25$
-most used codes out of a codebook of size $200$.
-Code from Figure \ref{fig:code} is the 37th most used code in that
-codebook. In Figure \ref{fig:closep}, we show a sample of patterns that
-fall into its cluster.
-
-\begin{figure}[htb]
-\begin{center}
-\includegraphics[width=.9\columnwidth]{codes1}
-\end{center}
-\caption{{$25$ most used codes on the training set.}}
-\label{fig:firstcodes}
-\end{figure}
-
-The most common pattern is a sustained note, as we can see in Figure
-\ref{fig:firstcodes}. Another common pattern is a sustained note
-with an harmonic, as seen in the LLE visualization (Figure \ref{fig:lle}).
-
-\begin{figure}[htb]
-\begin{center}
-\includegraphics[width=.9\columnwidth]{close_patterns1}
-\end{center}
-\caption{{Sample of the close patterns to code in Figure \ref{fig:code}.}}
-\label{fig:closep}
-\end{figure}
-
-\subsection{Specific Examples}
-
-\section{Links with NMF}\label{sec:nmf}
-In this section we discuss links with non-negative matrix factorization
-\cite{Lee2000} (NMF), another method to build a codebook representing some data.
-NMF finds two matrices $H$ and $W$ that approximate some data $D$ (one
-sample per line) by
-\begin{eqnarray}
-D \approx W^{m \times k} H^{k \times n}
-\end{eqnarray}
-where $\epsilon$ is some noise and $W$ and $H$ only have non-negative 
-elements. If $W$ is $\#samples \times k$, $k$ is the codebook size.
-
-Note that NMF does not solve the same problem. It tries to explain each pattern
-as a combination of codes. Thus it does not perform clustering, though you
-could cluster patterns using their encoding weights. Also, we do not fully
-compare our methods with NMF, it would be a paper on its own.
-
-In all our experiments we use the following NMF
-implementation\footnote{implementation: \url{http://www.ee.columbia.edu/~grindlay/code.html}}.
-
-\subsection{Training NMF on our Codebook}
-NMF provides us with a way of measuring how much information is retained
-in our codebook. More specifically, we can train a codebook of size $500$
-using our online VQ algorithm, then decompose that code into $10$ basis
-using NMF. We can compare that to $10$ basis from training NMF on the
-whole data.
-
-The idea is simple, if we encode our test set using basis from the
-codebook as well as if we were using basis from the whole data, our codes
-retain much of the information. That being said, the typical NMF implementations
-are not expected to scale to data that does not fit in memory. Online VQ
-could be an interesting preprocessing step. Future work include looking
-at online NMF \cite{Thurau2009}.
-
-\subsection{Experiment}
-Put figure, compare basis, yeah!
-
-\begin{table}
-\begin{center}
-\begin{tabular}{|l|l|c|c||c|}
-\hline
-\# codes & \# bars & full data & codebook & test set \\ \hline \hline
-10 & 1 & $$ & $$ & \\ \hline
-20 & 1 & $$ & $$ & \\ \hline
-10 & 2 & $\bf{1774788}$ & $$ & $1782164$\\ \hline
-20 & 2 & $$ & $$ & \\ \hline
-\end{tabular}
-\end{center}
-\caption{{Encoding error of the test data set using NMF basis learned
-on the whole data and on the codebook. We also encode the test set
-directly to give us an upper bound.}}
-\label{tab:nmf}
-\end{table}
-% RESULTS:         whole data             codebook     test set
-%10 codes 2 bars   1774788.2566 (200 it)               1739639.7304 (1000 it)
-%10 codes 1 bar
-
-\section{Conclusion}
-We presented a practical method to perform large-scale clustering of
-harmonic patterns. We assessed the basic properties of the method through
-experiments on a large collection of music. We suggested many ways
-to interpret the data through visualization (LLE) and further encoding
-(NMF). We discussed the possibility to move to even larger scales
-and we provided our source code\footnote{Code not yet released to preserve
-submission's anonymity}.
-
-As for enhancements, we specifically did not compress our features using
-gaussian mixtures or other generative model. That being said, these methods
-could be use to develop better distance measures between \textit{Shingles}.
-The use of the euclidean distance is arbitrary. Summarizing patches
-with gaussians, and then comparing the distance between those gaussians,
-could reduce the influence of the noise in the distance measure.
-
-Moving on to larger scales, we could imagine merging codes together and
-splitting some on two if the clusters become too small or too large.We could
-also cluster the codebook itself, in a similar fashion as hierarchical
-gaussian mixtures \cite{Vasconcelos2001}.
 

### 2010-03-20 Paper/Paper.tex
new draft improved

@@
   {Third author} {Affiliation3 \\ {\tt author3@ismir.edu}}
-% what order do we use? Thierry, Ron, Dan?
+% what order do we use? Thierry, Ron, Dan?  Dan, Thierry, Ron?
 
@@
 \begin{abstract}
-
+The goal is to analyze very large collections of music. We describe
+a clustering scheme of harmonic patterns that scales linearly in time
+with the number of songs analyzed. These harmonic patterns are related
+to the \textit{Shingles} idea, but we scale to larger sets and we do not
+perform $k$-nn with them. We create a codebook and assert
+its properties. In particular, we show that clustering with notions of
+bars and song segments helps, and that we get codes associated more
+with some artists than others.
 \end{abstract}
@@
 \section{Introduction}\label{sec:introduction}
+
+Large-scale \cite{Bertin-Mahieux2008}
+
+\section{Previous Work}\label{sec:prevwork}
+Casey and Slaney \cite{Casey2006,Casey2007,Casey2008}
+
+\section{Data}\label{sec:data}
+In this section we present our use of the Echo Nest features and the
+datasets we use for training and testing our model.
+
+\subsection{EchoNest API}
+
+\subsection{USPOP}
+
+\subsection{Artist Set}
+
+\section{Algorithm}
+Wth the scaling factor as our main concern, we present an online
+vector quantization algorithm to cluster the harmonic patterns.
+
+\begin{algorithm}
+%\caption{Pseudocode of Vector Quantization}
+\begin{algorithmic}
+\STATE$l$ learning rate
+\STATE$\{P_n\}$ set of patches
+\STATE$\{C_k\}$ codebook of $K$ codes
+%\STATE $m \leftarrow b$
+\REQUIRE $0 < l \leq 1$
+\FOR{$nIters$}
+\FOR{$p \in \{P_n\}$}
+\STATE$c \leftarrow min_{c \in C_k} dist(p,c)$
+\STATE$c \leftarrow c + (p - c) * l$
+\ENDFOR
+\ENDFOR
+\RETURN $\{C_k\}$
+\caption{{Pseudocode of Online Vector Quantization. Note that we can replace
+the number of iteration by a threshold on the distortion over some test set.}
+\label{algo:vq}}
+\end{algorithmic}
+\end{algorithm}
+
+\section{Experiments}\label{sec:experiments}
+
+\subsection{Basic Properties}
+This section present some fundamental results of the clustering algorithm.
+These are important, but do not change the larger picture. Thus we will not
+spend too much time describing them. In short: more training samples is
+better, more codes is better, larger patterns are more difficult.
+
+First, a larger dataset gives a better encoding codebook. See Figure
+\ref{fig unknown}.
+
+Secondly, a larger codebook better encodes new song. See Figure \ref{ejdue}.
+
+Finally, larger patterns are more difficult to encode, thus require
+larger codebooks. See Figure \ref{mouahahah}.
+
+These properties are important to verify and measure, but they are not
+surprising for a clustering algorithm. We move on to more interestng facts.
+
+\subsection{Music Randomness}
+It will not surprise any reader, music is not random. This fact should
+become apparent in our experiments in the following way. Information
+theory tells us that if the signal to encode doubles, we need to square
+the size of the codebook to encode it with the same distortion.
+
+To help the intuition, here is an example. Imagine a binary array of 
+length $1$. We need two
+codes to encode perfectly: $(0)$ and $(1)$. If the binary array is now of
+length $2$, we now need four codes: $(0,0)$, $(0,1)$, $(1,1)$ and $(1,0)$.
+
+Since music is not random, we expect that squaring the codebook when the
+pattern size doubles will rather give improved results than constant ones.
+See Figure \ref{it's comng...}.
+
+\section{Visualization}
+The goal here is to give an intuition to the reader of what kind of
+typical patterns do we get. We also want to have a feel for what the
+clusters look like, how many boring sustaned notes are encoded in the
+codebook, can the codebook describes the complexity of the music space, etc.
+
+Below we present some figures to help the reader with those questions.
+Please keep in mind the limited space.
+
+\subsection{Encoding of a Song}
+Here we present a song with its two encodings. First, the harmonic patterns
+given by the EchoNest. Secondly, each pattern is encoded by its closest
+code in the codebook. See Figure \ref{youppie}.
+
+\subsection{Codebook}
+We can present the most typical patterns (Figure \ref{1}). 
+Then, LLE? (Figure \ref{2}). 
+Then, the graph on variance. (Figure \ref{fig:code_var}).
+
+\begin{figure}[htb]
+\begin{center}
+\includegraphics[width=.9\columnwidth]{codes_lle}
+\end{center}
+\caption{Locally linear embedding (LLE) visualization of the codebook.}
+\label{fig:lle}
+\end{figure}
+
+\begin{figure}[htb]
+\begin{center}
+\includegraphics[width=.9\columnwidth]{code_variance}
+\end{center}
+\caption{Average variance in time of a codebook.}
+\label{fig:code_var}
+\end{figure}
+
+\subsection{Cluster}
+Here we present real patterns that are closer to the code presented in ...
+than any other code. The goal is to feel how much variance is present in
+a cluster.
+
+\section{Further Experiments}
+
+\subsection{Song Segmentation}
+We explained that we used bars as defined by the EchoNest algorithm to
+train our patterns. Let's assume once again that those bars are
+reasonable. We investigate whether we could recover the bar segmentation
+of new songs from our codebook. It would helps us assert that codes
+encodes something related to bars, for instance a strong beat at the
+beginning.
+
+\subsection{Artist Recognition}
+Let us be clear, the clustering method proposed in this work is not
+intended to perform well at artist recognition. An obvious fact is that
+we lose all timbral information which would be usefull. However, we can
+imagine that some artist use more often some patterns than others.
+See Figure \ref{fig:conf_mat}.
+
+\begin{figure}[htb]
+\begin{center}
+\includegraphics[width=.9\columnwidth]{conf_mat_per_artist}
+\end{center}
+\caption{Confusion matrix for artist recognition task.}
+\label{fig:conf_mat}
+\end{figure}
+
+\section{Conclusion and Future Work}
 

### 2010-03-20 Paper/draft_20_03_10.tex
most typical pattern of metallica

% -----------------------------------------------
% Template for ISMIR 2010
% (based on earlier ISMIR templates)
% -----------------------------------------------

\documentclass{article}
\usepackage{ismir2010,amsmath,cite}
\usepackage{graphicx}
\usepackage{url}
\usepackage{algorithm,algorithmic}

% Title.
% ------
\title{Large-scale harmonic patterns clustering}

% Single address
% To use with only one author or several with the same address
% ---------------
%\oneauthor
% {Names should be omitted for double-blind reviewing}
% {Affiliations should be omitted for double-blind reviewing}

% Two addresses
% --------------
%\twoauthors
%  {First author} {School \\ Department}
%  {Second author} {Company \\ Address}

% Three addresses
% --------------
\threeauthors
  {First author} {Affiliation1 \\ {\tt author1@ismir.edu}}
  {Second author} {\bf Retain these fake authors in\\\bf submission to preserve the formatting}
  {Third author} {Affiliation3 \\ {\tt author3@ismir.edu}}
% what order do we use? Thierry, Ron, Dan? is Graham on?

\begin{document}
%
\maketitle
%
\begin{abstract}
We present a method for discovering typical patterns in music.
Following up on the \textit{Shingle} idea, we bring these harmonic patterns
into the clustering framework. We use EchoNest API and a training set
of $43$K songs to assert our method. We restrain ourselves to an online
algorithm, thus it can scale to arbitrary data sizes.
We assess the basic good behaviours of our algorithm on this harmonic
data. We explain how to interpret the centroids of the clusters through
visualization and further encoding. We discuss ways to move to even larger
scales and we provide our code for reproducing the results.
TAKE-HOME MESSAGE: \textit{Shingles} idea can go large-scale?
\end{abstract}
%
\section{Introduction}\label{sec:introduction}

Working with large amount of audio data is a goal and a challenge at the
same time. One of the most common task is to find the nearest neighbor,
based on some distance, to a given segment of music. This can be viewed
as \textit{retrieval} as you are looking for some specific target
based on some query. $k$-nearest neighbors ($k$nn) algorithms have been
designed for that purpose, and recent improvements let them scale to
considerable data set sizes \cite{Datar2004}. The \textit{Shingles}
were introduced in that framework \cite{Casey2006}. A \textit{Shingle}
is a pattern of stacked chroma features. Similarity between
\textit{Shingles} coresponds to similarity between harmonic patterns. For this
reason, they were used for cover or remix recognition \cite{Casey2007}.

That being said, nearest neighbor search is only one aspect of the
analysis of large-scale music data sets. Finding $k$ neighbors for a
particular query does not tell you much about the music space.
Clustering is a larger and more flexible framework where the goal is
to divide the space in meaningful blocks. The starting point is not
a query, but all the data itself. In this work, we bring the
\textit{Shingle} idea into that framework in order to discover
meaningful and typical music harmonic patterns. The underlying assumption
is that the music space is articulated around a certain number of
frequent patterns.

Note that clustering and neighbor search are not unrelated.
In practise, we would hope that close neighbors live in one or two
close clusters, thus clustering can be a preprocess step. Also,
knowing all $k$ neighbors of each sample let you do graph clustering,
a subset of the clustering algorithms.

We investigate a vector quantization \cite{Datar2004} (VQ) method to
cluster the \textit{Shingles}. We create a codebook of typical patterns
that have many usages:
\begin{enumerate}
\item describing the music space
\item finding potential similar songs and/or covers
\item eliminating quickly very wrong matches in a query system
\end{enumerate}

A major component of this work is scalability. We experiment with a training
set of $43$ thousands song, but everything scales linear in time
relative to the number of samples. We could imagine, as an extension
of this work, an algorithm ``listening'' to all music it can find online.
Training time would be an issue, but not memory.

In this paper, we present a practical way to do such large-scale 
clustering using online vector quantization. We present clustering results 
of an online algorithm on a data set of $43$K songs, but the method applies 
to any data size. We show how to select different parameters such as the size 
of the \textit{Shingles} using an encoding scheme on a test set of $8$K songs. 
Finally, we illustrate the method with some specific examples.

\subsection{Definitions}
A pattern, or a patch, is a \textit{Shingle} taken from an actual song.
A code is similar to a pattern, but is the result of the clustering
algorithm and most likely does not come from an actual song.
Distortion is our measure of distance between a pattern and a code
(or between two patterns, see equation \ref{eq:dist} in Section 
\ref{sec:experiments}).

\section{Previous Work}\label{sec:prevwork}
Our work can be seen as an extension of the \textit{Shingles} idea
described by Casey and Slaney \cite{Casey2006,Casey2007,Casey2008}.
The two main differences are the clustering framework and the use of
beats and bars to select the position of the \textit{Shingles}.

Casey and Slaney \cite{Casey2006} use LSH \cite{Datar2004} to perform
fast approximate nearest neighbor search. Their implementation is E$^2$LSH
\cite{E2LSH}. This package can efficiently handle about $50$K or $100$K
vectors of size $200$. Even with a better implementation, the LSH
model still has some scaling problem. A large portion of the data
must be known to decide how many projections to use, and what guarantee
of finding the nearest neighbor it can provides. Also, we could fix the
number of projections and consider as clusters all the elements that
fall into the same bin. However, this does not give us a centroid, a
typical pattern representing that cluster.

Nearest neighbor and clustering methods are common in music similarity 
research, for instance \cite{Cano2004,Holzapfel2009}. The difference here 
are the actual patterns that we cluster.

Regarding similar representations of music,
Barrington et al. recently studied music texture using a set of ordered
MFCC vectors \cite{Barrington2009a}. This could be seen as complementary
to our work. Chromas capture the harmonic information and some of the
rythm while the MFCC are great timbral features.

Other large-scale work use summarization of the features. We mention 
gaussian mixtures \cite{Mandel2005}, 
aggregate features \cite{Bertin-Mahieux2008}
and linear dynamical systems \cite{Barrington2009a}. 
The advantage of \textit{Shingles} representation (stacked chroma vectors)
is that it is close from an harmonic point of view to the actual music, 
in particular
we can ``listen'' to the patches (see Section \ref{sec:experiments}).
This is an advantage for analyzing a large data set, thus
we restrain ourselves from using generative models to compress our features.

\section{Data}\label{sec:data}
In this section we present our use of the Echo Nest features and the
datasets we use for training and testing our model.

\subsection{Echo Nest features}
We use the Echo Nest analyze API \cite{EchoNest} throughout this work.
The API gives us a chroma vector (length $12$) for every music event, 
or ``segment'', for any song we upload to their platform. 
We also get an estimate of where the beats and the bars are. Beats span over
multiple segments, and bars span over multiple beats. 
We transform the chroma vectors per segment into per beat vectors using a 
simple average. We can then stack beat vectors for a number of bars 
(usually 1 or 2). 
We get a fixed-size patch by resampling the number of chroma vectors. Typically,
there are $4$ beats per bar. Thus an appropriate pattern size for one bar
is $12$x$4$ or $12$x$8$.

Note that we do not claim that any of these informations (segments, beats, bars)
are perfectly accurate (e.g. in \cite{Barrington2009a} authors obtain a better
song segmentation than Echonest). Practise showed us that they are reasonable, 
and the size of the data set should make up for the imperfections or noise.
We also believe that patches sizes based on a number of beats or bars are more
meaningful than an arbitrary time length. More on this in the experiments
(Section \ref{sec:experiments}).

\begin{figure}[htb]
\begin{center}
\includegraphics[width=.8\columnwidth]{code}
\end{center}
\caption{{A typical code from a codebook of size $200$. Patterns represented
$2$ bars and the pattern length was set to $16$.}}
\label{fig:code}
\end{figure}

\subsection{Cowbell Dataset}
We have $43,300$ songs, giving  $3,720,091$ non zero bars, a bar usually 
contains $4$ beats. WHERE DOES THIS DATA COMES FROM? GIFT FROM EN?

\subsection{uspop2002}
We had acces to a low quality (32kbps) version of the songs in the uspop 2002 
data set \cite{uspop2002}.
This well-known data show a great diversity of pop songs, but is not particular
otherwise. We also tried to use Tzanetakis dataset \cite{Tzanetakis2002a}, but 
the $30$ seconds segments seemed hard to analyze by the Echo Nest API, most did
not contain any bar.

uspop2002 serves as testing set to measure how well a codebook learned on
the Cowbell data set can represent new songs. We get Echo Nest features
for $8651$ songs from that dataset.

\section{Algorithm}\label{sec:algo}
In this section we present the creation of the harmonic patterns from
the Echo Nest data. Then, we explain the online vector quantization algorithm
used to learn a codebook of those patterns.

\subsection{Pattern Creation}
Here we should explain key invariance, resizing, etc, and not in the
experiment section.

We try to normalize the patterns in respect to the key. We roll the matrix
so the row with the more energy is always the lowest one. This can easily be
seen in Figure \ref{fig:code}. We considered the same technique for rolling
along the other axis and be invariant regarding the downbeat. However, it
seems less robust to noise. Note that we roll each pattern independentely.
Also, this implies that the pitch envelope is important rather than the actual 
pitches.

\subsection{Codebook Learning}
We use an online version of the vector quantization algorithm 
\cite{Gersho1991}. This can also be seen as an online K-means.
The idea is as follow: take a sample from your data, find the closest
code in the codebook, bring that code closer to the sample by some amount.
The details are explained in Algorithm \ref{algo:vq}. Note that this
algorithm, though not optimal, is $O(n)$ with $n$ number of songs seen.
Thus it scales linearly to any size of data set. The setting of the learning
rate might become trickier, but this is not fundamentally a problem.

We initialize the codebook by choosing $K$ random points from our dataset.

\begin{algorithm}
%\caption{Pseudocode of Vector Quantization}
\begin{algorithmic}
\STATE$l$ learning rate
\STATE$\{P_n\}$ set of patches
\STATE$\{C_k\}$ codebook of $K$ codes
%\STATE $m \leftarrow b$
\REQUIRE $0 < l \leq 1$
\FOR{$nIters$}
\FOR{$p \in \{P_n\}$}
\STATE$c \leftarrow min_{c \in C_k} dist(p,c)$
\STATE$c \leftarrow c + (p - c) * l$
\ENDFOR
\ENDFOR
\RETURN $\{C_k\}$
\caption{{Pseudocode of Online Vector Quantization. Note that we can replace
the number of iteration by a threshold on the distortion over some test set.}
\label{algo:vq}}
\end{algorithmic}
\end{algorithm}

\section{Experiments}\label{sec:experiments}
In this section we present general experiments that describe our method.
We start by giving some details on how we selected our training set and
how to measure error. Secondly, we experiment with different pattern sizes.
Thirdly, we look at the effect of the size of the
data set on the training. The next experiment look at the size of the
codebook. Then, we show that the patterns are not random,
which is expected from music data. Finally, we show examples of patterns,
codebooks and clusters.

\subsection{Setting}\label{ssec:setting}
We take one or two bars, normalize the patches to size 4, 8, or 16.
We roll the patch to be invariant to the key (on the patch level, not on
the song level). We learn a codebook of size $K$ over the cowbell dataset 
using the VQ algorithm (Algorithm \ref{algo:vq}). A typical learning rate 
is $0.01$ for $20$ iterations over the whole dataset.

We use the codebook to encode our testing set (see Section \ref{sec:data}).
Each pattern is encoded with only one code. We can measure the average
distance between a pattern and its encoding. We can also measure the use
of the different codes, i.e. the size of the cluster around a code.

As a distance measure (or distortion measure) between patterns, we use
the average squared euclidean distance. In details, if a pattern $p1$
is composed of elements $p1_{i,j}$, and similarly for a pattern $p2$,
the distance between $p1$ and $p2$ is:
\begin{eqnarray}
  dist(p1,p2) = \sum_{i,j} (p1_{i,j} - p2_{i,j})^2 / size(p1)  \label{eq:dist}
\end{eqnarray}
We assume $p1$ and $p2$ have the same size.

\subsection{Pattern length}
See encoding experiments in Table \ref{tab:psize}.

\begin{table}
\begin{center}
\begin{tabular}{|l|c|}
\hline
size & avg. dist \\ \hline \hline
1 bar (4) & $0.036722$\\
2 bars (8) & $0.044301$\\
2 bars (12) & $0.043583$\\
2 bars (16) & $0.044086$ \\ \hline
3 beats & $0.032688$\\
5 beats & $0.040858$\\
7 beats & $0.045515$\\ \hline
\end{tabular}
\end{center}
\caption{{Encoding score for different pattern sizes. We can select the
size of the patterns based on bars. In this case, we put the actual size
of the pattern in parentheses. If the pattern size is based on beats, the
size is the number of beats. Results are on the uspop dataset, not seen
during training.}}
\label{tab:psize}
\end{table}
% RESULTS
% 200codes, 2 bars, pSize=16, maxener, True, False, not scaled = 
% 200codes, 2 bars, pSize=8, maxener, True, False, not scaled  = 0.042432
% 200codes, 1 bars, pSize=4, maxener, True, False, not scaled  = 
% 200codes, 7 beats, pSize=7, maxener, True, False, not scaled = 
% 200codes, 5 beats, pSize=5, maxener, True, False, not scaled = 
% 200codes, 3 beats, pSize=3, maxener, True, False, not scaled = 
% NEW STUFF - experiment set 7, 100 codes! 
% 2 beats 0 bars       0.025346
% 4 beats 0 bars       0.037181
% 6 beats 0 bars       0.043237
% 8 beats 0 bars       0.046832
% 9 beats 0 bars       0.048579
% 10 beats 0 bars      0.049530
% 11 beats 0 bars      0.050717
% 12 beats 0 bars      0.052020

\subsection{Data Size}
We are interesting to see how the amount of training data changes
the quality of our codebook.
We experiment with different data sizes for a fixed size of codebook and a 
fixed code length. The data points are chose randomly from the whole
data set.
In the experiment a pattern size correspond to $1$ bar resized to $4$
beats, thus a pattern size of $12$x$4$. The codebook size is $100$.
We measure the distortion on our test set. 
See results in Figure \ref{fig:sizes}.

Our learning can overfit, which explain the bad scores of small datasets.
The larger the dataset, the smallest the chance to overfit.
We are looking for a size of dataset that guarantees us a good learning
process, while being unnecessary large.

$500$ samples per code in the codebook seems a good data size, see
Figure \ref{fig:sizes} for size $50$K.

\begin{figure}[htb]
\begin{center}
\includegraphics[width=.99\columnwidth]{data_sizes}
\end{center}
\caption{{Distortion for a codebook of size $100$ encoding one bar
at a time, a bar being represented by $4$ chroma feature vector.
Therefore, a code or a pattern is of size $12$ x $4$ = $48$.
Distortion measure on the test data set. Data size tested range
from $10$ thousands to $2$ millions. Patterns were selected at
random from the dataset of approximately $3,7$ million patterns.}}
\label{fig:sizes}
\end{figure}

\iffalse
\begin{table}
\begin{center}
\begin{tabular}{|l|c|c|c|c|c|c|}
\hline
size & $5$K & $10$K & $50$K & $100$K & $500$K & $2$M  \\ \hline
distortion & $$ & $0.143$ & $0.142$ & $0.142$ & $0.142$ & $$\\ \hline
\end{tabular}
\end{center}
\caption{{Distortion for a codebook of size $100$ encoding one bar
at a time, a bar being represented by $4$ chroma feature vector.
Therefore, a code or a pattern is of size $12$ x $4$ = $48$.
Distortion measure on the test data set. Data size tested range
from $10$ thousands to $2$ millions. Patterns were selected at
random from the dataset of approximately $3,7$ million patterns.}}
\label{tab:datasize}
\end{table}
\fi
% RESULTS:    
% 1bar s4     avg_distOLD    avg_dist
% 0                          0.052247
% 1K          0.150370       0.041941
% 5K          0.145327       0.038449
% 10K         0.143567       0.036892
% 50K         0.142257       0.036722
% 100K        0.142195       0.035973
% 250K        0.141796       0.035910
% 500K        0.142081       0.035904
% 1M          0.142118
% 2M          0.141596
% 1bar s4 first X samples
% 0
% 1K          0.157142       0.049282
% 5K          0.150045       0.046441
% 10K         0.148357       0.045507
% 50K         0.146069       0.044301
% 100K        0.145765       0.044183
% 250K        0.145819       0.043896
% 500K        0.145772       0.043906
% 1M          0.145924       
% 2M          0.146065       
% 0bar s4 (and test set same, we could do 1bar s4 in test set...)
% 0
% 1K          0.155474       0.042596
% 5K          0.149619       0.038971
% 10K         0.148056       0.038114
% 50K         0.146328       0.037181
% 100K        0.146004       0.037055
% 250K        0.146024       0.037014
% 500K        0.145962       0.037004
% 1M          0.145873
% 2M          0.145955
% 2bar s8
% 0                          0.066864179424
% 1K          0.171300       0.049282
% 5K          0.165960       0.046441
% 10K         0.165063       0.045507
% 50K         0.162439       0.044301
% 100K        0.161588       0.044183
% 250K        0.161910       0.043896
% 500K        0.161233       0.043906
% 1M          0.161057
% 2M          0.161068

\subsection{Codebook Size}
We experiment with the size of the codebook and its affect on the
performance of the encoding. Results are shown in Table \ref{tab:cbsize}.
Note that we trained on a number of samples depending on the codebook
size, i.e. we used $500$ samples per code to learn. This is consistent
with our findings from Figure \ref{fig:sizes}.

\begin{table}
\begin{center}
\begin{tabular}{|l|l|c|}
\hline
\# codes & \# samples & avg. dist \\ \hline \hline
1 & $500$ & $0.066079$ \\
10 & $5$K & $0.045690$ \\
50 & $25$K & $0.038456$ \\
100 & $50$K & $0.036725$ \\
200 & $100$K & $0.033923$ \\
500 & $250$K & $0.030973$ \\
1000 & $500$K & \\ \hline
\end{tabular}
\end{center}
\caption{{Experiment with the codebook size. We train on patterns
corresponding to $1$ bar resized to $4$ beats. The number of training
samples is set to $500$ times the number of codes, which is sufficient
according to Figure \ref{fig:sizes}.}}
\label{tab:cbsize}
\end{table}
% RESULTS:      avg_distsOLD    avg_dist
% code 1        0.195939        0.066079
% code 10       0.161812        0.045690
% code 50       0.147472        0.038456
% code 100      0.142194        0.036725
% code 200      0.137548        0.033923
% code 500      0.131211        0.030973
% REDONE WITH FIX NUMBER OF FEATURES - 500K (new correct dist)
% code 1        
% code 10       0.045579
% code 50       0.038302
% code 100      0.035904
% code 200      
% code 500

\subsection{Codebook Size in Function of Pattern Length}
We experiment on the randomness of the \textit{Shingles}, assuming
that there is some correlation between beats and bars.
The intuition is as follow: if the signal was totally random, every
time the signal size double, the codebook size must be squared.
To illustrate this, imagine a binary array of length $1$. We need two
codes to encode perfectly: $(0)$ and $(1)$. If the binary array is now of
length $2$, we need four codes: $(0,0)$, $(0,1)$, $(1,1)$ and $(1,0)$.

In Table \ref{tab:size_pattern}, we test that hypothesis. As expected,
the distortion improves with larger pattern sizes. The details
are difficult to analyze. The error jump from two beats to one bar
can probably be explained by the obvious case of sustained notes.
We also believe that smaller codebooks are more affected by their
initialization. Fortunately, we can conclude that music is not random,
which is always a good sanity check!

Then... 3 beats, 5 beats, 7 beats don't encode as well?

Special: 4 beats versus 1 bar - 4 beats, show the problem of being
not in sync with bars.

See codebook size experiments in Table \ref{tab:cbsize}.

\begin{table}
\begin{center}
\begin{tabular}{|l|c|c|}
\hline
\# codes & \# bars & avg. dist \\ \hline \hline
2 & $1$ & $0.058435$ \\
4 & $2$ & $0.054792$ \\
16 & $4$ & $0.055489$ \\
256 & $8$ & $0.054860$ \\ \hline
2 & $2$ beats* & $0.066416$ \\
4 & $1$ & $0.050712$ \\
16 & $2$ & $0.050172$ \\
256 & $4$ & $0.049894$ \\ \hline
\end{tabular}
\end{center}
\caption{{Codebook size in function of the number of bars to encode.
All experiments done wtih $256 \times 500 = 128$K samples randomly chosen.
* For the $2$ beats experiment, we take the first half of every beats.
It is different from ignoring bars and taking every $2$ beats. It is more
coherent with the other settings ($1-2-4-8$ bars).}}
\label{tab:size_pattern}
\end{table}
% RESULTS:           avg_distsOLD   avg_dist
%2 codes 1 bar       0.180750       0.058435
%4 codes 2 bars      0.183631       0.054792
%16 codes 4 bars     0.184078       0.055489
%256 codes 8 bars    0.183464       0.054860
%---------------
%2 codes 2 beats     0.166840       0.046955
%2 codes 2 beats                    0.066416
% based on bars
%4 codes 1 bar       0.171495       0.050712
%16 codes 2 bars     0.173767       0.050172
%256 codes 4 bars    0.174169       0.049894

% see generate_expo_codes_curves.py to generate the following figure:
\begin{figure}[htb]
\begin{center}
\includegraphics[width=.99\columnwidth]{codesize_patternsize}
\end{center}
\caption{{Possible figure to replace Table \ref{tab:size_pattern}.}}
\label{fig:size_pattern}
\end{figure}

\subsection{Bar offset}
We study how great are patterns based on bars by playing with an offset.
The offset can be $0$ (normal case) or a quarter of bar, half a bar, and
three quarters of a bar.

\begin{table}
\begin{center}
\begin{tabular}{|l|c|c|}
\hline
offset & avg. dist \\ \hline \hline
$0$ & $$ \\
$.25$ & $0.037130$ \\
$.50$ & $0.037241$ \\
$.75$ & $0.037337$ \\ \hline
\end{tabular}
\end{center}
\caption{{Results with different offset. Always $100$ codes.
\label{tab:offset}}}
\end{table}
% RESULTS             avg_results
% 0 offset            0.036722 (to be confirmed by set8exp0)
% .25 ofset           0.037130
% .50 ofset           0.037241
% .75 ofset           0.037337
% RESULTS FOR PARTIALBAR TOO
% 0 off .25 part      0.046304
% .25 off .25 part    0.044950
% .50 off .25 part    0.044395
% .75 off .25 part    0.044473

\subsection{Visualization}
Interpreting a codebook of size larger than $100$ is not obvious.
We can use an algorithm like locally linear embedding \cite{Roweis2000} (LLE) 
or multidimensional scaling \cite{Kruskal1964} to visualize the data in
2 dimensions. Here we use the former\footnote{implementation: 
\url{http://www.astro.washington.edu/users/vanderplas/coding/LLE/}}, the
result can be seen in Figure \ref{fig:lle}.

\begin{figure}[htb]
\begin{center}
\includegraphics[width=.9\columnwidth]{codes_lle}
\end{center}
\caption{{LLE visualization of the codebook.}}
\label{fig:lle}
\end{figure}

Another way is simply to present the codes in order of cluster size.
We encode the training or testing set as explained in Subsection 
\ref{ssec:setting} and we count how many patterns were closer to a particular
code than any other code. In Figure \ref{fig:firstcodes}, we show the $25$
most used codes out of a codebook of size $200$.
Code from Figure \ref{fig:code} is the 37th most used code in that
codebook. In Figure \ref{fig:closep}, we show a sample of patterns that
fall into its cluster.

\begin{figure}[htb]
\begin{center}
\includegraphics[width=.9\columnwidth]{codes1}
\end{center}
\caption{{$25$ most used codes on the training set.}}
\label{fig:firstcodes}
\end{figure}

The most common pattern is a sustained note, as we can see in Figure
\ref{fig:firstcodes}. Another common pattern is a sustained note
with an harmonic, as seen in the LLE visualization (Figure \ref{fig:lle}).

\begin{figure}[htb]
\begin{center}
\includegraphics[width=.9\columnwidth]{close_patterns1}
\end{center}
\caption{{Sample of the close patterns to code in Figure \ref{fig:code}.}}
\label{fig:closep}
\end{figure}

\subsection{Specific Examples}

\section{Conclusion}
We presented a practical method to perform large-scale clustering of
harmonic patterns. We assessed the basic properties of the method through
experiments on a large collection of music. We suggested many ways
to interpret the data through visualization (LLE) and further encoding
(NMF). We discussed the possibility to move to even larger scales
and we provided our source code\footnote{Code not yet released to preserve
submission's anonymity}.

As for enhancements, we specifically did not compress our features using
gaussian mixtures or other generative model. That being said, these methods
could be use to develop better distance measures between \textit{Shingles}.
The use of the euclidean distance is arbitrary. Summarizing patches
with gaussians, and then comparing the distance between those gaussians,
could reduce the influence of the noise in the distance measure.

Moving on to larger scales, we could imagine merging codes together and
splitting some on two if the clusters become too small or too large.We could
also cluster the codebook itself, in a similar fashion as hierarchical
gaussian mixtures \cite{Vasconcelos2001}.

\small
%\section{Acknowledgements}
%Thierry is NSERC graduate fellow, or some title like that.
% Graham for the NMF stuff, unless is an author.

%\begin{thebibliography}{citations}
%\bibitem{Someone:04} 
%X. Someone and Y. Someone:
%{\it Title of the Book},
%Editorial Acme, Utrecht, 2004.
%\end{thebibliography}

\bibliography{tbm_bib}

\end{document}

### 2010-03-20 Paper/Paper.tex
new draft improved
improved, almost all figures in

@@
 
-\section{Experiments}\label{sec:experiments}
+\section{Experiments - Fundamentals}\label{sec:experiments}
+Note that we will present two additional experiments in 
+Section \ref{sec:exps2}.
 
@@
 better, more codes is better, larger patterns are more difficult.
-
-First, a larger dataset gives a better encoding codebook. See Figure
-\ref{fig unknown}.
-
-Secondly, a larger codebook better encodes new song. See Figure \ref{ejdue}.
-
-Finally, larger patterns are more difficult to encode, thus require
-larger codebooks. See Figure \ref{mouahahah}.
+\begin{itemize}
+\item First, a larger dataset gives a better encoding codebook. See Figure
+\ref{fig:data_sizes}.
+\item Secondly, a larger codebook better encodes new song. See 
+Table \ref{tab:cbsize}.
+\item Finally, larger patterns are more difficult to encode, thus require
+larger codebooks. See Figure \ref{fig:perbeat}.
+\end{itemize}
 
@@
 surprising for a clustering algorithm. We move on to more interestng facts.
+
+\begin{figure}[htb]
+\begin{center}
+\includegraphics[width=.99\columnwidth]{data_sizes}
+\end{center}
+\caption{\small{Distortion for a codebook of size $100$ encoding one bar
+at a time, a bar being represented by $4$ chroma feature vector.
+Therefore, a code or a pattern is of size $12$ x $4$ = $48$.
+Distortion measure on the test data set. Data size tested range
+from $10$ thousands to $2$ millions. Patterns were selected at
+random from the dataset of approximately $3,7$ million patterns.}}
+\label{fig:data_sizes}
+\end{figure}
+
+% NEW RESULTS COMING SOON WITH FIXED SIZED DATASET
+\begin{table}
+\begin{center}
+\begin{tabular}{|l|l|c|}
+\hline
+\# codes & \# samples & avg. dist \\ \hline \hline
+1 & $500$ & $0.066079$ \\
+10 & $5$K & $0.045690$ \\
+50 & $25$K & $0.038456$ \\
+100 & $50$K & $0.036725$ \\
+200 & $100$K & $0.033923$ \\
+500 & $250$K & $0.030973$ \\
+1000 & $500$K & \\ \hline
+\end{tabular}
+\end{center}
+\caption{\small{Experiment with the codebook size. We train on patterns
+corresponding to $1$ bar resized to $4$ beats. The number of training
+samples is set to $500$ times the number of codes, which is sufficient
+according to Figure \ref{fig:sizes}. NEW RESULTS COMING SOON WITH FIXED 
+SIZED DATASET }}
+\label{tab:cbsize}
+\end{table}
+
+\begin{figure}[htb]
+\begin{center}
+\includegraphics[width=.9\columnwidth]{encoding_per_beat}
+\end{center}
+\caption{\small{Encoding larger and larger patterns with a fixed size
+codebook.}}
+\label{fig:perbeat}
+\end{figure}
 
@@
 given by the EchoNest. Secondly, each pattern is encoded by its closest
-code in the codebook. See Figure \ref{youppie}.
+code in the codebook. See Figure \ref{fig:encodesong}.
+
+\begin{figure}[htb]
+\begin{center}
+\includegraphics[width=.9\columnwidth]{song_encoded}
+\end{center}
+\caption{\small{Original song and its encoding. Must get a better one for
+the paper!!!!!!}}
+\label{fig:encodesong}
+\end{figure}
 
 \subsection{Codebook}
-We can present the most typical patterns (Figure \ref{1}). 
-Then, LLE? (Figure \ref{2}). 
+We can present the most typical patterns (Figure \ref{fig:codes1}). 
+Then, LLE? (Figure \ref{fig:lle}). 
 Then, the graph on variance. (Figure \ref{fig:code_var}).
+
+\begin{figure}[htb]
+\begin{center}
+\includegraphics[width=.9\columnwidth]{codes1}
+\end{center}
+\caption{\small{Most used codes in the codebook.}}
+\label{fig:codes1}
+\end{figure}
 
@@
 \end{center}
-\caption{Locally linear embedding (LLE) visualization of the codebook.}
+\caption{\small{Locally linear embedding (LLE) visualization of the codebook.}}
 \label{fig:lle}
@@
 \end{center}
-\caption{Average variance in time of a codebook.}
+\caption{\small{Average variance in time of a codebook.}}
 \label{fig:code_var}
@@
 than any other code. The goal is to feel how much variance is present in
-a cluster.
+a cluster. See Figure \ref{fig:cluster}.
 
-\section{Further Experiments}
+\begin{figure}[htb]
+\begin{center}
+\includegraphics[width=.9\columnwidth]{close_patterns1}
+\end{center}
+\caption{\small{Cluster around some pattern... TO BE REDONE}}
+\label{fig:cluster}
+\end{figure}
+
+\section{Further Experiments}\label{sec:exps2}
 
@@
 beginning.
+
+See Table \ref{tab:offset}.
+
+\begin{table}
+\begin{center}
+\begin{tabular}{c|c}
+%\hline
+offset & \% of times chosen \\ \hline
+0 & $\mathbf{62.6}$\\
+1 & $16.5$\\
+2 & $09.4$\\
+3 & $11.5$\\
+\end{tabular}
+\end{center}
+\caption{\small{
+Best offset according to our codebook. The real one is $0$.
+}}
+\label{tab:offset}
+\end{table}
 
@@
 \end{center}
-\caption{Confusion matrix for artist recognition task.}
+\caption{\small{Confusion matrix for artist recognition task.}}
 \label{fig:conf_mat}
+\end{figure}
+
+Looking at the most typical patterns per artist can be insightful.
+To find these patters, we compare its use by one artist and divide
+by its use by all artists. In Figure \ref{fig:metallica} we see 
+Metallica's pattern and in Figure \ref{fig:amosvega}, Tori Amos and
+Suzanne Vega's pattern. These three artists were easily identified.
+
+\begin{figure}[htb]
+\begin{center}
+\includegraphics[width=.4\columnwidth]{metallica_pattern}
+\end{center}
+\caption{\small{Most typical pattern for Metallica.}}
+\label{fig:metallica}
+\end{figure}
+
+\begin{figure}[htb]
+\begin{center}
+\includegraphics[width=.4\columnwidth]{toriamos_suzannevega_pattern}
+\end{center}
+\caption{\small{Most typical pattern for Tori Amos and Suzanne Vega.}}
+\label{fig:amosvega}
 \end{figure}

### 2010-03-20 Paper/Paper.tex
stuff added back from draft

@@
 
+\subsection{Definitions}
+A pattern, or a patch, is a \textit{Shingle} taken from an actual song.
+A code is similar to a pattern, but is the result of the clustering
+algorithm and most likely does not come from an actual song.
+Distortion is our measure of distance between a pattern and a code
+(or between two patterns, see equation \ref{eq:dist} in Section 
+\ref{sec:experiments}).
+
 \section{Previous Work}\label{sec:prevwork}
-Casey and Slaney \cite{Casey2006,Casey2007,Casey2008}
+Our work can be seen as an extension of the \textit{Shingles} idea
+described by Casey and Slaney \cite{Casey2006,Casey2007,Casey2008}.
+The two main differences are the clustering framework and the use of
+beats and bars to select the position of the \textit{Shingles}.
+
+Casey and Slaney \cite{Casey2006} use LSH \cite{Datar2004} to perform
+fast approximate nearest neighbor search. Their implementation is E$^2$LSH
+\cite{E2LSH}. This package can efficiently handle about $50$K or $100$K
+vectors of size $200$. Even with a better implementation, the LSH
+model still has some scaling problem. A large portion of the data
+must be known to decide how many projections to use, and what guarantee
+of finding the nearest neighbor it can provides. Also, we could fix the
+number of projections and consider as clusters all the elements that
+fall into the same bin. However, this does not give us a centroid, a
+typical pattern representing that cluster.
 
@@
 \subsection{EchoNest API}
+We use the Echo Nest analyze API \cite{EchoNest} throughout this work.
+The API gives us a chroma vector (length $12$) for every music event, 
+or ``segment'', for any song we upload to their platform. 
+We also get an estimate of where the beats and the bars are. Beats span over
+multiple segments, and bars span over multiple beats. 
+We transform the chroma vectors per segment into per beat vectors using a 
+simple average. We can then stack beat vectors for a number of bars 
+(usually 1 or 2). 
+We get a fixed-size patch by resampling the number of chroma vectors. Typically,
+there are $4$ beats per bar. Thus an appropriate pattern size for one bar
+is $12$x$4$ or $12$x$8$.
+
+Note that we do not claim that any of these informations (segments, beats, bars)
+are perfectly accurate (e.g. in \cite{Barrington2009a} authors obtain a better
+song segmentation than Echonest). Practise showed us that they are reasonable, 
+and the size of the data set should make up for the imperfections or noise.
+We also believe that patches sizes based on a number of beats or bars are more
+meaningful than an arbitrary time length. More on this in the experiments
+(Section \ref{sec:experiments}).
+
+\begin{figure}[htb]
+\begin{center}
+\includegraphics[width=.8\columnwidth]{code}
+\end{center}
+\caption{\small{A typical code from a codebook of size $200$. 
+Patterns represented
+$2$ bars and the pattern length was set to $16$.
+BLACK AND WHITE VERSION COMING.
+}}
+\label{fig:code}
+\end{figure}
 
@@
 \RETURN $\{C_k\}$
-\caption{{Pseudocode of Online Vector Quantization. Note that we can replace
-the number of iteration by a threshold on the distortion over some test set.}
+\caption{\small{Pseudocode of Online Vector Quantization. Note that we can 
+replace the number of iteration by a threshold on the distortion over some 
+test set.}
 \label{algo:vq}}
@@
 \section{Conclusion and Future Work}
+We presented a practical method to perform large-scale clustering of
+harmonic patterns. We assessed the basic properties of the method through
+experiments on a large collection of music. We suggested many ways
+to interpret the data through visualization (LLE) and further encoding
+(NMF). We discussed the possibility to move to even larger scales
+and we provided our source code\footnote{Code not yet released to preserve
+submission's anonymity}.
+
+As for enhancements, we specifically did not compress our features using
+gaussian mixtures or other generative model. That being said, these methods
+could be use to develop better distance measures between \textit{Shingles}.
+The use of the euclidean distance is arbitrary. Summarizing patches
+with gaussians, and then comparing the distance between those gaussians,
+could reduce the influence of the noise in the distance measure.
+
+Moving on to larger scales, we could imagine merging codes together and
+splitting some on two if the clusters become too small or too large.We could
+also cluster the codebook itself, in a similar fashion as hierarchical
+gaussian mixtures \cite{Vasconcelos2001}.
 

### 2010-03-20 Paper/Paper.tex
stuff added back from draft
improved

@@
 
+\subsection{Cowbell Dataset}
+We have $43,300$ songs, giving  $3,720,091$ non zero bars, a bar usually 
+contains $4$ beats. WHERE DOES THIS DATA COMES FROM? GIFT FROM EN?
+
 \subsection{USPOP}
+We had acces to a low quality (32kbps) version of the songs in the uspop 2002 
+data set \cite{uspop2002}.
+This well-known data show a great diversity of pop songs, but is not particular
+otherwise. We also tried to use Tzanetakis dataset \cite{Tzanetakis2002a}, but 
+the $30$ seconds segments seemed hard to analyze by the Echo Nest API, most did
+not contain any bar.
+
+uspop2002 serves as testing set to measure how well a codebook learned on
+the Cowbell data set can represent new songs. We get Echo Nest features
+for $8651$ songs from that dataset.
 
 \subsection{Artist Set}
+A set of songs from $20$ artists.
 
@@
 imagine that some artist use more often some patterns than others.
+Best accuracy reported is around $54.5\%$. Our result is $23.4\%$. Random
+is around $5\%$.
+
 See Figure \ref{fig:conf_mat}.
@@
 \end{center}
-\caption{\small{Most typical pattern for Metallica.}}
+\caption{\small{Most typical pattern for Metallica.
+COULD BE MIXED WITH TORI AMOS FIGURE.
+}}
 \label{fig:metallica}

### 2010-03-20 PythonSrc/for_ron.py
set 11 redone because wrong names...

@@ -216,6 +216,21 @@ experiment_args = [
      (os.path.join(outputDir, 'set14exp9'),4,1,130,50000,False,0,0.75,1),
      (os.path.join(outputDir, 'set14exp10'),4,1,145,50000,False,0,0.75,1),
      (os.path.join(outputDir, 'set14exp11'),4,1,160,50000,False,0,0.75,1)],
+    # EXPERIMENT SET 15: redoing set 11 with good names...!
+    [(os.path.join(outputDir, 'set15exp0'),1,1,2,128000,False,0,0.00,.25),
+     (os.path.join(outputDir, 'set15exp1'),1,1,2,128000,False,0,0.25,.25),
+     (os.path.join(outputDir, 'set15exp2'),1,1,2,128000,False,0,0.50,.25),
+     (os.path.join(outputDir, 'set15exp3'),1,1,2,128000,False,0,0.75,.25),
+     (os.path.join(outputDir, 'set15exp4'),2,1,4,128000,False,0,0.00,.50),
+     (os.path.join(outputDir, 'set15exp5'),2,1,4,128000,False,0,0.25,.50),
+     (os.path.join(outputDir, 'set15exp6'),2,1,4,128000,False,0,0.50,.50),
+     (os.path.join(outputDir, 'set15exp7'),2,1,4,128000,False,0,0.75,.50),
+     (os.path.join(outputDir, 'set15exp8'),4,1,16,128000,False,0),
+     (os.path.join(outputDir, 'set15exp9'),8,2,256,128000,False,0),
+     (os.path.join(outputDir, 'set15exp10 '),2,1,2,128000,False,0,0.00,.50),
+     (os.path.join(outputDir, 'set15exp11'),2,1,2,128000,False,0,0.25,.50),
+     (os.path.join(outputDir, 'set15exp12'),2,1,2,128000,False,0,0.50,.50),
+     (os.path.join(outputDir, 'set15exp13'),2,1,2,128000,False,0,0.75,.50)],
     ]
 
         

### 2010-03-20 PythonSrc/for_ron.py
set 11 redone because wrong names...

@@ -230,7 +230,12 @@ experiment_args = [
      (os.path.join(outputDir, 'set15exp10 '),2,1,2,128000,False,0,0.00,.50),
      (os.path.join(outputDir, 'set15exp11'),2,1,2,128000,False,0,0.25,.50),
      (os.path.join(outputDir, 'set15exp12'),2,1,2,128000,False,0,0.50,.50),
-     (os.path.join(outputDir, 'set15exp13'),2,1,2,128000,False,0,0.75,.50)],
+     (os.path.join(outputDir, 'set15exp13'),2,1,2,128000,False,0,0.75,.50),
+     (os.path.join(outputDir, 'set15exp14'),4,1,4,128000,False,0),
+     (os.path.join(outputDir, 'set15exp15'),8,2,16,128000,False,0),
+     (os.path.join(outputDir, 'set15exp16'),16,4,256,128000,False,0),
+     (os.path.join(outputDir, 'set15exp17'),4,1,16,128000,False,0),
+     (os.path.join(outputDir, 'set15exp18'),8,2,256,128000,False,0)],
     ]
 
         

### 2010-03-20 PythonSrc/for_ron.py
set 11 redone because wrong names...

@@ -235,7 +235,11 @@ experiment_args = [
      (os.path.join(outputDir, 'set15exp15'),8,2,16,128000,False,0),
      (os.path.join(outputDir, 'set15exp16'),16,4,256,128000,False,0),
      (os.path.join(outputDir, 'set15exp17'),4,1,16,128000,False,0),
-     (os.path.join(outputDir, 'set15exp18'),8,2,256,128000,False,0)],
+     (os.path.join(outputDir, 'set15exp18'),8,2,256,128000,False,0),
+     (os.path.join(outputDir, 'set15exp19'),4,1,2,128000,False,0),
+     (os.path.join(outputDir, 'set15exp20'),8,2,4,128000,False,0),
+     (os.path.join(outputDir, 'set15exp21'),16,4,16,128000,False,0),
+     (os.path.join(outputDir, 'set15exp22'),32,8,256,128000,False,0)],
     ]
 
         

### 2010-03-20 Paper/Paper.tex
improved

@@
 
-\section{Algorithm}
-Wth the scaling factor as our main concern, we present an online
-vector quantization algorithm to cluster the harmonic patterns.
+\section{Algorithm}\label{sec:algo}
+In this section we present the creation of the harmonic patterns from
+the Echo Nest data. Then, we explain the online vector quantization algorithm
+used to learn a codebook of those patterns.
+
+\subsection{Pattern Creation}
+Here we should explain key invariance, resizing, etc, and not in the
+experiment section.
+
+We try to normalize the patterns in respect to the key. We roll the matrix
+so the row with the more energy is always the lowest one. This can easily be
+seen in Figure \ref{fig:code}. We considered the same technique for rolling
+along the other axis and be invariant regarding the downbeat. However, it
+seems less robust to noise. Note that we roll each pattern independentely.
+Also, this implies that the pitch envelope is important rather than the actual 
+pitches.
+
+\subsection{Codebook Learning}
+We use an online version of the vector quantization algorithm 
+\cite{Gersho1991}. This can also be seen as an online K-means.
+The idea is as follow: take a sample from your data, find the closest
+code in the codebook, bring that code closer to the sample by some amount.
+The details are explained in Algorithm \ref{algo:vq}. Note that this
+algorithm, though not optimal, is $O(n)$ with $n$ number of songs seen.
+Thus it scales linearly to any size of data set. The setting of the learning
+rate might become trickier, but this is not fundamentally a problem.
+
+We initialize the codebook by choosing $K$ random points from our dataset.
 
@@
 Section \ref{sec:exps2}.
+
+\subsection{Setting}\label{ssec:setting}
+We take one or two bars, normalize the patches to size 4, 8, or 16.
+We roll the patch to be invariant to the key (on the patch level, not on
+the song level). We learn a codebook of size $K$ over the cowbell dataset 
+using the VQ algorithm (Algorithm \ref{algo:vq}). A typical learning rate 
+is $0.01$ for $200$ iterations over the whole dataset.
+
+We use the codebook to encode our testing set (see Section \ref{sec:data}).
+Each pattern is encoded with only one code. We can measure the average
+distance between a pattern and its encoding. We can also measure the use
+of the different codes, i.e. the size of the cluster around a code.
+
+As a distance measure (or distortion measure) between patterns, we use
+the average squared euclidean distance. In details, if a pattern $p1$
+is composed of elements $p1_{i,j}$, and similarly for a pattern $p2$,
+the distance between $p1$ and $p2$ is:
+\begin{eqnarray}
+  dist(p1,p2) = \sum_{i,j} (p1_{i,j} - p2_{i,j})^2 / size(p1)  \label{eq:dist}
+\end{eqnarray}
+We assume $p1$ and $p2$ have the same size.
 

### 2010-03-20 Paper/Paper.tex
improved

@@
 
-\section{Visualization}
+\section{Visualization} \label{sec:visu}
 The goal here is to give an intuition to the reader of what kind of
@@
 \section{Further Experiments}\label{sec:exps2}
+We have shown the basic properties of our clustering algorithm in
+Section \ref{sec:experiments}. In Section \ref{sec:visu}, we showed
+how to visualize the codebook and the clusters. We know move on to
+show other applications that can take advantage of our clustering.
+In the next two subsections, we talk about song segmentation and
+artist recognition. Our algorithm was not devised for any of these tasks,
+but some information contained in the centroids can be naturally used.
 
@@
 encodes something related to bars, for instance a strong beat at the
-beginning.
+beginning. Results are shown \ref{tab:offset}. We find the right segment
+with an accuracy of $62\%$ where random is $25\%$. Details of the
+experiment follow.
 
-See Table \ref{tab:offset}.
+We train a codebook of size $100$ on bars resized to $4$ beats. Then,
+we take the longest sequence of bars of $4$ beats in the test set.
+As we mentioned before, $4$ beats is the most common (by far!) bar size
+in terms of beats according to the EchoNest API. We then encode this
+sequence using an offset of $0$, $1$, $2$ or $3$ beats. The best offset
+is the one that gives the encoding with the lowest distortion. The
+good answer is an offset of $0$.
 
@@
 Best accuracy reported is around $54.5\%$. Our result is $23.4\%$. Random
-is around $5\%$.
+is around $5\%$. The confusion matrix can be seen in Figure \ref{fig:conf_mat}.
+Details of the experiment follow.
 
-See Figure \ref{fig:conf_mat}.
+All songs in the artist dataset are encoded as an histogram of the codes
+used in that song. The frequency values are normalized by the number
+of patterns in the song. We test each song in a leave-one-out setting.
+Leaving aside the test song, we represent the $20$ artists by the average of
+the song histograms of that artist. The test song is compared to
+the histogram of each artist using euclidean distance.
 

### 2010-03-21 Paper/Paper.tex
improved
improved by tbm

@@
 meaningful than an arbitrary time length. More on this in the experiments
-(Section \ref{sec:experiments}).
+(Section \ref{sec:experiments} TO BE CHECKED).
 
@@
 \section{Experiments - Fundamentals}\label{sec:experiments}
+In this section we assert the basic properties of our encoding scheme.
+The first subsection gives the details to reproduce the experiments,
+such as the distance measure and learning rate. In second subsection, we
+measure properties such as ``how the number of codes affect the distortion''
+and ``how many learning samples should we use''. Though important numbers
+to report, the busy reader could skip to subsection \ref{ssec:random}
+and the following sections.
 Note that we will present two additional experiments in 
@@
 corresponding to $1$ bar resized to $4$ beats. The number of training
-samples is set to $500$ times the number of codes, which is sufficient
-according to Figure \ref{fig:sizes}. NEW RESULTS COMING SOON WITH FIXED 
-SIZED DATASET }}
+samples is set to $500$ times the number of codes. 
+NEW RESULTS COMING SOON WITH FIXED SIZED DATASET }}
 \label{tab:cbsize}
@@
 
-\subsection{Music Randomness}
+\subsection{Music Randomness}\label{ssec:random}
 It will not surprise any reader, music is not random. This fact should
@@
 pattern size doubles will rather give improved results than constant ones.
-See Figure \ref{it's comng...}.
+See Figure \ref{fig:size_pattern}.
+
+% see generate_expo_codes_curves.py to generate the following figure:
+\begin{figure}[htb]
+\begin{center}
+\includegraphics[width=.99\columnwidth]{codesize_patternsize}
+\end{center}
+\caption{\small{NEW RESULTS COMING IN}}
+\label{fig:size_pattern}
+\end{figure}
 

### 2010-03-21 Paper/generate_barsize_barplot.py
to generate barsize bar plot

@@ -1,0 +1,56 @@
+
+import scipy
+import scipy.io
+import numpy as np
+import sys
+
+
+
+
+def generate():
+    """ generate the graph """
+    import pylab as P
+
+
+    # load data
+    mat = scipy.io.loadmat('/home/thierry/Desktop/tmpPaper/number_beats_per_bar.mat',struct_as_record=True)
+
+    numBars = mat['numBars']
+    numBars = numBars.flatten()
+
+    # rearrange data, from 1 beat to >9
+    res = numBars[1:10]
+    
+    # bar plot
+    P.bar(np.array(range(9))+.6,res)
+
+    # text
+    P.xlabel('number of beats per bar')
+    P.ylabel('frequency in training data')
+    P.title('number of beats per bar in the training data based on EchoNest API')
+
+    # x ticks
+    labels = [str(x) for x in (range(1,9))]
+    labels.append('>9')
+    P.xticks(P.arange(1,len(labels)+1),list(labels))
+
+
+    # done, show
+    P.show()
+
+
+def die_with_usage():
+    """ HELP MENU """
+    print 'to generate the bar plot that shows size per bar in paper:'
+    print '    python generate_barsize_barplot -go'
+    sys.exit(0)
+    
+
+
+
+if __name__ == '__main__':
+
+    if len(sys.argv) < 2:
+        die_with_usage()
+
+    generate()

### 2010-03-21 Paper/Paper.tex
improved by tbm, bar size bar plot added

@@
 We get a fixed-size patch by resampling the number of chroma vectors. Typically,
-there are $4$ beats per bar. Thus an appropriate pattern size for one bar
+there are $4$ beats per bar (Figure \ref{fig:barsize}. 
+Thus an appropriate pattern size for one bar
 is $12$x$4$ or $12$x$8$.
+
+\begin{figure}[htb]
+\begin{center}
+\includegraphics[width=.9\columnwidth]{bar_size_freq}
+\end{center}
+\caption{\small{
+Frequency of the different bar size in the training data.
+}}
+\label{fig:barsize}
+\end{figure}
 
@@
 typical patterns do we get. We also want to have a feel for what the
-clusters look like, how many boring sustaned notes are encoded in the
+clusters look like, how many boring sustained notes are encoded in the
 codebook, can the codebook describes the complexity of the music space, etc.
@@
 Below we present some figures to help the reader with those questions.
-Please keep in mind the limited space.
+Please bear in mind the limited space.
 

### 2010-03-21 Paper/Paper.tex
improved by tbm, patterns for metallica mixed with vega pattern

@@
 \usepackage{algorithm,algorithmic}
+\usepackage{subfloat,subfig}
 
@@
 We get a fixed-size patch by resampling the number of chroma vectors. Typically,
-there are $4$ beats per bar (Figure \ref{fig:barsize}. 
+there are $4$ beats per bar (Figure \ref{fig:barsize}). 
 Thus an appropriate pattern size for one bar
@@
 from $10$ thousands to $2$ millions. Patterns were selected at
-random from the dataset of approximately $3,7$ million patterns.}}
+random from the dataset of approximately $3,7$ million patterns.
+OLD GRAPH WITH OLD NUMBERS
+}}
 \label{fig:data_sizes}
@@
 To find these patters, we compare its use by one artist and divide
-by its use by all artists. In Figure \ref{fig:metallica} we see 
-Metallica's pattern and in Figure \ref{fig:amosvega}, Tori Amos and
+by its use by all artists. In Figure \ref{fig:typicalpat} we see 
+Metallica's pattern, then Tori Amos and
 Suzanne Vega's pattern. These three artists were easily identified.
 
+% ignoring the two following figures
+\iffalse
 \begin{figure}[htb]
@@
 \label{fig:amosvega}
+\end{figure}
+\fi
+
+\begin{figure}[htb]
+  \centering
+  \subfloat[][\small{Metallica}]{\label{fig:metallica2}\includegraphics[width=0.4\columnwidth]{metallica_pattern}}                
+  \subfloat[][\small{Tori Amos / Suzanne Vega}]{\label{fig:amosvega2}\includegraphics[width=0.4\columnwidth]{toriamos_suzannevega_pattern}}
+  \caption{\small{Typical patterns for different artists.}}
+  \label{fig:typicalpat}
 \end{figure}

### 2010-03-21 Paper/Paper.tex
improved by tbm, patterns for metallica mixed with vega pattern
improved

@@
 \usepackage{algorithm,algorithmic}
+
+% added by tbm
 \usepackage{subfloat,subfig}
+\setlength{\abovecaptionskip}{-3pt}
+\setlength{\belowcaptionskip}{10pt}
 
@@
 
-Large-scale \cite{Bertin-Mahieux2008}
+Large-scale \cite{Bertin-Mahieux2008}. See previous work for \textit{Shingles}
+papers.
 
@@
 The first subsection gives the details to reproduce the experiments,
-such as the distance measure and learning rate. In second subsection, we
+such as the distance measure and learning rate. In the second subsection, we
 measure properties such as ``how the number of codes affect the distortion''
@@
 \end{center}
-\caption{\small{Average variance in time of a codebook.}}
+\caption{\small{Average variance in time of a codebook. The vertical
+axis cuts at the $63$rd pattern, a rough estimate of the number of 
+sustained note patterns.
+}}
 \label{fig:code_var}
@@
 imagine that some artist use more often some patterns than others.
-Best accuracy reported is around $54.5\%$. Our result is $23.4\%$. Random
+Best accuracy reported\footnote{few details given to preserve anonymity} 
+is around $54.5\%$. Our result is $23.4\%$. Random
 is around $5\%$. The confusion matrix can be seen in Figure \ref{fig:conf_mat}.

### 2010-03-21 Paper/Paper.tex
improved

@@
 
-Large-scale \cite{Bertin-Mahieux2008}. See previous work for \textit{Shingles}
-papers.
+Working with large collections is a necessity.
+
+Computationally expensive, most of the techniques developed do not scale
+to hundred of thousands songs. Practical question, how to describe a music
+space made of millions of songs?
+
+Some work have tackled that problem. For instance, Bertin-Mahieux et al.
+\cite{Bertin-Mahieux2008} automatically tag a collection larger than $100$K
+songs. Casey and Slaney \cite{Casey2006} introduced their \textit{Shingles}
+representation and perform $k$-nn over thousands of examples.
 
@@
 We use the Echo Nest analyze API \cite{EchoNest} throughout this work.
-The API gives us a chroma vector (length $12$) for every music event, 
-or ``segment'', for any song we upload to their platform. 
+The API gives us a chroma vector (length $12$) for every music event (called 
+``segment'') for any song we upload to their platform. 
 We also get an estimate of where the beats and the bars are. Beats span over
@@
 \subsection{Artist Set}
-A set of songs from $20$ artists.
+This set ecompass $1402$ songs $20$ artists, mostly rock and pop of different 
+subgenres. Some results were available regarding the artist recognition
+task. Few details given to preserve anonymity.
 
@@
 We can present the most typical patterns (Figure \ref{fig:codes1}). 
-Then, LLE? (Figure \ref{fig:lle}). 
+Then, LLE\footnote{implementation: 
+\url{http://www.astro.washington.edu/users/vanderplas/coding/LLE/}} 
+\cite{Roweis2000}? (Figure \ref{fig:lle}). 
 Then, the graph on variance. (Figure \ref{fig:code_var}).
@@
 %Thierry is NSERC graduate fellow, or some title like that.
-% Graham for the NMF stuff, unless is an author.
+%Ron has a grant too.
+% Graham for the NMF stuff and numerous discussions.
 

### 2010-03-21 Paper/Paper.tex
improved

@@
 representation and perform $k$-nn over thousands of examples.
+
+We approach the clustering problem keeping in mind the scaling
+requirement. We use a representation similar to the \textit{Shingles}.
+We train on $43$K songs and test on more than $7$K ones.
 
@@
 \subsection{USPOP}
-We had acces to a low quality (32kbps) version of the songs in the uspop 2002 
+We had access to a low quality (32kbps) version of the songs in the uspop 2002 
 data set \cite{uspop2002}.
@@
 \subsection{Artist Set}
-This set ecompass $1402$ songs $20$ artists, mostly rock and pop of different 
+This set encompass $1402$ songs $20$ artists, mostly rock and pop of different 
 subgenres. Some results were available regarding the artist recognition
@@
 \subsection{Pattern Creation}
-Here we should explain key invariance, resizing, etc, and not in the
-experiment section.
+Patterns are defined as a number of beats or bars. In the case of beats,
+we simply cut the set of chromas per beat every $N$ beats. In the case
+of bars, we gather all beats corresponding to the $N$ bars requested.
+The number of chroma vectors we get can vary. We resize the features in the
+time axis to a fixed length. Resizing can be argued over, but it should
+preserve the main bar characteristics (such as a strong beat at the beginning
+or a note change in the middle).
 
@@
 along the other axis and be invariant regarding the downbeat. However, it
-seems less robust to noise. Note that we roll each pattern independentely.
+seems less robust to noise. Note that we roll each pattern independently.
 Also, this implies that the pitch envelope is important rather than the actual 
@@
 These properties are important to verify and measure, but they are not
-surprising for a clustering algorithm. We move on to more interestng facts.
+surprising for a clustering algorithm. We move on to more interesting facts.
 
@@
 intended to perform well at artist recognition. An obvious fact is that
-we lose all timbral information which would be usefull. However, we can
+we lose all timbral information which would be useful. However, we can
 imagine that some artist use more often some patterns than others.

### 2010-03-21 Paper/Paper.tex
improved

@@
 of patterns in the song. We test each song in a leave-one-out setting.
-Leaving aside the test song, we represent the $20$ artists by the average of
+Leaving aside the test song, we represent each of the $20$ artists by the 
+average of
 the song histograms of that artist. The test song is compared to
@@
 Looking at the most typical patterns per artist can be insightful.
-To find these patters, we compare its use by one artist and divide
+To find these patterns, we compare its use by one artist and divide
 by its use by all artists. In Figure \ref{fig:typicalpat} we see 
@@
 experiments on a large collection of music. We suggested many ways
-to interpret the data through visualization (LLE) and further encoding
-(NMF). We discussed the possibility to move to even larger scales
+to interpret the data (such as LLE) and assert the merit of the codebook
+through further experiments.
+We discussed the possibility to move to even larger scales
 and we provided our source code\footnote{Code not yet released to preserve

### 2010-03-21 Paper/Paper.tex
improved

@@
 \subsection{Cluster}
-Here we present real patterns that are closer to the code presented in ...
-than any other code. The goal is to feel how much variance is present in
-a cluster. See Figure \ref{fig:cluster}.
+Once we understand the kind of codebook we create, we naturally want
+to visualize the clusters, i.e. the pattersn from actual song that
+surround a given centroid. Unsurprisingly, real patterns tend to be
+noisier than codes. However, centroids seem to capture important aspects
+of the patterns. We could compare the centroids with the result of
+a lowpass filtering.
+
+See Figure \ref{fig:cluster}.
 

### 2010-03-21 Paper/Paper.tex
new black and white figures
improved

@@
 \begin{center}
-\includegraphics[width=.8\columnwidth]{code}
+\includegraphics[width=.6\columnwidth]{code}
 \end{center}
@@
 \subsection{Cowbell Dataset}
-We have $43,300$ songs, giving  $3,720,091$ non zero bars, a bar usually 
+We have $43,300$ songs, giving  
+%$\bf{3,720,091}$ 
+$\mathbf{3.7}$\textbf{M} 
+non zero bars, a bar usually 
 contains $4$ beats. WHERE DOES THIS DATA COMES FROM? GIFT FROM EN?
@@
 
-See Figure \ref{fig:cluster}.
+See Figure \ref{fig:cluster} and \ref{fig:cluster_diff}.
 
@@
 \label{fig:cluster}
+\end{figure}
+
+\begin{figure}[htb]
+\begin{center}
+\includegraphics[width=.9\columnwidth]{close_patterns_diff}
+\end{center}
+\caption{\small{Cluster around the pattern and the difference of both.}}
+\label{fig:cluster_diff}
 \end{figure}

### 2010-03-21 Paper/Paper.tex
improved

@@
 This well-known data show a great diversity of pop songs, but is not particular
-otherwise. We also tried to use Tzanetakis dataset \cite{Tzanetakis2002a}, but 
+otherwise\footnote{We 
+also tried to use Tzanetakis dataset \cite{Tzanetakis2002a}, but 
 the $30$ seconds segments seemed hard to analyze by the Echo Nest API, most did
-not contain any bar.
-
+not contain any bar.}.
 uspop2002 serves as testing set to measure how well a codebook learned on
 the Cowbell data set can represent new songs. We get Echo Nest features
-for $8651$ songs from that dataset.
+for $8651$ songs of that dataset.
 

### 2010-03-21 Paper/Paper.tex
improved

@@
 Patterns represented
-$2$ bars and the pattern length was set to $16$.
-BLACK AND WHITE VERSION COMING.
+$2$ bars and the pattern length was set to $8$.
 }}
@@
 \end{center}
-\caption{\small{Cluster around some pattern... TO BE REDONE}}
+\caption{\small{Cluster around centroid presented in
+Figure \ref{fig:code}. Taken from the artist20 dataset, the cluster
+size is actually $639$. Shown samples were randomly selected.
+}}
 \label{fig:cluster}
@@
 \end{center}
-\caption{\small{Cluster around the pattern and the difference of both.}}
+\caption{\small{First three patterns of figure \ref{fig:cluster}
+($2$nd line) presented with the centroid from Figure \ref{fig:code}
+($1$st line) and the difference between both ($3$rd line).
+}}
 \label{fig:cluster_diff}
@@
 good answer is an offset of $0$.
+
+The result presented here suggest a dynamic programming setting where
+we could find the optimal bar segmentation according to our codebook. 
+We could try
+all possible bar length (in terms of beats) and not restrict ourselves
+to $4$ beats bars and offsets between $0$ and $3$. This is part of
+our future work.
 

### 2010-03-21 Paper/Paper.tex
improved

@@
   \centering
-  \subfloat[][\small{Metallica}]{\label{fig:metallica2}\includegraphics[width=0.4\columnwidth]{metallica_pattern}}                
-  \subfloat[][\small{Tori Amos / Suzanne Vega}]{\label{fig:amosvega2}\includegraphics[width=0.4\columnwidth]{toriamos_suzannevega_pattern}}
+  \subfloat[][\small{Metallica}]{\label{fig:metallica2}\includegraphics[width=0.35\columnwidth]{metallica_pattern}}
+  \hspace{5mm}                
+  \subfloat[][\small{Tori Amos / Suzanne Vega}]{\label{fig:amosvega2}\includegraphics[width=0.35\columnwidth]{toriamos_suzannevega_pattern}}
   \caption{\small{Typical patterns for different artists.}}

### 2010-03-22 Paper/generate_datasize_curves.py
new results

@@ -17,17 +17,20 @@ def create_fig():
 
     # DATA
 
-    sizes = [1000,5000,10000,50000,100000,250000,500000,1000000,2000000]
-    labels = ['1K','5K','10K','50K','100K','250K','500K','1M','2M']
+    sizes = [1000,5000,10000,50000,100000,250000,500000]
+    labels = ['1K','5K','10K','50K','100K','250K','500K']
     
     # data 1 bar 4 beats
-    bar1beat4 = [0.150370,0.145327,0.143567,0.142257,0.142195,0.141796,0.142081,0.142118,0.141596]
+    bar1beat4 = [0.052247,0.041941,0.038449,0.036892,0.036722,
+                 0.035973,0.035910,0.035904]
     # data 2 bars 8 beats
-    bar2beat8 = [0.171300,0.165960,0.165063,0.162439,0.161588,0.161910,0.161233,0.161057,0.161068]
+    bar2beat8 = [0.066548,0.049282,0.046441,0.045507,0.044301,
+                 0.044183,0.043896,0.043906]
     # data 1 bar 4 beats not random! first K in data
-    bar1beat4FIRST = [0.157142,0.150045,0.148357,0.146069,0.145765,0.145819,0.145772,0.145924,0]
+    bar1beat4FIRST = [0.052247,0.044359,0.039107,0.038257,0.036157,
+                      0.036697,0.036598,0.035935]
     # data 0 bar 4 beats
-    bar0beat4 = [0.155474,0.149619,0.148056,0.146328,0.146004,0,0,0,0]
+    #bar0beat4 = [0.155474,0.149619,0.148056,0.146328,0.146004,0.146024,0.145962,0.145873,0.145955]
 
 
     # START DISPLAYING STUFF
@@ -37,8 +40,8 @@ def create_fig():
     # axis
     xmin = 0
     xmax = len(labels)
-    ymin = 0.13
-    ymax = 0.17
+    ymin = 0.035
+    ymax = 0.067
     P.axis([xmin,xmax,ymin,ymax])
 
     # plot1
@@ -48,7 +51,7 @@ def create_fig():
     # plot3
     P.plot(bar1beat4FIRST,'--',label='1 bar 4 beats non random')
     # plot4
-    P.plot(bar0beat4,'-.',label='0 bars 4 beats')
+    #P.plot(bar0beat4,'-.',label='0 bars 4 beats')
     
     # labels
[…]

### 2010-03-22 Paper/generate_expo_codes_curves.py
new results

@@ -16,13 +16,17 @@ import pylab as P
 def create_fig():
 
     # DATA
-    labels = ['2','4','16','256']
+    #labels = ['2','4','16','256']
+    labels = ['1/4','1/2','1','2','4','8']
 
+    data1 = [0.027235,0.030586,0.043468]
+    range1 = [0,1,2]
     # 2 beats - 1 bar - 2 bars - 4 bars
-    data1 = [0.066416,0.050712,0.050172,0.049894]
+    data2 = [0.032700,0.054496,0.049674,0.049894]
+    range2 = [1,2,3,4]
     # 1 bar - 2 bars - 4 bars - 8 bars
-    data2 = [0.058435,0.054792,0.055489,0.054860]
-
+    data3 = [0.058189,0.055909,0.055811,0.054860]
+    range3 = [2,3,4,5]
     
     # START DISPLAYING STUFF
     
@@ -31,15 +35,18 @@ def create_fig():
     # axis
     xmin = 0
     xmax = len(labels)
-    ymin = 0.49
-    ymax = 0.67
+    ymin = 0.027
+    ymax = 0.059
     P.axis([xmin,xmax,ymin,ymax])
 
     # plot1
-    P.plot(data1,'o-',label='2 beats, 1 bar, 2 bars, 4 bars')
+    P.plot(range1,data1,'o-',label='1/4, 1/2, 1, 2 bars')
     # plot2
-    P.plot(data2,'x-',label='1 bar, 2 bars, 4 bars, 8 bars')
+    P.plot(range2,data2,'x-',label='1/2, 1, 2, 4 bars')
+    # plot3
+    P.plot(range3,data3,'<-',label='1, 2, 4, 8 bars')
 
+    """
     # ANNOTATION
     gcf = P.gca()
     # data1 pt1
@@ -90,16 +97,17 @@ def create_fig():
                 arrowprops=dict(arrowstyle="->",
                                 connectionstyle="arc3,rad=-.2")
                 )
-    
+    """
+                
     # labels
     P.xticks(P.arange(len(labels)),list(labels))
     
     # x titles and y titles
-    P.xlabel('number of codes')
+    P.xlabel('bars')
     P.ylabel('average error')
 
     # legend
-    P.legend()
+    P.legend(loc='lower right')
 
     # main title
     P.title('Encoding error per code and pattern size')

### 2010-03-23 Paper/Paper.tex
improved
added title and abstract from dan

@@
 % ------
-\title{Large-scale harmonic patterns clustering}
+\title{Clustering Beat-Chroma Patterns in a Large Music Database}
 
@@
 \begin{abstract}
-The goal is to analyze very large collections of music. We describe
-a clustering scheme of harmonic patterns that scales linearly in time
-with the number of songs analyzed. These harmonic patterns are related
-to the \textit{Shingles} idea, but we scale to larger sets and we do not
-perform $k$-nn with them. We create a codebook and assert
-its properties. In particular, we show that clustering with notions of
-bars and song segments helps, and that we get codes associated more
-with some artists than others.
+A musical style or genre implies a set of common conventions
+and patterns combined and deployed in different ways to make
+individual musical pieces; for instance, most would agree that
+contemporary pop music is assembled from a relatively small
+palette of harmonic and melodic patterns.  The purpose of this
+paper is to use a database of tens of thousands of songs
+in combination with a compact representation of melodic-harmonic
+content, the beat-synchronous chromagram, and data-mining
+tools (clustering) to attempt to explicitly catalog this palette --
+at least within the limitations of the beat-chroma representation.
+We use online $k$-means clustering to summarize 3.7 million
+4-beat bars in a codebook of a few hundred prototypes.
+By measuring how accurately such a quantized codebook
+can reconstruct the original data, we can quantify the degree
+of diversity (distortion as a function of codebook size) and
+temporal structure (i.e. the advantage gained
+by joint quantizing multiple frames) in this music.  The most
+popular codewords themselves reveal the common chords
+used in the music.  Finally, the quantized representation of
+music can be used for music retrieval tasks such as artist
+and genre classification, and identifying songs that are
+similar in terms of their melodic-harmonic content.
 \end{abstract}
@@
 \caption{\small{Encoding larger and larger patterns with a fixed size
-codebook.}}
+codebook. The size of the pattern is defined by the number of beats,
+the bar information was not taken into account for that experiment.
+}}
 \label{fig:perbeat}
@@
 \begin{center}
-\includegraphics[width=.9\columnwidth]{close_patterns_diff}
+\includegraphics[width=.8\columnwidth]{close_patterns_diff}
 \end{center}
@@
 \begin{figure}[htb]
-\begin{center}
+%\begin{center}
 \includegraphics[width=.9\columnwidth]{conf_mat_per_artist}
-\end{center}
+%\end{center}
 \caption{\small{Confusion matrix for artist recognition task.}}

### 2010-03-23 Paper/Paper.tex
added emacs latex mode header

@@
+% -*- TeX-PDF-mode: t -*-
+
 % -----------------------------------------------
@@
 
-\section{Previous Work}\label{sec:prevwork}
+\subsection{Previous Work}\label{sec:prevwork}
 Our work can be seen as an extension of the \textit{Shingles} idea
@@
 
-\section{Data}\label{sec:data}
+\section{Approach}\label{sec:approach}
 In this section we present our use of the Echo Nest features and the
@@
 
-\subsection{EchoNest API}
+\subsection{Features}
 We use the Echo Nest analyze API \cite{EchoNest} throughout this work.
@@
 
-\subsection{Cowbell Dataset}
-We have $43,300$ songs, giving  
-%$\bf{3,720,091}$ 
-$\mathbf{3.7}$\textbf{M} 
-non zero bars, a bar usually 
-contains $4$ beats. WHERE DOES THIS DATA COMES FROM? GIFT FROM EN?
-
-\subsection{USPOP}
-We had access to a low quality (32kbps) version of the songs in the uspop 2002 
-data set \cite{uspop2002}.
-This well-known data show a great diversity of pop songs, but is not particular
-otherwise\footnote{We 
-also tried to use Tzanetakis dataset \cite{Tzanetakis2002a}, but 
-the $30$ seconds segments seemed hard to analyze by the Echo Nest API, most did
-not contain any bar.}.
-uspop2002 serves as testing set to measure how well a codebook learned on
-the Cowbell data set can represent new songs. We get Echo Nest features
-for $8651$ songs of that dataset.
-
-\subsection{Artist Set}
-This set encompass $1402$ songs $20$ artists, mostly rock and pop of different 
-subgenres. Some results were available regarding the artist recognition
-task. Few details given to preserve anonymity.
-
-\section{Algorithm}\label{sec:algo}
+%\section{Algorithm}\label{sec:algo}
 In this section we present the creation of the harmonic patterns from
@@
 so the row with the more energy is always the lowest one. This can easily be
-seen in Figure \ref{fig:code}. We considered the same technique for rolling
+seen in Figure \ref{fig:code}. We considered the same techniqueh for rolling
 along the other axis and be invariant regarding the downbeat. However, it
@@
 
-\section{Experiments - Fundamentals}\label{sec:experiments}
+\section{Experiments}\label{sec:experiments}
 In this section we assert the basic properties of our encoding scheme.
@@
 
+\subsection{Training: Cowbell Dataset}\label{sec:traindata}
+We have $43,300$ songs, giving  
+%$\bf{3,720,091}$ 
+$\mathbf{3.7}$\textbf{M} 
+non silent bars, a bar usually 
+contains $4$ beats. WHERE DOES THIS DATA COMES FROM? GIFT FROM EN?
+\footnote{http://www.morecowbell.dj/}
+
+\subsection{Testing: USPOP}\label{sec:testdata}
+We had access to a low quality (32kbps) version of the songs in the uspop 2002 
+data set \cite{uspop2002}.
+This well-known data show a great diversity of pop songs, but is not particular
+otherwise\footnote{We 
+also tried to use Tzanetakis dataset \cite{Tzanetakis2002a}, but 
+the $30$ seconds segments seemed hard to analyze by the Echo Nest API, most did
+not contain any bar.}.
+uspop2002 serves as testing set to measure how well a codebook learned on
+the Cowbell data set can represent new songs. We get Echo Nest features
+for $8651$ songs of that dataset.
+
 \subsection{Setting}\label{ssec:setting}
@@
 
-We use the codebook to encode our testing set (see Section \ref{sec:data}).
+We use the codebook to encode our testing set (see Section \ref{sec:testdata}).
 Each pattern is encoded with only one code. We can measure the average
@@
 \caption{\small{Original song and its encoding. Must get a better one for
-the paper!!!!!!}}
+the paper!!!!!!  What happened to good day sunshine?}}
 \label{fig:encodesong}
 \end{figure}
+
+Crazy idea: chord recognition via quantization: transcribe codewords,
+then just encode song to get chords. would this work?
 
@@
 
-\section{Further Experiments}\label{sec:exps2}
+\section{Applications}\label{sec:exps2}
 We have shown the basic properties of our clustering algorithm in
@@
 \subsection{Artist Recognition}
+
+Data:
+This set encompass $1402$ songs $20$ artists, mostly rock and pop of different 
+subgenres. Some results were available regarding the artist recognition
+task. Few details given to preserve anonymity.
+
 Let us be clear, the clustering method proposed in this work is not
@@
 
+Artists like metallica are characterized by ``wideband'' patterns -
+featured muddled due to noisy acoustics - lots of transients,
+distortion, etc.  (not by cleaned up by EN's feature analysis)
+
 % ignoring the two following figures
@@
 \end{center}
-\caption{\small{Most typical pattern for Tori Amos and Suzanne Vega.}}
+\caption{\small{Most typical pattern for Tori Amos and Suzanne Vega.
+    Can we show the top 2 or 4 for each to get a better idea?}}
 \label{fig:amosvega}
@@
 gaussian mixtures \cite{Vasconcelos2001}.
+%  (similar idea used for indexing \cite{Muja2009})
 

### 2010-03-23 Paper/generate_expo_codes_curves.py
new result

@@ -19,8 +19,8 @@ def create_fig():
     #labels = ['2','4','16','256']
     labels = ['1/4','1/2','1','2','4','8']
 
-    data1 = [0.027235,0.030586,0.043468]
-    range1 = [0,1,2]
+    data1 = [0.027235,0.030586,0.043468,0.041904045]
+    range1 = [0,1,2,3]
     # 2 beats - 1 bar - 2 bars - 4 bars
     data2 = [0.032700,0.054496,0.049674,0.049894]
     range2 = [1,2,3,4]

### 2010-03-23 Paper/generate_datasize_curves.py
typo solved

@@ -17,8 +17,8 @@ def create_fig():
 
     # DATA
 
-    sizes = [1000,5000,10000,50000,100000,250000,500000]
-    labels = ['1K','5K','10K','50K','100K','250K','500K']
+    sizes = [0,1000,5000,10000,50000,100000,250000,500000]
+    labels = ['0','1K','5K','10K','50K','100K','250K','500K']
     
     # data 1 bar 4 beats
     bar1beat4 = [0.052247,0.041941,0.038449,0.036892,0.036722,

### 2010-03-23 Paper/Paper.tex
artist recognition paragraph improved

@@
 
-Data:
-This set encompass $1402$ songs $20$ artists, mostly rock and pop of different 
-subgenres. Some results were available regarding the artist recognition
-task. Few details given to preserve anonymity.
+We apply our codebook to the artist recognition task.
+Our data set encompass $1402$ songs from $20$ artists, 
+mostly rock and pop of different subgenres. Results exist in the
+litterature for that set \cite{Ellis2007}. The best accuracy is
+$59\%$, whereas using only chroma as a representation yields an
+accuracy of $33\%$. The latter results is obtained through a
+gaussian mixture model.
 
-Let us be clear, the clustering method proposed in this work is not
+Once again, the clustering method proposed in this work is not
 intended to perform well at artist recognition. An obvious fact is that
@@
 imagine that some artist use more often some patterns than others.
-Best accuracy reported\footnote{few details given to preserve anonymity} 
-is around $54.5\%$. Our result is $23.4\%$. Random
-is around $5\%$. The confusion matrix can be seen in Figure \ref{fig:conf_mat}.
+Our accuracy result is $23.4\%$ while random
+is around $5\%$. 
+The confusion matrix can be seen in Figure \ref{fig:conf_mat}.
 Details of the experiment follow.

### 2010-03-23 Paper/Paper.tex
good day sunshine added
improved caption for experiment where we square codebook size

@@
 Distortion measure on the test data set. Data size tested range
-from $10$ thousands to $2$ millions. Patterns were selected at
-random from the dataset of approximately $3,7$ million patterns.
-OLD GRAPH WITH OLD NUMBERS
+from $0$ (just initialization) to $500$ thousand. Patterns were selected at
+random from the dataset of approximately $3,7$ million patterns, except
+for the dotted line (non random in the legend). In that case we took the
+first $N$ samples in our dataset. The difference is only visible for small
+training sets ($1$K).
 }}
@@
 \end{center}
-\caption{\small{NEW RESULTS COMING IN}}
+\caption{\small{
+Experiment on squaring the codebook size when the pattern size doubles.
+Each line represent $4$ codebook sizes: $2$, $4$, $16$ and $256$.
+The pattern size range from $1$ for $1/4$ of a bar to $32$ for $8$ bars.
+If everything was random, each of these lines should be flat.
+}}
 \label{fig:size_pattern}
@@
 \end{center}
-\caption{\small{Original song and its encoding. Must get a better one for
-the paper!!!!!!  What happened to good day sunshine?}}
+\caption{\small{Good Day Sunshine by \textit{The Beatles}.
+Original song and its encoding using a $200$ codes codebook learning
+$2$ bars at a time.
+}}
 \label{fig:encodesong}

### 2010-03-23 Paper/Paper.tex
improved caption for experiment where we square codebook size
more captions improved

@@
 The pattern size range from $1$ for $1/4$ of a bar to $32$ for $8$ bars.
-If everything was random, each of these lines should be flat.
+If everything was random, each of these lines should be flat. The training
+size was fixed to $128$K samples.
 }}
@@
 \end{center}
-\caption{\small{Most used codes in the codebook.}}
+\caption{\small{
+Most used codes in the codebook in term of frequency of use in the test
+set. This codebook was of size $200$ and trained on $2$ bars, pattern
+size is $12$x$8$. Code in Figure \ref{fig:code} is the $8$th most used
+here (second row, third column).
+}}
 \label{fig:codes1}
@@
 \end{center}
-\caption{\small{Locally linear embedding (LLE) visualization of the codebook.}}
+\caption{\small{
+Locally linear embedding (LLE) visualization of the codebook.
+LLE tries to keep near neighbors (here according to euclidean distance)
+in the same $2$D neighborhood. Shown patterns were randomly selected
+in each neighborhood.
+}}
 \label{fig:lle}

### 2010-03-23 Paper/Paper.tex
intro to experiemnts section shorten

@@
 \caption{\small{Pseudocode of Online Vector Quantization. Note that we can 
-replace the number of iteration by a threshold on the distortion over some 
+replace the number of iterations by a threshold on the distortion over some 
 test set.}
@@
 \section{Experiments}\label{sec:experiments}
-In this section we assert the basic properties of our encoding scheme.
-The first subsection gives the details to reproduce the experiments,
-such as the distance measure and learning rate. In the second subsection, we
-measure properties such as ``how the number of codes affect the distortion''
-and ``how many learning samples should we use''. Though important numbers
-to report, the busy reader could skip to subsection \ref{ssec:random}
-and the following sections.
+In this section we present different clustering experiments. We introduce
+our training data and our main testing data. Some detailed settings
+of our algorithm are also provided. Then, we answer some common questions
+about any clustering algorithms, for instance 
+``how the number of codes affect the distortion''
+and ``how many learning samples should we use''.
 Note that we will present two additional experiments in 

### 2010-03-23 Paper/Paper.tex
codebook visualization section improved

@@
 \subsection{Codebook}
-We can present the most typical patterns (Figure \ref{fig:codes1}). 
-Then, LLE\footnote{implementation: 
+Visualizing codewords in a meaningful way is always a challenge, 
+even if we have only a one or two hundreds of them.
+We present $3$ figures in this section that should help the reader
+foster his intuition. First, we display the most used codewords in
+the testing set (Figure \ref{fig:codes1}). Then, we use the
+ LLE\footnote{implementation: 
 \url{http://www.astro.washington.edu/users/vanderplas/coding/LLE/}} 
-\cite{Roweis2000}? (Figure \ref{fig:lle}). 
-Then, the graph on variance. (Figure \ref{fig:code_var}).
+algorithm \cite{Roweis2000} to visualize codewords arranged in
+coherent groups (Figure \ref{fig:lle}).
+
+Finally, more related to our particular data, we try to assess how many
+sustained patterns are represented in the codebook. We look at
+the average variance over time, and we get an estimate of $60\%$ of the
+codebook (Figure \ref{fig:code_var}).
 

### 2010-03-23 PythonSrc/for_ron.py
set16 added, redoing set 15 without keyinvariance

@@ -18,7 +18,7 @@ featsDir = os.path.expanduser('~/projects/ismir10-patterns/beatFeats')
 testFeatsDir = os.path.expanduser('~/projects/ismir10-patterns/uspop_mat')
 outputDir = os.path.expanduser('~/projects/ismir10-patterns/experiments')
 
-def do_experiment(experiment_dir,beats,bars,nCodes,nSamples=0,useFirsts=False,seed=0,offset=0,partialbar=1):
+def do_experiment(experiment_dir,beats,bars,nCodes,nSamples=0,useFirsts=False,seed=0,offset=0,partialbar=1,keyinv=True):
     """
     Performs an independant experiment!!!!
     """
@@ -31,7 +31,7 @@ def do_experiment(experiment_dir,beats,bars,nCodes,nSamples=0,useFirsts=False,se
 
     args = dict(experiment_dir=experiment_dir, beats=beats, bars=bars,
                 nCodes=nCodes, nSamples=nSamples, useFirsts=useFirsts,seed=seed,
-                offset=offset,partialbar=partialbar)
+                offset=offset,partialbar=partialbar,keyinv=keyinv)
     sp.io.savemat(os.path.join(experiment_dir, 'args.mat'), args)
 
     # TRAINING
@@ -41,7 +41,7 @@ def do_experiment(experiment_dir,beats,bars,nCodes,nSamples=0,useFirsts=False,se
     if not os.path.exists(os.path.join(experiment_dir, 'codebook.mat')):
         # load everything, unit: 1 bar, resized to 4 beats
         # key invariant, not downbeatinvariant
-        featsNorm = demos.get_data_maxener(pSize=beats,keyInv=True,downBeatInv=False,bars=bars,offset=offset,partialbar=partialbar)
+        featsNorm = demos.get_data_maxener(pSize=beats,keyInv=keyinv,downBeatInv=False,bars=bars,offset=offset,partialbar=partialbar)
         
         # select nSamples random samples out of it
         if nSamples == 0:
@@ -74,7 +74,7 @@ def do_experiment(experiment_dir,beats,bars,nCodes,nSamples=0,useFirsts=False,se
     
     # load and test
     dists,avg_dists = demos.load_and_encode_data(codebook,pSize=beats,
-                                                 keyInv=True,
[…]

### 2010-03-23 Paper/generate_expo_codes_curves.py
modifs... not very good

@@ -13,9 +13,10 @@ import pylab as P
 
 
 
-def create_fig():
+def create_fig(penalty=True):
 
     # DATA
+    sizes = [1,2,4,8,16,32]
     #labels = ['2','4','16','256']
     labels = ['1/4','1/2','1','2','4','8']
 
@@ -27,6 +28,24 @@ def create_fig():
     # 1 bar - 2 bars - 4 bars - 8 bars
     data3 = [0.058189,0.055909,0.055811,0.054860]
     range3 = [2,3,4,5]
+
+
+    # PENALTY
+    if penalty:
+        """
+        for k in range(len(data1)):
+            data1[k] = data1[k] / sizes[range1[k]] * np.log2(12)
+        for k in range(len(data2)):
+            data2[k] = data2[k] / sizes[range2[k]] * np.log2(12)
+        for k in range(len(data3)):
+            data3[k] = data3[k] / sizes[range3[k]] * np.log2(12)
+        """
+        for k in range(len(data1)):
+            data1[k] = data1[k] / np.log2(pow(12,range1[k])+1)
+        for k in range(len(data2)):
+            data2[k] = data2[k] / np.log2(pow(12,range2[k])+1)
+        for k in range(len(data3)):
+            data3[k] = data3[k] / np.log2(pow(12,range3[k])+1)
     
     # START DISPLAYING STUFF
     
@@ -35,8 +54,12 @@ def create_fig():
     # axis
     xmin = 0
     xmax = len(labels)
-    ymin = 0.027
-    ymax = 0.059
+    if not penalty:
+        ymin = 0.027
+        ymax = 0.059
+    else:
+        ymin = min(min(data1),min(data2),min(data3))
+        ymax = max(max(data1),max(data2),max(data3))
     P.axis([xmin,xmax,ymin,ymax])
 
     # plot1
@@ -107,13 +130,13 @@ def create_fig():
     P.ylabel('average error')
 
     # legend
-    P.legend(loc='lower right')
-
+    if not penalty:
+        P.legend(loc='lower right')
+    else:
+        P.legend(loc='upper right')
+        
     # main title
     P.title('Encoding error per code and pattern size')
-
-
-
 
     P.show()
 
@@ -123,7 +146,7 @@ def create_fig():
 def die_with_usage():
     """ help menu """
     print 'create figure'
-    print 'usage: python generate_expo_codes_curves.py -go'
[…]

### 2010-03-23 Paper/generate_rate_distortion_curves.py
better rate distortion curves

@@ -1,0 +1,64 @@
+"""
+Code to generate the curve where the number of codes is squared
+when the pattern size is doubled
+"""
+
+
+
+import os
+import sys
+import time
+import numpy as np
+import pylab as P
+
+
+
+def create_fig():
+
+    # DATA
+    sizes = [1,2,4,8,16,32]
+    #labels = ['2','4','16','256']
+    labels = ['1/4','1/2','1','2','4','8']
+
+    data1 = [0.027235,0.030586,0.043468,0.041904045]
+    range1 = [0,1,2,3]
+    # 2 beats - 1 bar - 2 bars - 4 bars
+    data2 = [0.032700,0.054496,0.049674,0.049894]
+    range2 = [1,2,3,4]
+    # 1 bar - 2 bars - 4 bars - 8 bars
+    data3 = [0.058189,0.055909,0.055811,0.054860]
+    range3 = [2,3,4,5]
+
+    # RATE
+    # R = (#code index + #roll index) / (pattern size in beats)
+    # R = (#codes used + 12 * #codes used) / (pattern size in beats)
+    # R = (32/pattern_size * 13) / (pattern size in beats)
+    # R = (32 * 13) / (pattern size in beats)^2
+    rates = np.ones(len(sizes)) * 32 * 13
+    rates /= np.array(sizes)
+
+    # plot1
+    P.plot(rates[range1],data1,'o-',label='1/4, 1/2, 1, 2 bars')
+
+
+    # main title
+    P.title('Encoding error per code and pattern size')
+
+
+    P.show()
+
+    
+
+def die_with_usage():
+    """ help menu """
+    print 'create figure'
+    print 'usage: python generate_expo_codes_curves.py -go'
+    sys.exit(0)
+
+
+if __name__ == '__main__':
+
+    if len(sys.argv) < 2:
+        die_with_usage()
+
+    create_fig()

### 2010-03-23 Paper/generate_rate_distortion_curves.py
better rate distortion curves

@@ -39,12 +39,16 @@ def create_fig():
 
     # plot1
     P.plot(rates[range1],data1,'o-',label='1/4, 1/2, 1, 2 bars')
+    # plot2
+    P.plot(rates[range2],data2,'o-',label='1/2, 1, 2, 4 bars')
+    # plot3
+    P.plot(rates[range3],data3,'o-',label='1, 2, 4, 8 bars')
 
-
+    # legend
+    P.legend(loc='upper right')
     # main title
     P.title('Encoding error per code and pattern size')
-
-
+    # done, show
     P.show()
 
     

### 2010-03-23 Paper/generate_rate_distortion_curves.py
better rate distortion curves

@@ -35,19 +35,41 @@ def create_fig():
     # R = (32/pattern_size * 13) / (pattern size in beats)
     # R = (32 * 13) / (pattern size in beats)^2
     rates = np.ones(len(sizes)) * 32 * 13
-    rates /= np.array(sizes)
+    rates /= (np.array(sizes) * np.array(sizes))
+
+    # data per codebook size
+    # and corresponding rates
+    cb2 = [data1[0],data2[0],data3[0]]
+    rates2 = [rates[k] for k in range(0,3)]
+    cb4 = [data1[1],data2[1],data3[1]]
+    rates4 = [rates[k] for k in range(1,4)]
+    cb16 = [data1[2],data2[2],data3[2]]
+    rates16 = [rates[k] for k in range(2,5)]
+    cb256 = [data1[3],data2[3],data3[3]]
+    rates256 = [rates[k] for k in range(3,6)]
+
 
     # plot1
-    P.plot(rates[range1],data1,'o-',label='1/4, 1/2, 1, 2 bars')
+    #P.plot(rates[range1],data1,'o-',label='1/4, 1/2, 1, 2 bars')
     # plot2
-    P.plot(rates[range2],data2,'o-',label='1/2, 1, 2, 4 bars')
+    #P.plot(rates[range2],data2,'o-',label='1/2, 1, 2, 4 bars')
     # plot3
-    P.plot(rates[range3],data3,'o-',label='1, 2, 4, 8 bars')
+    #P.plot(rates[range3],data3,'o-',label='1, 2, 4, 8 bars')
 
+    # plot1
+    P.plot(rates2,cb2,'o-',label='2 codewords')
+    P.plot(rates4,cb4,'o-',label='4 codewords')
+    P.plot(rates16,cb16,'o-',label='16 codewords')
+    P.plot(rates256,cb256,'o-',label='256 codewords')        
+
+
+    # labels
+    P.xlabel('rate')
+    P.ylabel('distortion')
     # legend
     P.legend(loc='upper right')
     # main title
-    P.title('Encoding error per code and pattern size')
+    P.title('Distortion rate for different codebook sizes')
     # done, show
     P.show()
 

### 2010-03-23 Paper/Paper.tex
dan stuff added, citations not worked out yet

@@
 \section{Introduction}\label{sec:introduction}
+The availability of very large collections of music audio present
+many interesting research opportunities.  Given millions of examples
+from a single, broad class (e.g., contemporary commercial pop music),
+can we infer anything about the underlying structure and common
+features of this class?  This paper describes our work in this direction.
 
-Working with large collections is a necessity.
+What are the common features of pop music?  There are conventions
+of subject matter, instrumentation, form, rhythm, harmony, and melody, among
+others.  Our interest here is in the tonal content of the music -- i.e. the
+harmony and melody.  As a computationally convenient proxy for a richer
+description of the tonal content of audio, we use the popular Chroma
+representation, which collapses an acoustic spectrum into a 12-dimensional
+description, with one bin for each semitone of the western musical octave.
+In addition, we simplify the time axis of our representation to take advantage
+of the strong beat present in most pop music, and record just one
+chroma vector per beat.  This beat-synchronous chromagram
+representation represents a typical music track
+in a few thousand values, yet when resynthesized back into audio via
+modulation of octave-invariant ``Shepard tones'', the melody and
+chord sequences of the original music typically remain recognizable
+\cite{EllisPoliner-icassp07}.  To the extent, then, that beat-chroma
+representations preserve tonal structure, they
+are an interesting domain in which to search for patterns -- rich enough
+to generate musically-relevant results, but simplified enough to
+abstract away some other aspects of the original audio, such as
+instrumentation and other stylistic details.
 
-Computationally expensive, most of the techniques developed do not scale
-to hundred of thousands songs. Practical question, how to describe a music
-space made of millions of songs?
+Specifically, this paper identifies common patterns in beat-synchronous
+chromagrams by learning codebooks from a large set of examples.
+The individual codewords consist of short beat-chroma patches, typically
+between 1 and 8 beats, which are able to represent the entire dataset
+of millions of patterns with the minimum error given a small codebook of
+a few hundred entries.  Our goal is to identify meaningful information
+about the musical structure represented in the entire database by
+examining individual entries in this codebook.  Since the common
+patterns represent a higher-level description of the musical content,
+we also expect them to be useful in other applications, such as
+music classification and retrieving tonally-related items.
 
-Some work have tackled that problem. For instance, Bertin-Mahieux et al.
-\cite{Bertin-Mahieux2008} automatically tag a collection larger than $100$K
-songs. Casey and Slaney \cite{Casey2006} introduced their \textit{Shingles}
-representation and perform $k$-nn over thousands of examples.
+One major issue in working with small patches concerns temporal
+alignment.  [explain bar alignment]
 
-We approach the clustering problem keeping in mind the scaling
-requirement. We use a representation similar to the \textit{Shingles}.
-We train on $43$K songs and test on more than $7$K ones.
+Prior work using small patches of chroma features includes ...
+[casey \& slaney] [ellis \& poliner] [ellis ismir 07].
 

### 2010-03-23 Paper/generate_rate_distortion_curves.py
better rate distortion curves

@@ -62,7 +62,20 @@ def create_fig():
     P.plot(rates16,cb16,'o-',label='16 codewords')
     P.plot(rates256,cb256,'o-',label='256 codewords')        
 
-
+    # ANNOTATION
+    gcf = P.gca()
+    # data1 pt1
+    gcf.annotate('1/4 bar', xy=(rates2[0], cb2[0]),  xycoords='data',
+                xytext=(-50, 30), textcoords='offset points',
+                arrowprops=dict(arrowstyle="->",
+                                connectionstyle="arc3,rad=.2")
+                )
+    gcf.annotate('1 bar', xy=(rates2[2], cb2[2]),  xycoords='data',
+                xytext=(50, -30), textcoords='offset points',
+                arrowprops=dict(arrowstyle="->",
+                                connectionstyle="arc3,rad=-.2")
+                )
+    
     # labels
     P.xlabel('rate')
     P.ylabel('distortion')

### 2010-03-23 Paper/Paper.tex
citations resolved
interpretation of fig 7

@@
 % what order do we use? Thierry, Ron, Dan?  Dan, Thierry, Ron?
+
+% Q: Why are there long spaces after "e.g." or "i.e."?
+% A: LaTeX interprets a dot and a blank as the end of a sentence. To
+%    prevent this, use "e.g.\ " and "i.e.\ " or "e.g.~" and "i.e.~".
+\newcommand{\ie}{i.e.~}
+\newcommand{\Ie}{I.e.~}
+\newcommand{\eg}{e.g.~}
+\newcommand{\Eg}{E.g.~}
 
@@
 chord sequences of the original music typically remain recognizable
-\cite{EllisPoliner-icassp07}.  To the extent, then, that beat-chroma
+\cite{Ellis2007a}.  To the extent, then, that beat-chroma
 representations preserve tonal structure, they
@@
 Prior work using small patches of chroma features includes ...
-[casey \& slaney] [ellis \& poliner] [ellis ismir 07].
+\cite{Casey2006,Ellis2007,Ellis2007a}.
 
@@
 \subsection{Codebook}
-Visualizing codewords in a meaningful way is always a challenge, 
-even if we have only a one or two hundreds of them.
-We present $3$ figures in this section that should help the reader
-foster his intuition. First, we display the most used codewords in
-the testing set (Figure \ref{fig:codes1}). Then, we use the
- LLE\footnote{implementation: 
+Visualizing codewords in a meaningful way is always a challenge, even
+if we have only a one or two hundreds of them.  We present $3$ figures
+in this section that should help the reader foster his
+intuition.
+
+FIXME: fix the labels in the figure (label pitch class on the y axis,
+and maybe only beats [0 2 4 6] on the ax axis, add codeword numbers to
+title), etc.
+
+The 25 most frequently used codewords in the testing set are shown in
+Figure \ref{fig:codes1}.  The codewords primarily consist of sustained
+notes and simple chords.  Since they are designed to be key-invariant,
+specific notes do not appear.  Instead the first 7 codewords
+correspond to a single note sustained across two bars (codeword 1),
+perfect fifth (codes 2 and 3) and fourth intervals (codes 4 7), and a
+major and minor triads (code 6 and 5).  Many of the remaining
+codewords correspond to transitions from one chord to another, \eg a
+transition from a major triad to a minor triad in code 8.
+
+(It is kind of interesting that sustained minor triad is more common
+than the major triad (codes 4 and 5 in figure 7).  Don't fear the
+reaper is in Amin, could it be related to this?)
+
+Then, we use the LLE\footnote{implementation:
   \url{http://www.astro.washington.edu/users/vanderplas/coding/LLE/}}

### 2010-03-23 Paper/Paper.tex
fixed i.e.'s and e.g.'s

@@
 of diversity (distortion as a function of codebook size) and
-temporal structure (i.e. the advantage gained
+temporal structure (\ie the advantage gained
 by joint quantizing multiple frames) in this music.  The most
@@
 many interesting research opportunities.  Given millions of examples
-from a single, broad class (e.g., contemporary commercial pop music),
+from a single, broad class (\eg, contemporary commercial pop music),
 can we infer anything about the underlying structure and common
@@
 of subject matter, instrumentation, form, rhythm, harmony, and melody, among
-others.  Our interest here is in the tonal content of the music -- i.e. the
+others.  Our interest here is in the tonal content of the music -- \ie the
 harmony and melody.  As a computationally convenient proxy for a richer
@@
 Note that we do not claim that any of these informations (segments, beats, bars)
-are perfectly accurate (e.g. in \cite{Barrington2009a} authors obtain a better
+are perfectly accurate (\eg in \cite{Barrington2009a} authors obtain a better
 song segmentation than Echonest). Practise showed us that they are reasonable, 
@@
 distance between a pattern and its encoding. We can also measure the use
-of the different codes, i.e. the size of the cluster around a code.
+of the different codes, \ie the size of the cluster around a code.
 
@@
 in each neighborhood.
-}}
+}
+ronw: I think this figure would be better if it only included the
+bottom subplot.
+}
 \label{fig:lle}
@@
 Once we understand the kind of codebook we create, we naturally want
-to visualize the clusters, i.e. the pattersn from actual song that
+to visualize the clusters, \ie the pattersn from actual song that
 surround a given centroid. Unsurprisingly, real patterns tend to be

### 2010-03-23 Paper/Paper.tex
fixed comment about code 8

@@
 codewords correspond to transitions from one chord to another, \eg a
-transition from a major triad to a minor triad in code 8.
+transition from one major triad to another (\ie Cmaj to Fmaj assuming
+that chroma 0 is C) in code 8.  Does this transition have a better
+name like ``relative x''?
 

### 2010-03-23 Paper/Paper.tex
typo

--- 2010-03-23 Paper/Paper.tex
@@
 % what order do we use? Thierry, Ron, Dan?  Dan, Thierry, Ron?
-
-% Q: Why are there long spaces after "e.g." or "i.e."?
-% A: LaTeX interprets a dot and a blank as the end of a sentence. To
-%    prevent this, use "e.g.\ " and "i.e.\ " or "e.g.~" and "i.e.~".
-\newcommand{\ie}{i.e.~}
-\newcommand{\Ie}{I.e.~}
-\newcommand{\eg}{e.g.~}
-\newcommand{\Eg}{E.g.~}
 
@@
 The number of chroma vectors we get can vary. We resize the features in the
-time axis to a fixed length. Resizing can be argued over, but it should
+time axis to a fixed length. Resizing instead of clipping can be argued over, 
+but it should
 preserve the main bar characteristics (such as a strong beat at the beginning
@@
 \subsection{Codebook}
-Visualizing codewords in a meaningful way is always a challenge, even
-if we have only a one or two hundreds of them.  We present $3$ figures
-in this section that should help the reader foster his
-intuition.
-
-FIXME: fix the labels in the figure (label pitch class on the y axis,
-and maybe only beats [0 2 4 6] on the ax axis, add codeword numbers to
-title), etc.
-
-The 25 most frequently used codewords in the testing set are shown in
-Figure \ref{fig:codes1}.  The codewords primarily consist of sustained
-notes and simple chords.  Since they are designed to be key-invariant,
-specific notes do not appear.  Instead the first 7 codewords
-correspond to a single note sustained across two bars (codeword 1),
-perfect fifth (codes 2 and 3) and fourth intervals (codes 4 7), and a
-major and minor triads (code 6 and 5).  Many of the remaining
-codewords correspond to transitions from one chord to another, \eg a
-transition from a major triad to a minor triad in code 8.
-
-(It is kind of interesting that sustained minor triad is more common
-than the major triad (codes 4 and 5 in figure 7).  Don't fear the
-reaper is in Amin, could it be related to this?)
-
-Then, we use the LLE\footnote{implementation:
+Visualizing codewords in a meaningful way is always a challenge, 
+even if we have only a one or two hundreds of them.
+We present $3$ figures in this section that should help the reader
+foster his intuition. First, we display the most used codewords in
+the testing set (Figure \ref{fig:codes1}). Then, we use the
+ LLE\footnote{implementation: 
 \url{http://www.astro.washington.edu/users/vanderplas/coding/LLE/}} 

### 2010-03-23 Paper/Paper.tex
fleshed out/reorganized section 4 a bit more

--- 2010-03-23 Paper/Paper.tex
@@
 \section{Visualization} \label{sec:visu}
-The goal here is to give an intuition to the reader of what kind of
-typical patterns do we get. We also want to have a feel for what the
-clusters look like, how many boring sustained notes are encoded in the
-codebook, can the codebook describes the complexity of the music space, etc.
+% The goal here is to give an intuition to the reader of what kind of
+% typical patterns do we get. We also want to have a feel for what the
+% clusters look like, how many boring sustained notes are encoded in the
+% codebook, can the codebook describes the complexity of the music space, etc.
 
-Below we present some figures to help the reader with those questions.
-Please bear in mind the limited space.
+\subsection{Codebook}\label{sec:codebook}
 
-\subsection{Encoding of a Song}
-Here we present a song with its two encodings. First, the harmonic patterns
-given by the EchoNest. Secondly, each pattern is encoded by its closest
-code in the codebook. See Figure \ref{fig:encodesong}.
-
-\begin{figure}[htb]
-\begin{center}
-\includegraphics[width=.9\columnwidth]{song_encoded}
-\end{center}
-\caption{\small{Good Day Sunshine by \textit{The Beatles}.
-Original song and its encoding using a $200$ codes codebook learning
-$2$ bars at a time.
-}}
-\label{fig:encodesong}
-\end{figure}
-
-Crazy idea: chord recognition via quantization: transcribe codewords,
-then just encode song to get chords. would this work?
-
-\subsection{Codebook}
-Visualizing codewords in a meaningful way is always a challenge, even
-if we have only a one or two hundreds of them.  We present $3$ figures
-in this section that should help the reader foster his
-intuition.
+How was this codebook trained?
+What do we expect it to capture?
 
@@
 
-The 25 most frequently used codewords in the testing set are shown in
+The 25 most frequently used codewords in the test set are shown in
 Figure \ref{fig:codes1}.  The codewords primarily consist of sustained
@@
 major and minor triads (code 6 and 5).  Many of the remaining
-codewords correspond to transitions from one chord to another, \eg a
-transition from one major triad to another (\ie Cmaj to Fmaj assuming
-that chroma 0 is C) in code 8.  Does this transition have a better
-name like ``relative x''?
+codewords correspond to common transitions from one chord to another,
+\eg a transition from one major triad to another (\ie Cmaj
+$\rightarrow$ Fmaj assuming that chroma 0 is C) in codes 8 and 10 and
+the reverse transition in code (\ie Fmaj $\rightarrow$ Cmaj) in code
+22.  Does this transition have a better name like ``relative x''?
 
@@
 the average variance over time, and we get an estimate of $60\%$ of the
-codebook (Figure \ref{fig:code_var}).
+codebook (Figure \ref{fig:code_var}).  25\% of the codebook correspond
+to very simple sustained patterns similar to the top row of Figure
+\ref{fig:codes1}.  How is this different from the next 35\%?
 
@@
 Once we understand the kind of codebook we create, we naturally want
-to visualize the clusters, \ie the pattersn from actual song that
+to visualize the clusters, \ie the patterns from actual song that
 surround a given centroid. Unsurprisingly, real patterns tend to be
@@
 ($1$st line) and the difference between both ($3$rd line).
-}}
+}
+ronw: does this figure really add all that much?
+}
 \label{fig:cluster_diff}
 \end{figure}
+
+\subsection{Example song encoding}
+
+Figure \ref{fig:encodesong} shows a typical example of encoding a song
+using the codebook described in Section \ref{sec:codebook}.  Both the
+Echo Nest chroma features for an example song and the corresponding
+quantization are displayed.  The quantized representation retains
+dominant harmonic structure of the original features, but is smoothed
+in time due to the use of 2 bar long codewords.
+%
+This is consistent with the codebook described in Section
+\ref{sec:codebook} which primarily consists of sustained chords across
+4 or 8 beats.
+
+\begin{figure}[htb]
+\begin{center}
+\includegraphics[width=.9\columnwidth]{song_encoded}
+\end{center}
+\caption{\small{Good Day Sunshine by \textit{The Beatles}.
+Original song and its encoding using a $200$ codes codebook learning
+$2$ bars at a time.
+}}
+\label{fig:encodesong}
+\end{figure}
+
+Crazy idea: chord recognition via quantization: transcribe codewords,
+then just encode song to get chords. would this work?
 

### 2010-03-23 Paper/Paper.tex
minor, see if merged done
Dan text added, citations resolved

@@
 
-Prior work using small patches of chroma features includes ...
-\cite{Casey2006,Ellis2007,Ellis2007a}.
+Prior work using small patches of chroma features includes
+the ``shingles'' of \cite{Casey2007}, which they used to
+identify ``remixes'' i.e. music based on some of the same
+underlying instrument tracks, as well as for matching
+performances of Mazurkas \cite{Casey2008}.  That work,
+however, was not concerned with extracting the deeper
+common patterns underlying different pieces (and did not
+use beat- or bar-synchronous features).  Earlier work in
+beat-synchronous analysis includes \cite{Bartsch2001},
+who looked for repeated patterns within single songs to
+identify the chorus, and \cite{Ellis2007a}, who
+cross-correlated beat-chroma matrices to match cover
+versions of pop music tracks.  None of these works
+examined or interpreted the content of the chroma
+matrices in any detail.  In contrast, here we hope to
+develop a codebook whose entries are of interest
+in their own right.
 
@@
 
+>>>>>>>>>>>>>>>>>>>> File 1
+>>>>>>>>>>>>>>>>>>>> File 2
+Crazy idea: chord recognition via quantization: transcribe codewords,
+then just encode song to get chords. would this work?
+>>>>>>>>>>>>>>>>>>>> File 3
+<<<<<<<<<<<<<<<<<<<<
+
 Then, we use the LLE\footnote{implementation:

### 2010-03-23 Paper/Paper.tex
experiment with canging the codebook size updated
remove merge problem

@@
 \hline
-\# codes & \# samples & avg. dist \\ \hline \hline
-1 & $500$ & $0.066079$ \\
-10 & $5$K & $0.045690$ \\
-50 & $25$K & $0.038456$ \\
-100 & $50$K & $0.036725$ \\
-200 & $100$K & $0.033923$ \\
-500 & $250$K & $0.030973$ \\
-1000 & $500$K & \\ \hline
+\# codes & distortion \\ \hline \hline
+1 & $0.066081$ \\
+10 & $0.045579$ \\
+50 & $0.038302$ \\
+100 & $0.035904$ \\
+500 & $0.030841$ \\ \hline
 \end{tabular}
@@
 \caption{\small{Experiment with the codebook size. We train on patterns
-corresponding to $1$ bar resized to $4$ beats. The number of training
-samples is set to $500$ times the number of codes. 
-NEW RESULTS COMING SOON WITH FIXED SIZED DATASET }}
+corresponding to $1$ bar resized to $4$ beats. 
+The number of training samples is $50000$.
+}}
 \label{tab:cbsize}
@@
 
->>>>>>>>>>>>>>>>>>>> File 1
->>>>>>>>>>>>>>>>>>>> File 2
 Crazy idea: chord recognition via quantization: transcribe codewords,
 then just encode song to get chords. would this work?
->>>>>>>>>>>>>>>>>>>> File 3
-<<<<<<<<<<<<<<<<<<<<
 

### 2010-03-23 Paper/Paper.tex
commenting out subsec previous work

@@
 
+% COMMENTING PREVIOUS WORK
+\iffalse
 \subsection{Previous Work}\label{sec:prevwork}
@@
 typical pattern representing that cluster.
+\fi
 
@@
 Most used codes in the codebook in term of frequency of use in the test
-set. This codebook was of size $200$ and trained on $2$ bars, pattern
+set (artist20, see Subsection \ref{ssec:artist}). 
+This codebook was of size $200$ and trained on $2$ bars, pattern
 size is $12$x$8$. Code in Figure \ref{fig:code} is the $8$th most used
-here (second row, third column).
+here (second row, third column). Over the patterns, we display the number of
+times it was used out of $71832$ total patterns in the dataset.
 }}
@@
 
-\subsection{Artist Recognition}
+\subsection{Artist Recognition} \label{ssec:artist}
 

### 2010-03-23 Paper/Paper.tex
figures improved

@@
 ($2$nd line) presented with the centroid from Figure \ref{fig:code}
-($1$st line) and the difference between both ($3$rd line).
+($1$st line) and the absolute difference between both ($3$rd line).
 }

### 2010-03-23 Paper/Paper.tex
better codes displayed
code display improved
new figure: freqs, added

@@
 The 25 most frequently used codewords in the test set are shown in
-Figure \ref{fig:codes1}.  The codewords primarily consist of sustained
+Figure \ref{fig:codes1}.  
+The frequency of use of these codewords are shown in Figure \ref{fig:freqs}.
+The codewords primarily consist of sustained
 notes and simple chords.  Since they are designed to be key-invariant,
@@
 size is $12$x$8$. Code in Figure \ref{fig:code} is the $8$th most used
-here (second row, third column). Over the patterns, we display the number of
-times it was used out of $71832$ total patterns in the dataset.
+here (second row, third column). 
+Numbers display are the index (from most used to least) and in parentheses
+the percentage of use in the whole collection times $100$.
+There is $71832$ patterns in total in the dataset.
 }}
 \label{fig:codes1}
+\end{figure}
+
+\begin{figure}[htb]
+\begin{center}
+\includegraphics[width=.7\columnwidth]{freqs}
+\end{center}
+\caption{\small{
+Frequency of use of the patterns in Figure \ref{fig:codes1} on the
+artist20 dataset.
+There is $71832$ patterns in total in the dataset.
+}}
+\label{fig:freqs}
 \end{figure}

### 2010-03-23 Paper/Paper.tex
new figure: freqs, added
caption improved

@@
 Numbers display are the index (from most used to least) and in parentheses
-the percentage of use in the whole collection times $100$.
+the percentage of use in the whole collection times.
 There is $71832$ patterns in total in the dataset.
@@
 There is $71832$ patterns in total in the dataset.
+Note that all codes were initialized with a random
+sample from the training set, some might not just be present in this
+test set.
 }}

### 2010-03-23 Paper/Paper.tex
improved

@@
 meaningful than an arbitrary time length. More on this in the experiments
-(Section \ref{sec:experiments} TO BE CHECKED).
+(Subsection \ref{ssec:segment}).
 
@@
 The number of chroma vectors we get can vary. We resize the features in the
-time axis to a fixed length. Resizing can be argued over, but it should
+time axis to a fixed length. Resizing over clipping can be argued over, 
+but it should
 preserve the main bar characteristics (such as a strong beat at the beginning
@@
 otherwise\footnote{We 
-also tried to use Tzanetakis dataset \cite{Tzanetakis2002a}, but 
-the $30$ seconds segments seemed hard to analyze by the Echo Nest API, most did
-not contain any bar.}.
+also tried to use Tzanetakis dataset, 
+%\cite{Tzanetakis2002a},
+but the $30$ seconds segments seemed hard to analyze by the Echo Nest API, 
+most did not contain any bar.}.
 uspop2002 serves as testing set to measure how well a codebook learned on
@@
 \subsection{Setting}\label{ssec:setting}
-We take one or two bars, normalize the patches to size 4, 8, or 16.
+We take one or two bars and normalize the patches to size 4 or 8 respectively.
 We roll the patch to be invariant to the key (on the patch level, not on
@@
 
-% NEW RESULTS COMING SOON WITH FIXED SIZED DATASET
 \begin{table}
@@
 
-\subsection{Song Segmentation}
+\subsection{Song Segmentation} \label{ssec:segment}
 We explained that we used bars as defined by the EchoNest algorithm to

### 2010-03-23 Paper/Paper.tex
improved

@@
 This section present some fundamental results of the clustering algorithm.
-These are important, but do not change the larger picture. Thus we will not
-spend too much time describing them. In short: more training samples is
-better, more codes is better, larger patterns are more difficult.
+These are important, but do not change the larger picture. Due to space
+constraints, we give few details and refer to the figures.
 \begin{itemize}
-\item First, a larger dataset gives a better encoding codebook. See Figure
+\item A larger dataset gives a better encoding codebook. See Figure
 \ref{fig:data_sizes}.
-\item Secondly, a larger codebook better encodes new song. See 
+\item A larger codebook better encodes new songs. See 
 Table \ref{tab:cbsize}.
-\item Finally, larger patterns are more difficult to encode, thus require
+\item Larger patterns are more difficult to encode, thus require
 larger codebooks. See Figure \ref{fig:perbeat}.
@@
 
-These properties are important to verify and measure, but they are not
-surprising for a clustering algorithm. We move on to more interesting facts.
+These properties are important to verify and for reproducing the experiments, 
+but they are not surprising for a clustering algorithm.
 

### 2010-03-23 Paper/Paper.tex
improved, removed rate distortion section
cowbell data description added

@@
 \subsection{Training: Cowbell Dataset}\label{sec:traindata}
-We have $43,300$ songs, giving  
-%$\bf{3,720,091}$ 
-$\mathbf{3.7}$\textbf{M} 
+We have $43,300$ tracks that were uploaded
+to morecowbell.dj\footnote{http://www.morecowbell.dj/}, 
+an online service that uses the Echo
+Nest to produce a remixed version of
+an uploaded music file that adds cowbell and other sound
+effects in time with the music.
+The $43.3$K songs, giving $\mathbf{3.7}$\textbf{M}
 non silent bars, a bar usually
-contains $4$ beats. WHERE DOES THIS DATA COMES FROM? GIFT FROM EN?
-\footnote{http://www.morecowbell.dj/}
+contains $4$ beats (Figure \ref{fig:barsize}).
 
@@
 
+% REMOVING THE FOLLOWING SUBSECTION ON RATE - DISTORTION
+\iffalse
 \subsection{Music Randomness}\label{ssec:random}
@@
 \end{figure}
+\fi
 
@@
 of the patterns. We could compare the centroids with the result of
-a lowpass filtering.
-
-See Figure \ref{fig:cluster} and \ref{fig:cluster_diff}.
+a lowpass filtering. In Figure \ref{fig:cluster} we show elements of the same
+cluster, and in Figure \ref{fig:cluster_diff}, we emphasize the difference
+with the centroid.
 

### 2010-03-23 Paper/Paper.tex
improved
improved, new figure added

@@
 % ------
-\title{Clustering Beat-Chroma Patterns in a Large Music Database}
+\title{Clustering Beat-Chroma Patterns\\ in a Large Music Database}
 
@@
 
-How was this codebook trained?
-What do we expect it to capture?
-
-FIXME: fix the labels in the figure (label pitch class on the y axis,
-and maybe only beats [0 2 4 6] on the ax axis, add codeword numbers to
-title), etc.
+We train a codebook of size $200$ over patterns of size $12$x$8$ covering
+$2$ bars at a time. Results shown are on a second testing set, see
+Subsection \ref{ssec:artist}.
 
@@
 
-(It is kind of interesting that sustained minor triad is more common
-than the major triad (codes 4 and 5 in figure 7).  Don't fear the
-reaper is in Amin, could it be related to this?)
-
-Crazy idea: chord recognition via quantization: transcribe codewords,
-then just encode song to get chords. would this work?
-
 Then, we use the LLE\footnote{implementation:
@@
 \ref{fig:codes1}.  How is this different from the next 35\%?
+
+Showing codewords of $2$ bars is practical and consistent with our experiments
+in Section \ref{sec:exps2}. However, longer codewors can also reveal
+interesting structures. A random samples of such codewords is presented
+in Figure \ref{fig:codes8bars}
 
@@
 \label{fig:cluster_diff}
+\end{figure}
+
+\begin{figure}[htb]
+\begin{center}
+\includegraphics[width=.8\columnwidth]{codes_8bars}
+\end{center}
+\caption{\small{Sample of longer codewords spanning over $8$ bars. Codewords
+selected randomly from a codebook of size $200$. As usual, we see sustained
+notes, but also some long-term structures.
+}}
+\label{fig:codes8bars}
 \end{figure}
@@
 gaussian mixtures or other generative model. That being said, these methods
-could be use to develop better distance measures between \textit{Shingles}.
+could be use to develop better distance measures between patterns.
 The use of the euclidean distance is arbitrary. Summarizing patches

### 2010-03-23 Paper/Paper.tex
improved

@@
 
-Then, we use the LLE\footnote{implementation:
-  \url{http://www.astro.washington.edu/users/vanderplas/coding/LLE/}}
-algorithm \cite{Roweis2000} to visualize codewords arranged in
-coherent groups (Figure \ref{fig:lle}).
-
-Finally, more related to our particular data, we try to assess how many
-sustained patterns are represented in the codebook. We look at
-the average variance over time, and we get an estimate of $60\%$ of the
-codebook (Figure \ref{fig:code_var}).  25\% of the codebook correspond
-to very simple sustained patterns similar to the top row of Figure
-\ref{fig:codes1}.  How is this different from the next 35\%?
-
-Showing codewords of $2$ bars is practical and consistent with our experiments
-in Section \ref{sec:exps2}. However, longer codewors can also reveal
-interesting structures. A random samples of such codewords is presented
-in Figure \ref{fig:codes8bars}
-
-\begin{figure}[htb]
+\begin{figure}[h]
 \begin{center}
@@
 \end{figure}
+
+Another interesting visualization uses LLE\footnote{implementation:
+  \url{http://www.astro.washington.edu/users/vanderplas/coding/LLE/}}
+algorithm \cite{Roweis2000} to display codewords arranged in
+coherent groups. See Figure \ref{fig:lle}.
+
+Finally, more related to our particular data, we try to assess how many
+sustained patterns are represented in the codebook. We look at
+the average variance over time, and we get an estimate of $60\%$ of the
+codebook (Figure \ref{fig:code_var}).  25\% of the codebook correspond
+to very simple sustained patterns similar to the top row of Figure
+\ref{fig:codes1}.  How is this different from the next 35\%?
+
+Showing codewords of $2$ bars is practical and consistent with our experiments
+in Section \ref{sec:exps2}. However, longer codewors can also reveal
+interesting structures. A random samples of such codewords is presented
+in Figure \ref{fig:codes8bars}
 

### 2010-03-23 Paper/Paper.tex
improved

@@
 \subsection{Definitions}
-A pattern, or a patch, is a \textit{Shingle} taken from an actual song.
-A code is similar to a pattern, but is the result of the clustering
+A pattern, or a patch, is a beat-chroma pattern taken from an actual song.
+A codeword (sometimes just code) is similar to a pattern, but is the result 
+of the clustering
 algorithm and most likely does not come from an actual song.
@@
 \end{center}
-\caption{\small{A typical code from a codebook of size $200$. 
+\caption{\small{A typical codeword from a codebook of size $200$. 
 Patterns represented
@@
 The idea is as follow: take a sample from your data, find the closest
-code in the codebook, bring that code closer to the sample by some amount.
+codeword in the codebook, bring that codeword closer to the 
+sample by some amount.
 The details are explained in Algorithm \ref{algo:vq}. Note that this
@@
 at a time, a bar being represented by $4$ chroma feature vector.
-Therefore, a code or a pattern is of size $12$ x $4$ = $48$.
+Therefore, a codeword or a pattern is of size $12$ x $4$ = $48$.
 Distortion measure on the test data set. Data size tested range
@@
 perfect fifth (codes 2 and 3) and fourth intervals (codes 4 7), and a
-major and minor triads (code 6 and 5).  Many of the remaining
+major and minor triads (codeword 6 and 5).  Many of the remaining
 codewords correspond to common transitions from one chord to another,
@@
 
-Crazy idea: chord recognition via quantization: transcribe codewords,
-then just encode song to get chords. would this work?
-
 \section{Applications}\label{sec:exps2}
@@
 Best offset according to our codebook. The real one is $0$.
+Random guess gives an accuracy of $25$\%.
 }}

### 2010-03-23 Paper/Paper.tex
improved

@@
 sustained patterns are represented in the codebook. We look at
-the average variance over time, and we get an estimate of $60\%$ of the
-codebook (Figure \ref{fig:code_var}).  25\% of the codebook correspond
+the average variance over time in Figure \ref{fig:code_var}.
+25\% of the codebook correspond
 to very simple sustained patterns similar to the top row of Figure
-\ref{fig:codes1}.  How is this different from the next 35\%?
+\ref{fig:codes1}. For $2$ bars codewords, it seems that interesting
+patterns appear in only $33$\% of the data.
 

### 2010-03-24 Paper/Paper.tex
eddited section 2

@@
 many interesting research opportunities.  Given millions of examples
-from a single, broad class (\eg, contemporary commercial pop music),
+from a single, broad class (\eg contemporary commercial pop music),
 can we infer anything about the underlying structure and common
@@
 In this section we present the creation of the harmonic patterns from
-the Echo Nest data. Then, we explain the online vector quantization algorithm
+the Echo Nest data.  Then we explain the online vector quantization algorithm
 used to learn a codebook of those patterns.
 
-\subsection{Pattern Creation}
-Patterns are defined as a number of beats or bars. In the case of beats,
-we simply cut the set of chromas per beat every $N$ beats. In the case
-of bars, we gather all beats corresponding to the $N$ bars requested.
-The number of chroma vectors we get can vary. We resize the features in the
-time axis to a fixed length. Resizing over clipping can be argued over, 
-but it should
-preserve the main bar characteristics (such as a strong beat at the beginning
-or a note change in the middle).
+%\subsection{Harmonic Patterns}
+\subsection{Beat-Chroma Patches}
+% Patterns are defined as a number of beats or bars. In the case of beats,
+% we simply cut the set of chromas per beat every $N$ beats. In the case
+% of bars, we gather all beats corresponding to the $N$ bars requested.
+% The number of chroma vectors we get can vary. We resize the features in the
+% time axis to a fixed length. Resizing over clipping can be argued over, 
+% but it should
+% preserve the main bar characteristics (such as a strong beat at the beginning
+% or a note change in the middle).
 
-We try to normalize the patterns in respect to the key. We roll the matrix
-so the row with the more energy is always the lowest one. This can easily be
-seen in Figure \ref{fig:code}. We considered the same techniqueh for rolling
-along the other axis and be invariant regarding the downbeat. However, it
-seems less robust to noise. Note that we roll each pattern independently.
-Also, this implies that the pitch envelope is important rather than the actual 
-pitches.
+We use the bar segmentation obtained from the analysis described in
+the previous section to break a song into a collection of beat-chroma
+``patches'', typically one or two bars in length. 
+%
+Because the bar length varies, depending on the meter of a particular
+song (see Figure \ref{fig:barsize}), we resample each patch to a fixed
+length of typically 4 to 8 beats.
+% although the majority of the data contains bars that are 4 beats long
 
-\subsection{Codebook Learning}
+Finally, we normalize the patches with respect to the key by rotating
+the pattern matrix so that the first row contains the most
+energy. This can easily be seen in Figure \ref{fig:code}.
+% We considered the same technique for rolling along the other axis and
+% be invariant regarding the downbeat. However, it seems less robust to
+% noise.
+Each patch within a song is normalized independently, so exact
+reconstruction of the original song requires knowledge of the
+rotation index (key) for each patch.
+
+The representation resulting from this process is invariant to both
+the key and meter of the original song.  This enables the study of
+broad harmonic patterns found throughout the data, without regard for
+the specific musical context.
+%
+% In the context of clustering, this enables us to study harmonic
+% pattern usage without maintaining clusters for the specific context a
+% particular patterns occurs in,
+In the context of clustering this avoids \eg obtaining separate
+clusters for every major triad in duple or triple meter.
+% Also, this implies that the pitch envelope is important rather than
+% the actual pitches.
+Something about why we're not interested in specific chords?
+
+\subsection{Clustering}
+
 We use an online version of the vector quantization algorithm
-\cite{Gersho1991}. This can also be seen as an online K-means.
-The idea is as follow: take a sample from your data, find the closest
-codeword in the codebook, bring that codeword closer to the 
-sample by some amount.
-The details are explained in Algorithm \ref{algo:vq}. Note that this
-algorithm, though not optimal, is $O(n)$ with $n$ number of songs seen.
-Thus it scales linearly to any size of data set. The setting of the learning
-rate might become trickier, but this is not fundamentally a problem.
-
-We initialize the codebook by choosing $K$ random points from our dataset.
+\cite{Gersho1991}, to cluster the beat-chroma patches described in the
+previous section.
+% This can also be seen as an online K-means.
+For each sample from the data, the algorithm finds the closest cluster
+in the codebook and updates the cluster centroid (codeword) to be
+closer to the sample based on some learning rate $l$.
+The clusters are updated as each data point is seen, as opposed to
+once per iteration in the standard k-means algorithm.
+The details are explained in Algorithm \ref{algo:vq}.
+%
+As in standard k-means clustering, the codebook is initialized by
+choosing $K$ random points from our dataset.
+%
+Note that this algorithm, although not optimal, scales linearly with
+the number of patches seen and can be interrupted at any time
+while still resulting in an updated codebook. %not that this is important...
+% The setting of the learning rate might become trickier, but this is
+% not fundamentally a problem.
 

### 2010-03-24 Paper/Paper.tex
made math prettier

@@
 \setlength{\belowcaptionskip}{10pt}
+
+\DeclareMathOperator*{\dist}{dist}
+\DeclareMathOperator*{\size}{size}
 
@@
 % the actual pitches.
-Something about why we're not interested in specific chords?
+FIXME: Something about why we're not interested in specific chords?
 
@@
 in the codebook and updates the cluster centroid (codeword) to be
-closer to the sample based on some learning rate $l$.
+closer to the sample based on some learning rate $\ell$.
 The clusters are updated as each data point is seen, as opposed to
@@
 \begin{algorithmic}
-\STATE$l$ learning rate
+\STATE$\ell$ learning rate
 \STATE$\{P_n\}$ set of patches
@@
 %\STATE $m \leftarrow b$
-\REQUIRE $0 < l \leq 1$
+\REQUIRE $0 < \ell \leq 1$
 \FOR{$nIters$}
 \FOR{$p \in \{P_n\}$}
-\STATE$c \leftarrow min_{c \in C_k} dist(p,c)$
-\STATE$c \leftarrow c + (p - c) * l$
+\STATE$c \leftarrow min_{c \in C_k} \dist(p,c)$
+\STATE$c \leftarrow c + (p - c) * \ell$
 \ENDFOR
@@
 \RETURN $\{C_k\}$
-\caption{\small{Pseudocode of Online Vector Quantization. Note that we can 
-replace the number of iterations by a threshold on the distortion over some 
-test set.}
+\caption{\small{Pseudocode for the online vector quantization
+    algorithm. Note that we can replace the number of iterations by a
+    threshold on the distortion over some test set.}
 \label{algo:vq}}
@@
 As a distance measure (or distortion measure) between patterns, we use
-the average squared euclidean distance. In details, if a pattern $p1$
-is composed of elements $p1_{i,j}$, and similarly for a pattern $p2$,
-the distance between $p1$ and $p2$ is:
+the average squared euclidean distance. In details, if a pattern $p_1$
+is composed of elements $p_{1ij}$, and similarly for a pattern $p_2$,
+the distance between $p_1$ and $p_2$ is:
 \begin{eqnarray}
-  dist(p1,p2) = \sum_{i,j} (p1_{i,j} - p2_{i,j})^2 / size(p1)  \label{eq:dist}
+  \dist(p_1,p_2) = \sum_{i,j} \frac{(p_{1ij} - p_{2ij})^2}{\size(p_1)}
+  \label{eq:dist}
 \end{eqnarray}
-We assume $p1$ and $p2$ have the same size.
+We assume $p_1$ and $p_2$ have the same size.
 

### 2010-03-24 Paper/Paper.tex
more updates to section 2

@@
 
+% COMMENTING PREVIOUS WORK AND DEFINITIONS
+\iffalse
 \subsection{Definitions}
@@
 
-% COMMENTING PREVIOUS WORK
-\iffalse
 \subsection{Previous Work}\label{sec:prevwork}
@@
 \section{Approach}\label{sec:approach}
-In this section we present our use of the Echo Nest features and the
-datasets we use for training and testing our model.
+% In this section we review the creation of the beat-chroma patterns using
+% the Echo Nest data.  Then we explain the online vector quantization algorithm
+% used to learn a codebook of those patterns.
 
 \subsection{Features}
-We use the Echo Nest analyze API \cite{EchoNest} throughout this work.
-The API gives us a chroma vector (length $12$) for every music event (called 
-``segment'') for any song we upload to their platform. 
-We also get an estimate of where the beats and the bars are. Beats span over
-multiple segments, and bars span over multiple beats. 
-We transform the chroma vectors per segment into per beat vectors using a 
-simple average. We can then stack beat vectors for a number of bars 
-(usually 1 or 2). 
-We get a fixed-size patch by resampling the number of chroma vectors. Typically,
-there are $4$ beats per bar (Figure \ref{fig:barsize}). 
-Thus an appropriate pattern size for one bar
-is $12$x$4$ or $12$x$8$.
+The feature analysis used throughout this work is based on Echo Nest
+analyze API \cite{EchoNest}.  
+%
+For any song uploaded to their platform this analysis returns a chroma
+vector (length $12$) for every music event (called ``segment''), and a
+segmentation of the song into beats and bars. Beats span multiple
+segments, and bars span multiple beats.
+%
+Averaging the per-segment chroma over beat times results in a
+beat-synchronous chroma feature representation similar to that used in
+\cite{Ellis2007a}.
 
-\begin{figure}[htb]
-\begin{center}
-\includegraphics[width=.9\columnwidth]{bar_size_freq}
-\end{center}
-\caption{\small{
-Frequency of the different bar size in the training data.
-}}
-\label{fig:barsize}
-\end{figure}
+% \begin{figure}[htb]
+% \begin{center}
+% \includegraphics[width=.9\columnwidth]{bar_size_freq}
+% \end{center}
+% \caption{\small{
+% Frequency of the different bar size in the training data.
+% }}
+% \label{fig:barsize}
+% \end{figure}
 
@@
 are perfectly accurate (\eg in \cite{Barrington2009a} authors obtain a better
-song segmentation than Echonest). Practise showed us that they are reasonable, 
+song segmentation than Echonest). Practice showed us that they are reasonable, 
 and the size of the data set should make up for the imperfections or noise.
@@
 (Subsection \ref{ssec:segment}).
-
-\begin{figure}[htb]
-\begin{center}
-\includegraphics[width=.6\columnwidth]{code}
-\end{center}
-\caption{\small{A typical codeword from a codebook of size $200$. 
-Patterns represented
-$2$ bars and the pattern length was set to $8$.
-}}
-\label{fig:code}
-\end{figure}
-
-%\section{Algorithm}\label{sec:algo}
-In this section we present the creation of the harmonic patterns from
-the Echo Nest data.  Then we explain the online vector quantization algorithm
-used to learn a codebook of those patterns.
 
@@
 
-We use the bar segmentation obtained from the analysis described in
-the previous section to break a song into a collection of beat-chroma
-``patches'', typically one or two bars in length. 
+We use the bar segmentation obtained from the Echo Nest analysis to
+break a song into a collection of beat-chroma ``patches'', typically
+one or two bars in length.
 %
 Because the bar length varies, depending on the meter of a particular
-song (see Figure \ref{fig:barsize}), we resample each patch to a fixed
-length of typically 4 to 8 beats.
+song, we resample each patch to a fixed length of typically 4 beats
+per bar.  Since the overwhelming majority (85\%) of our training data
+consisted of bars that were 4 beats long, this resampling was not
+often required.  Most of the remaining bars (14\%) were 3 beats in
+length.  (FIXME: thierry, can you check these numbers, I'm making them
+up based on the deletedfigure)
 % although the majority of the data contains bars that are 4 beats long
+The resulting patches consist of $12 \times 4$ or $12 \times 8$ matrices.
+
+\begin{figure}[thb]
+\begin{center}
+\includegraphics[width=.6\columnwidth]{code}
+\end{center}
+\caption{\small{A typical codeword from a codebook of size $200$. 
+Patterns represented
+$2$ bars and the pattern length was set to $8$.
+}}
+\label{fig:code}
+\end{figure}
 
@@
 \subsection{Clustering}
+
+% We can then stack beat vectors for a number of bars (usually 1 or 2).
 

### 2010-03-24 Paper/Paper.tex
lots of tweaks to section 4

@@
 \end{center}
-\caption{\small{A typical codeword from a codebook of size $200$. 
+\caption{\small{A typical codeword from a codebook of size $200$ (code
+    7 in Figure \ref{fig:codes1})
 Patterns represented
@@
 and ``how many learning samples should we use''.
-Note that we will present two additional experiments in 
-Section \ref{sec:exps2}.
+% Note that we will present two additional experiments in 
+% Section \ref{sec:exps2}.
 
-\subsection{Training: Cowbell Dataset}\label{sec:traindata}
-We have $43,300$ tracks that were uploaded
-to morecowbell.dj\footnote{http://www.morecowbell.dj/}, 
-an online service that uses the Echo
-Nest to produce a remixed version of
-an uploaded music file that adds cowbell and other sound
-effects in time with the music.
-The $43.3$K songs, giving $\mathbf{3.7}$\textbf{M}
-non silent bars, a bar usually
-contains $4$ beats (Figure \ref{fig:barsize}).
+\subsection{Data}
+%\subsection{Training data: Cowbell Dataset}
+\label{sec:traindata}
+%\subsection{Test data: USPOP}
+\label{sec:testdata}
+Our training data consists of $43,300$ tracks that were uploaded to
+morecowbell.dj,\footnote{http://www.morecowbell.dj/} an online service
+based on the Echo Nest analyze API which remixes uploaded music by
+adding cowbell and other sound effects synchronized in time with the
+music.  The $43.3$K songs contain $\mathbf{3.7}$ million non silent
+bars which we cluster using the approach described in the previous
+section.
 
-\subsection{Testing: USPOP}\label{sec:testdata}
-We had access to a low quality (32kbps) version of the songs in the uspop 2002 
-data set \cite{uspop2002}.
-This well-known data show a great diversity of pop songs, but is not particular
-otherwise\footnote{We 
-also tried to use Tzanetakis dataset, 
-%\cite{Tzanetakis2002a},
-but the $30$ seconds segments seemed hard to analyze by the Echo Nest API, 
-most did not contain any bar.}.
-uspop2002 serves as testing set to measure how well a codebook learned on
-the Cowbell data set can represent new songs. We get Echo Nest features
-for $8651$ songs of that dataset.
+For testing, we made use of low quality (32kbps) version of the songs
+from the uspop2002 data set \cite{uspop2002}.  This well-known data
+contains a wide variety of pop songs, but is not particular otherwise.
+% \footnote{We also tried to use Tzanetakis dataset,
+% %\cite{Tzanetakis2002a},
+% but the $30$ seconds segments seemed hard to analyze by the Echo Nest API, 
+% most did not contain any bar.}.
+uspop2002 serves as test set to measure how well a codebook learned on
+the Cowbell data set can represent new songs.  We obtained Echo Nest features
+for the $8651$ songs contained in the dataset.
 
 \subsection{Setting}\label{ssec:setting}
-We take one or two bars and normalize the patches to size 4 or 8 respectively.
-We roll the patch to be invariant to the key (on the patch level, not on
-the song level). We learn a codebook of size $K$ over the cowbell dataset 
-using the VQ algorithm (Algorithm \ref{algo:vq}). A typical learning rate 
-is $0.01$ for $200$ iterations over the whole dataset.
-
-We use the codebook to encode our testing set (see Section \ref{sec:testdata}).
+We take one or two bars and normalize the patches to size 4 or 8
+respectively.  We learn a codebook of size $K$ over the Cowbell
+dataset using the VQ algorithm (Algorithm \ref{algo:vq}). We used a
+learning rate of $\ell=0.01$ for $200$ iterations over the whole
+dataset.
+%
+We use the codebook to encode the test set.
 Each pattern is encoded with only one code. We can measure the average
@@
 
-As a distance measure (or distortion measure) between patterns, we use
-the average squared euclidean distance. In details, if a pattern $p_1$
-is composed of elements $p_{1ij}$, and similarly for a pattern $p_2$,
-the distance between $p_1$ and $p_2$ is:
+We use the average squared Euclidean distance as the distortion
+measure between chroma patches.  Given a pattern $p_1$
+composed of elements $p_{1ij}$, and a similar pattern $p_2$, the
+distance between them is:
 \begin{eqnarray}
@@
 \end{eqnarray}
-We assume $p_1$ and $p_2$ have the same size.
+We assume $p_1$ and $p_2$ have the same size.  This is enforced by the
+meter normalization procedure described in Section \ref{sec:approach}.
 
 \subsection{Basic Properties}
+FIXME: flesh this out a \emph{little} more?
+
 This section present some fundamental results of the clustering algorithm.
@@
 \begin{itemize}
-\item A larger dataset gives a better encoding codebook. See Figure
+\item Encoding performance improves with increasing training data. See Figure
 \ref{fig:data_sizes}.
-\item A larger codebook better encodes new songs. See 
+\item Encoding performance improves with increasing codebook size. See 
 Table \ref{tab:cbsize}.
-\item Larger patterns are more difficult to encode, thus require
+\item Larger patterns are more difficult to encode, thus requiring
 larger codebooks. See Figure \ref{fig:perbeat}.
@@
 training sets ($1$K).
+%
+FIXME: Can we kill the ``non-random'' line.  It doesn't really add anything.
 }}
@@
 codebook. The size of the pattern is defined by the number of beats,
-the bar information was not taken into account for that experiment.
+the bar information was not taken into account for this experiment.
 }}
@@
 We train a codebook of size $200$ over patterns of size $12$x$8$ covering
-$2$ bars at a time. Results shown are on a second testing set, see
-Subsection \ref{ssec:artist}.
+$2$ bars at a time.  The results shown are on the test set from
+Section \ref{ssec:artist}.
 
@@
 correspond to a single note sustained across two bars (codeword 1),
-perfect fifth (codes 2 and 3) and fourth intervals (codes 4 7), and a
-major and minor triads (codeword 6 and 5).  Many of the remaining
+perfect fifth (codewords 2 and 3) and fourth intervals (codewords 4
+and 7), and a
+major and minor triads (codewords 6 and 5, respectively).  Many of the remaining
 codewords correspond to common transitions from one chord to another,
-\eg a transition from one major triad to another (\ie Cmaj
-$\rightarrow$ Fmaj assuming that chroma 0 is C) in codes 8 and 10 and
-the reverse transition in code (\ie Fmaj $\rightarrow$ Cmaj) in code
-22.  Does this transition have a better name like ``relative x''?
+\eg a transitions between major triads in codes 8 and 10 ( Cmaj
+$\rightarrow$ Fmaj)% assuming that chroma 0 is C) 
+and the reverse transition in code 22 (\ie Fmaj $\rightarrow$ Cmaj).
+FIXME: Does this transition have a better name like ``relative x''?
 
@@
 \caption{\small{
-Most used codes in the codebook in term of frequency of use in the test
-set (artist20, see Subsection \ref{ssec:artist}). 
-This codebook was of size $200$ and trained on $2$ bars, pattern
-size is $12$x$8$. Code in Figure \ref{fig:code} is the $8$th most used
-here (second row, third column). 
-Numbers display are the index (from most used to least) and in parentheses
-the percentage of use in the whole collection times.
-There is $71832$ patterns in total in the dataset.
+Most frequently used codes in the codebook %in term of frequency of use
+in the artist20 test set. 
+This codebook contains $200$ codes trained on $2$ bar, $12 \times 8$
+patches.
+The numbers 
+%above each panel correspond to the display are the index (from most used to least) and
+in parentheses correspond to 
+the percentage of use in the test set.
+%There are $71832$ patterns in total in the dataset.
 }}
@@
 to very simple sustained patterns similar to the top row of Figure
-\ref{fig:codes1}. For $2$ bars codewords, it seems that interesting
-patterns appear in only $33$\% of the data.
+\ref{fig:codes1}. For $2$ bars codewords, it seems that patterns
+containing more complex temporal activity account for only $33$\% of
+the data.
 
 Showing codewords of $2$ bars is practical and consistent with our experiments
-in Section \ref{sec:exps2}. However, longer codewors can also reveal
-interesting structures. A random samples of such codewords is presented
-in Figure \ref{fig:codes8bars}
+in Section \ref{sec:exps2}. However, longer codewords can also reveal
+interesting structure. A random sample of such codewords is presented
+in Figure \ref{fig:codes8bars}.  FIXME: why a random sample?  Can we
+show their usage in the test (or train) set as in fig \ref{fig:codes1}?
 
@@
 }
-ronw: I think this figure would be better if it only included the
-bottom subplot.
+FIXME:Can we make this a little bit bigger so the patterns are
+actually visible?  Does the figure add much over Fig \ref{fig:codes1}?
+Maybe it has some more complex patterns? 
 }
@@
 with the centroid.
+
+FIXME: we want the differences to be close to uniformly random, right
+(hence centroid)?
+And that's kind of what they look like.  Does Fig
+\ref{fig:cluster_diff} really add much?
 
@@
 accuracy of $33\%$. The latter results is obtained through a
-gaussian mixture model.
+Gaussian mixture model.
 
@@
 As for enhancements, we specifically did not compress our features using
-gaussian mixtures or other generative model. That being said, these methods
+Gaussian mixtures or other generative model. That being said, these methods
 could be use to develop better distance measures between patterns.
 The use of the euclidean distance is arbitrary. Summarizing patches
-with gaussians, and then comparing the distance between those gaussians,
+with Gaussians, and then comparing the distance between those Gaussians,
 could reduce the influence of the noise in the distance measure.
@@
 also cluster the codebook itself, in a similar fashion as hierarchical
-gaussian mixtures \cite{Vasconcelos2001}.
+Gaussian mixtures \cite{Vasconcelos2001}.
 %  (similar idea used for indexing \cite{Muja2009})

### 2010-03-24 Paper/Paper.tex
more comments and tweaks to sections 4 and 5

@@
 axis cuts at the $63$rd pattern, a rough estimate of the number of 
-sustained note patterns.
+codewords consisting of sustained chords.
 }}
@@
 selected randomly from a codebook of size $200$. As usual, we see sustained
-notes, but also some long-term structures.
+notes, but also some long-term structure.
+FIXME: I (ronw) like the middle-left pattern, but I'm totally confused
+by those in the top right and bottom row.  Are they ever used?  They
+don't seem to be smoothed out at all, and are probably outliers
+containing only one or two training points.
 }}
@@
 \section{Applications}\label{sec:exps2}
-We have shown the basic properties of our clustering algorithm in
-Section \ref{sec:experiments}. In Section \ref{sec:visu}, we showed
-how to visualize the codebook and the clusters. We know move on to
-show other applications that can take advantage of our clustering.
-In the next two subsections, we talk about song segmentation and
-artist recognition. Our algorithm was not devised for any of these tasks,
-but some information contained in the centroids can be naturally used.
+In this section we describe two other applications of the clustering
+described in this paper: the song segmentation, and artist
+recognition.  This demonstrates the utility of unsupervised clustering
+to FIXME ... 
+% Although our algorithm was not devised for any of these
+% tasks, but some information contained in the centroids can be
+% naturally used.
 

### 2010-03-24 Paper/Paper.tex
typos

@@
 
-Artists like metallica are characterized by ``wideband'' patterns -
+Artists like Metallica are characterized by ``wideband'' patterns -
 featured muddled due to noisy acoustics - lots of transients,
@@
 We discussed the possibility to move to even larger scales
-and we provided our source code\footnote{Code not yet released to preserve
-submission's anonymity}.
+and we provided our source code.\footnote{Code not yet released to preserve
+submission's anonymity}
 

### 2010-03-24 Paper/Paper.tex
quick caption edits

@@
 \end{center}
-\caption{\small{Experiment with the codebook size. We train on patterns
-corresponding to $1$ bar resized to $4$ beats. 
-The number of training samples is $50000$.
+\caption{\small{Varying codebook size using $50,000$ training samples.
+    We train on patterns corresponding to $1$ bar resized to $4$ beats.
   }}
@@
 \end{center}
-\caption{\small{
-Frequency of use of the patterns in Figure \ref{fig:codes1} on the
-artist20 dataset. 
-There is $71832$ patterns in total in the dataset.
-Note that all codes were initialized with a random
-sample from the training set, some might not just be present in this
-test set.
+\caption{\small{ Usage frequency of the patterns in the codebook from
+    Figure \ref{fig:codes1} on the artist20 dataset.  There are $71,832$
+    patterns in total in the dataset.  Note that all codes were
+    initialized with a random sample from the training set and might
+    not be used in this test set.
+    FIXME: so why not show usage in training data?  I think test.mat
+    should contain this information.
   }}

### 2010-03-24 Paper/Paper.tex
number added for bar length

@@
 song, we resample each patch to a fixed length of typically 4 beats
-per bar.  Since the overwhelming majority (85\%) of our training data
+per bar.  Since the overwhelming majority (82\%) of our training data
 consisted of bars that were 4 beats long, this resampling was not
-often required.  Most of the remaining bars (14\%) were 3 beats in
-length.  (FIXME: thierry, can you check these numbers, I'm making them
-up based on the deletedfigure)
+often required.  Most of the remaining bars (10\%) were 3 beats in
+length.
 % although the majority of the data contains bars that are 4 beats long

### 2010-03-24 Paper/generate_datasize_curves.py
random line removed

@@ -49,7 +49,7 @@ def create_fig():
     # plot2
     P.plot(bar2beat8,'x-',label='2 bars 8 beats')
     # plot3
-    P.plot(bar1beat4FIRST,'--',label='1 bar 4 beats non random')
+    #P.plot(bar1beat4FIRST,'--',label='1 bar 4 beats non random')
     # plot4
     #P.plot(bar0beat4,'-.',label='0 bars 4 beats')
     

### 2010-03-24 Paper/Paper.tex
codewords number -1

@@
 from $0$ (just initialization) to $500$ thousand. Patterns were selected at
-random from the dataset of approximately $3,7$ million patterns, except
-for the dotted line (non random in the legend). In that case we took the
-first $N$ samples in our dataset. The difference is only visible for small
-training sets ($1$K).
-%
-FIXME: Can we kill the ``non-random'' line.  It doesn't really add anything.
+random from the dataset of approximately $3,7$ million patterns.
 }}
@@
 specific notes do not appear.  Instead the first 7 codewords
-correspond to a single note sustained across two bars (codeword 1),
-perfect fifth (codewords 2 and 3) and fourth intervals (codewords 4
-and 7), and a
-major and minor triads (codewords 6 and 5, respectively).  Many of the remaining
+correspond to a single note sustained across two bars (codeword 0),
+perfect fifth (codewords 1 and 2) and fourth intervals (codewords 3
+and 6), and a
+major and minor triads (codewords 5 and 4, respectively).  Many of the remaining
 codewords correspond to common transitions from one chord to another,
-\eg a transitions between major triads in codes 8 and 10 ( Cmaj
+\eg a transitions between major triads in codes 7 and 9 ( Cmaj
 $\rightarrow$ Fmaj)% assuming that chroma 0 is C) 
-and the reverse transition in code 22 (\ie Fmaj $\rightarrow$ Cmaj).
+and the reverse transition in code 21 (\ie Fmaj $\rightarrow$ Cmaj).
 FIXME: Does this transition have a better name like ``relative x''?

### 2010-03-24 Paper/Paper.tex
improved

@@
 \eg a transitions between major triads in codes 7 and 9 ( Cmaj
-$\rightarrow$ Fmaj)% assuming that chroma 0 is C) 
+$\rightarrow$ Fmaj or C5 $\rightarrow$ F5 as a guitar power chord)
+% assuming that chroma 0 is C) 
 and the reverse transition in code 21 (\ie Fmaj $\rightarrow$ Cmaj).
-FIXME: Does this transition have a better name like ``relative x''?
+%FIXME: Does this transition have a better name like ``relative x''?
+% C5 -> F5 in 'guitar chords', basically a power chord
 
@@
 interesting structure. A random sample of such codewords is presented
-in Figure \ref{fig:codes8bars}.  FIXME: why a random sample?  Can we
-show their usage in the test (or train) set as in fig \ref{fig:codes1}?
+in Figure \ref{fig:codes8bars}. Most used codewords would be less interesting,
+as they all are sustained patterns.
 

### 2010-03-24 Paper/Paper.tex
improved

@@
 in each neighborhood.
-}
-FIXME:Can we make this a little bit bigger so the patterns are
-actually visible?  Does the figure add much over Fig \ref{fig:codes1}?
-Maybe it has some more complex patterns? 
-}
+}}
 \label{fig:lle}
@@
 \caption{\small{Sample of longer codewords spanning over $8$ bars. Codewords
-selected randomly from a codebook of size $200$. As usual, we see sustained
-notes, but also some long-term structure.
-FIXME: I (ronw) like the middle-left pattern, but I'm totally confused
-by those in the top right and bottom row.  Are they ever used?  They
-don't seem to be smoothed out at all, and are probably outliers
-containing only one or two training points.
+selected randomly from the codewords used in the test set. Percentage of
+used display in parentheses.
+Most patterns are sustained notes, but codeword 0 shows a repeted two notes
+pattern.
 }}

### 2010-03-24 Paper/Paper.tex
added names for Cmaj->Fmaj chord transition

@@
 codewords correspond to common transitions from one chord to another,
-\eg a transitions between major triads in codes 7 and 9 ( Cmaj
+\eg a tonic-subdominant transition in codes 7 and 9
+(Cmaj
 $\rightarrow$ Fmaj or C5 $\rightarrow$ F5 as a guitar power chord)
 % assuming that chroma 0 is C) 
-and the reverse transition in code 21 (\ie Fmaj $\rightarrow$ Cmaj).
+and the reverse tonic-dominant transition in code 21 (\ie Fmaj
+$\rightarrow$ Cmaj).
 %FIXME: Does this transition have a better name like ``relative x''?
@@
 \caption{\small{
-Most frequently used codes in the codebook %in term of frequency of use
+25 most frequently used codes in the codebook %in term of frequency of use
 in the artist20 test set. 
@@
 selected randomly from the codewords used in the test set. Percentage of
-used display in parentheses.
-Most patterns are sustained notes, but codeword 0 shows a repeted two notes
+use display in parentheses.
+Most patterns are sustained notes, but code 0 shows a repeating two chord
 pattern.
@@
 \caption{\small{Good Day Sunshine by \textit{The Beatles}.
-Original song and its encoding using a $200$ codes codebook learning
+Original song and its encoding using a $200$ entry codebook learning
 $2$ bars at a time.

### 2010-03-24 Paper/Paper.tex
comment on figure 10
revised caption for fig 1

@@
 \iffalse
-\subsection{Definitions}
+\subsectionDefinitions}
 A pattern, or a patch, is a beat-chroma pattern taken from an actual song.
@@
 \caption{\small{A typical codeword from a codebook of size $200$ (code
-    7 in Figure \ref{fig:codes1})
-Patterns represented
-$2$ bars and the pattern length was set to $8$.
-}}
+    7 in Figure \ref{fig:codes1}), corresponding to a
+    tonic-subdominant chord progression.  The patch is composed of $2$
+    bars and the pattern length was set to $8$ beats.  }}
 \label{fig:code}
@@
 \end{center}
-\caption{\small{Sample of longer codewords spanning over $8$ bars. Codewords
-selected randomly from the codewords used in the test set. Percentage of
-use display in parentheses.
-Most patterns are sustained notes, but code 0 shows a repeating two chord
-pattern.
-}}
+\caption{\small{Sample of longer codewords spanning over $8$
+    bars. Codewords selected randomly from the codewords used in the
+    test set. Percentage of use display in parentheses.  Most patterns
+    consist of sustained note or chords, but code 0 shows a repeating
+    two chord pattern and code 4 contains two bars of a
+    1$\rightarrow$1$\rightarrow$1$\rightarrow$2 progression.}}
 \label{fig:codes8bars}

### 2010-03-24 Paper/Paper.tex
improved
reworded section 5.2

@@
 In this section we describe two other applications of the clustering
-described in this paper: the song segmentation, and artist
+described in this paper: song segmentation, and artist
 recognition.  This demonstrates the utility of unsupervised clustering
@@
 \subsection{Song Segmentation} \label{ssec:segment}
-We explained that we used bars as defined by the EchoNest algorithm to
-train our patterns. Let's assume once again that those bars are
-reasonable. We investigate whether we could recover the bar segmentation
-of new songs from our codebook. It would helps us assert that codes
-encodes something related to bars, for instance a strong beat at the
-beginning. Results are shown \ref{tab:offset}. We find the right segment
-with an accuracy of $62\%$ where random is $25\%$. Details of the
-experiment follow.
+
+Since the clustering described in Section \ref{sec:approach} is based
+on the segmentation of the signal in to bars, it stands to reason that
+the cluster centroids should contain information related to bar
+alignment, such as the presence of a strong beat on the first beat.
+%
+In this section we investigate whether it is possible to recover the
+bar segmentation of new songs using the information contained in the
+cluster centroids.  The results are shown \ref{tab:offset}. We find
+the right segment with an accuracy of $62\%$ where random is
+$25\%$. Details of the experiment follow.
 
 We train a codebook of size $100$ on bars resized to $4$ beats. Then,
-we take the longest sequence of bars of $4$ beats in the test set.
-As we mentioned before, $4$ beats is the most common (by far!) bar size
-in terms of beats according to the EchoNest API. We then encode this
-sequence using an offset of $0$, $1$, $2$ or $3$ beats. The best offset
-is the one that gives the encoding with the lowest distortion. The
-good answer is an offset of $0$.
+we take the longest sequence of bars of $4$ beats in the test set.  As
+we mentioned before, $4$ beats is the most common bar size using the
+Echo Nest segmentation. We then encode this sequence using an offset
+of $0$, $1$, $2$ or $3$ beats. The best offset in terms of distortion
+should correspond to the bar alignment relative to the original
+feature matrix.
+
+FIXME: Actually I'm (ronw) not sure how this was done.  Every 4 beat
+segment was treated separately?  Its not looking for a per-song alignment?
+%The good answer is an offset of $0$.
 

### 2010-03-24 Paper/Paper.tex
fixed section 5.1
prettier tables
comment on specific chords not interesting

--- 2010-03-24 Paper/Paper.tex
@@
 are perfectly accurate (\eg in \cite{Barrington2009a} authors obtain a better
-song segmentation than Echonest). Practice showed us that they are reasonable, 
+song segmentation than the Echo Nest). 
+Practice showed us that they are reasonable, 
 and the size of the data set should make up for the imperfections or noise.
@@
 % the actual pitches.
-FIXME: Something about why we're not interested in specific chords?
+%FIXME: Something about why we're not interested in specific chords?
+Specific chords do not add much information about a music collection.
+A cover song can be performed in a different key, but the motifs it
+contains remain the same.
 
@@
 \caption{\small{
-Best offset according to our codebook. The real one is $0$.
+Song segmentation experiment:
+best offset according to our codebook. The real one is $0$.
 Random guess gives an accuracy of $25$\%.

### 2010-03-24 Paper/Paper.tex

--- 2010-03-24 Paper/Paper.tex
@@
 \setlength{\belowcaptionskip}{10pt}
+
+\usepackage{booktabs}
 
@@
 \begin{center}
-\begin{tabular}{|l|l|c|}
-\hline
-\# codes & distortion \\ \hline \hline
+\begin{tabular}{llc}
+\toprule
+Codebook size & Distortion \\
+\midrule
 1 & $0.066081$ \\
@@
 100 & $0.035904$ \\
-500 & $0.030841$ \\ \hline
+500 & $0.030841$ \\
+\bottomrule
 \end{tabular}
@@
 of $0$, $1$, $2$ or $3$ beats. The best offset in terms of distortion
-should correspond to the bar alignment relative to the original
-feature matrix.
+should correspond to the bar alignment in the original
+feature matrix.  The correct answer is an offset of $0$.
 
@@
 segment was treated separately?  Its not looking for a per-song alignment?
-%The good answer is an offset of $0$.
 
@@
 \begin{center}
-\begin{tabular}{c|c}
-%\hline
-offset & \% of times chosen \\ \hline
+\begin{tabular}{cc}
+\toprule
+Offset & \% of times chosen \\
+\midrule
 0 & $\mathbf{62.6}$\\
@@
 3 & $11.5$\\
+\bottomrule
 \end{tabular}

### 2010-03-24 Paper/Paper.tex
diff figure better introduced, kept in
added FIXME in intro
improved bar alignment explanation

@@
 One major issue in working with small patches concerns temporal
-alignment.  [explain bar alignment]
+alignment.  FIXME: [explain bar alignment]
 
@@
 cluster, and in Figure \ref{fig:cluster_diff}, we emphasize the difference
-with the centroid.
+with the centroid. We hope for the difference to look uniformly random,
+the result seems to fit that assumption.
 
-FIXME: we want the differences to be close to uniformly random, right
-(hence centroid)?
-And that's kind of what they look like.  Does Fig
-\ref{fig:cluster_diff} really add much?
+%FIXME: we want the differences to be close to uniformly random, right
+%(hence centroid)?
+%And that's kind of what they look like.  Does Fig
+%\ref{fig:cluster_diff} really add much?
 
@@
 recognition.  This demonstrates the utility of unsupervised clustering
-to FIXME ... 
+to supervised tasks. The representation created from large collections
+captures meaningful information without being forced to do so.
 % Although our algorithm was not devised for any of these
@@
 We train a codebook of size $100$ on bars resized to $4$ beats. Then,
-we take the longest sequence of bars of $4$ beats in the test set.  As
+we take the longest sequence of bars of $4$ beats for each song
+in the test set.  As
 we mentioned before, $4$ beats is the most common bar size using the
-Echo Nest segmentation. We then encode this sequence using an offset
-of $0$, $1$, $2$ or $3$ beats. The best offset in terms of distortion
+Echo Nest segmentation. We then encode each of these sequences using an offset
+of $0$, $1$, $2$ or $3$ beats. The best offset for a test sequence
+(the one with the lowest distortion)
 should correspond to the bar alignment in the original
@@
 
-FIXME: Actually I'm (ronw) not sure how this was done.  Every 4 beat
-segment was treated separately?  Its not looking for a per-song alignment?
+%FIXME: Actually I'm (ronw) not sure how this was done.  Every 4 beat
+%segment was treated separately?  Its not looking for a per-song alignment?
+%TBM: yes, I simply summed the divergence from all patterns for a given
+% divergence, the lowest sum is the winner. So it's a per-song alignment.
 

### 2010-03-24 Paper/Paper.tex
intro thing about beat alignment added

@@
 One major issue in working with small patches concerns temporal
-alignment.  FIXME: [explain bar alignment]
+alignment. Learning the same motifs with different offset
+(of a few seconds, or audio frames) is redundant. We take advantage
+of rythmic knowledge about the song to align patches with beats.
+We can further group beats into bars, meaningful music segments.
+The use of the Echo Nest API \cite{EchoNest} for that purpose is 
+explained in Subsection \ref{ssec:beatpatch}. Patches built around
+those temporal landmarks should capture music motifs more easily.
 
@@
 %\subsection{Harmonic Patterns}
-\subsection{Beat-Chroma Patches}
+\subsection{Beat-Chroma Patches} \label{ssec:beatpatch}
 % Patterns are defined as a number of beats or bars. In the case of beats,
@@
 algorithm \cite{Roweis2000} to display codewords arranged in
-coherent groups. See Figure \ref{fig:lle}.
+coherent groups. See Figure \ref{fig:lle}. FOr instance, we see
+``wideband'' patterns grouping to the right.
 

### 2010-03-24 Paper/Paper.tex
basic properties fleshed out

@@
 \subsection{Basic Properties}
-FIXME: flesh this out a \emph{little} more?
-
 This section present some fundamental results of the clustering algorithm.
-These are important, but do not change the larger picture. Due to space
-constraints, we give few details and refer to the figures.
+None of the result appear surprising,  but they are important for 
+reproducing the experiments.
 \begin{itemize}
-\item Encoding performance improves with increasing training data. See Figure
-\ref{fig:data_sizes}.
-\item Encoding performance improves with increasing codebook size. See 
-Table \ref{tab:cbsize}.
+\item Encoding performance improves with increasing training data (Figure
+\ref{fig:data_sizes}). Regarding the covergence of the performance,
+we can imagine using $500$ or $1$K samples per codeword to learn if the
+training set size is problematic.
+\item Encoding performance improves with increasing codebook size
+(Table \ref{tab:cbsize}). The increase seems significant, and
+increasing the size of the codebook in an efficient way
+is definitely part of our future work.
 \item Larger patterns are more difficult to encode, thus requiring
-larger codebooks. See Figure \ref{fig:perbeat}.
+larger codebooks. See Figure \ref{fig:perbeat}. The increase is steeper
+until $4$ beats which usually corresponds to one bar.
 \end{itemize}
-
-These properties are important to verify and for reproducing the experiments, 
-but they are not surprising for a clustering algorithm.
 

### 2010-03-24 Paper/Paper.tex
submittable... almost, missing new frequency figure

@@
 (of a few seconds, or audio frames) is redundant. We take advantage
-of rythmic knowledge about the song to align patches with beats.
+of rhythmic knowledge about the song to align patches with beats.
 We can further group beats into bars, meaningful music segments.
@@
 common patterns underlying different pieces (and did not
-use beat- or bar-synchronous features).  Earlier work in
+use either beat- or bar-synchronous features).  Earlier work in
 beat-synchronous analysis includes \cite{Bartsch2001},
@@
 \item Encoding performance improves with increasing training data (Figure
-\ref{fig:data_sizes}). Regarding the covergence of the performance,
+\ref{fig:data_sizes}). Regarding the convergence of the performance,
 we can imagine using $500$ or $1$K samples per codeword to learn if the
@@
 algorithm \cite{Roweis2000} to display codewords arranged in
-coherent groups. See Figure \ref{fig:lle}. FOr instance, we see
+coherent groups. See Figure \ref{fig:lle}. For instance, we see
 ``wideband'' patterns grouping to the right.
@@
 mostly rock and pop of different subgenres. Results exist in the
-litterature for that set \cite{Ellis2007}. The best accuracy is
+literature for that set \cite{Ellis2007}. The best accuracy is
 $59\%$, whereas using only chroma as a representation yields an
@@
 featured muddled due to noisy acoustics - lots of transients,
-distortion, etc.  (not by cleaned up by EN's feature analysis)
+distortion, etc, which can be seen in Figure \ref{fig:metallica2}.
+%FIXME: ron, what di you mean: (not by cleaned up by EN's feature analysis)
 

### 2010-03-24 Paper/Paper.tex
figures moved around

@@
 \setlength{\abovecaptionskip}{-3pt}
-\setlength{\belowcaptionskip}{10pt}
+\setlength{\belowcaptionskip}{6pt}
 
@@
 of rhythmic knowledge about the song to align patches with beats.
-We can further group beats into bars, meaningful music segments.
+We can further group beats into bars to produces patches
+for meaningful music segments.
 The use of the Echo Nest API \cite{EchoNest} for that purpose is 
@@
 \fi
+
+\begin{figure}[thb]
+\begin{center}
+\includegraphics[width=.6\columnwidth]{code}
+\end{center}
+\caption{\small{A typical codeword from a codebook of size $200$ (code
+    7 in Figure \ref{fig:codes1}), corresponding to a
+    tonic-subdominant chord progression.  The patch is composed of $2$
+    bars and the pattern length was set to $8$ beats.  }}
+\label{fig:code}
+\end{figure}
 
@@
 The resulting patches consist of $12 \times 4$ or $12 \times 8$ matrices.
-
-\begin{figure}[thb]
-\begin{center}
-\includegraphics[width=.6\columnwidth]{code}
-\end{center}
-\caption{\small{A typical codeword from a codebook of size $200$ (code
-    7 in Figure \ref{fig:codes1}), corresponding to a
-    tonic-subdominant chord progression.  The patch is composed of $2$
-    bars and the pattern length was set to $8$ beats.  }}
-\label{fig:code}
-\end{figure}
 
@@
 
+\begin{figure}[htb]
+\begin{center}
+\includegraphics[width=.8\columnwidth]{data_sizes}
+\end{center}
+\caption{\small{Distortion for a codebook of size $100$ encoding one bar
+at a time, a bar being represented by $4$ chroma feature vector.
+Therefore, a codeword or a pattern is of size $12$ x $4$ = $48$.
+Distortion measure on the test data set. Data size tested range
+from $0$ (just initialization) to $500$ thousand. Patterns were selected at
+random from the dataset of approximately $3,7$ million patterns.
+}}
+\label{fig:data_sizes}
+\end{figure}
+
 \subsection{Basic Properties}
@@
 \end{itemize}
-
-\begin{figure}[htb]
-\begin{center}
-\includegraphics[width=.99\columnwidth]{data_sizes}
-\end{center}
-\caption{\small{Distortion for a codebook of size $100$ encoding one bar
-at a time, a bar being represented by $4$ chroma feature vector.
-Therefore, a codeword or a pattern is of size $12$ x $4$ = $48$.
-Distortion measure on the test data set. Data size tested range
-from $0$ (just initialization) to $500$ thousand. Patterns were selected at
-random from the dataset of approximately $3,7$ million patterns.
-}}
-\label{fig:data_sizes}
-\end{figure}
 
@@
 \begin{center}
-\includegraphics[width=.9\columnwidth]{encoding_per_beat}
+\includegraphics[width=.8\columnwidth]{encoding_per_beat}
 \end{center}
@@
 \begin{center}
-\includegraphics[width=.7\columnwidth]{freqs}
+\includegraphics[width=.6\columnwidth]{freqs}
 \end{center}
@@
 
+\begin{figure}[htb]
+\begin{center}
+\includegraphics[width=.9\columnwidth]{close_patterns1}
+\end{center}
+\caption{\small{Cluster around centroid presented in
+Figure \ref{fig:code}. Taken from the artist20 dataset, the cluster
+size is actually $639$. Shown samples were randomly selected.
+}}
+\label{fig:cluster}
+\end{figure}
+
 \subsection{Cluster}
@@
 %\ref{fig:cluster_diff} really add much?
-
-\begin{figure}[htb]
-\begin{center}
-\includegraphics[width=.9\columnwidth]{close_patterns1}
-\end{center}
-\caption{\small{Cluster around centroid presented in
-Figure \ref{fig:code}. Taken from the artist20 dataset, the cluster
-size is actually $639$. Shown samples were randomly selected.
-}}
-\label{fig:cluster}
-\end{figure}
 
@@
   \centering
-  \subfloat[][\small{Metallica}]{\label{fig:metallica2}\includegraphics[width=0.35\columnwidth]{metallica_pattern}}
+  \subfloat[][\small{Metallica}]{\label{fig:metallica2}\includegraphics[width=0.32\columnwidth]{metallica_pattern}}
   \hspace{5mm}                
-  \subfloat[][\small{Tori Amos / Suzanne Vega}]{\label{fig:amosvega2}\includegraphics[width=0.35\columnwidth]{toriamos_suzannevega_pattern}}
+  \subfloat[][\small{Tori Amos / Suzanne Vega}]{\label{fig:amosvega2}\includegraphics[width=0.32\columnwidth]{toriamos_suzannevega_pattern}}
   \caption{\small{Typical patterns for different artists.}}

### 2010-03-24 Paper/Paper.tex
tiny improvements

@@
 We use an online version of the vector quantization algorithm
-\cite{Gersho1991}, to cluster the beat-chroma patches described in the
+\cite{Gersho1991} to cluster the beat-chroma patches described in the
 previous section.
@@
 ($1$st line) and the absolute difference between both ($3$rd line).
-}
-ronw: does this figure really add all that much?
-}
+}}
 \label{fig:cluster_diff}
@@
 bar segmentation of new songs using the information contained in the
-cluster centroids.  The results are shown \ref{tab:offset}. We find
+cluster centroids.  The results are shown Table \ref{tab:offset}. We find
 the right segment with an accuracy of $62\%$ where random is
@@
 Suzanne Vega's pattern. These three artists were easily identified.
-
 Artists like Metallica are characterized by ``wideband'' patterns -

### 2010-03-24 Paper/Paper.tex
tiny improvements

@@
 \subsection{Cluster}
-Once we understand the kind of codebook we create, we naturally want
-to visualize the clusters, \ie the patterns from actual song that
-surround a given centroid. Unsurprisingly, real patterns tend to be
+Visualizing a codebook is half of the story, we also want to understand
+clusters, \ie the patterns from actual song that
+surround a given centroid (codeword). Unsurprisingly, real patterns are
 noisier than codes. However, centroids seem to capture important aspects
@@
 splitting some on two if the clusters become too small or too large.We could
-also cluster the codebook itself, in a similar fashion as hierarchical
+also cluster codebooks themselves, in a similar fashion as hierarchical
 Gaussian mixtures \cite{Vasconcelos2001}.

### 2010-03-24 Paper/Paper.tex
SUBMITTABLE :)

@@
 \begin{center}
-\includegraphics[width=.6\columnwidth]{freqs}
+\includegraphics[width=.99\columnwidth]{freqs}
 \end{center}
-\caption{\small{ Usage frequency of the patterns in the codebook from
+\caption{\small{Usage frequency of the patterns in the codebook from
 Figure \ref{fig:codes1} on the artist20 dataset.  There are $71,832$
-    patterns in total in the dataset.  Note that all codes were
-    initialized with a random sample from the training set and might
-    not be used in this test set.
-    FIXME: so why not show usage in training data?  I think test.mat
-    should contain this information.
+patterns in total in the dataset. Also frequency on the training set.
+Note that even though all codewords are initialized from sample, some
+are used only once (training set) or never (test set).
 }}

### 2010-03-24 Paper/Paper.tex
minor

@@
 
-% COMMENTING PREVIOUS WORK AND DEFINITIONS
-\iffalse
-\subsectionDefinitions}
-A pattern, or a patch, is a beat-chroma pattern taken from an actual song.
-A codeword (sometimes just code) is similar to a pattern, but is the result 
-of the clustering
-algorithm and most likely does not come from an actual song.
-Distortion is our measure of distance between a pattern and a code
-(or between two patterns, see equation \ref{eq:dist} in Section 
-\ref{sec:experiments}).
-
-\subsection{Previous Work}\label{sec:prevwork}
-Our work can be seen as an extension of the \textit{Shingles} idea
-described by Casey and Slaney \cite{Casey2006,Casey2007,Casey2008}.
-The two main differences are the clustering framework and the use of
-beats and bars to select the position of the \textit{Shingles}.
-
-Casey and Slaney \cite{Casey2006} use LSH \cite{Datar2004} to perform
-fast approximate nearest neighbor search. Their implementation is E$^2$LSH
-\cite{E2LSH}. This package can efficiently handle about $50$K or $100$K
-vectors of size $200$. Even with a better implementation, the LSH
-model still has some scaling problem. A large portion of the data
-must be known to decide how many projections to use, and what guarantee
-of finding the nearest neighbor it can provides. Also, we could fix the
-number of projections and consider as clusters all the elements that
-fall into the same bin. However, this does not give us a centroid, a
-typical pattern representing that cluster.
-\fi
-
 \begin{figure}[thb]
@@
 our training data and our main testing data. Some detailed settings
-of our algorithm are also provided. Then, we answer some common questions
-about any clustering algorithms, for instance 
-``how the number of codes affect the distortion''
-and ``how many learning samples should we use''.
-% Note that we will present two additional experiments in 
-% Section \ref{sec:exps2}.
+of our algorithm are also provided. As for any clustering algorithm, we
+measure the influence of the number of codewords and the training set size.
 
@@
 \end{figure}
-
-% REMOVING THE FOLLOWING SUBSECTION ON RATE - DISTORTION
-\iffalse
-\subsection{Music Randomness}\label{ssec:random}
-It will not surprise any reader, music is not random. This fact should
-become apparent in our experiments in the following way. Information
-theory tells us that if the signal to encode doubles, we need to square
-the size of the codebook to encode it with the same distortion.
-
-To help the intuition, here is an example. Imagine a binary array of 
-length $1$. We need two
-codes to encode perfectly: $(0)$ and $(1)$. If the binary array is now of
-length $2$, we now need four codes: $(0,0)$, $(0,1)$, $(1,1)$ and $(1,0)$.
-
-Since music is not random, we expect that squaring the codebook when the
-pattern size doubles will rather give improved results than constant ones.
-See Figure \ref{fig:size_pattern}.
-
-% see generate_expo_codes_curves.py to generate the following figure:
-\begin{figure}[htb]
-\begin{center}
-\includegraphics[width=.99\columnwidth]{codesize_patternsize}
-\end{center}
-\caption{\small{
-Experiment on squaring the codebook size when the pattern size doubles.
-Each line represent $4$ codebook sizes: $2$, $4$, $16$ and $256$.
-The pattern size range from $1$ for $1/4$ of a bar to $32$ for $8$ bars.
-If everything was random, each of these lines should be flat. The training
-size was fixed to $128$K samples.
-}}
-\label{fig:size_pattern}
-\end{figure}
-\fi
 

### 2010-03-24 Paper/Paper.tex
spacing in last figures caption improved
typo
typo in algorithm
artist20 now officially named

@@
 \FOR{$p \in \{P_n\}$}
-\STATE$c \leftarrow min_{c \in C_k} \dist(p,c)$
+\STATE$c \leftarrow \min_{c \in C_k} \dist(p,c)$
 \STATE$c \leftarrow c + (p - c) * \ell$
@@
 patterns in total in the dataset. Also frequency on the training set.
-Note that even though all codewords are initialized from sample, some
+Note that even though all codewords are initialized from samples, some
 are used only once (training set) or never (test set).
@@
 We apply our codebook to the artist recognition task.
-Our data set encompass $1402$ songs from $20$ artists, 
+Our data set, ``artist20'', encompass $1402$ songs from $20$ artists, 
 mostly rock and pop of different subgenres. Results exist in the
@@
   \subfloat[][\small{Tori Amos / Suzanne Vega}]{\label{fig:amosvega2}\includegraphics[width=0.32\columnwidth]{toriamos_suzannevega_pattern}}
+  \vspace{2mm}
   \caption{\small{Typical patterns for different artists.}}

### 2010-03-24 Paper/Paper.tex
figure 1 changed, no grid
dpwe's changes up to section 2.2

@@
 in combination with a compact representation of melodic-harmonic
-content, the beat-synchronous chromagram, and data-mining
+content (the beat-synchronous chromagram) and data-mining
 tools (clustering) to attempt to explicitly catalog this palette --
@@
 others.  Our interest here is in the tonal content of the music -- \ie the
-harmony and melody.  As a computationally convenient proxy for a richer
+harmony and melody.  As a computationally-convenient proxy for a richer
 description of the tonal content of audio, we use the popular Chroma
@@
 modulation of octave-invariant ``Shepard tones'', the melody and
-chord sequences of the original music typically remain recognizable
+chord sequences of the original music usually remain recognizable
 \cite{Ellis2007a}.  To the extent, then, that beat-chroma
-representations preserve tonal structure, they
+representations preserve tonal content, they
 are an interesting domain in which to search for patterns -- rich enough
 to generate musically-relevant results, but simplified enough to
-abstract away some other aspects of the original audio, such as
+abstract away aspects of the original audio such as
 instrumentation and other stylistic details.
@@
 chromagrams by learning codebooks from a large set of examples.
-The individual codewords consist of short beat-chroma patches, typically
+The individual codewords consist of short beat-chroma patches of
 between 1 and 8 beats, which are able to represent the entire dataset
-of millions of patterns with the minimum error given a small codebook of
+of millions of patterns with minimum error given a small codebook of
 a few hundred entries.  Our goal is to identify meaningful information
@@
 examining individual entries in this codebook.  Since the common
-patterns represent a higher-level description of the musical content,
+patterns represent a higher-level description of the musical content
+than the raw chroma,
 we also expect them to be useful in other applications, such as
@@
 One major issue in working with small patches concerns temporal
-alignment. Learning the same motifs with different offset
-(of a few seconds, or audio frames) is redundant. We take advantage
+alignment. Learning the same motifs with different offsets 
+(in time or beats) is redundant. We take advantage
 of rhythmic knowledge about the song to align patches with beats.
@@
 for meaningful music segments.
-The use of the Echo Nest API \cite{EchoNest} for that purpose is 
-explained in Subsection \ref{ssec:beatpatch}. Patches built around
-those temporal landmarks should capture music motifs more easily.
+Our use of the Echo Nest API \cite{EchoNest} for this purpose is 
+explained in Section \ref{ssec:beatpatch}. Patches built around
+these temporal landmarks should capture music motifs more easily.
 
 Prior work using small patches of chroma features includes
-the ``shingles'' of \cite{Casey2007}, which they used to
-identify ``remixes'' i.e. music based on some of the same
-underlying instrument tracks, as well as for matching
+the ``shingles'' of \cite{Casey2007}, which were used to
+identify ``remixes'', i.e., music based on some of the same
+underlying instrument tracks, and also for matching
 performances of Mazurkas \cite{Casey2008}.  That work,
@@
 beat-synchronous analysis includes \cite{Bartsch2001},
-who looked for repeated patterns within single songs to
-identify the chorus, and \cite{Ellis2007a}, who
+which looked for repeated patterns within single songs to
+identify the chorus, and \cite{Ellis2007a}, which 
 cross-correlated beat-chroma matrices to match cover
@@
 vector (length $12$) for every music event (called ``segment''), and a
-segmentation of the song into beats and bars. Beats span multiple
-segments, and bars span multiple beats.
+segmentation of the song into beats and bars. Beats may span or 
+subdivide segments; bars span multiple beats.
 %
@@
 beat-synchronous chroma feature representation similar to that used in
-\cite{Ellis2007a}.
+\cite{Ellis2007a}.  Echo Nest chroma vectors are normalized to have the 
+largest value in each column equal to 1.
 
@@
 
-Note that we do not claim that any of these informations (segments, beats, bars)
-are perfectly accurate (\eg in \cite{Barrington2009a} authors obtain a better
-song segmentation than the Echo Nest). 
-Practice showed us that they are reasonable, 
-and the size of the data set should make up for the imperfections or noise.
-We also believe that patches sizes based on a number of beats or bars are more
-meaningful than an arbitrary time length. More on this in the experiments
-(Subsection \ref{ssec:segment}).
+Note that none of this information (segments, beats, bars)
+can be assumed perfectly accurate.
+%(\eg in \cite{Barrington2009a} authors obtain a better
+%song segmentation than the Echo Nest). 
+In practice, we have found them reasonable, 
+and given the size of the data set, any rare imperfections or noise
+can be diluted to irrelevance by the good examples.  
+We also believe that patch sizes based on a number of beats or bars are more
+meaningful than an arbitrary time length. This is discussed further in 
+Section \ref{ssec:segment}.
 
@@
 %
-Because the bar length varies, depending on the meter of a particular
+Because the bar length is not guaranteed to be 4 beats, 
+depending on the meter of a particular
 song, we resample each patch to a fixed length of typically 4 beats
-per bar.  Since the overwhelming majority (82\%) of our training data
-consisted of bars that were 4 beats long, this resampling was not
-often required.  Most of the remaining bars (10\%) were 3 beats in
+per bar.  However, the majority (82\%) of our training data
+consisted of bars that were 4 beats long, so this resampling 
+usually had no effect.  Most of the remaining bars (10\%) were 3 beats in
 length.
@@
 
-Finally, we normalize the patches with respect to the key by rotating
+Finally, we normalize the patches with respect to transposition by rotating
 the pattern matrix so that the first row contains the most
-energy. This can easily be seen in Figure \ref{fig:code}.
+energy. This can be seen in the example codeword of Figure \ref{fig:code}.
 % We considered the same technique for rolling along the other axis and
@@
 % noise.
-Each patch within a song is normalized independently, so exact
+Each patch within a song is normalized independently, so
 reconstruction of the original song requires knowledge of the
@@
 % particular patterns occurs in,
-In the context of clustering this avoids \eg obtaining separate
+In the context of clustering this avoids problems such as obtaining separate
 clusters for every major triad in duple or triple meter.
@@
 %FIXME: Something about why we're not interested in specific chords?
-Specific chords do not add much information about a music collection.
-A cover song can be performed in a different key, but the motifs it
-contains remain the same.
+%
+%Specific chords do not add much information about a music collection.
+%A cover song can be performed in a different key, but the motifs it
+%contains remain the same.
+%%%% dpwe: I don't agree with this, I'd rather say nothing
 

### 2010-03-24 Paper/Paper.tex
Dan edits added
changes through section 3.2

@@
 in the codebook and updates the cluster centroid (codeword) to be
-closer to the sample based on some learning rate $\ell$.
+closer to the sample, using a learning rate parameter $\ell$.
 The clusters are updated as each data point is seen, as opposed to
 once per iteration in the standard k-means algorithm.
-The details are explained in Algorithm \ref{algo:vq}.
+The details can be found in Algorithm \ref{algo:vq}.
 %
@@
 \section{Experiments}\label{sec:experiments}
-In this section we present different clustering experiments. We introduce
-our training data and our main testing data. Some detailed settings
-of our algorithm are also provided. As for any clustering algorithm, we
-measure the influence of the number of codewords and the training set size.
+
+% In this section we present different clustering experiments. We introduce
+% our training data and our main testing data. Some detailed settings
+% of our algorithm are also provided. As for any clustering algorithm, we
+% measure the influence of the number of codewords and the training set size.
+In this section we present a set of experiments designed to evaluate
+quantization quality as a function of size of training set, codebook
+size, and patch length.
 
@@
 \label{sec:testdata}
-Our training data consists of $43,300$ tracks that were uploaded to
+The training data consists of $43,300$ tracks that were uploaded to
 morecowbell.dj,\footnote{http://www.morecowbell.dj/} an online service
-based on the Echo Nest analyze API which remixes uploaded music by
+based on the Echo Nest analyze API which remixes music by
 adding cowbell and other sound effects synchronized in time with the
@@
 
-For testing, we made use of low quality (32kbps) version of the songs
-from the uspop2002 data set \cite{uspop2002}.  This well-known data
-contains a wide variety of pop songs, but is not particular otherwise.
+For testing, we made use of low quality (32kbps MP3) audio 
+from the uspop2002 data set \cite{uspop2002}.  The collections
+includes a wide variety of popular music from 10 different sub-genres.
 % \footnote{We also tried to use Tzanetakis dataset,
@@
 % most did not contain any bar.}.
-uspop2002 serves as test set to measure how well a codebook learned on
-the Cowbell data set can represent new songs.  We obtained Echo Nest features
-for the $8651$ songs contained in the dataset.
+We use this data to measure how well a codebook learned on
+the Cowbell data set can represent new songs.  
+We obtained Echo Nest features for the $8651$ songs contained in the data set.
 
 \subsection{Setting}\label{ssec:setting}
-We take one or two bars and normalize the patches to size 4 or 8
-respectively.  We learn a codebook of size $K$ over the Cowbell
-dataset using the VQ algorithm (Algorithm \ref{algo:vq}). We used a
-learning rate of $\ell=0.01$ for $200$ iterations over the whole
-dataset.
+We train codebooks based on beat-chroma patches composed of 4 or 8
+beats (one or two bars).
+Codebooks over the Cowbell
+data set are trained using the VQ algorithm (Algorithm \ref{algo:vq}). We used a
+learning rate of $\ell=0.01$ for $200$ iterations over the data.
 %
-We use the codebook to encode the test set.
-Each pattern is encoded with only one code. We can measure the average
-distance between a pattern and its encoding. We can also measure the use
-of the different codes, \ie the size of the cluster around a code.
+Given these codebooks, we quantize the test set by assigning each
+patch to a single codeword.  We report the average
+distance between patches and their encodings and describe the usage
+frequency of the different codes. %, \ie the size of the cluster around a code.
 
@@
 measure between chroma patches.  Given a pattern $p_1$
-composed of elements $p_{1ij}$, and a similar pattern $p_2$, the
+composed of elements $p_{1ij}$, and a second pattern $p_2$, the
 distance between them is:
@@
 \end{eqnarray}
-We assume $p_1$ and $p_2$ have the same size.  This is enforced by the
+We assume that $p_1$ and $p_2$ have the same size.  This is enforced by the
 meter normalization procedure described in Section \ref{sec:approach}.

### 2010-03-24 Paper/Paper.tex
more small tweaks

@@
 \begin{center}
-\includegraphics[width=.6\columnwidth]{code}
+\includegraphics[width=.4\columnwidth]{code}
 \end{center}
@@
 perfect fifth (codewords 1 and 2) and fourth intervals (codewords 3
-and 6), and a
+and 6), and 
 major and minor triads (codewords 5 and 4, respectively).  Many of the remaining
@@
 Locally linear embedding (LLE) visualization of the codebook.
-LLE tries to keep near neighbors (here according to euclidean distance)
+LLE tries to keep near neighbors (here according to Euclidean distance)
 in the same $2$D neighborhood. Shown patterns were randomly selected
@@
 \caption{\small{
-Song segmentation experiment:
-best offset according to our codebook. The real one is $0$.
-Random guess gives an accuracy of $25$\%.
+Results from the song segmentation experiment.
+The correct offset is $0$.
+Baseline performance by choosing the offset randomly gives an accuracy of $25$\%.
 }}
@@
 the song histograms of that artist. The test song is compared to
-the histogram of each artist using euclidean distance.
+the histogram of each artist using Euclidean distance.
 
@@
 Suzanne Vega's pattern. These three artists were easily identified.
-Artists like Metallica are characterized by ``wideband'' patterns -
-featured muddled due to noisy acoustics - lots of transients,
+Artists such as Metallica are characterized by ``wideband'' patterns,
+containing features muddled by the noisy acoustics containing many
+transients,
 distortion, etc, which can be seen in Figure \ref{fig:metallica2}.
@@
 could be use to develop better distance measures between patterns.
-The use of the euclidean distance is arbitrary. Summarizing patches
+The use of the Euclidean distance is arbitrary. Summarizing patches
 with Gaussians, and then comparing the distance between those Gaussians,
@@
 Moving on to larger scales, we could imagine merging codes together and
-splitting some on two if the clusters become too small or too large.We could
+splitting some on two if the clusters become too small or too large.  We could
 also cluster codebooks themselves, in a similar fashion as hierarchical

### 2010-03-24 Paper/Paper.tex
clustering intro slightly improved

--- 2010-03-24 Paper/Paper.tex
@@
 \subsection{Cluster}
-Visualizing a codebook is half of the story, we also want to understand
-clusters, \ie the patterns from actual song that
-surround a given centroid (codeword). Unsurprisingly, real patterns are
-noisier than codes. However, centroids seem to capture important aspects
-of the patterns. We could compare the centroids with the result of
-a lowpass filtering. In Figure \ref{fig:cluster} we show elements of the same
+Clusters are formed around a centroid (codeword) by all patterns closer
+to that codeword than any other. Since the centroid is somewhat an
+average of those patterns, the latter tend to be noisier. Hence, codewords
+can be seen as the output of a lowpass filtering mecanism.
+In Figure \ref{fig:cluster} we show elements of one given
 cluster, and in Figure \ref{fig:cluster_diff}, we emphasize the difference

### 2010-03-24 Paper/Paper.tex
dataset now always in one word

@@
 In practice, we have found them reasonable, 
-and given the size of the data set, any rare imperfections or noise
+and given the size of the dataset, any rare imperfections or noise
 can be diluted to irrelevance by the good examples.  
@@
 For testing, we made use of low quality (32kbps MP3) audio 
-from the uspop2002 data set \cite{uspop2002}.  The collections
+from the uspop2002 dataset \cite{uspop2002}.  The collections
 includes a wide variety of popular music from 10 different sub-genres.
@@
 We use this data to measure how well a codebook learned on
-the Cowbell data set can represent new songs.  
-We obtained Echo Nest features for the $8651$ songs contained in the data set.
+the Cowbell dataset can represent new songs.  
+We obtained Echo Nest features for the $8651$ songs contained in the dataset.
 
@@
 Codebooks over the Cowbell
-data set are trained using the VQ algorithm (Algorithm \ref{algo:vq}). We used a
+dataset are trained using the VQ algorithm (Algorithm \ref{algo:vq}). We used a
 learning rate of $\ell=0.01$ for $200$ iterations over the data.
@@
 Therefore, a codeword or a pattern is of size $12$ x $4$ = $48$.
-Distortion measure on the test data set. Data size tested range
+Distortion measure on the test dataset. Data size tested range
 from $0$ (just initialization) to $500$ thousand. Patterns were selected at
@@
 We apply our codebook to the artist recognition task.
-Our data set, ``artist20'', encompass $1402$ songs from $20$ artists, 
+Our dataset, ``artist20'', encompass $1402$ songs from $20$ artists, 
 mostly rock and pop of different subgenres. Results exist in the

### 2010-03-24 Paper/Paper.tex
slightly better explanation of song seg exp
slight improvement

@@
 Echo Nest segmentation. We then encode each of these sequences using an offset
-of $0$, $1$, $2$ or $3$ beats. The best offset for a test sequence
-(the one with the lowest distortion)
-should correspond to the bar alignment in the original
-feature matrix.  The correct answer is an offset of $0$.
+of $0$, $1$, $2$ or $3$ beats. The offset on a test sequence
+with the lowest distortion is our predicted alignment.
+The correct answer is an offset of $0$.
 
@@
 
-Once again, the clustering method proposed in this work is not
+The clustering method proposed in this work is not
 intended to perform well at artist recognition. An obvious fact is that

### 2010-03-25 Paper/Paper.tex
slight improvement
application intro redone

@@
 \section{Applications}\label{sec:exps2}
-In this section we describe two other applications of the clustering
-described in this paper: song segmentation, and artist
-recognition.  This demonstrates the utility of unsupervised clustering
-to supervised tasks. The representation created from large collections
-captures meaningful information without being forced to do so.
+In this section we apply the result of our method to two task
+not related to clustering. We try to demonstrate that the codewords
+we learned capture musically meaningful patterns, at least to
+contemporary pop music. This also demonstrates the utility of unsupervised 
+clustering to other MIR tasks usually tackled in a supervised
+framework. In the first experiment, we show that bar information is retained 
+in the codewords. Thus we can discover the bar boundaries for new songs.
+The second experiment shows we can discriminate artists
+based on the encoding of their songs, hence we can associate artists
+with specific codes.
+
 % Although our algorithm was not devised for any of these
@@
 The clustering method proposed in this work is not
-intended to perform well at artist recognition. An obvious fact is that
-we lose all timbral information which would be useful. However, we can
+intended to perform well at artist recognition. 
+In particular,
+we lose all timbral information which would be useful. 
+However, we can
 imagine that some artist use more often some patterns than others.

### 2010-03-25 Paper/Paper.tex
slight improvement

@@
 
-Figure \ref{fig:encodesong} shows a typical example of encoding a song
+Figure \ref{fig:encodesong} shows an example of a song encoded
 using the codebook described in Section \ref{sec:codebook}.  Both the
@@
 %
-In this section we investigate whether it is possible to recover the
+We investigate whether it is possible to recover the
 bar segmentation of new songs using the information contained in the
@@
 literature for that set \cite{Ellis2007}. The best accuracy is
-$59\%$, whereas using only chroma as a representation yields an
-accuracy of $33\%$. The latter results is obtained through a
-Gaussian mixture model.
+$59\%$, whereas using only chroma and a Gaussian mixture model
+yields an accuracy of $33\%$.
 

### 2010-03-25 Paper/Paper.tex
slight improvement
Dan stuff added

@@
   {Third author} {Affiliation3 \\ {\tt author3@ismir.edu}}
-% what order do we use? Thierry, Ron, Dan?  Dan, Thierry, Ron?
+% Thierry, Ron, Dan
 
@@
 \subsection{Clustering}
-
 % We can then stack beat vectors for a number of bars (usually 1 or 2).
@@
 in the codebook and updates the cluster centroid (codeword) to be
-closer to the sample, using a learning rate parameter $\ell$.
+closer to the sample according to a learning rate $\ell$.
 The clusters are updated as each data point is seen, as opposed to
-once per iteration in the standard k-means algorithm.
-The details can be found in Algorithm \ref{algo:vq}.
+once per iteration in the standard $k$-means algorithm.
+The details are explained in Algorithm \ref{algo:vq}.
 %
-As in standard k-means clustering, the codebook is initialized by
+As in standard $k$-means clustering, the codebook is initialized by
 choosing $K$ random points from our dataset.
@@
 the number of patches seen and can be interrupted at any time
-while still resulting in an updated codebook. %not that this is important...
+to obtain an updated codebook. %not that this is important...
 % The setting of the learning rate might become trickier, but this is
@@
 \section{Experiments}\label{sec:experiments}
-
-% In this section we present different clustering experiments. We introduce
-% our training data and our main testing data. Some detailed settings
-% of our algorithm are also provided. As for any clustering algorithm, we
-% measure the influence of the number of codewords and the training set size.
-In this section we present a set of experiments designed to evaluate
-quantization quality as a function of size of training set, codebook
-size, and patch length.
+In this section we present different clustering experiments and introduce
+our principal training and test data. Some detailed settings
+of our algorithm are also provided. As for any clustering algorithm, we
+measure the influence of the number of codewords and the training set size.
 
@@
 \label{sec:testdata}
-The training data consists of $43,300$ tracks that were uploaded to
+Our training data consists of $43,300$ tracks that were uploaded to
 morecowbell.dj,\footnote{http://www.morecowbell.dj/} an online service
-based on the Echo Nest analyze API which remixes music by
+based on the Echo Nest analyze API which remixes uploaded music by
 adding cowbell and other sound effects synchronized in time with the
-music.  The $43.3$K songs contain $\mathbf{3.7}$ million non silent
-bars which we cluster using the approach described in the previous
+music.  The $43.3$K songs contain $3.7$ million non-silent
+bars which we clustered using the approach described in the previous
 section.
 
-For testing, we made use of low quality (32kbps MP3) audio 
-from the uspop2002 dataset \cite{uspop2002}.  The collections
-includes a wide variety of popular music from 10 different sub-genres.
+For testing, we made use of low quality (32kbps, 8 kHz bandwidth mono MP3) versions of the songs
+from the {\it uspop2002} data set \cite{uspop2002}.  This well-known data set 
+contains pop songs from a range of artists and styles.
 % \footnote{We also tried to use Tzanetakis dataset,
@@
 % most did not contain any bar.}.
-We use this data to measure how well a codebook learned on
-the Cowbell dataset can represent new songs.  
-We obtained Echo Nest features for the $8651$ songs contained in the dataset.
+{\it uspop2002} serves as test set to measure how well a codebook learned on
+the Cowbell data set can represent new songs.  We obtained Echo Nest features
+for the $8651$ songs contained in the dataset.
 
 \subsection{Setting}\label{ssec:setting}
-We train codebooks based on beat-chroma patches composed of 4 or 8
-beats (one or two bars).
-Codebooks over the Cowbell
-dataset are trained using the VQ algorithm (Algorithm \ref{algo:vq}). We used a
-learning rate of $\ell=0.01$ for $200$ iterations over the data.
+We take one or two bars and resample the patches to 4 or 8 columns 
+respectively.  We learn a codebook of size $K$ over the Cowbell
+dataset using the online VQ algorithm (Algorithm \ref{algo:vq}). We use a
+learning rate of $\ell=0.01$ for $200$ iterations over the whole
+dataset.
 %
-Given these codebooks, we quantize the test set by assigning each
-patch to a single codeword.  We report the average
-distance between patches and their encodings and describe the usage
-frequency of the different codes. %, \ie the size of the cluster around a code.
+We then use the resulting codebook to encode the test set.
+Each pattern is encoded with only one code. We can measure the average
+distance between a pattern and its encoding. We can also measure the use
+of the different codes, 
+%\ie the size of the cluster around a code.
+i.e., the proportion of patterns that quantize to each code.
 
@@
 measure between chroma patches.  Given a pattern $p_1$
-composed of elements $p_{1ij}$, and a second pattern $p_2$, the
+composed of elements $p_1(i,j)$, and a similar pattern $p_2$, the
 distance between them is:
 \begin{eqnarray}
-  \dist(p_1,p_2) = \sum_{i,j} \frac{(p_{1ij} - p_{2ij})^2}{\size(p_1)}
+  \dist(p_1,p_2) = \sum_{i,j} \frac{(p_1(i,j) - p_2(i,j))^2}{\size(p_1)}
   \label{eq:dist}
 \end{eqnarray}
-We assume that $p_1$ and $p_2$ have the same size.  This is enforced by the
-meter normalization procedure described in Section \ref{sec:approach}.
+We assume $p_1$ and $p_2$ have the same size.  This is enforced by the
+resampling procedure described in Section \ref{sec:approach}.
 
-\begin{figure}[htb]
+\begin{figure}[t]
 \begin{center}
@@
 \caption{\small{Distortion for a codebook of size $100$ encoding one bar
-at a time, a bar being represented by $4$ chroma feature vector.
-Therefore, a codeword or a pattern is of size $12$ x $4$ = $48$.
-Distortion measure on the test dataset. Data size tested range
-from $0$ (just initialization) to $500$ thousand. Patterns were selected at
-random from the dataset of approximately $3,7$ million patterns.
+at a time with by $4$ columns.
+Therefore, each codeword has $12 \times 4 = 48$ elements.
+Distortion is measurde on the test set.  Training data sizes range
+from $0$ (just initialization) to $500,000$. Patterns were selected at
+random from the dataset of approximately $3.7$ million patterns.
 }}
@@
 
-\subsection{Basic Properties}
-This section present some fundamental results of the clustering algorithm.
-None of the result appear surprising,  but they are important for 
-reproducing the experiments.
+\subsection{Codebook properties}
+This section presents some basic results of the clustering.
+While unsurprising, these results may be useful for comparison 
+when reproducing this work.
 \begin{itemize}
 \item Encoding performance improves with increasing training data (Figure
-\ref{fig:data_sizes}). Regarding the convergence of the performance,
-we can imagine using $500$ or $1$K samples per codeword to learn if the
-training set size is problematic.
+\ref{fig:data_sizes}). Distortion improvements plateau by around $1000$ samples 
+per codeword ($100,000$ samples for the $100$-entry codebook of the figure).
 \item Encoding performance improves with increasing codebook size
-(Table \ref{tab:cbsize}). The increase seems significant, and
-increasing the size of the codebook in an efficient way
-is definitely part of our future work.
+(Table \ref{tab:cbsize}).  Computation costs scales with codebook size, 
+which limited the largest codebooks used in this work, but larger codebooks 
+(and more efficient algorithms to enable them) are clearly a promising future 
+direction.
 \item Larger patterns are more difficult to encode, thus requiring
-larger codebooks. See Figure \ref{fig:perbeat}. The increase is steeper
-until $4$ beats which usually corresponds to one bar.
+larger codebooks. See Figure \ref{fig:perbeat}. The increase is steepest
+below $4$ beats (1 bar), although there is no dramatic change at 
+this threshold.
 \end{itemize}
@@
 \end{center}
-\caption{\small{Varying codebook size using $50,000$ training samples.
-    We train on patterns corresponding to $1$ bar resized to $4$ beats.
+\caption{\small{Distortion as a function of codebook size for a fixed training set 
+of $50,000$ samples.  Codebook consists of $1$ bar ($4$ beat) patterns.
   }}
@@
 
-\begin{figure}[htb]
+\begin{figure}[b]
 \begin{center}
@@
 \end{center}
-\caption{\small{Encoding larger and larger patterns with a fixed size
-codebook. The size of the pattern is defined by the number of beats,
-the bar information was not taken into account for this experiment.
+\caption{\small{Encoding patterns of different sizes with a fixed size
+codebook of 100 patterns. The size of the pattern is defined by the number of beats. 
+Downbeat (bar alignment) information was not used for this experiment. % why not?
 }}
@@
 
-We train a codebook of size $200$ over patterns of size $12$x$8$ covering
-$2$ bars at a time.  The results shown are on the test set from
+We trained a codebook $200$ entries of patterns sized $12\times8$, covering
+$2$ bars at a time.  The results shown are on the {\it artist20} test set described in 
 Section \ref{ssec:artist}.
@@
 Figure \ref{fig:codes1}.  
-The frequency of use of these codewords are shown in Figure \ref{fig:freqs}.
+The frequency of use of these codewords is shown in Figure \ref{fig:freqs}.
 The codewords primarily consist of sustained
@@
 perfect fifth (codewords 1 and 2) and fourth intervals (codewords 3
-and 6), and 
-major and minor triads (codewords 5 and 4, respectively).  Many of the remaining
+and 6, noting that the fourth occurs when the per-pattern transposition 
+detects the fifth rather than the root as the strongest chroma bin, and vice-versa), and a
+major triads transposed to the root and fifth (codewords 5 and 4, respectively).  Many of the remaining
 codewords correspond to common transitions from one chord to another,
-\eg a tonic-subdominant transition in codes 7 and 9
-(Cmaj
-$\rightarrow$ Fmaj or C5 $\rightarrow$ F5 as a guitar power chord)
+\eg a V-I transition in codes 7 and 9
+(e.g., Gmaj
+$\rightarrow$ Cmaj or G5 $\rightarrow$ C5 as a guitar power chord)
 % assuming that chroma 0 is C) 
-and the reverse tonic-dominant transition in code 21 (\ie Fmaj
-$\rightarrow$ Cmaj).
+and the reverse I-V transition in code 21 (e.g., Cmaj
+$\rightarrow$ Gmaj).
 %FIXME: Does this transition have a better name like ``relative x''?
@@
 
-\begin{figure}[h]
+\begin{figure}[t]
 \begin{center}
@@
 \caption{\small{
-25 most frequently used codes in the codebook %in term of frequency of use
-in the artist20 test set. 
-This codebook contains $200$ codes trained on $2$ bar, $12 \times 8$
-patches.
-The numbers 
-%above each panel correspond to the display are the index (from most used to least) and
-in parentheses correspond to 
-the percentage of use in the test set.
+The 25 codes that are 
+most commonly used for the {\it artist20} test set. 
+Codes are from the $200$-entry codebook trained on $2$ bar, $12 \times 8$
+patches.  The proportion of patches accounted for by each pattern is shown 
+in parentheses.
 %There are $71832$ patterns in total in the dataset.
@@
 
-\begin{figure}[htb]
+\begin{figure}[bt]
 \begin{center}
@@
 \end{center}
-\caption{\small{Usage frequency of the patterns in the codebook from
-Figure \ref{fig:codes1} on the artist20 dataset.  There are $71,832$
-patterns in total in the dataset. Also frequency on the training set.
+\caption{\small{Usage proportions for all 200 codewords 
+on the {\it artist20} test set (which comprises $71,832$
+patterns).  Also shown are the usage proportions for the training set (``cowbell''), 
+which are similar.
 Note that even though all codewords are initialized from samples, some
-are used only once (training set) or never (test set).
+are used only once in the training set, or not at all for test set.
 }}
@@
 using the codebook described in Section \ref{sec:codebook}.  Both the
-Echo Nest chroma features for an example song and the corresponding
+Echo Nest chroma features for the example song and the corresponding
 quantization are displayed.  The quantized representation retains

### 2010-03-25 Paper/Paper.tex
line added about initialization
typo
expanded on LLE

@@
 choosing $K$ random points from our dataset.
+% ADDED BY TBM AFTER DAN
+Initialization had no significant effect in our experiments.
 %
@@
 Therefore, each codeword has $12 \times 4 = 48$ elements.
-Distortion is measurde on the test set.  Training data sizes range
+Distortion is measured on the test set.  Training data sizes range
 from $0$ (just initialization) to $500,000$. Patterns were selected at
@@
 
-Another interesting visualization uses LLE\footnote{implementation:
-  \url{http://www.astro.washington.edu/users/vanderplas/coding/LLE/}}
-algorithm \cite{Roweis2000} to display codewords arranged in
-coherent groups. See Figure \ref{fig:lle}. For instance, we see
-``wideband'' patterns grouping to the right.
+Figure \ref{fig:lle} visualizes the codebook using locally linear
+embedding \cite{Roweis2000} to group near neighbors on a 2 dimensional
+plane\footnote{We used the implementation from:
+  \url{http://www.astro.washington.edu/users/vanderplas/coding/LLE/}}.
+This visualization highlights different groups of codewords: sustained
+chords on the top left, chord changes on the bottom left, and more
+complex sustained chords and ``wideband'' noisy patterns grouping to
+the right.
 
@@
 \end{center}
-\caption{\small{
-Locally linear embedding (LLE) visualization of the codebook.
-LLE tries to keep near neighbors (here according to Euclidean distance)
-in the same $2$D neighborhood. Shown patterns were randomly selected
-in each neighborhood.
+\caption{\small{ Locally linear embedding (LLE) visualization of the
+    codebook described in Section \ref{sec:visu}.  Each codeword is
+    represented by a point on the scatter plot.  Randomly selected
+    codewords from the embedded space are displayed over the plot.
+    %FIXME: does this figure really provide any new insight?
   }}

### 2010-03-25 Paper/Paper.tex
added write up of figure 7

@@
 
-Figure \ref{fig:lle} visualizes the codebook using locally linear
-embedding \cite{Roweis2000} to group near neighbors on a 2 dimensional
-plane\footnote{We used the implementation from:
+An alternate visualization of the codebook is shown in Figure
+\ref{fig:lle}.  We use locally linear embedding \cite{Roweis2000} to
+group near neighbors on a 2 dimensional plane\footnote{We used the
+  implementation from:
   \url{http://www.astro.washington.edu/users/vanderplas/coding/LLE/}}.
@@
 complex sustained chords and ``wideband'' noisy patterns grouping to
-the right.
+the right of the figure.
 
-Finally, more related to our particular data, we try to assess how many
-sustained patterns are represented in the codebook. We look at
-the average variance over time in Figure \ref{fig:code_var}.
-25\% of the codebook correspond
-to very simple sustained patterns similar to the top row of Figure
-\ref{fig:codes1}. For $2$ bars codewords, it seems that patterns
+Given the prevalence of simple codewords in Figures \ref{fig:codes1}
+and \ref{fig:lle}, we assess the proportion of the codebook devoted to
+representing sustained chord patterns.  The average variance over time
+within each codeword is plotted in Figure \ref{fig:code_var}.  From
+this figure it is clear that 25\% of the codebook corresponds to very
+simple sustained patterns similar to the top row of Figure
+\ref{fig:codes1}. For $2$ bar codewords, it seems that patterns
 containing more complex temporal activity account for only $33$\% of
@@
 average of those patterns, the latter tend to be noisier. Hence, codewords
-can be seen as the output of a lowpass filtering mecanism.
+can be seen as the output of a lowpass filtering mechanism.
 In Figure \ref{fig:cluster} we show elements of one given

### 2010-03-25 Paper/Paper.tex
fixed figure ordering
used small refs

@@
 chord sequences of the original music usually remain recognizable
-\cite{Ellis2007a}.  To the extent, then, that beat-chroma
+\cite{Ellis2007a_small}.  To the extent, then, that beat-chroma
 representations preserve tonal content, they
@@
 Prior work using small patches of chroma features includes
-the ``shingles'' of \cite{Casey2007}, which were used to
+the ``shingles'' of \cite{Casey2007_small}, which were used to
 identify ``remixes'', i.e., music based on some of the same
@@
 use either beat- or bar-synchronous features).  Earlier work in
-beat-synchronous analysis includes \cite{Bartsch2001},
+beat-synchronous analysis includes \cite{Bartsch2001_small},
 which looked for repeated patterns within single songs to
-identify the chorus, and \cite{Ellis2007a}, which 
+identify the chorus, and \cite{Ellis2007a_small}, which 
 cross-correlated beat-chroma matrices to match cover
@@
 beat-synchronous chroma feature representation similar to that used in
-\cite{Ellis2007a}.  Echo Nest chroma vectors are normalized to have the 
+\cite{Ellis2007a_small}.  Echo Nest chroma vectors are normalized to have the 
 largest value in each column equal to 1.

### 2010-03-25 Paper/Paper.tex
prettier equation
new shorter conclusion

@@
 \begin{eqnarray}
-  \dist(p_1,p_2) = \sum_{i,j} \frac{(p_1(i,j) - p_2(i,j))^2}{\size(p_1)}
+  \dist(p_1,p_2) = \sum_{i,j} \frac{\left(p_1(i,j) - p_2(i,j)\right)^2}{\size(p_1)}
   \label{eq:dist}
@@
 \begin{center}
-\includegraphics[width=.9\columnwidth]{close_patterns1}
+\includegraphics[width=.8\columnwidth]{codes_8bars}
 \end{center}
-\caption{\small{Cluster around centroid presented in
-Figure \ref{fig:code}. Taken from the artist20 dataset, the cluster
-size is actually $639$. Shown samples were randomly selected.
-}}
-\label{fig:cluster}
+\caption{\small{Sample of longer codewords spanning over $8$
+    bars. Codewords selected randomly from the codewords used in the
+    test set. Percentage of use display in parentheses.  Most patterns
+    consist of sustained note or chords, but code 0 shows a repeating
+    two chord pattern and code 4 contains two bars of a
+    1$\rightarrow$1$\rightarrow$1$\rightarrow$2 progression.}}
+\label{fig:codes8bars}
 \end{figure}
@@
 \begin{center}
+\includegraphics[width=.9\columnwidth]{close_patterns1}
+\end{center}
+\caption{\small{Cluster around centroid presented in
+Figure \ref{fig:code}. Taken from the artist20 dataset, the cluster
+size is actually $639$. Shown samples were randomly selected.
+}}
+\label{fig:cluster}
+\end{figure}
+
+\begin{figure}[htb]
+\begin{center}
 \includegraphics[width=.8\columnwidth]{close_patterns_diff}
@@
 \label{fig:cluster_diff}
-\end{figure}
-
-\begin{figure}[htb]
-\begin{center}
-\includegraphics[width=.8\columnwidth]{codes_8bars}
-\end{center}
-\caption{\small{Sample of longer codewords spanning over $8$
-    bars. Codewords selected randomly from the codewords used in the
-    test set. Percentage of use display in parentheses.  Most patterns
-    consist of sustained note or chords, but code 0 shows a repeating
-    two chord pattern and code 4 contains two bars of a
-    1$\rightarrow$1$\rightarrow$1$\rightarrow$2 progression.}}
-\label{fig:codes8bars}
 \end{figure}
@@
 \section{Applications}\label{sec:exps2}
-In this section we apply the result of our method to two task
-not related to clustering. We try to demonstrate that the codewords
+In this section we apply the result of our method to two tasks
+unrelated to clustering. We try to demonstrate that the codewords
 we learned capture musically meaningful patterns, at least to
@@
 
-As for enhancements, we specifically did not compress our features using
-Gaussian mixtures or other generative model. That being said, these methods
-could be use to develop better distance measures between patterns.
-The use of the Euclidean distance is arbitrary. Summarizing patches
-with Gaussians, and then comparing the distance between those Gaussians,
-could reduce the influence of the noise in the distance measure.
+The main future direction is larger scales. This implies $3$ improvements.
+First, gather more songs from the Echo Nest API. Second, efficiently work
+with larger codebooks. Some improvements have been published for $k$-means
+that avoid comparing every code to every sample. Third, the learning
+algorithm must not converge too fast. We could add a filter that selects
+increasingly hard example to encode, our merge many codebooks in a similar
+fashion as hierarchical Gaussian mixtures \cite{Vasconcelos2001}.
 
-Moving on to larger scales, we could imagine merging codes together and
-splitting some on two if the clusters become too small or too large.  We could
-also cluster codebooks themselves, in a similar fashion as hierarchical
-Gaussian mixtures \cite{Vasconcelos2001}.
+%As for enhancements, we specifically did not compress our features using
+%Gaussian mixtures or other generative model. That being said, these methods
+%could be use to develop better distance measures between patterns.
+%The use of the Euclidean distance is arbitrary. Summarizing patches
+%with Gaussians, and then comparing the distance between those Gaussians,
+%could reduce the influence of the noise in the distance measure.
+
+%Moving on to larger scales, we could imagine merging codes together and
+%splitting some on two if the clusters become too small or too large.  We could
+%also cluster codebooks themselves, in a similar fashion as hierarchical
+%Gaussian mixtures \cite{Vasconcelos2001}.
 %  (similar idea used for indexing \cite{Muja2009})

### 2010-03-25 Paper/Paper.tex
minor
conclusion improved

@@
 % ADDED BY TBM AFTER DAN
-Initialization had no significant effect in our experiments.
+Initialization had no significant effect on the distortion
+in our experiments.
 %
@@
 with larger codebooks. Some improvements have been published for $k$-means
-that avoid comparing every code to every sample. Third, the learning
+that avoid comparing every code to every sample. Initialization
+should also be modified, so every codeword is used (unlike in
+Figure \ref{fig:freqs}). Third, the learning
 algorithm must not converge too fast. We could add a filter that selects
-increasingly hard example to encode, our merge many codebooks in a similar
+increasingly hard example to encode, or merge many codebooks in a similar
 fashion as hierarchical Gaussian mixtures \cite{Vasconcelos2001}.

### 2010-03-25 Paper/Paper.tex
reworked bar alignment into intro

--- 2010-03-25 Paper/Paper.tex
@@
 The individual codewords consist of short beat-chroma patches of
-between 1 and 8 beats, which are able to represent the entire dataset
+between 1 and 8 beats, optionally aligned to bar boundaries.
+%
+The additional temporal alignment eliminates redundancy that would be
+created by learning multiple codewords to represent the same motive at
+multiple beat offsets.
+%
+The codewords are able to represent the entire dataset
 of millions of patterns with minimum error given a small codebook of
-a few hundred entries.  Our goal is to identify meaningful information
+a few hundred entries.
+%
+Our goal is to identify meaningful information
 about the musical structure represented in the entire database by
@@
 music classification and retrieving tonally-related items.
-
-One major issue in working with small patches concerns temporal
-alignment. Learning the same motifs with different offsets 
-(in time or beats) is redundant. We take advantage
-of rhythmic knowledge about the song to align patches with beats.
-We can further group beats into bars to produces patches
-for meaningful music segments.
-Our use of the Echo Nest API \cite{EchoNest} for this purpose is 
-explained in Section \ref{ssec:beatpatch}. Patches built around
-these temporal landmarks should capture music motifs more easily.
 
@@
 Our dataset, ``artist20'', encompass $1402$ songs from $20$ artists, 
-mostly rock and pop of different subgenres. Results exist in the
-literature for that set \cite{Ellis2007}. The best accuracy is
-$59\%$, whereas using only chroma and a Gaussian mixture model
-yields an accuracy of $33\%$.
+mostly rock and pop of different subgenres.
+The best accuracy on the dataset in the literature is $59\%$, whereas
+using only chroma and a Gaussian mixture model yields an accuracy of
+$33\%$ \cite{Ellis2007}.
 

### 2010-03-25 Paper/Paper.tex
figure figure 1 placement
changed all figure placement to [t]
more rewording to sections 5 and 6

@@
 
-\begin{figure}[thb]
+\begin{figure}[t]
 \begin{center}
@@
 
-\begin{figure}[b]
+\begin{figure}[t]
 \begin{center}
@@
 
-\begin{figure}[bt]
+\begin{figure}[t]
 \begin{center}
@@
 
-\begin{figure}[htb]
+\begin{figure}[t]
 \begin{center}
@@
 \section{Applications}\label{sec:exps2}
-In this section we apply the result of our method to two tasks
-unrelated to clustering. We try to demonstrate that the codewords
-we learned capture musically meaningful patterns, at least to
-contemporary pop music. This also demonstrates the utility of unsupervised 
-clustering to other MIR tasks usually tackled in a supervised
-framework. In the first experiment, we show that bar information is retained 
-in the codewords. Thus we can discover the bar boundaries for new songs.
-The second experiment shows we can discriminate artists
-based on the encoding of their songs, hence we can associate artists
-with specific codes.
+In this section we investigate the application of beat-chroma
+codebooks to two tasks unrelated to clustering.
+%
+In the first experiment, we show that the codebooks retrain
+information related to bar alignment and can therefore be used to
+discover the bar boundaries for new songs.  The second experiment
+shows that we can discriminate artists based on the encoding of their
+songs, implying that some artists have a tendency to use different
+beat-chroma patterns than others.
 
@@
 
-\begin{table}
+\begin{table}[t]
 \begin{center}
@@
 
-\begin{figure}[htb]
+\begin{figure}[t]
 %\begin{center}
@@
 \iffalse
-\begin{figure}[htb]
+\begin{figure}[t]
 \begin{center}
@@
 
-\begin{figure}[htb]
+\begin{figure}[t]
   \centering
@@
 harmonic patterns. We assessed the basic properties of the method through
-experiments on a large collection of music. We suggested many ways
-to interpret the data (such as LLE) and assert the merit of the codebook
+experiments on a large collection of music, and 
+demonstrated that these codebooks capture musically meaningful
+patterns, at least to contemporary pop music.
+%
+ We suggested many ways
+to interpret the data %(such as LLE)
+and assert the merits of the codebook
 through further experiments.
 We discussed the possibility to move to even larger scales
-and we provided our source code.\footnote{Code not yet released to preserve
+and provided our source code.\footnote{Code not yet released to preserve
 submission's anonymity}
+%
+Finally, we demonstrated the utility of unsupervised clustering to MIR
+tasks that are typically tackled in a supervised setting.
 

### 2010-03-25 Paper/Paper.tex
conclusion slightly improved

@@
 
-The main future direction is larger scales. This implies $3$ improvements.
-First, gather more songs from the Echo Nest API. Second, efficiently work
-with larger codebooks. Some improvements have been published for $k$-means
-that avoid comparing every code to every sample. Initialization
+For even larger scales, we will focus on three facets.
+First, gather more songs from the Echo Nest API. 
+%
+Second,
+larger codebooks paired with computationally efficient algorithms.
+Improvements to $k$-means
+avoid comparing every code to every new sample. Initialization
 should also be modified, so every codeword is used (unlike in
-Figure \ref{fig:freqs}). Third, the learning
+Figure \ref{fig:freqs}). 
+%
+Third, the learning
 algorithm must not converge too fast. We could add a filter that selects
-increasingly hard example to encode, or merge many codebooks in a similar
-fashion as hierarchical Gaussian mixtures \cite{Vasconcelos2001}.
+increasingly difficult examples to encode, or merge many codebooks in a 
+similar fashion as hierarchical Gaussian mixtures \cite{Vasconcelos2001}.
 

### 2010-03-25 Paper/Paper.tex
typo

@@
 Since the clustering described in Section \ref{sec:approach} is based
-on the segmentation of the signal in to bars, it stands to reason that
+on the segmentation of the signal into bars, it stands to reason that
 the cluster centroids should contain information related to bar

### 2010-03-25 Paper/Paper.tex
typo

@@
 bar segmentation of new songs using the information contained in the
-cluster centroids.  The results are shown Table \ref{tab:offset}. We find
+cluster centroids.  The results are shown in Table \ref{tab:offset}. We find
 the right segment with an accuracy of $62\%$ where random is
@@
 all possible bar length (in terms of beats) and not restrict ourselves
-to $4$ beats bars and offsets between $0$ and $3$. This is part of
+to bars of $4$ beats and offsets between $0$ and $3$. This is part of
 our future work.

### 2010-03-25 Paper/Paper.tex
minor
MAJOR fnal version of Dan

--- 2010-03-25 Paper/Paper.tex
@@
   {Third author} {Affiliation3 \\ {\tt author3@ismir.edu}}
-% Thierry, Ron, Dan
+% what order do we use? Thierry, Ron, Dan?  Dan, Thierry, Ron?
 
@@
 chord sequences of the original music usually remain recognizable
-\cite{Ellis2007a_small}.  To the extent, then, that beat-chroma
+\cite{Ellis2007a}.  To the extent, then, that beat-chroma
 representations preserve tonal content, they
@@
 The individual codewords consist of short beat-chroma patches of
-between 1 and 8 beats, optionally aligned to bar boundaries.
-%
-The additional temporal alignment eliminates redundancy that would be
-created by learning multiple codewords to represent the same motive at
-multiple beat offsets.
-%
-The codewords are able to represent the entire dataset
+between 1 and 8 beats, which are able to represent the entire dataset
 of millions of patterns with minimum error given a small codebook of
-a few hundred entries.
-%
-Our goal is to identify meaningful information
+a few hundred entries.  Our goal is to identify meaningful information
 about the musical structure represented in the entire database by
@@
 
+One major issue in working with small patches concerns temporal
+alignment. Learning the same motifs with different offsets 
+(in time or beats) is redundant. We take advantage
+of rhythmic knowledge about the song to align patches with beats.
+We can further group beats into bars to produces patches
+for meaningful music segments.
+Our use of the Echo Nest API \cite{EchoNest} for this purpose is 
+explained in Section \ref{ssec:beatpatch}. Patches built around
+these temporal landmarks should capture music motifs more easily.
+
 Prior work using small patches of chroma features includes
-the ``shingles'' of \cite{Casey2007_small}, which were used to
+the ``shingles'' of \cite{Casey2007}, which were used to
 identify ``remixes'', i.e., music based on some of the same
@@
 use either beat- or bar-synchronous features).  Earlier work in
-beat-synchronous analysis includes \cite{Bartsch2001_small},
+beat-synchronous analysis includes \cite{Bartsch2001},
 which looked for repeated patterns within single songs to
-identify the chorus, and \cite{Ellis2007a_small}, which 
+identify the chorus, and \cite{Ellis2007a}, which 
 cross-correlated beat-chroma matrices to match cover
@@
 
-\begin{figure}[t]
+\begin{figure}[thb]
 \begin{center}
-\includegraphics[width=.4\columnwidth]{code}
+\includegraphics[width=.6\columnwidth]{code}
 \end{center}
@@
 beat-synchronous chroma feature representation similar to that used in
-\cite{Ellis2007a_small}.  Echo Nest chroma vectors are normalized to have the 
+\cite{Ellis2007a}.  Echo Nest chroma vectors are normalized to have the 
 largest value in each column equal to 1.
@@
 In practice, we have found them reasonable, 
-and given the size of the dataset, any rare imperfections or noise
+and given the size of the data set, any rare imperfections or noise
 can be diluted to irrelevance by the good examples.  
@@
 depending on the meter of a particular
-song, we resample each patch to a fixed length of typically 4 beats
-per bar.  However, the majority (82\%) of our training data
+song, we resample each patch to a fixed length of 4 beats
+per bar (except where noted).  However, the majority (82\%) of our training data
 consisted of bars that were 4 beats long, so this resampling 
@@
 In the context of clustering this avoids problems such as obtaining separate
-clusters for every major triad in duple or triple meter.
+clusters for every major triad for both duple and triple meters.
 % Also, this implies that the pitch envelope is important rather than
@@
 \subsection{Clustering}
+
 % We can then stack beat vectors for a number of bars (usually 1 or 2).
@@
 choosing $K$ random points from our dataset.
-% ADDED BY TBM AFTER DAN
-Initialization had no significant effect on the distortion
-in our experiments.
 %
@@
 
-\subsection{Setting}\label{ssec:setting}
+\subsection{Settings}\label{ssec:setting}
 We take one or two bars and resample the patches to 4 or 8 columns 
@@
 \begin{eqnarray}
-  \dist(p_1,p_2) = \sum_{i,j} \frac{\left(p_1(i,j) - p_2(i,j)\right)^2}{\size(p_1)}
+  \dist(p_1,p_2) = \sum_{i,j} \frac{(p_1(i,j) - p_2(i,j))^2}{\size(p_1)}
   \label{eq:dist}
@@
 Therefore, each codeword has $12 \times 4 = 48$ elements.
-Distortion is measured on the test set.  Training data sizes range
+Distortion is measurde on the test set.  Training data sizes range
 from $0$ (just initialization) to $500,000$. Patterns were selected at
@@
 
-\begin{figure}[t]
+\begin{figure}[b]
 \begin{center}
@@
 
+\begin{figure}[t]
+\begin{center}
+\includegraphics[width=\columnwidth]{codes1}
+\end{center}
+\caption{\small{
+The 25 codes that are 
+most commonly used for the {\it artist20} test set. 
+Codes are from the $200$-entry codebook trained on $2$ bar, $12 \times 8$
+patches.  The proportion of patches accounted for by each pattern is shown 
+in parentheses.
+%There are $71832$ patterns in total in the dataset.
+}}
+\label{fig:codes1}
+\end{figure}
+
+\begin{figure}[bt]
+\begin{center}
+\includegraphics[width=.99\columnwidth]{freqs}
+\end{center}
+\caption{\small{Usage proportions for all 200 codewords 
+on the {\it artist20} test set (which comprises $71,832$
+patterns).  Also shown are the usage proportions for the training set (``cowbell''), 
+which are similar.
+Note that even though all codewords are initialized from samples, some
+are used only once in the training set, or not at all for test set.
+}}
+\label{fig:freqs}
+\end{figure}
+
 The 25 most frequently used codewords in the test set are shown in
@@
 (e.g., Gmaj
-$\rightarrow$ Cmaj or G5 $\rightarrow$ C5 as a guitar power chord)
+$\rightarrow$ Cmaj, or G5 $\rightarrow$ C5 as a guitar power chord)
 % assuming that chroma 0 is C) 
@@
 \begin{center}
-\includegraphics[width=.9\columnwidth]{codes1}
+\includegraphics[width=.9\columnwidth]{codes_lle}
 \end{center}
 \caption{\small{
-The 25 codes that are 
-most commonly used for the {\it artist20} test set. 
-Codes are from the $200$-entry codebook trained on $2$ bar, $12 \times 8$
-patches.  The proportion of patches accounted for by each pattern is shown 
-in parentheses.
-%There are $71832$ patterns in total in the dataset.
-}}
-\label{fig:codes1}
-\end{figure}
-
-\begin{figure}[t]
-\begin{center}
-\includegraphics[width=.99\columnwidth]{freqs}
-\end{center}
-\caption{\small{Usage proportions for all 200 codewords 
-on the {\it artist20} test set (which comprises $71,832$
-patterns).  Also shown are the usage proportions for the training set (``cowbell''), 
-which are similar.
-Note that even though all codewords are initialized from samples, some
-are used only once in the training set, or not at all for test set.
-}}
-\label{fig:freqs}
-\end{figure}
-
-An alternate visualization of the codebook is shown in Figure
-\ref{fig:lle}.  We use locally linear embedding \cite{Roweis2000} to
-group near neighbors on a 2 dimensional plane\footnote{We used the
-  implementation from:
-  \url{http://www.astro.washington.edu/users/vanderplas/coding/LLE/}}.
-This visualization highlights different groups of codewords: sustained
-chords on the top left, chord changes on the bottom left, and more
-complex sustained chords and ``wideband'' noisy patterns grouping to
-the right of the figure.
-
-Given the prevalence of simple codewords in Figures \ref{fig:codes1}
-and \ref{fig:lle}, we assess the proportion of the codebook devoted to
-representing sustained chord patterns.  The average variance over time
-within each codeword is plotted in Figure \ref{fig:code_var}.  From
-this figure it is clear that 25\% of the codebook corresponds to very
-simple sustained patterns similar to the top row of Figure
-\ref{fig:codes1}. For $2$ bar codewords, it seems that patterns
-containing more complex temporal activity account for only $33$\% of
-the data.
-
-Showing codewords of $2$ bars is practical and consistent with our experiments
-in Section \ref{sec:exps2}. However, longer codewords can also reveal
-interesting structure. A random sample of such codewords is presented
-in Figure \ref{fig:codes8bars}. Most used codewords would be less interesting,
-as they all are sustained patterns.
-
-\begin{figure}[t]
-\begin{center}
-\includegraphics[width=.9\columnwidth]{codes_lle}
-\end{center}
-\caption{\small{ Locally linear embedding (LLE) visualization of the
-    codebook described in Section \ref{sec:visu}.  Each codeword is
-    represented by a point on the scatter plot.  Randomly selected
-    codewords from the embedded space are displayed over the plot.
-    %FIXME: does this figure really provide any new insight?
+LLE visualization of the codebook.
+%LLE tries to keep near neighbors (here according to euclidean distance)
+%in the same $2$D neighborhood. 
+Shown patterns are randomly selected
+from each neighborhood.
 }}
@@
 \end{figure}
+
+In an effort to visualize the span of the entire codebook, we used 
+Locally linear embedding (LLE) \cite{Roweis2000}\footnote{implementation:
+  \url{http://www.astro.washington.edu/users/vanderplas/coding/LLE/}}
+to arrange the codewords on a 2D plane while keeping similar 
+patterns as neighbors.  Figure \ref{fig:lle} shows the resulting 
+distribution along with a sampling of patterns; notice trends such as the 
+more ``wideband'' patterns to the right.
+
+%Another interesting visualization uses LLE\footnote{implementation:
+%  \url{http://www.astro.washington.edu/users/vanderplas/coding/LLE/}}
+%algorithm \cite{Roweis2000} to display codewords arranged in
+%coherent groups. See Figure \ref{fig:lle}. For instance, we see
+%``wideband'' patterns grouping to the right.
 
@@
 \end{center}
-\caption{\small{Average variance in time of a codebook. The vertical
-axis cuts at the $63$rd pattern, a rough estimate of the number of 
-codewords consisting of sustained chords.
+\caption{\small{Average variance of codewords along the time dimension. The vertical
+axis cuts at the $53$rd pattern, roughly the number of 
+codewords consisting entirely of sustained chords.  Representative patterns are shown 
+in each range.
 }}
@@
 
-\begin{figure}[t]
+Noting that many of the codewords reflect sustained patterns with 
+little temporal variation, Figure \ref{fig:code_var} plots the average 
+variance along time of all 200 patterns.  
+Some 26\% of the codewords have very low variance, corresponding 
+to stationary patterns similar to the top row of Figure
+\ref{fig:codes1}. 
+%For $2$ bars codewords, it seems that patterns
+%containing more complex temporal activity account for only $33$\% of
+%the data.
+%%%% dpwe: I can't tell what this means, so drop it.
+
+\begin{figure}[htb]
 \begin{center}
@@
 \end{center}
-\caption{\small{Sample of longer codewords spanning over $8$
-    bars. Codewords selected randomly from the codewords used in the
-    test set. Percentage of use display in parentheses.  Most patterns
-    consist of sustained note or chords, but code 0 shows a repeating
-    two chord pattern and code 4 contains two bars of a
+\caption{\small{Sample of longer codewords spanning $8$
+    bars. Codewords were randomly selected from a 100-entry 
+    codebook. Percentages of use are shown in parentheses.  Most patterns
+    consist of sustained notes or chords, but code 0 shows one-bar 
+    alternations between two chords, and code 4 contains two cycles of a
     1$\rightarrow$1$\rightarrow$1$\rightarrow$2 progression.}}
@@
 
-\subsection{Cluster}
-Clusters are formed around a centroid (codeword) by all patterns closer
-to that codeword than any other. Since the centroid is somewhat an
-average of those patterns, the latter tend to be noisier. Hence, codewords
-can be seen as the output of a lowpass filtering mechanism.
-In Figure \ref{fig:cluster} we show elements of one given
-cluster, and in Figure \ref{fig:cluster_diff}, we emphasize the difference
-with the centroid. We hope for the difference to look uniformly random,
-the result seems to fit that assumption.
+We made some preliminary experiments with codebooks based on 
+longer patches.  Figure \ref{fig:codes8bars} presents a codewords 
+from an 8 bar (32 beat) codebook.  We show a random selection 
+since all the most-common codewords were less interesting, sustained 
+patterns.
+
+%Showing codewords of $2$ bars is practical and consistent with our experiments
+%in Section \ref{sec:exps2}. However, longer codewords can also reveal
+%interesting structure. A random sample of such codewords is presented
+%in Figure \ref{fig:codes8bars}. Most used codewords would be less interesting,
+%as they all are sustained patterns.
+
+\begin{figure}[htb]
+\begin{center}
+\includegraphics[width=.9\columnwidth]{close_patterns1}
+\end{center}
+\caption{\small{Cluster around centroid presented in
+Figure \ref{fig:code}. Taken from the {\it artist20} dataset, the cluster
+size is actually $639$. Shown samples were randomly selected.
+}}
+\label{fig:cluster}
+\end{figure}
+
+\subsection{Within-cluster behavior}
+In addition to inspecting the codewords, it is important to understand 
+the nature of the cluster of patterns represented by each codeword, i.e., 
+how well the centroid describes them, and the kind of detail that has 
+been left out of the codebook.  Figure \ref{fig:cluster} shows a random 
+selection of the $639$ patterns from the {\it artist20} test set that 
+were quantized to codeword 7 from Figure \ref{fig:codes1}, the V-I cadence. 
+Figure \ref{fig:cluster_diff} illustrates the difference between the actual 
+patterns and the quantized codeword for the first three patterns; although 
+there are substantial differences, they are largely unstructured, indicating 
+that the codebook has captured the important underlying trend.
+
+%Visualizing a codebook is half of the story, we also want to understand
+%clusters, \ie the patterns from actual song that
+%surround a given centroid (codeword). Unsurprisingly, real patterns are
+%noisier than codes. However, centroids seem to capture important aspects
+%of the patterns. We could compare the centroids with the result of
+%a lowpass filtering. In Figure \ref{fig:cluster} we show elements of the same
+%cluster, and in Figure \ref{fig:cluster_diff}, we emphasize the difference
+%with the centroid. We hope for the difference to look uniformly random,
+%the result seems to fit that assumption.
 
@@
 
-\begin{figure}[t]
-\begin{center}
-\includegraphics[width=.9\columnwidth]{close_patterns1}
-\end{center}
-\caption{\small{Cluster around centroid presented in
-Figure \ref{fig:code}. Taken from the artist20 dataset, the cluster
-size is actually $639$. Shown samples were randomly selected.
-}}
-\label{fig:cluster}
-\end{figure}
-
-\begin{figure}[t]
+\begin{figure}[htb]
 \begin{center}
@@
 
-Figure \ref{fig:encodesong} shows an example of a song encoded
-using the codebook described in Section \ref{sec:codebook}.  Both the
-Echo Nest chroma features for the example song and the corresponding
-quantization are displayed.  The quantized representation retains
-dominant harmonic structure of the original features, but is smoothed
-in time due to the use of 2 bar long codewords.
+Figure \ref{fig:encodesong} gives an example of encoding a song
+using the codebook, showing both the full, original data, and the 
+reconstruction using only the quantized codewords (at the correct transpositions).  
+The quantized representation retains
+the essential harmonic structure of the original features, but has smoothed
+away much of the detail below the level of the 2 bar codewords.
 %
-This is consistent with the codebook described in Section
-\ref{sec:codebook} which primarily consists of sustained chords across
-4 or 8 beats.
+%This is consistent with the codebook described in Section
+%\ref{sec:codebook} which primarily consists of sustained chords across
+%4 or 8 beats.
 
-\begin{figure}[t]
+\begin{figure}[htb]
 \begin{center}
@@
 \caption{\small{Good Day Sunshine by \textit{The Beatles}.
-Original song and its encoding using a $200$ entry codebook learning
-$2$ bars at a time.
+Original song and encoding with a $200$ entry codebook of 
+$2$ bar patterns.
 }}
@@
 \section{Applications}\label{sec:exps2}
-In this section we investigate the application of beat-chroma
-codebooks to two tasks unrelated to clustering.
+We present two applications of the beat-chroma codebooks
+to illustrate how the ``natural'' structure identified via 
+unsupervised clustering can provide useful 
+features for subsequent supervised tasks.  We will discuss how 
+the codewords can be used in song segmentation, and for artist
+recognition.  
+%The representation created from large collections
+%captures meaningful information without being forced to do so.
 %
-In the first experiment, we show that the codebooks retrain
-information related to bar alignment and can therefore be used to
-discover the bar boundaries for new songs.  The second experiment
-shows that we can discriminate artists based on the encoding of their
-songs, implying that some artists have a tendency to use different
-beat-chroma patterns than others.
-
 % Although our algorithm was not devised for any of these
@@
 Since the clustering described in Section \ref{sec:approach} is based
-on the segmentation of the signal into bars, it stands to reason that
-the cluster centroids should contain information related to bar
+on the segmentation of the signal in to bars, 
+the codewords should contain information related to bar
 alignment, such as the presence of a strong beat on the first beat.
 %
-We investigate whether it is possible to recover the
-bar segmentation of new songs using the information contained in the
-cluster centroids.  The results are shown Table \ref{tab:offset}. We find
-the right segment with an accuracy of $62\%$ where random is
-$25\%$. Details of the experiment follow.
-
-We train a codebook of size $100$ on bars resized to $4$ beats. Then,
+In this section we investigate using the codebook to identify the
+bar segmentation of new songs.  
+% using the information contained in the cluster centroids.  
+We train a codebook of size $100$ on bars resampled to $4$ beats. Then,
 we take the longest sequence of bars of $4$ beats for each song
-in the test set.  As
-we mentioned before, $4$ beats is the most common bar size using the
-Echo Nest segmentation. We then encode each of these sequences using an offset
-of $0$, $1$, $2$ or $3$ beats. The offset on a test sequence
-with the lowest distortion is our predicted alignment.
-The correct answer is an offset of $0$.
+in the test set (to avoid the alignment skew that results from spanning 
+irregularly-sized bars).  
+%As we mentioned before, $4$ beats is the most common bar size using the
+%Echo Nest segmentation. 
+We then encode each of these sequences using an offset
+of $0$, $1$, $2$ or $3$ beats, and record for each song 
+the offset giving the lowest distortion.  
+%The best offset for a test sequence
+%(the one with the lowest distortion)
+%should correspond to the bar alignment in the original
+%feature matrix.  
+%The correct answer is an offset of $0$.
+%
+The results in Table \ref{tab:offset}
+show that the ``true'' offset of $0$ beats is chosen in $62\%$ of cases (where a random 
+guess would yield $25\%$).  Thus, the codebook is useful for identifying bar (downbeat) 
+alignments.  A more flexible implementation of this idea would use dynamic 
+programming to align bar-length patterns to the entire piece, including the 
+possibility of 3- or 5-beat bars (as are sometimes needed to accommodate 
+beat tracking errors) with an appropriate associated penalty.
 
@@
 
-The result presented here suggest a dynamic programming setting where
-we could find the optimal bar segmentation according to our codebook. 
-We could try
-all possible bar length (in terms of beats) and not restrict ourselves
-to $4$ beats bars and offsets between $0$ and $3$. This is part of
-our future work.
+%The result presented here suggest a dynamic programming setting where
+%we could find the optimal bar segmentation according to our codebook. 
+%We could try
+%all possible bar length (in terms of beats) and not restrict ourselves
+%to $4$ beats bars and offsets between $0$ and $3$. This is part of
+%our future work.
 
-\begin{table}[t]
+\begin{table}
 \begin{center}
@@
 1 & $16.5$\\
-2 & $09.4$\\
+2 & $\text{ }9.4$\\
 3 & $11.5$\\
@@
 \caption{\small{
-Results from the song segmentation experiment.
-The correct offset is $0$.
-Baseline performance by choosing the offset randomly gives an accuracy of $25$\%.
+Song segmentation experiment:
+offsets relative to ground-truth 4-beat bar boundaries 
+that gave minimum distortion encodings from the bar-aligned codebook.
+%The real one is $0$.
+%Random guess gives an accuracy of $25$\%.
 }}
@@
 
-We apply our codebook to the artist recognition task.
-Our dataset, ``artist20'', encompass $1402$ songs from $20$ artists, 
+We apply our codebook to a simple artist recognition task.
+We use the {\it artist20} data set, composed of $1402$ songs from $20$ artists, 
 mostly rock and pop of different subgenres.  
-The best accuracy on the dataset in the literature is $59\%$, whereas
-using only chroma and a Gaussian mixture model yields an accuracy of
-$33\%$ \cite{Ellis2007}.
+Previously published results using GMMs on MFCC features achieve an 
+accuracy of $59\%$, whereas using only chroma as a representation yields an
+accuracy of $33\%$  \cite{Ellis2007}.
 
-The clustering method proposed in this work is not
-intended to perform well at artist recognition. 
-In particular,
-we lose all timbral information which would be useful. 
-However, we can
-imagine that some artist use more often some patterns than others.
-Our accuracy result is $23.4\%$ while random
+Although beat-chroma quantization naturally discards information 
+that could be useful in artist classification, it is interesting to 
+investigate whether some artist use certain patterns more often than others.
+
+The dataset is encoded as histograms of the codewords 
+used for each song, with frequency values normalized by the number
+of patterns in the song. We test each song in a leave-one-out setting, 
+and represent each of the $20$ artists by the average of their (remaining) 
+song-histograms.  The test song is matched to an artist based on 
+minimum Euclidean distance to these per-artist averages.
+This gives an accuracy of $23.4\%$, where the random baseline 
 is around $5\%$. 
-The confusion matrix can be seen in Figure \ref{fig:conf_mat}.
-Details of the experiment follow.
-
-All songs in the artist dataset are encoded as an histogram of the codes
-used in that song. The frequency values are normalized by the number
-of patterns in the song. We test each song in a leave-one-out setting.
-Leaving aside the test song, we represent each of the $20$ artists by the 
-average of
-the song histograms of that artist. The test song is compared to
-the histogram of each artist using Euclidean distance.
+The confusion matrix can be seen in Figure \ref{fig:conf_mat}, showing 
+that certain artists are recognized at an accuracy far above the average.
 
@@
 %\end{center}
-\caption{\small{Confusion matrix for artist recognition task.}}
+\caption{\small{Confusion matrix for the artist recognition task.}}
 \label{fig:conf_mat}
@@
 
-Looking at the most typical patterns per artist can be insightful.
-To find these patterns, we compare its use by one artist and divide
-by its use by all artists. In Figure \ref{fig:typicalpat} we see 
-Metallica's pattern, then Tori Amos and
-Suzanne Vega's pattern. These three artists were easily identified.
-Artists such as Metallica are characterized by ``wideband'' patterns,
-containing features muddled by the noisy acoustics containing many
-transients,
-distortion, etc, which can be seen in Figure \ref{fig:metallica2}.
+%Looking at the most typical patterns per artist can be insightful.
+It is interesting to inspect the most ``discriminative'' patterns for 
+individual artists.  
+To find these patterns, we compare a pattern's use by one artist and divide
+by its use across all artists. Figure \ref{fig:typicalpat} 
+shows the dominant patterns for Metallica, and for Tori Amos and
+Suzanne Vega (who shared a `favorite' pattern). 
+These three artists were easily identified.
+Artists like Metallica are characterized by ``wideband'' patterns, 
+with energy spread across multiple adjacent chroma bins, 
+indicative of noise-like transients in the audio.
 %FIXME: ron, what di you mean: (not by cleaned up by EN's feature analysis)
@@
 \iffalse
-\begin{figure}[t]
+\begin{figure}[htb]
 \begin{center}
@@
 
-\begin{figure}[t]
+\begin{figure}[htb]
 \begin{center}
@@
 
-\begin{figure}[t]
+\begin{figure}[h]
   \centering
@@
 \section{Conclusion and Future Work}
-We presented a practical method to perform large-scale clustering of
-harmonic patterns. We assessed the basic properties of the method through
-experiments on a large collection of music, and 
-demonstrated that these codebooks capture musically meaningful
-patterns, at least to contemporary pop music.
-%
- We suggested many ways
-to interpret the data %(such as LLE)
-and assert the merits of the codebook
+We have presented a practical method to perform large-scale clustering of
+tonal patterns, and assessed the basic properties of the method through
+experiments on a large collection of music. We have explored several ways
+to inspect and interpret the data 
+%(such as LLE) 
+and suggested the merit of the representation
 through further experiments.
-We discussed the possibility to move to even larger scales
-and provided our source code.\footnote{Code not yet released to preserve
-submission's anonymity}
-%
-Finally, we demonstrated the utility of unsupervised clustering to MIR
-tasks that are typically tackled in a supervised setting.
+We have discussed the possibility to move to even larger scales
+and we provide our source code for other interested researchers.
+\footnote{Code not yet released to preserve submission's anonymity}
 
-For even larger scales, we will focus on three facets.
-First, gather more songs from the Echo Nest API. 
-%
-Second,
-larger codebooks paired with computationally efficient algorithms.
-Improvements to $k$-means
-avoid comparing every code to every new sample. Initialization
-should also be modified, so every codeword is used (unlike in
-Figure \ref{fig:freqs}). 
-%
-Third, the learning
-algorithm must not converge too fast. We could add a filter that selects
-increasingly difficult examples to encode, or merge many codebooks in a 
-similar fashion as hierarchical Gaussian mixtures \cite{Vasconcelos2001}.
+Future work may include more sophisticated clustering that moves 
+beyond simple Euclidean-distance-based quantization, perhaps 
+by separately modeling the spread within each cluster (i.e., a Gaussian 
+mixture or other generative model). 
+%That being said, these methods
+%could be use to develop better distance measures between patterns.
+%The use of the euclidean distance is arbitrary. 
+Summarizing patches
+with Gaussians, and then comparing the distance between those Gaussians,
+could reduce the influence of the noise in the distance measure.
 
-%As for enhancements, we specifically did not compress our features using
-%Gaussian mixtures or other generative model. That being said, these methods
-%could be use to develop better distance measures between patterns.
-%The use of the Euclidean distance is arbitrary. Summarizing patches
-%with Gaussians, and then comparing the distance between those Gaussians,
-%could reduce the influence of the noise in the distance measure.
-
-%Moving on to larger scales, we could imagine merging codes together and
-%splitting some on two if the clusters become too small or too large.  We could
-%also cluster codebooks themselves, in a similar fashion as hierarchical
-%Gaussian mixtures \cite{Vasconcelos2001}.
+Moving on to larger scales, we would like to pursue a scheme of incrementally 
+splitting and merging codewords in response to a continuous, online stream 
+of features, to create an increasingly-detailed, dynamic model.  We could
+also cluster codebooks themselves, in a fashion similar to hierarchical
+Gaussian mixtures \cite{Vasconcelos2001}.
 %  (similar idea used for indexing \cite{Muja2009})

### 2010-03-25 Paper/Paper.tex
minor

@@
 
-\begin{figure}[thb]
+\begin{figure}[t]
 \begin{center}
-\includegraphics[width=.6\columnwidth]{code}
+\includegraphics[width=.4\columnwidth]{code}
 \end{center}
@@
 
-\small
+%\small
 %\section{Acknowledgements}

### 2010-03-25 Paper/Paper.tex
ron comment on LLE added back

@@
 patterns as neighbors.  Figure \ref{fig:lle} shows the resulting 
-distribution along with a sampling of patterns; notice trends such as the 
-more ``wideband'' patterns to the right.
+distribution along with a sampling of patterns; notice 
+% ron's text instead of Dan
+sustained
+chords on the top left, chord changes on the bottom left, and more
+complex sustained chords and ``wideband'' noisy patterns grouping to
+the right of the figure.
+
+%trends such as the 
+%more ``wideband'' patterns to the right.
 

### 2010-05-18 Paper/Paper.tex
acknowledgements and authors updated

@@
 \threeauthors
-  {First author} {Affiliation1 \\ {\tt author1@ismir.edu}}
-  {Second author} {\bf Retain these fake authors in\\\bf submission to preserve the formatting}
-  {Third author} {Affiliation3 \\ {\tt author3@ismir.edu}}
+  {Thierry Bertin-Mahieux} {Columbia University\\ {\tt tb2332@columbia.edu}}
+  {Ron Weiss} {New York University \\ {\tt ronw@nyu.edu}}
+  {Dan Ellis} {Columbia University \\ {\tt dpwe@ee.columbia.edu}}
 % what order do we use? Thierry, Ron, Dan?  Dan, Thierry, Ron?
@@
 %\small
-%\section{Acknowledgements}
-%Thierry is NSERC graduate fellow, or some title like that.
-%Ron has a grant too.
-% Graham for the NMF stuff and numerous discussions.
+\section{Acknowledgements}
+Thanks to Graham Grindlay for numerous discussions and helpful comments.
+T. Bertin-Mahieux is supported by a NSERC scholarship.
+This material is based upon work supported by the IMLS (grant LG-06-08-0073-08).
 

### 2010-05-18 Paper/Paper.tex
fixed my name!

@@
   {Thierry Bertin-Mahieux} {Columbia University\\ {\tt tb2332@columbia.edu}}
-  {Ron Weiss} {New York University \\ {\tt ronw@nyu.edu}}
+  {Ron J. Weiss} {New York University \\ {\tt ronw@nyu.edu}}
   {Dan Ellis} {Columbia University \\ {\tt dpwe@ee.columbia.edu}}

### 2010-05-27 Paper/Paper.tex
some modifs done for final version

--- 2010-05-18 Paper/Paper.tex
@@
 reconstruction of the original song requires knowledge of the
-rotation index (key) for each patch.
+rotation index for each patch. This rotation removes the key information
+of a patch.
 
@@
 are used only once in the training set, or not at all for test set.
+This explains why the curves drop to $0$.
 }}
@@
 
-\begin{figure}[htb]
+\begin{figure}[t]
 \begin{center}
-\includegraphics[width=.9\columnwidth]{close_patterns1}
+\includegraphics[width=.95\columnwidth]{close_patterns1}
 \end{center}
@@
 size is actually $639$. Shown samples were randomly selected.
+This gives an intuition of the variance
+in a given cluster.
 }}
@@
 
-\begin{figure}[htb]
+\begin{figure}[t]
 \begin{center}
-\includegraphics[width=.8\columnwidth]{close_patterns_diff}
+\includegraphics[width=.95\columnwidth]{close_patterns_diff}
 \end{center}
@@
 
+\small
 \bibliography{tbm_bib}

### 2010-05-27 Paper/Paper.tex
code url added

@@
 and we provide our source code for other interested researchers.
-\footnote{Code not yet released to preserve submission's anonymity}
+\footnote{See \textit{Papers} section at \url{www.columbia.edu/~tb2332/}}
 

### 2010-05-28 Paper/Paper.tex
more modifs

@@
 \usepackage{subfloat,subfig}
-\setlength{\abovecaptionskip}{-3pt}
+\setlength{\abovecaptionskip}{0pt}
 \setlength{\belowcaptionskip}{6pt}
@@
   {Ron J. Weiss} {New York University \\ {\tt ronw@nyu.edu}}
-  {Dan Ellis} {Columbia University \\ {\tt dpwe@ee.columbia.edu}}
+  {Daniel P. W. Ellis} {Columbia University \\ {\tt dpwe@ee.columbia.edu}}
 % what order do we use? Thierry, Ron, Dan?  Dan, Thierry, Ron?
@@
 \begin{center}
-\includegraphics[width=.4\columnwidth]{code}
+\includegraphics[width=.4\columnwidth]{code_axis}
 \end{center}
@@
 features for subsequent supervised tasks.  We will discuss how 
-the codewords can be used in song segmentation, and for artist
+the codewords can be used in bar alignment, and for artist
 recognition.  
@@
 
-\subsection{Song Segmentation} \label{ssec:segment}
+\subsection{Bar Alignment} \label{ssec:segment}
 
@@
 \caption{\small{
-Song segmentation experiment:
+Bar alignment experiment:
 offsets relative to ground-truth 4-beat bar boundaries 
@@
 We have discussed the possibility to move to even larger scales
-and we provide our source code for other interested researchers.
-\footnote{See \textit{Papers} section at \url{www.columbia.edu/~tb2332/}}
+and we provide our source code for other interested researchers
+\footnote{See \textit{Papers} section at \url{www.columbia.edu/~tb2332/}}.
 
@@
 T. Bertin-Mahieux is supported by a NSERC scholarship.
-This material is based upon work supported by the IMLS (grant LG-06-08-0073-08).
+This material is based upon work supported by the IMLS (grant LG-06-08-0073-08)
+and by NSF grant IIS-0713334. Any opinions, findings and
+conclusions or recommendations expressed in this material are those of
+the authors and do not necessarily reflect the views of the sponsors
 

### 2010-06-01 Paper/Paper.tex
minor
minor tweaks.  made paper fit in 6 pages

@@
 harmony and melody.  As a computationally-convenient proxy for a richer
-description of the tonal content of audio, we use the popular Chroma
+description of the tonal content of audio, we use the popular chroma
 representation, which collapses an acoustic spectrum into a 12-dimensional
@@
 reconstruction of the original song requires knowledge of the
-rotation index for each patch. This rotation removes the key information
-of a patch.
+rotation index for each patch.
+%This rotation removes the key information of a patch.
 
@@
 For testing, we made use of low quality (32kbps, 8 kHz bandwidth mono MP3) versions of the songs
-from the {\it uspop2002} data set \cite{uspop2002}.  This well-known data set 
+from the {\it uspop2002} data set \cite{uspop2002}.  This data set 
 contains pop songs from a range of artists and styles.
@@
 We then use the resulting codebook to encode the test set.
-Each pattern is encoded with only one code. We can measure the average
+Each pattern is encoded with a single code. We can measure the average
 distance between a pattern and its encoding. We can also measure the use
@@
 
-We trained a codebook $200$ entries of patterns sized $12\times8$, covering
+We trained a codebook containing $200$ patterns sized $12\times8$, covering
 $2$ bars at a time.  The results shown are on the {\it artist20} test set described in 
@@
 features for subsequent supervised tasks.  We will discuss how 
-the codewords can be used in bar alignment, and for artist
+the codewords can be used in bar alignment, and artist
 recognition.  
@@
 
-\begin{figure}[h]
+\begin{figure}[t]
   \centering
@@
 We have discussed the possibility to move to even larger scales
-and we provide our source code for other interested researchers
+and we provide our source code for other interested 
+researchers
 \footnote{See \textit{Papers} section at \url{www.columbia.edu/~tb2332/}}.
@@
 T. Bertin-Mahieux is supported by a NSERC scholarship.
-This material is based upon work supported by the IMLS (grant LG-06-08-0073-08)
+This material is based upon work supported by IMLS grant LG-06-08-0073-08
 and by NSF grant IIS-0713334. Any opinions, findings and
 conclusions or recommendations expressed in this material are those of
-the authors and do not necessarily reflect the views of the sponsors
+the authors and do not necessarily reflect the views of the sponsors.
 

### 2010-06-01 Paper/Poster/poster.tex
template from Graham

\documentclass[landscape,a0b,final]{a0poster}

\usepackage{epsfig}
\usepackage{multicol}
\usepackage{pstricks,pst-grad}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Definition of some variables and colors
\setlength{\columnsep}{3cm}
\setlength{\columnseprule}{2mm}
\setlength{\parindent}{0.0cm}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%               Background                     %%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\newcommand{\background}[3]{
  \newrgbcolor{cgradbegin}{#1}
  \newrgbcolor{cgradend}{#2}
  \psframe[fillstyle=gradient,gradend=cgradend,
  gradbegin=cgradbegin,gradmidpoint=#3](0.,0.)(1.\textwidth,-1.\textheight)
}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%                Poster                        %%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\newenvironment{poster}{
  \begin{center}
  \begin{minipage}[c]{0.98\textwidth}
}{
  \end{minipage} 
  \end{center}
}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%                pcolumn                       %%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\newenvironment{pcolumn}[1]{
  \begin{minipage}{#1\textwidth}
  \begin{center}
}{
  \end{center}
  \end{minipage}
}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%                pbox                          %%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\newrgbcolor{lcolor}{00. 0. 0.80}
\newrgbcolor{gcolor1}{1. 1. 1.}
\newrgbcolor{gcolor2}{.80 .80 1.}

\newcommand{\pbox}[4]{
\psshadowbox[#3]{
\begin{minipage}[t][#2][t]{#1}
#4
\end{minipage}
}}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%                myfig                         %%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% \myfig - replacement for \figure
% necessary, since in multicol-environment 
% \figure won't work
\newcommand{\myfig}[3][0]{
\begin{center}
  \vspace{1.5cm}
  \includegraphics[width=#3\hsize,angle=#1]{#2}
  \nobreak\medskip
\end{center}}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%                mytable                       %%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% \mytable - replacement for \table
% necessary, since in multicol-environment 
% \table won't work
\newcommand{\mytable}[3][0]{
\begin{center}
  \vspace{1.5cm}
  \includegraphics[width=#3\hsize,angle=#1]{#2}
  \nobreak\medskip
\end{center}}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%                mycaption                     %%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% \mycaption - replacement for \caption
% necessary, since in multicol-environment \figure and
% therefore \caption won't work

%\newcounter{figure}
\setcounter{figure}{1}
\newcommand{\mycaption}[1]{
  \vspace{0.5cm}
  \begin{quote}
    {{\sc Figure} \arabic{figure}: #1}
  \end{quote}
  \vspace{1cm}
  \stepcounter{figure}
}

%\newcounter{table}
\setcounter{table}{1}
\newcommand{\mytablecaption}[1]{
  \vspace{1.0cm}
  \begin{quote}
    {{\sc Table} \arabic{table}: #1}
  \end{quote}
  \vspace{1cm}
  \stepcounter{table}
}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%% Begin of Document
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{document}

\background{1.0 1.0 1.0}{1.0 1.0 1.0}{0.3}

\vspace*{2cm}

\newrgbcolor{lightred}{0.7 0.0 0.0}
\newrgbcolor{white}{1.0 1.0 1.0}
\newrgbcolor{whitepink}{1.0 0.8 0.9}

\begin{poster}

%%%%%%%%%%%%%%%%%%%%%
%%% Header
%%%%%%%%%%%%%%%%%%%%%
\begin{center}
\begin{pcolumn}{0.98}

\pbox{0.957\textwidth}{}{linewidth=2mm,framearc=0.3,linecolor=lightred,fillstyle=gradient,gradangle=0,gradbegin=white,gradend=whitepink,gradmidpoint=1.0,framesep=1em}{

%%% Unisiegel
\begin{minipage}[c][9cm][c]{0.1\textwidth}
  \begin{center}
    \includegraphics[width=7cm,angle=0]{labrosa-logo.eps}
  \end{center}
\end{minipage}
%%% Titel
\begin{minipage}[c][9cm][c]{0.78\textwidth}
  \begin{center}
    {\sc \Huge A Ground-breaking Work of Epic Proportions}\\[10mm]
    {\Large Phil D. Student and Daniel P. W. Ellis\\[3.5mm]
     \large LabROSA, Dept. of Electrical Engineering, Columbia University \\
    \tt{\{phds,dpwe\}@ee.columbia.edu}}
  \end{center}
\end{minipage}
%%% GK-Logo
\begin{minipage}[c][9cm][c]{0.1\textwidth}
  \begin{center}
    \includegraphics[width=7cm,angle=0]{seas-logo.eps}
  \end{center}
\end{minipage}

}
\end{pcolumn}
\end{center}

\vspace*{0.5cm}

%%% Begin of Multicols-Enviroment
\begin{multicols}{3}

%%% Section 1
\begin{center}
  \pbox{0.8\columnwidth}{}{linewidth=2mm,framearc=0.1,linecolor=lightred,fillstyle=gradient,gradangle=0,gradbegin=white,gradend=whitepink,gradmidpoint=1.0,framesep=1em}{
    \begin{center}
      \large Section 1
    \end{center}}
\end{center}

\vspace{1.0cm}

\begin{list}{\labelitemi}{\leftmargin=2em \labelsep=1em}
\item A bullet point
\item A bullet point
\item A bullet point
\item A bullet point
\end{list}

%%% Section 2
\vspace{1.5cm}
\begin{center}
  \pbox{0.8\columnwidth}{}{linewidth=2mm,framearc=0.1,linecolor=lightred,fillstyle=gradient,gradangle=0,gradbegin=white,gradend=whitepink,gradmidpoint=1.0,framesep=1em}{
    \begin{center}
      \large Section 2
    \end{center}}
\end{center}

\vspace{1.0cm}

\begin{list}{\labelitemi}{\leftmargin=2em \labelsep=1em}
\item A bullet point
\item A bullet point
\item A bullet point
\end{list}

%%% Figure example
%\begin{center}
  % first argument: eps-file
  % second argument: stretching-factor relative to Column-width (<1)
  % optional argument: rotation angle (0-360), default=0
  %\myfig[0]{myfig.eps}{0.8}
  %\mycaption{Some caption}
%\end{center}

% force a new column
\newpage

%%% Section 3
\vspace{2cm}
\begin{center}
  \pbox{0.8\columnwidth}{}{linewidth=2mm,framearc=0.1,linecolor=lightred,fillstyle=gradient,gradangle=0,gradbegin=white,gradend=whitepink,gradmidpoint=1.0,framesep=1em}{
    \begin{center}
      \large Section 3
    \end{center}}
\end{center}

\vspace{1.0cm}

\begin{list}{\labelitemi}{\leftmargin=2em \labelsep=1em}
\item A bullet point
\item A bullet point
\item A bullet point
\item A bullet point
\end{list}

% force a new column
\newpage

%%% Section 4
\vspace{2cm}
\begin{center}
  \pbox{0.8\columnwidth}{}{linewidth=2mm,framearc=0.1,linecolor=lightred,fillstyle=gradient,gradangle=0,gradbegin=white,gradend=whitepink,gradmidpoint=1.0,framesep=1em}{
    \begin{center}
      \large Section 4
    \end{center}}
\end{center}

\vspace{1.0cm}

\begin{list}{\labelitemi}{\leftmargin=2em \labelsep=1em}
\item A bullet point
\item A bullet point
\item A bullet point
\item A bullet point
\end{list}

%%% Section 5
\vspace{0cm}
\begin{center}
  \pbox{0.8\columnwidth}{}{linewidth=2mm,framearc=0.1,linecolor=lightred,fillstyle=gradient,gradangle=0,gradbegin=white,gradend=whitepink,gradmidpoint=1.0,framesep=1em}{
    \begin{center}
      \large Section 5
    \end{center}}
\end{center}

\vspace{1.0cm}

\begin{list}{\labelitemi}{\leftmargin=2em \labelsep=1em}
\item A bullet point
\item A bullet point
\end{list}

%%% References
\bibliographystyle{IEEEtran}
\bibliography{refs.bib}

\end{multicols}
\end{poster}
\end{document}

### 2010-06-01 Paper/Poster/mkpdf.sh

@@ -1,0 +1,4 @@
+#/usr/bash
+latex poster.tex
+dvips -Ppdf poster.dvi -o poster.ps
+ps2pdf poster.ps poster.pdf

### 2010-06-14 Paper/Poster/poster.tex
title
working on experiments

@@
   \begin{center}
-    {\sc \Huge A Ground-breaking Work of Epic Proportions}\\[10mm]
-    {\Large Phil D. Student and Daniel P. W. Ellis\\[3.5mm]
-     \large LabROSA, Dept. of Electrical Engineering, Columbia University \\
-    \tt{\{phds,dpwe\}@ee.columbia.edu}}
+    {\sc \Huge Clustering Beat-Chroma Patterns in a Large Music Database}\\[10mm]
+    {\Large Thierry Bertin-Mahieux $^\dagger$, Ron J. Weiss $^\ddagger$ and Daniel P. W. Ellis $^\dagger$\\[3.5mm]
+     \large $\dagger$ LabROSA, Dept. of Electrical Engineering, Columbia University \\
+     \large $\ddagger$ MARL, New York University \\
+    \tt{\{thierry,ronw,dpwe\}@ee.columbia.edu}}
   \end{center}
@@
     \begin{center}
-      \large Section 1
+      \large Introduction
     \end{center}}
@@
     \begin{center}
-      \large Section 2
+      \large Audio Features - Echo Nest
     \end{center}}
@@
     \begin{center}
-      \large Section 3
+      \large Vector Quantization
     \end{center}}
@@
 
+%%% Section 6
+\vspace{0cm}
+\begin{center}
+  \pbox{0.8\columnwidth}{}{linewidth=2mm,framearc=0.1,linecolor=lightred,fillstyle=gradient,gradangle=0,gradbegin=white,gradend=whitepink,gradmidpoint=1.0,framesep=1em}{
+    \begin{center}
+      \large Pattern Analysis
+    \end{center}}
+\end{center}
+
+\vspace{1.0cm}
+
 % force a new column
@@
 
-%%% Section 4
+%%% Section 5
 \vspace{2cm}
@@
     \begin{center}
-      \large Section 4
+      \large Experiments
     \end{center}}
@@
 \vspace{1.0cm}
+
+We present two applications of the beat-chroma codebooks
+to illustrate how the ``natural'' structure identified via 
+unsupervised clustering can provide useful 
+features for subsequent supervised tasks.
+
+\vspace{1cm}
+
+%%% CONFUSION MATRIX
+\begin{minipage}[c][9cm][c]{0.1\textwidth}
+  \begin{center}
+    \includegraphics[width=15cm,angle=0]{conf_mat_per_artist.ps}
+  \end{center}
+\end{minipage}
+\begin{minipage}[c][9cm][c]{0.21\textwidth}
+  \begin{center}
+    fmkefeifiwefi comment on confusion matrix\\
+    cjsbdqwb\\
+    cecewe
+  \end{center}
+\end{minipage}
 
@@
 
-%%% Section 5
+%%% Section 6
 \vspace{0cm}
@@
     \begin{center}
-      \large Section 5
+      \large Conclusion
     \end{center}}

### 2010-06-14 Paper/Poster/poster.tex
improved...

@@
 \usepackage{pstricks,pst-grad}
+\usepackage{url}
+\usepackage{wrapfig}
 
@@
 
-%%% Section 6
+%%% Section 4
 \vspace{0cm}
@@
 
-\vspace{1.0cm}
+\vspace{4.0cm}
+
+\begin{minipage}[c][9cm][c]{\columnwidth}
 
@@
 features for subsequent supervised tasks.
-
 \vspace{1cm}
 
-%%% CONFUSION MATRIX
-\begin{minipage}[c][9cm][c]{0.1\textwidth}
+\begin{wrapfigure}{r}{0.5\textwidth}
+  \vspace{-20pt}
   \begin{center}
-    \includegraphics[width=15cm,angle=0]{conf_mat_per_artist.ps}
+    \includegraphics[width=0.48\textwidth]{conf_mat_per_artist.ps}
   \end{center}
+  \vspace{-20pt}
+\end{wrapfigure}
+
+\textbf{Artist recognition task}.
+We use the {\it artist20} data set: $1402$ songs from $20$ artists, 
+mostly rock and pop of different subgenres.  
+Previously published results using GMMs on MFCC features achieve an 
+accuracy of $59\%$, whereas using only chroma as a representation yields an
+accuracy of $33\%$  \cite{Ellis2007}.
+
+We get an accuracy of $\mathbf{23.4\%}$, random baseline 
+is around $5\%$. 
+The confusion matrix is shown here, note 
+that certain artists are recognized at an accuracy far above the average.
+
+\end{minipage}
+
+\vspace{2cm}
+
+\begin{minipage}[c][9cm][c]{\columnwidth}
+  %\begin{center}
+\textbf{Bar alignment task}.
+Since the clustering described is based
+on the segmentation of the signal in to bars, 
+the codewords should contain information related to bar
+alignment, such as the presence of a strong beat on the first beat.\\
+  %\end{center}
 \end{minipage}
@@
   \begin{center}
-    fmkefeifiwefi comment on confusion matrix\\
-    cjsbdqwb\\
-    cecewe
+      \begin{tabular}{cc}
+        Offset & \% of times chosen \\ \hline \hline
+        0 & $\mathbf{62.6}$\\
+        1 & $16.5$\\
+        2 & $9.4$\\
+        3 & $11.5$
+      \end{tabular}
+      %\mytablecaption{\small{
+      %    Bar alignment experiment:
+      %    offsets relative to ground-truth 4-beat bar boundaries 
+      %    that gave minimum distortion encodings from the bar-aligned codebook.
+      %}}
+      \label{tab:offset}
   \end{center}
@@
 \end{multicols}
+
+% FOOTNOTE
+%\vspace{10cm}
+\begin{minipage}{\textwidth}
+\begin{flushright}
+Poster presented at ISMIR 2010 (Utrecht) for the paper:
+Clustering Beat-Chroma Patterns in a Large Music Database by
+T. Bertin-Mahieux, R. J. Weiss and D. P. W. Ellis. Contact:
+\url{tb2332@columbia.edu}, code: \url{http://www.columbia.edu/~tb2332/}.
+\end{flushright}
+\end{minipage}
+
 \end{poster}

### 2010-07-06 Paper/Poster/poster.tex
improving intro

@@
 \begin{list}{\labelitemi}{\leftmargin=2em \labelsep=1em}
+\item Availability of very large collections of music audio:
+can we infer anything about the underlying structure and common
+features of e.g. commercial pop music?
+\item Our interest: tonal content of the music -- i.e. the
+harmony and melody.
+\item Beat-synchronous chromagrams: rich enough
+to generate musically-relevant results, simplified enough to
+abstract away instrumentation and other stylistic details.
+\item This paper identifies common patterns in beat-synchronous
+chromagrams by learning codebooks from a large set of examples.
+The individual codewords consist of short beat-chroma patches of
+between 1 and 8 beats, optionally aligned to bar boundaries.
+\end{list}
+
+\myfig[0]{code_axis.ps}{.2}
+
+%
+Our goal is to identify meaningful information
+about the musical structure represented in the entire database by
+examining individual entries in this codebook.  Since the common
+patterns represent a higher-level description of the musical content
+than the raw chroma,
+we also expect them to be useful in other applications, such as
+music classification and retrieving tonally-related items.
+
+Prior work using small patches of chroma features includes
+the ``shingles'' of \cite{Casey2007}, which were used to
+identify ``remixes'', i.e., music based on some of the same
+underlying instrument tracks, and also for matching
+performances of Mazurkas \cite{Casey2008}.  That work,
+however, was not concerned with extracting the deeper
+common patterns underlying different pieces (and did not
+use either beat- or bar-synchronous features).  Earlier work in
+beat-synchronous analysis includes \cite{Bartsch2001},
+which looked for repeated patterns within single songs to
+identify the chorus, and \cite{Ellis2007a}, which 
+cross-correlated beat-chroma matrices to match cover
+versions of pop music tracks.  None of these works
+examined or interpreted the content of the chroma
+matrices in any detail.  In contrast, here we hope to
+develop a codebook whose entries are of interest
+in their own right.
+
+\begin{list}{\labelitemi}{\leftmargin=2em \labelsep=1em}
 \item A bullet point
@@
 \bibliographystyle{IEEEtran}
-\bibliography{refs.bib}
+\bibliography{tbm_bib.bib}
 

### 2010-07-06 Paper/Poster/poster.tex
bib added

--- 2010-06-14 Paper/Poster/poster.tex
@@
 \begin{list}{\labelitemi}{\leftmargin=2em \labelsep=1em}
+\item Availability of very large collections of music audio:
+can we infer anything about the underlying structure and common
+features of e.g. commercial pop music?
+\item Our interest: tonal content of the music -- i.e. the
+harmony and melody.
+\item Beat-synchronous chromagrams: rich enough
+to generate musically-relevant results, simplified enough to
+abstract away instrumentation and other stylistic details.
+\item This paper identifies common patterns in beat-synchronous
+chromagrams by learning codebooks from a large set of examples.
+The individual codewords consist of short beat-chroma patches of
+between 1 and 8 beats, optionally aligned to bar boundaries.
+\end{list}
+
+\myfig[0]{code_axis.ps}{.2}
+
+%
+\begin{list}{\labelitemi}{\leftmargin=2em \labelsep=1em}
+\item Goal: identify meaningful information
+about the musical structure represented in the entire database by
+examining individual entries in this codebook.
+\item Prior work: ``shingles'' of \cite{Casey2007}, 
+beat-synchronous analysis to indeitfy the chorus by \cite{Bartsch2001},
+and cover recognition by \cite{Ellis2007a}.
+\end{list}
+
+\begin{list}{\labelitemi}{\leftmargin=2em \labelsep=1em}
 \item A bullet point
@@
 \bibliographystyle{IEEEtran}
-\bibliography{refs.bib}
+\bibliography{tbm_bib}
 
@@
 % FOOTNOTE
-%\vspace{10cm}
+\vspace{1cm}
 \begin{minipage}{\textwidth}
+\begin{small}
 \begin{flushright}
@@
 \end{flushright}
+\end{small}
 \end{minipage}

### 2010-07-07 Paper/Poster/poster.tex
better

@@
 \usepackage{wrapfig}
+\usepackage{algorithm,algorithmic}
 
@@
 abstract away instrumentation and other stylistic details.
-\item This paper identifies common patterns in beat-synchronous
-chromagrams by learning codebooks from a large set of examples.
-The individual codewords consist of short beat-chroma patches of
-between 1 and 8 beats, optionally aligned to bar boundaries.
 \end{list}
@@
 examining individual entries in this codebook.
+\item Method: identify common patterns in beat-synchronous
+chromagrams by learning codebooks from a large set of examples.
+Individual codewords consist of short beat-chroma patches of
+between 1 and 8 beats, optionally aligned to bar boundaries.
 \item Prior work: ``shingles'' of \cite{Casey2007}, 
@@
 and cover recognition by \cite{Ellis2007a}.
-\end{list}
-
-\begin{list}{\labelitemi}{\leftmargin=2em \labelsep=1em}
-\item A bullet point
-\item A bullet point
-\item A bullet point
-\item A bullet point
 \end{list}
@@
 
+\textbf{Chromagrams from the Echo Nest online API.}
 \begin{list}{\labelitemi}{\leftmargin=2em \labelsep=1em}
-\item A bullet point
-\item A bullet point
-\item A bullet point
+\item Feature analysis based on Echo Nest analyze API \cite{EchoNest}.
+\item For any song, EchoNest provides a chroma
+vector (length $12$) for every music event (called ``segment''), and a
+segmentation of the song into beats and bars. Beats may span or 
+subdivide segments; bars span multiple beats.
+\item Averaging per-segment chroma over beat times results in a
+beat-synchronous chroma feature representation.
+\item Dataset size: $43,000$ songs.
+\end{list}
+
+\textbf{Beat-Chroma Patches.}
+\begin{list}{\labelitemi}{\leftmargin=2em \labelsep=1em}
+\item We use Echo Nest analysis to break a song into a collection of 
+beat-chroma ``patches'', typically one or two bars in length.
+\item $82\%$ of the bars in our data were 4 beats long.
+\item Other cases: we resample patches to a fixed length of 4 beats.
+\item We rotate patches so that the first row contains the most energy.
+Each patch is normalized independently.
 \end{list}
@@
 \begin{list}{\labelitemi}{\leftmargin=2em \labelsep=1em}
-\item A bullet point
-\item A bullet point
-\item A bullet point
-\item A bullet point
+\item Vector Quantization algorithm \cite{Gersho1991} to cluster beat-chroma
+patches
+\item VQ can be seen as online $k$-means
+\item VQ initialized with $k$ random patches from the data
+\item VQ, although not optimal, scales linearly with
+the number of patches seen
 \end{list}
+
+\begin{minipage}{\columnwidth}
+%\caption{Pseudocode of Vector Quantization}
+\begin{algorithmic}
+\begin{algorithm}[H]
+\STATE$\ell$ learning rate
+\STATE$\{P_n\}$ set of patches
+\STATE$\{C_k\}$ codebook of $K$ codes
+\REQUIRE $0 < \ell \leq 1$
+\FOR{$nIters$}
+\FOR{$p \in \{P_n\}$}
+\STATE{$c \leftarrow \min_{c \in C_k} \mbox{dist}(p,c)$}
+\STATE{$c \leftarrow c + (p - c) * \ell$}
+\ENDFOR
+\ENDFOR
+\RETURN $\{C_k\}$
+\end{algorithm}
+\end{algorithmic}
+\end{minipage}
 
@@
 %%% References
+\begin{small}
 \bibliographystyle{IEEEtran}
 \bibliography{tbm_bib}
+\end{small}
 

### 2010-07-08 Paper/Poster/poster.tex
fixed the algorithm
good orientation

@@
-\documentclass[landscape,a0b,final]{a0poster}
+%\documentclass[landscape,a0b,final]{a0poster}
+\documentclass[portrait,a0b,final]{a0poster}
 
@@
 %%% Unisiegel
-\begin{minipage}[c][9cm][c]{0.1\textwidth}
+\begin{minipage}[c][10cm][c]{0.1\textwidth}
   \begin{center}
@@
 \end{minipage}
-%%% Titel
+%%% Title
 \begin{minipage}[c][9cm][c]{0.78\textwidth}
   \begin{center}
-    {\sc \Huge Clustering Beat-Chroma Patterns in a Large Music Database}\\[10mm]
+    {\sc \Huge Clustering Beat-Chroma Patterns}\\[5mm]
+  {\sc \Huge in a Large Music Database}\\[10mm]
     {\Large Thierry Bertin-Mahieux $^\dagger$, Ron J. Weiss $^\ddagger$ and Daniel P. W. Ellis $^\dagger$\\[3.5mm]
@@
 
-\myfig[0]{code_axis.ps}{.2}
+\myfig[0]{code_axis.ps}{.4}
 
@@
 
-\begin{minipage}{\columnwidth}
+\vspace{1cm}
+
+\makebox[\columnwidth]{ % replace framebox by makebox
+
+\begin{minipage}[l]{.4\columnwidth}
 %\caption{Pseudocode of Vector Quantization}
@@
 \begin{algorithm}[H]
+\vspace{.7cm}
 \STATE$\ell$ learning rate
@@
 \STATE$\{C_k\}$ codebook of $K$ codes
-\REQUIRE $0 < \ell \leq 1$
+%\REQUIRE $0 < \ell \leq 1$
 \FOR{$nIters$}
@@
 \RETURN $\{C_k\}$
+\vspace{.7cm}
 \end{algorithm}
@@
 
+\hspace{1.5cm}
+\begin{minipage}[r]{.4\columnwidth}
+Intuitively:
+\begin{list}{\labelitemi}{\leftmargin=1em \labelsep=1em}
+\item For each new patch, find the closest codeword.
+\item Bring that codewords closer to the patch by some learning
+rate.
+\item Iterate.
+\end{list}
+\end{minipage}
+
+}
+
 %%% Section 4
-\vspace{0cm}
+\vspace{2cm}
 \begin{center}
@@
 
-\vspace{4.0cm}
+\vspace{1.0cm}
 
-\begin{minipage}[c][9cm][c]{\columnwidth}
+\begin{minipage}[c]{\columnwidth}
 
@@
 
-\begin{wrapfigure}{r}{0.5\textwidth}
+\begin{wrapfigure}{r}{0.6\textwidth}
   \vspace{-20pt}
   \begin{center}
-    \includegraphics[width=0.48\textwidth]{conf_mat_per_artist.ps}
+    \includegraphics[width=0.58\textwidth]{conf_mat_per_artist.ps}
   \end{center}

### 2010-07-20 Paper/Poster/poster.tex
new figs
nyu logo added

@@
 %%% Unisiegel
-\begin{minipage}[c][10cm][c]{0.1\textwidth}
+\begin{minipage}[c][10cm][c]{0.2\textwidth}
   \begin{center}
-    \includegraphics[width=7cm,angle=0]{labrosa-logo.eps}
+    \includegraphics[width=7cm,angle=0]{labrosa-logo.eps} \\
+    \vspace{1cm}
+    \includegraphics[width=7cm,angle=0]{seas-logo.eps}
   \end{center}
 \end{minipage}
+%\begin{minipage}[c][9cm][c]{0.1\textwidth}
+%  \begin{center}
+%    \includegraphics[width=7cm,angle=0]{seas-logo.eps}
+%  \end{center}
+%\end{minipage}
 %%% Title
-\begin{minipage}[c][9cm][c]{0.78\textwidth}
+\begin{minipage}[c][9cm][c]{0.58\textwidth}
   \begin{center}
@@
 %%% GK-Logo
-\begin{minipage}[c][9cm][c]{0.1\textwidth}
+\begin{minipage}[c][7cm][c]{0.2\textwidth}
   \begin{center}
-    \includegraphics[width=7cm,angle=0]{seas-logo.eps}
+    \includegraphics[width=10cm,angle=0]{MARL_logo_high_res.eps} \\
+    \vspace{1.5cm}
+    \includegraphics[width=8cm,angle=0]{nyu_music_tech_logo_2_2_bw.eps}
   \end{center}
@@
 
+\myfig[0]{song_encoded.ps}{.9}
+
+\myfig[0]{codes_8bars.ps}{.8}
+
 % force a new column
@@
 \vspace{1.0cm}
+
+Boring graphs on distortion, codeword size, etc, can be seen in the paper.
+
+\myfig[0]{codes1.ps}{.8}
+
+\myfig[0]{close_patterns1.ps}{.8}
+
+\myfig[0]{close_patterns_diff.ps}{.8}
+
+\myfig[0]{codes_lle.ps}{.8}
 

### 2010-07-21 Paper/Poster/poster.tex
getting close to be ready

@@
     \includegraphics[width=7cm,angle=0]{labrosa-logo.eps} \\
-    \vspace{1cm}
+    \vspace{.5cm}
     \includegraphics[width=7cm,angle=0]{seas-logo.eps}
@@
     \includegraphics[width=10cm,angle=0]{MARL_logo_high_res.eps} \\
-    \vspace{1.5cm}
+    \vspace{1.8cm}
     \includegraphics[width=8cm,angle=0]{nyu_music_tech_logo_2_2_bw.eps}
@@
 \begin{list}{\labelitemi}{\leftmargin=2em \labelsep=1em}
-\item Goal: identify meaningful information
+\item \textbf{Goal}: identify meaningful information
 about the musical structure represented in the entire database by
 examining individual entries in this codebook.
-\item Method: identify common patterns in beat-synchronous
+\item \textbf{Method}: identify common patterns in beat-synchronous
 chromagrams by learning codebooks from a large set of examples.
@@
 \item Prior work: ``shingles'' of \cite{Casey2007}, 
-beat-synchronous analysis to indeitfy the chorus by \cite{Bartsch2001},
+beat-synchronous analysis to identify the chorus by \cite{Bartsch2001},
 and cover recognition by \cite{Ellis2007a}.
@@
 
-Boring graphs on distortion, codeword size, etc, can be seen in the paper.
+Additional graphs on distortion, codeword size, etc, can be seen in the paper.
+As expected, a larger codebook improves distortion, longer codewords are more
+difficult to learn, and a larger training set improves the results.
 
@@
 
+% force a new column
+\newpage
+
+\mbox{} % dumb, but need something before vspace
+\vspace{-1cm}
 \myfig[0]{codes_lle.ps}{.8}
 
-% force a new column
-\newpage
+% figure caption
+\begin{center}
+\begin{minipage}[c]{.5\columnwidth}
+\begin{small}
+Possible visualization of the codebook using locally linear
+embedding (LLE).
+\end{small}
+\end{minipage}
+\end{center}
 
@@
 
-\vspace{2cm}
+\vspace{1.5cm}
 
-\begin{minipage}[c][9cm][c]{\columnwidth}
-  %\begin{center}
-\textbf{Bar alignment task}.
-Since the clustering described is based
-on the segmentation of the signal in to bars, 
-the codewords should contain information related to bar
-alignment, such as the presence of a strong beat on the first beat.\\
-  %\end{center}
-\end{minipage}
-\begin{minipage}[c][9cm][c]{0.21\textwidth}
+\makebox[\columnwidth]{ % replace framebox by makebox for no edges
+
+\begin{minipage}[l]{0.21\columnwidth}
   \begin{center}
@@
       \end{tabular}
-      %\mytablecaption{\small{
-      %    Bar alignment experiment:
-      %    offsets relative to ground-truth 4-beat bar boundaries 
-      %    that gave minimum distortion encodings from the bar-aligned codebook.
-      %}}
-      \label{tab:offset}
   \end{center}
@@
 
-\begin{list}{\labelitemi}{\leftmargin=2em \labelsep=1em}
-\item A bullet point
-\item A bullet point
-\item A bullet point
-\item A bullet point
-\end{list}
+\hspace{5cm}
+
+\begin{minipage}[r]{.5\columnwidth}
+\textbf{Bar alignment task}.
+Since the clustering described is based
+on the segmentation of the signal in to bars, 
+the codewords should contain information related to bar
+alignment, such as the presence of a strong beat on the first beat.
+It is the case $62\%$ of the times, see paper for details.
+\end{minipage}
+
+} % end framebox / makebox
+
+\vspace{1.5cm}
 
@@
 \begin{list}{\labelitemi}{\leftmargin=2em \labelsep=1em}
-\item A bullet point
-\item A bullet point
+\item Practical method for large-scale clustering of tonal patterns.
+\item Ways to inspect the resulting codebook.
+\item Links to existing MIR tasks.
+\item \textbf{Python code available.}
 \end{list}
@@
 \begin{flushright}
+Acknowledgements: thanks to G. Grindlay for
+helpful comments. TBM is supported by a
+NSERC scholarship. This material is based upon work
+supported by IMLS grant LG-06-08-0073-08 and by NSF
+grant IIS-0713334. \\
+%Any opinions, findings and conclu-
+%sions or recommendations expressed in this material are
+%those of the authors and do not necessarily reflect the views
+%of the sponsors.\\
 Poster presented at ISMIR 2010 (Utrecht) for the paper:

### 2010-07-21 Paper/Poster/poster.tex
getting close to be ready

@@
 
-\myfig[0]{code_axis.ps}{.4}
+\vspace{-.3cm}
+\makebox[\columnwidth]{ % replace framebox by makebox
+\hspace{-4cm}
+\begin{minipage}[l]{.78\columnwidth}
+\myfig[0]{code_axis.ps}{.45}
+\end{minipage}
+\hspace{-3cm}
+\begin{minipage}[l]{.4\columnwidth}
+Typical learned codeword spanning $2$ bars, each bar
+considered to be $4$ beats. Codebook size was $200$.
+\end{minipage}
+} % end framebox / makebox
+\vspace{.5cm}
 
@@
 
+\vspace{.5cm}
 \textbf{Beat-Chroma Patches.}
@@
 
-\myfig[0]{song_encoded.ps}{.9}
+\myfig[0]{song_encoded.ps}{.7}
+% figure caption
+\begin{center}
+\begin{minipage}[c]{.7\columnwidth}
+\begin{small}
+Exemple of an encoded song (Good Day Sunshine
+by \textit{The Beatles}) using a codebook. Codebook size:
+$200$, codewords: $2$ bars.
+\end{small}
+\end{minipage}
+\end{center}
 
-\myfig[0]{codes_8bars.ps}{.8}
+\myfig[0]{codes_8bars.ps}{.7}
+% figure caption
+\begin{center}
+\begin{minipage}[c]{.7\columnwidth}
+\begin{small}
+Longer codewords spanning $8$ bars, randomly selected from
+a $100$-entry codebook. Most patterns consist of sustained notes 
+or chords, but code 0 shows one-bar alternations between 
+two chords, and code 4 contains two cycles of a
+$1\rightarrow1\rightarrow1\rightarrow2$ progression.
+
+\end{small}
+\end{minipage}
+\end{center}
 
@@
 \begin{center}
-\begin{minipage}[c]{.5\columnwidth}
+\begin{minipage}[c]{.7\columnwidth}
 \begin{small}

### 2010-07-21 Paper/Poster/poster.tex
done maybe...

@@
 \end{minipage}
-\hspace{-3cm}
+\hspace{-3.5cm}
 \begin{minipage}[l]{.4\columnwidth}
 Typical learned codeword spanning $2$ bars, each bar
-considered to be $4$ beats. Codebook size was $200$.
+considered to be $4$ beats. Codebook size was $200$.\\
+
+A codewords can also be seen as the centroid of a cluster.
 \end{minipage}
@@
 \vspace{1.0cm}
-
 Additional graphs on distortion, codeword size, etc, can be seen in the paper.
 As expected, a larger codebook improves distortion, longer codewords are more
-difficult to learn, and a larger training set improves the results.
+difficult to learn, and a larger training set improves the results.\\
+
+Below are samples from actual codebooks and centroids.
 
 \myfig[0]{codes1.ps}{.8}
+% figure caption
+\begin{center}
+\begin{minipage}[c]{.7\columnwidth}
+\begin{small}
+The 25 codes most commonly used. 
+Codes are from a $200$-entry codebook trained
+on $2$ bars, $12 × 8$ patches. The proportion of patches accounted
+for by each pattern is shown in parentheses.
+\end{small}
+\end{minipage}
+\end{center}
 
 \myfig[0]{close_patterns1.ps}{.8}
+% figure caption
+\begin{center}
+\begin{minipage}[c]{.7\columnwidth}
+\begin{small}
+Cluster around first centroid (in \textit{introduction} section).
+Random selection from a cluster of size $639$. 
+This gives an intuition of the
+variance in one cluster.
+\end{small}
+\end{minipage}
+\end{center}
 
 \myfig[0]{close_patterns_diff.ps}{.8}
+% figure caption
+\begin{center}
+\begin{minipage}[c]{.7\columnwidth}
+\begin{small}
+First three patterns from figure above ($2$nd row) presented
+with the centroid ($1$st row) and the absolute difference between 
+both ($3$ row).
+\end{small}
+\end{minipage}
+\end{center}
 
@@
 
+% authors
+\vspace{-2cm}
+\myfig[0]{authors.eps}{.8}
+% figure caption
+\vspace{-2cm}
+\begin{center}
+\begin{minipage}[c]{.7\columnwidth}
+\begin{center}
+\begin{small}
+Looking forward to discuss our work with you!
+\end{small}
+\end{center}
+\end{minipage}
+\end{center}
+
 \end{multicols}

### 2010-07-22 Paper/Poster/poster.tex
getting there, some feedback from Ron included

@@
 
-\pbox{0.957\textwidth}{}{linewidth=2mm,framearc=0.3,linecolor=lightred,fillstyle=gradient,gradangle=0,gradbegin=white,gradend=whitepink,gradmidpoint=1.0,framesep=1em}{
+\makebox[0.957\textwidth]{
 
 %%% Unisiegel
-\begin{minipage}[c][10cm][c]{0.2\textwidth}
+\begin{minipage}[c][7cm][c]{0.2\textwidth}
   \begin{center}
@@
 %\end{minipage}
+
+\pbox{0.57\textwidth}{}{linewidth=2mm,framearc=0.3,linecolor=lightred,fillstyle=gradient,gradangle=0,gradbegin=white,gradend=whitepink,gradmidpoint=1.0,framesep=1em}{
 %%% Title
-\begin{minipage}[c][9cm][c]{0.58\textwidth}
+\begin{center}
+\begin{minipage}[c][10cm][c]{0.90\textwidth}
   \begin{center}
@@
      \large $\dagger$ LabROSA, Dept. of Electrical Engineering, Columbia University \\
-     \large $\ddagger$ MARL, New York University \\
+     \large $\ddagger$ Music and Audio Research Laboratory, New York University \\
     \tt{\{thierry,ronw,dpwe\}@ee.columbia.edu}}
@@
 \end{minipage}
+\end{center}
+}
+
 %%% GK-Logo
@@
     \vspace{1.8cm}
-    \includegraphics[width=8cm,angle=0]{nyu_music_tech_logo_2_2_bw.eps}
+    %\includegraphics[width=8cm,angle=0]{nyu_music_tech_logo_2_2_bw.eps}
+    \includegraphics[width=14cm,angle=0]{NYU_logo.eps}
   \end{center}
@@
     \begin{center}
-      \large Introduction
+      \LARGE Introduction
     \end{center}}
@@
 \begin{list}{\labelitemi}{\leftmargin=2em \labelsep=1em}
-\item Availability of very large collections of music audio:
-can we infer anything about the underlying structure and common
-features of e.g. commercial pop music?
+\item Can we infer anything about the underlying structure and common
+features of commercial pop music by analyzing very large collections
+of music audio?
 \item Our interest: tonal content of the music -- i.e. the
@@
 \begin{minipage}[l]{.78\columnwidth}
-\myfig[0]{code_axis.ps}{.45}
+\myfig[0]{code_axis.ps}{.40}
 \end{minipage}
@@
 \begin{minipage}[l]{.4\columnwidth}
-Typical learned codeword spanning $2$ bars, each bar
+Typical %learned 
+codeword spanning $2$ bars, each bar
 considered to be $4$ beats. Codebook size was $200$.\\
@@
     \begin{center}
-      \large Audio Features - Echo Nest
+      \LARGE Audio Features - Echo Nest
     \end{center}}
@@
 vector (length $12$) for every music event (called ``segment''), and a
-segmentation of the song into beats and bars. Beats may span or 
-subdivide segments; bars span multiple beats.
+segmentation of the song into beats and bars.
 \item Averaging per-segment chroma over beat times results in a
@@
 \item We rotate patches so that the first row contains the most energy.
-Each patch is normalized independently.
+This is analogous to key normalization, but
+each patch is normalized independently.
 \end{list}
@@
     \begin{center}
-      \large Vector Quantization
+      \LARGE Vector Quantization
     \end{center}}
@@
 patches
-\item VQ can be seen as online $k$-means
-\item VQ initialized with $k$ random patches from the data
-\item VQ, although not optimal, scales linearly with
+\item Can be seen as online $k$-means
+\item Initialized with $k$ random patches from the data
+\item Although not optimal, scales linearly with
 the number of patches seen
@@
     \begin{center}
-      \large Pattern Analysis
+      \LARGE Pattern Analysis
     \end{center}}
@@
 
+\myfig[0]{close_patterns1.ps}{.8}
+% figure caption
+\begin{center}
+\begin{minipage}[c]{.7\columnwidth}
+\begin{small}
+Cluster around first centroid (in \textit{introduction} section).
+Random selection from a cluster of size $639$. 
+This gives an intuition of the
+variance in one cluster.
+\end{small}
+\end{minipage}
+\end{center}
+
 \myfig[0]{codes1.ps}{.8}
@@
 for by each pattern is shown in parentheses.
-\end{small}
-\end{minipage}
-\end{center}
-
-\myfig[0]{close_patterns1.ps}{.8}
-% figure caption
-\begin{center}
-\begin{minipage}[c]{.7\columnwidth}
-\begin{small}
-Cluster around first centroid (in \textit{introduction} section).
-Random selection from a cluster of size $639$. 
-This gives an intuition of the
-variance in one cluster.
 \end{small}
@@
     \begin{center}
-      \large Experiments
+      \LARGE Experiments
     \end{center}}
@@
 \textbf{Artist recognition task}.
-We use the {\it artist20} data set: $1402$ songs from $20$ artists, 
-mostly rock and pop of different subgenres.  
-Previously published results using GMMs on MFCC features achieve an 
-accuracy of $59\%$, whereas using only chroma as a representation yields an
-accuracy of $33\%$  \cite{Ellis2007}.
-
-We get an accuracy of $\mathbf{23.4\%}$, random baseline 
-is around $5\%$. 
-The confusion matrix is shown here, note 
-that certain artists are recognized at an accuracy far above the average.
+\begin{list}{\labelitemi}{\leftmargin=2em \labelsep=1em}
+\item {\it artist20} data set: $1402$ songs from $20$ artists.
+\item Baseline: GMM on MFCC ($59\%$), GMM on chromas ($33\%$),
+random ($5\%$).
+\item We get an accuracy of $\mathbf{23.4\%}$.
+\end{list}
+%We use the {\it artist20} data set: $1402$ songs from $20$ artists, 
+%mostly rock and pop of different subgenres.  
+%Previously published results using GMMs on MFCC features achieve an 
+%accuracy of $59\%$, whereas using only chroma as a representation yields an
+%accuracy of $33\%$  \cite{Ellis2007}.
+%
+%We get an accuracy of $\mathbf{23.4\%}$, random baseline 
+%is around $5\%$. 
+%The confusion matrix is shown here, note 
+%that certain artists are recognized at an accuracy far above the average.
 
@@
 \textbf{Bar alignment task}.
-Since the clustering described is based
-on the segmentation of the signal in to bars, 
-the codewords should contain information related to bar
-alignment, such as the presence of a strong beat on the first beat.
-It is the case $62\%$ of the times, see paper for details.
+\begin{list}{\labelitemi}{\leftmargin=2em \labelsep=1em}
+\item Clustering based on the signal segmented in bars.
+\item Codewords should contain bar information (e.g. a strong
+first beat).
+\item Our encoding find the right alignment $62\%$ of the time.
+\end{list}
+%Since the clustering described is based
+%on the segmentation of the signal in to bars, 
+%the codewords should contain information related to bar
+%alignment, such as the presence of a strong beat on the first beat.
+%It is the case $62\%$ of the times, see paper for details.
 \end{minipage}
@@
     \begin{center}
-      \large Conclusion
+      \LARGE Conclusion
     \end{center}}
@@
 %%% References
-\begin{small}
+%\begin{small}
 \bibliographystyle{IEEEtran}
 \bibliography{tbm_bib}
-\end{small}
+%\end{small}
 
+\iffalse
 % authors
@@
 \end{center}
+\fi
 

### 2010-07-26 Paper/Poster/poster.tex
getting there, some feedback from Ron included
final

@@
 \usepackage{algorithm,algorithmic}
+%\usepackage{times} % recommended by Ron
 
@@
 \begin{list}{\labelitemi}{\leftmargin=2em \labelsep=1em}
+\item Availability of very large collections of music.
 \item Can we infer anything about the underlying structure and common
-features of commercial pop music by analyzing very large collections
-of music audio?
+features of pop music?
 \item Our interest: tonal content of the music -- i.e. the
@@
 \item Beat-synchronous chromagrams: rich enough
-to generate musically-relevant results, simplified enough to
-abstract away instrumentation and other stylistic details.
+to generate musically-relevant results,
+abstract away instrumentation and stylistic details.
 \end{list}
@@
 \item \textbf{Goal}: identify meaningful information
-about the musical structure represented in the entire database by
-examining individual entries in this codebook.
+about the musical structure
 \item \textbf{Method}: identify common patterns in beat-synchronous
 chromagrams by learning codebooks from a large set of examples.
-Individual codewords consist of short beat-chroma patches of
+\item Entire database
+represented by individual entries in this codebook.
+\item Individual codewords consist of short beat-chroma patches of
 between 1 and 8 beats, optionally aligned to bar boundaries.
@@
 \item For any song, EchoNest provides a chroma
-vector (length $12$) for every music event (called ``segment''), and a
+vector (length $12$) for every music event (called ``segment'')
+\item Also provides a
 segmentation of the song into beats and bars.
@@
 \item Other cases: we resample patches to a fixed length of 4 beats.
-\item We rotate patches so that the first row contains the most energy.
-This is analogous to key normalization, but
-each patch is normalized independently.
+\item We rotate patches so that the first row contains the most energy
+(this is analogous to key normalization).
 \end{list}
@@
 
-\myfig[0]{song_encoded.ps}{.7}
+\myfig[0]{song_encoded.ps}{.68}
 % figure caption
@@
 \begin{small}
-Exemple of an encoded song (Good Day Sunshine
+Example of an \textbf{encoded song} (Good Day Sunshine
 by \textit{The Beatles}) using a codebook. Codebook size:
@@
 
-\myfig[0]{codes_8bars.ps}{.7}
+\myfig[0]{codes_8bars.ps}{.68}
 % figure caption
@@
 \begin{small}
-Longer codewords spanning $8$ bars, randomly selected from
+Longer \textbf{codewords spanning $\mathbf{8}$ bars}, randomly selected from
 a $100$-entry codebook. Most patterns consist of sustained notes 
@@
 
-Below are samples from actual codebooks and centroids.
+Below are samples from \textbf{actual codebooks and centroids}.
 
@@
 \begin{small}
-Cluster around first centroid (in \textit{introduction} section).
+\textbf{Cluster around first centroid} (in \textit{introduction} section).
 Random selection from a cluster of size $639$. 
@@
 \begin{small}
-The 25 codes most commonly used. 
+The 25 \textbf{codes most commonly used}. 
 Codes are from a $200$-entry codebook trained
@@
 First three patterns from figure above ($2$nd row) presented
-with the centroid ($1$st row) and the absolute difference between 
-both ($3$ row).
+with the centroid ($1$st row) and the \textbf{absolute difference} between 
+patterns and centroid ($3$ row).
 \end{small}
@@
 \vspace{-1cm}
-\myfig[0]{codes_lle.ps}{.8}
+\myfig[0]{codes_lle.ps}{.9}
 
@@
 
-\begin{wrapfigure}{r}{0.6\textwidth}
-  \vspace{-20pt}
-  \begin{center}
-    \includegraphics[width=0.58\textwidth]{conf_mat_per_artist.ps}
-  \end{center}
-  \vspace{-20pt}
-\end{wrapfigure}
+%\begin{wrapfigure}{r}{0.5\textwidth}
+%  \vspace{-20pt}
+%  \begin{center}
+%    \hspace{-1cm}
+%    \includegraphics[width=0.52\textwidth]{conf_mat_per_artist.ps}
+%  \end{center}
+%  \vspace{-20pt}
+%\end{wrapfigure}
 
@@
 \item {\it artist20} data set: $1402$ songs from $20$ artists.
-\item Baseline: GMM on MFCC ($59\%$), GMM on chromas ($33\%$),
-random ($5\%$).
+\item Baseline: GMM - MFCC $59\%$, GMM - chromas $33\%$,
+random $5\%$.
 \item We get an accuracy of $\mathbf{23.4\%}$.
 \end{list}
+
+\myfig[0]{conf_mat_per_artist.ps}{.55}
+
 %We use the {\it artist20} data set: $1402$ songs from $20$ artists, 
@@
 % FOOTNOTE
-\vspace{1cm}
+\vspace{1.5cm}
 \begin{minipage}{\textwidth}
